### 2020-06-07 outputs/documents/paper.Rmd
First commit

---
output: 
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    toc: false
    fig_caption: true
    latex_engine: pdflatex
    template: templates/svm-latex-ms.tex
bibliography: "references.bib"
header-includes:
  -  \usepackage{hyperref}
biblio-style: apalike
title: "Consistency scores in text data"
thanks: "Our code and datasets are available at: X. Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca."
author:
- name: Rohan Alexander
  affiliation: University of Toronto
- name: Ke-Li Chiu
  affiliation: University of Toronto  
abstract: "..."
keywords: "..."
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
# spacing: double
endnote: no
graphics: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

When we think about a data science project, we may like to think that our job is to 'let the data speak'. But this is rarely the case in practise. Datasets can have errors, be biased, incomplete, or messy. In any case, it is the underlying statistical process of which the dataset is an artifact that is typically of interest. In order to use statistical models to understand that process we often need to prepare the dataset in some way. This is particularly the case when working with text. However, this process requires many decisions. Should we correct obvious errors? What about slightly-less-obvious errors? To what extent have we introduced new errors? Have we made decisions that have affected, or even driven, our results?

In this paper we introduce the concept of consistency for a text corpus. Consistency refers to the proportion of words that are able to be predicted by a trained model based on the previous words and surrounding context. Further, we define internal consistency as when the model is trained on the corpus itself, and external consistency as when the model is trained on a more general corpus. Together, these concepts provide a guide to the cleanliness and consistency of a text dataset. This can be important when deciding whether a dataset is fit for purpose; when carrying out data cleaning and preparation tasks; and as a comparison between datasets.

To provide an example, consider the sentence, 'the cat in the...'. A child who has read this book could tell you that the next word should be 'hat'. Hence if the sentence was actually 'the cat in the bat', then that child would know something was likely wrong. A consistency score would likely be lower than if the sentence were 'the cat in the hat'. After the researcher corrects this error, a consistency score would likely increase. By following how the consistency scores change during the data preparation and cleaning stages the researcher can better understand the effect of the changes. Including consistency scores when corpuses are shared allows researchers to be more transparent about their corpus. And finally, the use of consistency scores allows for automation in the cleaning process.

We apply our approach to X (one option is a Hansard, but it's really big, so yeah, I don't really want to do that). Additionally, we construct a Shiny app that computes internal and external consistency scores for smaller corpuses and allows the researcher to make changes and see how it updates.

The remainder of our paper is structured as follows...

# Background

A typical data science workflow involves

Internal and external consistency

# Stylized example

# Application

# Discussion

Internal validity vs external- one is their own words the other is a general set of words.

In the same way that precision and recall provide important measures…

# References

### 2020-07-06 outputs/documents/paper.Rmd
Add jane eyre example

@@
 
+As a stylized example, let's consider the following actual paragraph from Jane Eyre, by Charlotte Bronte:
+
+> There was no possibility of taking a walk that day.
+We had been wandering, indeed, in the leafless shrubbery an hour in the morning; but since dinner (Mrs.
+Reed, when there was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, that further out-door exercise was now out of the question.
+
+Let's pretend that this text had been created from optical character recognition and that it had the following errors: some 'h' were replaced with 'b'; and some 'd' have been replaced with 'al':
+
+> There was no possibility of taking a walk that day.
+We had been wandering, indeed, in tbe leafless shrubbery an hour in tbe morning; but since dinner (Mrs.
+Reeal, when tbere was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, tbat further out-door exercise was now out of the question.
+
+Assume a model that is trained to perfectly forecast the next word in Jane Eyre.
+For this fragment there are 62 words, comprising 5 errors and 57 correct words.
+So the internal consistency score of this fragment would be: 57/62 = 0.919.
+When we recognise and correct the errors, this consistency score would increase to 1.
+
+Similarly, assume a model that is trained on an external data source.
+This means that it will recognise the the cases where some 'h' were replaced with 'b', but not recognise that 'Reed' has become 'Reeal'.
+Hence, the external consistency score would be 58/62 = 0.935.
+
+# Models
+
+Various models can be used....
+
+## ngrams
+
+## Something
+
+## Soemthing else
+
+# Data
+
+Various data can be used....
+
 # Application

### 2020-07-08 outputs/documents/paper.Rmd
uploaded literatures
more literatures and one-liners of currently reading articles
added literature about generative pre-training

@@
 
+### Article: Improving Language Understanding by Generative Pre-Training (2018)
+Introduction of a semi-supervised approach for language understanding machine learning tasks. The approach is the combinaiton of unsupervised pre-training and supervised fine-tuning.
+
+OpenAI GPT-3 is
+
 # Stylized example
@@
 
+### Article: Class-Based n-gram Models of Natural Language (1992)
+Syntax-based (grammartical) and semantic-based (sensible) word classifications in n-gram models.
+
+### Article: A Statistical Approach to Automatic OCR Error Correction in Context (1996)
+Context-sensitive correction system to both non-word and real-word errors based on n-gram models applied to OCR postprocessing.
+
+### Article: Unsupervised Learning of Sentence Embeddings using Compositional n-Gram Features (2017)
+Embedding of sentences (sequence of words with semantic representations) compositional n-gram features.
+
 ## Something

### 2020-07-08 outputs/documents/paper.Rmd
write up of GPT 3 WIP

@@
 
-OpenAI GPT-3 is
+### Article: Language Models are Few-Shot Learners (2020)
+Official paper from OpenAI to introduce OpenAI GPT-3. Emphasizes on the "fewer-shots" and "task-agnostic" aspects of the latest langual model compared to its predecessor GPT-2.
+
+In recent years, pre-trained language models have tremendous contributions in advancing a wide range of diverse natural language processing (NLP) tasks such as reading comprehension, text generation, and many others.
+The pre-trained models enables (citation).
+As the pre-trainde models remove the need for researchers and technologists to train models from scratch, they promote faster growth and advancement int he field of NLP.
+OpenAI GPT-3 is the latest generative pre-trained language model released by OpenAI in 2020.
+It has 175 billion parameters and is trained to [do that and that].
+GPT-2, the predecessor of GPT-3, employs a semi-supervised approach that combines unsupervised pre-training and supervised fine-tuning where task-specific data set and task-specific fine-tuning were still needed in achieving these tasks until the birth of GPT-3 (citation).
+GPT-3 removes such needs, and its task-agnostic nature enables the accessible creation of cutting-edge NLP applications.
+
+GPT-3 learns like human do — with brief directive and simple instruction...
+
+In our application, we deploy OpenAI GPT-3 to generate text and ...
 

### 2020-07-08 outputs/documents/paper.Rmd
ngram writeup WIP

@@
 
+Without exception, all human languages have some words that co-occur more frequently with others.
+Under this assumption, we can regard the production of English text as a set of conditional probabilities, written as Pr($w_{k}$ | $w^{k-1}_{1}$), where $w_{k}$ is the predicted word and $w^{k-1}_{1}$ is the history of the word occuring in a sequence (citation).
+In other word, the generation of prediction $w_{k}$ is based on the history $w^{k-1}_{1}$.
+[Add more to link to n-gram]
+
+N-gram is a sequence of n words. Taking the first sentence from the Jane Eyre excerpt as an example:
+
+> There was no possibility of...
+
+"There" is an unigram, "There was" is a bigram, "There was no" is a trigram, and "There was no possibility of" is a 5-gram.
+On the other hand, the trigrams of the exerpt are "There was no", "was no possibility" and "no possibility of".
+The overlapping chaining of words reveals the statstical behaviour in human language.
+Moreover, N-gram then enables us to assign a probability to the occurrence of a sequence of words or the probability of a word occurring next in a sequence of words.
+
+[Include N to do more statistics write up]
+
+The application of n-gram is versatile. N-gram is widely applied in text prediction and spelling-correction
+
+> There was no possibility of taking a walk that day
+
+This is useful because...
+
+N-gram is a sequence of words that has n words in it. The n-gram model is widely applied to 
+Since n-gram 
+N-gram is a model that is widely applied to 
+N-gram is under the assumtion that the production of English text can be characterized by a set of conditional probabilities, Pr($w_{k}$ | $w^{k-1}_{1}$), where $w_{k}$ is the prediction and $w^{k-1}_{1}$ is the history (Citation).
+
 ### Article: Class-Based n-gram Models of Natural Language (1992)

### 2020-07-08 outputs/documents/paper.Rmd
Write-up fixes, still WIP

--- 2020-07-08 outputs/documents/paper.Rmd
@@
 
+/// Start of  OpenAI GPT-3 section ///
+
+In recent years, pre-trained language models have tremendous contributions in advancing natural language processing (NLP) tasks such as reading comprehension, text generation, and many others (citation).
+As the pre-trained models remove the need for researchers and technologists to train models from scratch, they promote faster growth and advancement int the field of NLP.
+OpenAI GPT-3 is the latest generative pre-trained language model released by OpenAI in 2020.
+It has 175 billion parameters and is trained to [do that and that].
+GPT-2, the predecessor of GPT-3, employs a semi-supervised approach that combines unsupervised pre-training and supervised fine-tuning where task-specific data set and task-specific fine-tuning were still needed in achieving these tasks until the birth of GPT-3 (citation).
+GPT-3 removes such needs, and its task-agnostic nature enables the accessible creation of cutting-edge NLP applications.
+
+GPT-3 learns like humans do — with brief directives and simple instruction...
+
+In our application, we deploy OpenAI GPT-3 to generate text and ...
+
+/// End of  OpenAI GPT-3 section ///
+
 ### Article: Improving Language Understanding by Generative Pre-Training (2018)
@@
 Official paper from OpenAI to introduce OpenAI GPT-3. Emphasizes on the "fewer-shots" and "task-agnostic" aspects of the latest langual model compared to its predecessor GPT-2.
-
-In recent years, pre-trained language models have tremendous contributions in advancing a wide range of diverse natural language processing (NLP) tasks such as reading comprehension, text generation, and many others.
-The pre-trained models enables (citation).
-As the pre-trainde models remove the need for researchers and technologists to train models from scratch, they promote faster growth and advancement int he field of NLP.
-OpenAI GPT-3 is the latest generative pre-trained language model released by OpenAI in 2020.
-It has 175 billion parameters and is trained to [do that and that].
-GPT-2, the predecessor of GPT-3, employs a semi-supervised approach that combines unsupervised pre-training and supervised fine-tuning where task-specific data set and task-specific fine-tuning were still needed in achieving these tasks until the birth of GPT-3 (citation).
-GPT-3 removes such needs, and its task-agnostic nature enables the accessible creation of cutting-edge NLP applications.
-
-GPT-3 learns like human do — with brief directive and simple instruction...
-
-In our application, we deploy OpenAI GPT-3 to generate text and ...
 
@@
 
+Without exception, all human languages have some words that co-occur more frequently with others.
+Under this assumption, we can regard the production of English text as a set of conditional probabilities, written as Pr($w_{k}$ | $w^{k-1}_{1}$), where $w_{k}$ is the predicted word, and $w^{k-1}_{1}$ is the history of the word occurring in a sequence (citation).
+In other word, the generation of prediction $w_{k}$ is based on the history $w^{k-1}_{1}$.
+[Add a sentence linking to n-gram]
+
+N-gram is a sequence of n words. Taking the imcomplete first sentence from the Jane Eyre excerpt as an example:
+
+> "There was no possibility of..."
+
+"There" is a unigram, "There was" is a bigram, "There was no" is a trigram, and "There was no possibility of" is a 5-gram.
+On the other hand, the trigrams of the excerpt are "There was no," "was no possibility," and "no possibility of." The overlapping chaining of words reveals the statistical behaviour in human language.
+Moreover, N-gram then enables us to assign a probability to the occurrence of a sequence of words or the likelihood of a word occurring next.
+
+[Maybe include N to do more statistics write up]
+
+The application of n-gram is versatile. N-gram is widely applied in text prediction and spelling-correction...
+
 ### Article: Class-Based n-gram Models of Natural Language (1992)

### 2020-07-09 outputs/documents/paper.Rmd
write-ups updated, WIP

@@
 OpenAI GPT-3 is the latest generative pre-trained language model released by OpenAI in 2020.
-It has 175 billion parameters and is trained to [do that and that].
+With 175 billion parameters, GPT-3 is claimed to have the capacity to generate text that ressembles human creation (citation).
 GPT-2, the predecessor of GPT-3, employs a semi-supervised approach that combines unsupervised pre-training and supervised fine-tuning where task-specific data set and task-specific fine-tuning were still needed in achieving these tasks until the birth of GPT-3 (citation).
@@
 
-GPT-3 learns like humans do — with brief directives and simple instruction...
+GPT-3 also learns like humans do — with brief directives and simple instructions...
+
+The task-agnostic feature also permits it to adapt to the style of the data being fed to it (citation).
 
@@
 
+Traditionally, word-correction techniques evaluate errors one by one without considering the context of the surrounding words.
+This was no longer the case in modern correction techniques as statistical language models (SLMs) and feature-based methods have been used for context-sensitive correction (citation).
 Without exception, all human languages have some words that co-occur more frequently with others.
-Under this assumption, we can regard the production of English text as a set of conditional probabilities, written as Pr($w_{k}$ | $w^{k-1}_{1}$), where $w_{k}$ is the predicted word, and $w^{k-1}_{1}$ is the history of the word occurring in a sequence (citation).
+Under this assumption, we can regard the production of English text as a set of conditional probabilities, written as Pr($w_{k}$ | $w^{k-1}_{1}$), where k is the number of words in a sentence, $w_{k}$ is the predicted word, and $w^{k-1}_{1}$ is the history of the word occurring in a sequence (citation).
 In other word, the generation of prediction $w_{k}$ is based on the history $w^{k-1}_{1}$.
-[Add a sentence linking to n-gram]
+This conditional probability is the foundation of an n-gram language model.
 
-N-gram is a sequence of n words. Taking the imcomplete first sentence from the Jane Eyre excerpt as an example:
+An n-gram model is a probabilistic language model that predicts the next word in a sequence of words.
+The $n$ in n-gram represents the number of words in a sequence.
+Taking an incomplete sentence from the Jane Eyre excerpt as an example:
 
-> "There was no possibility of..."
+> "We had been wandering,"
 
-"There" is a unigram, "There was" is a bigram, "There was no" is a trigram, and "There was no possibility of" is a 5-gram.
-On the other hand, the trigrams of the excerpt are "There was no," "was no possibility," and "no possibility of." The overlapping chaining of words reveals the statistical behaviour in human language.
+"We" is a unigram, "We had" is a bigram, "We had been" is a trigram, and "We had been wandering" is a 4-gram.
+On the other hand, the trigrams of the excerpt are "We had been," and "had been wandering." The overlapping chaining of words reveals the statistical behaviour in human language.
 Moreover, N-gram then enables us to assign a probability to the occurrence of a sequence of words or the likelihood of a word occurring next.
+Consider the two sentences: "We had been wandering," and "We had been wangling." The former is likely to be more frequently encountered in a training corpus.
+Thus, the n-gram would assign a higher probability to "wandering" than to "wangling."
 
-[Maybe include N to do more statistics write up]
+By including the $n$ into the previous conditional probability to be 
+Pr($w_{k}$ | $w^{k-1}_{1}$)
 
-The application of n-gram is versatile. N-gram is widely applied in text prediction and spelling-correction...
+The application of n-gram is versatile. N-gram is widely applied in text prediction, spelling-correction, machine translation...
 

### 2020-07-09 outputs/documents/paper.Rmd
write-ups updated, WIP
Add abstract.

@@
   affiliation: University of Toronto  
-abstract: "..."
+abstract: "In this paper we use GPT-3 to clean the text from PDFs by comparing the original text with the generated text.
+We also introduce the notion of a consistency score, which refers to the proporation of text that is unchanged by the model, and is used to monitor changes during an iterative cleaning process.
+We illustrate our process on text from the Canadian Hansard and develop a Shiny application and R package to make our procedure easier for others to use."
 keywords: "..."
@@
 
-By including the $n$ into the previous conditional probability to be 
-Pr($w_{k}$ | $w^{k-1}_{1}$)
+[Statistic notation of n-gram?]
+
+[Character-level n-gram?]
 

### 2020-07-09 shinyapp/app.R
Shiny App setup success

@@ -1,0 +1,61 @@
+library(shiny)
+
+# Hello Shiny example set up code
+
+# Define UI for app that draws a histogram ----
+ui <- fluidPage(
+
+  # App title ----
+  titlePanel("Hello Shiny!"),
+
+  # Sidebar layout with input and output definitions ----
+  sidebarLayout(
+
+    # Sidebar panel for inputs ----
+    sidebarPanel(
+
+      # Input: Slider for the number of bins ----
+      sliderInput(inputId = "bins",
+                  label = "Number of bins:",
+                  min = 1,
+                  max = 50,
+                  value = 30)
+
+    ),
+
+    # Main panel for displaying outputs ----
+    mainPanel(
+
+      # Output: Histogram ----
+      plotOutput(outputId = "distPlot")
+
+    )
+  )
+)
+
+
+# Define server logic required to draw a histogram ----
+server <- function(input, output) {
+
+  # Histogram of the Old Faithful Geyser Data ----
+  # with requested number of bins
+  # This expression that generates a histogram is wrapped in a call
+  # to renderPlot to indicate that:
+  #
+  # 1. It is "reactive" and therefore should be automatically
+  #    re-executed when inputs (input$bins) change
+  # 2. Its output type is a plot
+  output$distPlot <- renderPlot({
+
+    x    <- faithful$waiting
+    bins <- seq(min(x), max(x), length.out = input$bins + 1)
+
+    hist(x, breaks = bins, col = "#75AADB", border = "white",
+         xlab = "Waiting time to next eruption (in mins)",
+         main = "Histogram of waiting times")
+
+  })
+
+}
+
+shinyApp(ui = ui, server = server)

### 2020-07-09 shinyapp/ui.R

@@ -1,12 +1,0 @@
-library(shiny)
-
-# Define UI for miles per gallon application
-shinyUI(pageWithSidebar(
-
-  # Application title
-  headerPanel("Miles Per Gallon"),
-
-  sidebarPanel(),
-
-  mainPanel()
-))

### 2020-07-09 shinyapp/app.R
input box and output box sync

@@ -1,61 +1,0 @@
-library(shiny)
-
-# Hello Shiny example set up code
-
-# Define UI for app that draws a histogram ----
-ui <- fluidPage(
-
-  # App title ----
-  titlePanel("Hello Shiny!"),
-
-  # Sidebar layout with input and output definitions ----
-  sidebarLayout(
-
-    # Sidebar panel for inputs ----
-    sidebarPanel(
-
-      # Input: Slider for the number of bins ----
-      sliderInput(inputId = "bins",
-                  label = "Number of bins:",
-                  min = 1,
-                  max = 50,
-                  value = 30)
-
-    ),
-
-    # Main panel for displaying outputs ----
-    mainPanel(
-
-      # Output: Histogram ----
-      plotOutput(outputId = "distPlot")
-
-    )
-  )
-)
-
-
-# Define server logic required to draw a histogram ----
-server <- function(input, output) {
-
-  # Histogram of the Old Faithful Geyser Data ----
-  # with requested number of bins
-  # This expression that generates a histogram is wrapped in a call
-  # to renderPlot to indicate that:
-  #
-  # 1. It is "reactive" and therefore should be automatically
-  #    re-executed when inputs (input$bins) change
-  # 2. Its output type is a plot
-  output$distPlot <- renderPlot({
-
-    x    <- faithful$waiting
-    bins <- seq(min(x), max(x), length.out = input$bins + 1)
-
-    hist(x, breaks = bins, col = "#75AADB", border = "white",
-         xlab = "Waiting time to next eruption (in mins)",
-         main = "Histogram of waiting times")
-
-  })
-
-}
-
-shinyApp(ui = ui, server = server)

### 2020-07-09 shinyapp/server.R

@@ -1,0 +1,14 @@
+function(input, output, session) {
+
+  output$outputText <- renderText({
+    as.character(input$my_text)
+  })
+
+  observe({
+    # Run whenever clear button is pressed
+    input$clear
+
+    # Send an update to my_text, resetting its value
+    updateTextInput(session, "my_text", value = "")
+  })
+}

### 2020-07-09 shinyapp/text-input.R

@@ -1,0 +1,31 @@
+# This function generates the client-side HTML for a URL input
+textInput <- function(inputId, label, value = "") {
+  tagList(
+    # This makes web page load the JS file in the HTML head.
+    # The call to singleton ensures it's only included once
+    # in a page.
+    #shiny::singleton(
+      #shiny::tags$head(
+        #shiny::tags$script(src = "url-input-binding.js")
+      #)
+    #),
+    shiny::tags$label(label, `for` = inputId),
+    shiny::tags$input(id = inputId, type = "url", value = value)
+  )
+}
+
+
+# Send an update message to a URL input on the client.
+# This update message can change the value and/or label.
+updateTextInput <- function(session, inputId,
+                           label = NULL, value = NULL) {
+  message <- dropNulls(list(label = label, value = value))
+  session$sendInputMessage(inputId, message)
+}
+
+
+# Given a vector or list, drop all the NULL items in it
+dropNulls <- function(x) {
+  x[!vapply(x, is.null, FUN.VALUE=logical(1))]
+}
+

### 2020-07-09 shinyapp/ui.R

@@ -1,0 +1,15 @@
+source("text-input.R")
+
+fluidPage(
+  titlePanel("Text Consistency Score"),
+
+  fluidRow(
+    column(6, wellPanel(
+      textInput("my_text", label = h3("Text input"), value = ""),
+      actionButton("clear", "Clear")
+    )),
+    column(6, wellPanel(
+      verbatimTextOutput("outputText")
+    ))
+  )
+)

### 2020-07-10 shinyapp/server.R
hardcoded text correction

@@ -1,14 +1,45 @@
+source("text-input.R")
+
+# tidyverse is needed for hardcoding string replacement, may be removed once we move to next phase
+library(tidyverse)
+
 function(input, output, session) {
-
+  # Fucntion that displays input text when get result button is clicked
+  getText <- eventReactive(input$getResult, {
+    # Hardcode text correction by replacing typos manually
+    str_replace_all(input$myText, c("Reeal" = "Reed", "tbere" = "there", "tbat" = "that"))
+  })
+  # Displays input text
   output$outputText <- renderText({
-    as.character(input$my_text)
+    getText()
   })
 
+  # Fucntion that displays scores when get result button is clicked
+    getScores <- eventReactive(input$getResult, {
+    "Internal consistency score: 0.919 <br> External consistency score: 0.956"
+  })
+  # Displays scores
+  output$outputScores <- renderText({
+    getScores()
+  })
+
+  # Disable get result button and clear button when there is no text data
   observe({
-    # Run whenever clear button is pressed
-    input$clear
-
-    # Send an update to my_text, resetting its value
-    updateTextInput(session, "my_text", value = "")
+    toggleState(id = "getResult", condition = nchar(input$myText) > 0)
+    toggleState(id = "clear", condition = nchar(input$myText) > 0)
   })
+
+  # Load demo data when the corresponding button is clicked
+  observe({
+    input$loadDemo
+    updateTextInput(session, "myText", value = demoInput)
+  })
+
+  # Clear text input area
+  observe({
+    input$clear
+    updateTextInput(session, "myText", value = "")
+  })
+
+
 }

### 2020-07-10 shinyapp/ui.R

@@ -1,15 +1,25 @@
-source("text-input.R")
+library(shinythemes)
+library(shinyjs)
 
 fluidPage(
+  useShinyjs(),
+  theme = shinytheme("cosmo"),
   titlePanel("Text Consistency Score"),
 
   fluidRow(
     column(6, wellPanel(
-      textInput("my_text", label = h3("Text input"), value = ""),
-      actionButton("clear", "Clear")
+      h4("Place your data here"),
+      actionButton("loadDemo", "Load Demo Data", class = "btn-sm btn-primary"),
+      textAreaInput("myText", "", value = "", height = "500px"),
+      actionButton("getResult", "Get Result", class = "btn-primary"),
+      actionButton("clear", "Clear", class = "btn-outline-secondary"),
+
     )),
     column(6, wellPanel(
-      verbatimTextOutput("outputText")
+      h4("See your result here"),
+      tableOutput("outputText"),
+      hr(),
+      tableOutput("outputScores"),
     ))
   )
 )

### 2020-07-10 outputs/documents/paper.Rmd

--- 2020-07-09 outputs/documents/paper.Rmd
@@
 title: "Consistency scores in text data"
-thanks: "Our code and datasets are available at: X. Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca."
+thanks: "A Shiny app for interactive and small-scale examples is available: HERE.
+An R package for larger scale applications is available: HERE.
+Our code and datasets are available: HERE.
+Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca."
 author:
@@
   affiliation: University of Toronto  
-abstract: "..."
-keywords: "..."
+abstract: "In this paper we introduce a process to clean the text extracted from PDFs using GPT-3.
+Our approach compares originally extracted text with the text generated from GPT-3 using earlier text as stimulus.
+To guide this process, we introduce the notion of a consistency score, which refers to the proporation of text that is unchanged by the model.
+This is used to monitor changes during the cleaning process, and to compare the messiness of different texts.
+We illustrate our process on text from the Canadian Hansard and introduce both a Shiny application and an R package to make our process easier for others to adopt."
+keywords: "text-as-data; GPT-3; quantitative analysis; cleaning."
 date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
 
-When we think about a data science project, we may like to think that our job is to 'let the data speak'.
+When we think of quantitative analysis, we may like to think that our job is to 'let the data speak'.
 But this is rarely the case in practise.
@@
 In any case, it is the underlying statistical process of which the dataset is an artifact that is typically of interest.
-In order to use statistical models to understand that process we often need to prepare the dataset in some way.
-This is particularly the case when working with text.
-However, this process requires many decisions.
-Should we correct obvious errors?
+In order to use statistical models to understand that process, we typically need to clean and prepare the dataset in some way.
+This is especially the case when we work with text data.
+But this cleaning and preparation requires us to make many decisions.
+To what extent should we correct obvious errors?
 What about slightly-less-obvious errors?
-To what extent have we introduced new errors?
-Have we made decisions that have affected, or even driven, our results?
+Although cleaning and preparation is a necessary step, we may be concerned about the extent to which have we introduced new errors, and the possibility that we have made decisions that have affected, or even driven, our results.
 
-In this paper we introduce the concept of consistency for a text corpus.
-Consistency refers to the proportion of words that are able to be predicted by a trained model based on the previous words and surrounding context.
+In this paper we introduce the concept of consistency in a text corpus.
+Consistency refers to the proportion of words that are able to be forecast by a statistical model, based on preceeding words and surrounding context.
 Further, we define internal consistency as when the model is trained on the corpus itself, and external consistency as when the model is trained on a more general corpus.
@@
 
-To provide an example, consider the sentence, 'the cat in the...'.
+To provide an example, consider the sentence, 'the cat in the\dots'.
 A child who has read this book could tell you that the next word should be 'hat'.
 Hence if the sentence was actually 'the cat in the bat', then that child would know something was likely wrong.
-A consistency score would likely be lower than if the sentence were 'the cat in the hat'.
-After the researcher corrects this error, a consistency score would likely increase.
-By following how the consistency scores change during the data preparation and cleaning stages the researcher can better understand the effect of the changes.
-Including consistency scores when corpuses are shared allows researchers to be more transparent about their corpus.
-And finally, the use of consistency scores allows for automation in the cleaning process.
+The consistency score would likely be lower than if the sentence were 'the cat in the hat'.
+After we correct this error, the consistency score would likely increase.
+By examining how the consistency scores evolve in response to changes made to the text during the data preparation and cleaning stages we can better understand the effect of the changes.
+Including consistency scores when text corpuses are shared allows researchers to be more transparent about their corpus.
+And finally, the use of consistency scores allows for an increased level of automation in the cleaning process.
 
-We apply our approach to X (one option is a Hansard, but it's really big, so yeah, I don't really want to do that).
-Additionally, we construct a Shiny app that computes internal and external consistency scores for smaller corpuses and allows the researcher to make changes and see how it updates.
+The model that we use to forecast is GPT-3 by OpenAI.
+OpenAI is...
+**(Ke-Li to add a two sentence summary in boring academic language)**.
+GPT-3 is...
+**(Ke-Li to add a three sentence summary in boring academic language)**.
 
-The remainder of our paper is structured as follows...
+We apply our approach to the Canadian Hansard.
+The Canadian Hansard was digitised by **(Rohan add ref)**.
+This was a process by which PDFs were scanned, put through optical character recognition (OCR), and then corrected to a reasonable extent.
+The dataset is extensive, fit-for-purpose, and, appropriately, it has been used considerably **(Rohan add refs)**.
+However, there are many issues with the dataset in terms of reading it.
+We focus on one particular year - **(Rohan pick one)** - and show that our approach can be used to relatively quickly improve the quality of the dataset in a reproducible and consistent manner.
+
+The remainder of our paper is structured as follows: **(Rohan add structure)**.
+Additionally, we construct a Shiny app available at: **(Rohan add link)**.
+That app computes internal and external consistency scores for corpus excerpts, and have developed an R Package that allows our approach to be used on larger datasets, which is available at: **(Rohan add link)**.
 
@@
 
-A typical data science workflow involves
+## PDF text extraction and n-grams
 
-Internal and external consistency
+PDFs contain a lot of information, but typically that information is not able to be analysed directly by statistical models.
+It needs to first be extracted from the PDF, often by optical character recognition (OCR), and then cleaned, prepared and made into some type of tabular dataset.
 
-/// Start of  OpenAI GPT-3 section ///
+One issue in this process is that OCR is not perfect.
+Hence there is a need to correct the output before it can be analysed.
+Traditionally, word-correction techniques evaluate errors one by one without considering the context of the surrounding words.
+This was no longer the case in modern correction techniques as statistical language models (SLMs) and feature-based methods have been used for context-sensitive correction (citation).
+Without exception, all human languages have some words that co-occur more frequently with others.
+Under this assumption, we can regard the production of English text as a set of conditional probabilities, written as Pr($w_{k}$ | $w^{k-1}_{1}$), where k is the number of words in a sentence, $w_{k}$ is the predicted word, and $w^{k-1}_{1}$ is the history of the word occurring in a sequence (citation).
+In other word, the generation of prediction $w_{k}$ is based on the history $w^{k-1}_{1}$.
+This conditional probability is the foundation of an n-gram language model.
 
-In recent years, pre-trained language models have tremendous contributions in advancing natural language processing (NLP) tasks such as reading comprehension, text generation, and many others (citation).
-As the pre-trained models remove the need for researchers and technologists to train models from scratch, they promote faster growth and advancement int the field of NLP.
-OpenAI GPT-3 is the latest generative pre-trained language model released by OpenAI in 2020.
-With 175 billion parameters, GPT-3 is claimed to have the capacity to generate text that ressembles human creation (citation).
-GPT-2, the predecessor of GPT-3, employs a semi-supervised approach that combines unsupervised pre-training and supervised fine-tuning where task-specific data set and task-specific fine-tuning were still needed in achieving these tasks until the birth of GPT-3 (citation).
-GPT-3 removes such needs, and its task-agnostic nature enables the accessible creation of cutting-edge NLP applications.
+An n-gram model is a probabilistic language model that predicts the next word in a sequence of words.
+The $n$ in n-gram represents the number of words in a sequence.
+Taking an incomplete sentence from the Jane Eyre excerpt as an example:
 
-GPT-3 also learns like humans do — with brief directives and simple instructions...
+> "We had been wandering,"
 
-The task-agnostic feature also permits it to adapt to the style of the data being fed to it (citation).
+"We" is a unigram, "We had" is a bigram, "We had been" is a trigram, and "We had been wandering" is a 4-gram.
+On the other hand, the trigrams of the excerpt are "We had been," and "had been wandering." The overlapping chaining of words reveals the statistical behaviour in human language.
+Moreover, N-gram then enables us to assign a probability to the occurrence of a sequence of words or the likelihood of a word occurring next.
+Consider the two sentences: "We had been wandering," and "We had been wangling." The former is likely to be more frequently encountered in a training corpus.
+Thus, the n-gram would assign a higher probability to "wandering" than to "wangling."
+
+[Statistic notation of n-gram?]
+
+[Character-level n-gram?]
+
+The application of n-gram is versatile. N-gram is widely applied in text prediction, spelling-correction, machine translation...
+
+### Article: Class-Based n-gram Models of Natural Language (1992)
+Syntax-based (grammartical) and semantic-based (sensible) word classifications in n-gram models.
+
+### Article: A Statistical Approach to Automatic OCR Error Correction in Context (1996)
+Context-sensitive correction system to both non-word and real-word errors based on n-gram models applied to OCR postprocessing.
+
+### Article: Unsupervised Learning of Sentence Embeddings using Compositional n-Gram Features (2017)
+Embedding of sentences (sequence of words with semantic representations) compositional n-gram features.
+
+## Pre-trained language models
+
+Draw on and cite: 'Language Models are Few-Shot Learners'
+
+In recent years, the use of pre-trained language models has played a significant role in advancing natural language processing (NLP) tasks such as reading comprehension, text generation, and **X** **(Ke-Li add ref)**.
+Pre-trained models remove the need for researchers to create training datasets or devote computational resources to this stage.
+Because of this, there has been substantial development in NLP research.
+
+OpenAI's GPT-3 is a generative, pre-trained, task-agnostic, language model that was released by OpenAI in **Ke-Li add month** 2020 **(Ke-Li add ref)**.
+OpenAI is **(Ke-Li add background)**.
+GPT-3 is the **Ke-Li add number X** model publicly released by OpenAI.
+It has '175 billion parameters' and OpenAI claim that it as the 'capacity to generate text that ressembles human creation' **(Ke-Li add ref)**.
+Initial research has found...
+**(Ke-Li please find papers that use it, describe them, and add refs - look on arXiv)**.
+
+GPT-3 is pre-trained, and takes a sample of text as an input **(Ke-Li add more detail about what it needs and what it outputs and how it works).**.
+It adapts to the the style of the data that is used an input **(Ke-Li please find the language that they use for this and then we need to make ours consistent with theirs).**
+
+# Stylized example
+
+**(TODO: This stylized example need to be changed and updated to actually use GPT-3 at some point.)**
 
 In our application, we deploy OpenAI GPT-3 to generate text and ...
-
-/// End of  OpenAI GPT-3 section ///
-
-### Article: Improving Language Understanding by Generative Pre-Training (2018)
-Introduction of a semi-supervised approach for language understanding machine learning tasks. The approach is the combinaiton of unsupervised pre-training and supervised fine-tuning.
-
-### Article: Language Models are Few-Shot Learners (2020)
-Official paper from OpenAI to introduce OpenAI GPT-3. Emphasizes on the "fewer-shots" and "task-agnostic" aspects of the latest langual model compared to its predecessor GPT-2.
-
-# Stylized example
 
@@
 
-# Models
-
-Various models can be used....
-
-## ngrams
-
-Traditionally, word-correction techniques evaluate errors one by one without considering the context of the surrounding words.
-This was no longer the case in modern correction techniques as statistical language models (SLMs) and feature-based methods have been used for context-sensitive correction (citation).
-Without exception, all human languages have some words that co-occur more frequently with others.
-Under this assumption, we can regard the production of English text as a set of conditional probabilities, written as Pr($w_{k}$ | $w^{k-1}_{1}$), where k is the number of words in a sentence, $w_{k}$ is the predicted word, and $w^{k-1}_{1}$ is the history of the word occurring in a sequence (citation).
-In other word, the generation of prediction $w_{k}$ is based on the history $w^{k-1}_{1}$.
-This conditional probability is the foundation of an n-gram language model.
-
-An n-gram model is a probabilistic language model that predicts the next word in a sequence of words.
-The $n$ in n-gram represents the number of words in a sequence.
-Taking an incomplete sentence from the Jane Eyre excerpt as an example:
-
-> "We had been wandering,"
-
-"We" is a unigram, "We had" is a bigram, "We had been" is a trigram, and "We had been wandering" is a 4-gram.
-On the other hand, the trigrams of the excerpt are "We had been," and "had been wandering." The overlapping chaining of words reveals the statistical behaviour in human language.
-Moreover, N-gram then enables us to assign a probability to the occurrence of a sequence of words or the likelihood of a word occurring next.
-Consider the two sentences: "We had been wandering," and "We had been wangling." The former is likely to be more frequently encountered in a training corpus.
-Thus, the n-gram would assign a higher probability to "wandering" than to "wangling."
-
-By including the $n$ into the previous conditional probability to be 
-Pr($w_{k}$ | $w^{k-1}_{1}$)
-
-The application of n-gram is versatile. N-gram is widely applied in text prediction, spelling-correction, machine translation...
-
-### Article: Class-Based n-gram Models of Natural Language (1992)
-Syntax-based (grammartical) and semantic-based (sensible) word classifications in n-gram models.
-
-### Article: A Statistical Approach to Automatic OCR Error Correction in Context (1996)
-Context-sensitive correction system to both non-word and real-word errors based on n-gram models applied to OCR postprocessing.
-
-### Article: Unsupervised Learning of Sentence Embeddings using Compositional n-Gram Features (2017)
-Embedding of sentences (sequence of words with semantic representations) compositional n-gram features.
-
-## Something
-
-## Soemthing else
-
 # Data
@@
 
+# Appendix
+
+## Literature summaries
+
+(These are just here for now - we'll remove later.)
+
+### Article: Improving Language Understanding by Generative Pre-Training (2018)
+Introduction of a semi-supervised approach for language understanding machine learning tasks. The approach is the combinaiton of unsupervised pre-training and supervised fine-tuning.
+
+### Article: Language Models are Few-Shot Learners (2020)
+Official paper from OpenAI to introduce OpenAI GPT-3. Emphasizes on the "fewer-shots" and "task-agnostic" aspects of the latest langual model compared to its predecessor GPT-2.
+
 # References

### 2020-07-10 shinyapp/server.R
add github script

@@ -1,7 +1,6 @@
+library(shiny)
+library(tidyverse) # tidyverse is needed for hardcoding string replacement, may be removed once we move to next phase
 source("text-input.R")
-
-# tidyverse is needed for hardcoding string replacement, may be removed once we move to next phase
-library(tidyverse)
 
 function(input, output, session) {
   # Fucntion that displays input text when get result button is clicked
@@ -43,3 +42,5 @@ function(input, output, session) {
 
 
 }
+library(shiny)
+runGitHub("consistency_scores_in_text_datasets", "RohanAlexander", subdir = "shinyapp/")

### 2020-07-10 shinyapp/ui.R

@@ -1,3 +1,4 @@
+library(shiny)
 library(shinythemes)
 library(shinyjs)
 

### 2020-07-10 shinyapp/server.R
created a zip file for hosting

@@ -43,4 +43,4 @@ function(input, output, session) {
 
 }
 library(shiny)
-runGitHub("consistency_scores_in_text_datasets", "RohanAlexander", subdir = "shinyapp/")
+runUrl("https://github.com/RohanAlexander/consistency_scores_in_text_datasets/tree/master/shinyapp.zip")

### 2020-07-10 shinyapp/server.R
undo gihub hosting, hosting on shinyapp.io instead

@@ -39,8 +39,4 @@ function(input, output, session) {
     input$clear
     updateTextInput(session, "myText", value = "")
   })
-
-
 }
-library(shiny)
-runUrl("https://github.com/RohanAlexander/consistency_scores_in_text_datasets/tree/master/shinyapp.zip")

### 2020-07-14 outputs/documents/paper.Rmd
Add plan for appendix
introduction GPT-3 part revision

@@
 The model that we use to forecast is GPT-3 by OpenAI.
-OpenAI is...
-**(Ke-Li to add a two sentence summary in boring academic language)**.
-GPT-3 is...
-**(Ke-Li to add a three sentence summary in boring academic language)**.
+OpenAI is an artificial intelligent research group who claims to be an pursuer of beneficial artificial general intelligence.
+Their works, including GPT-3, have a focus on generative models who have the capacity to generate data tha resemble human creation.
+GPT-3, the Generative Pretrained Transformer 3, is an unsupervised language model that has more than 175 billions of parameters, referring to the weights of the connections that enable the neural network to learn **(Keli add citation)**.
+GPT-3 can learn NLP tasks with few examples and without task-specific dataset **(Keli add citation)**.
+This means GPT-3 can learns like human do, which makes the creation of NLP technologies more accessible.
 
@@
 
+## Discussion of how GPT-3 works
+
+**(Ke-Li - it would be good for us to go through how it actually works in as much detail as we can manage (in our own words).
+This is mostly just to force us to learn how it works and may be taken out of the final version of the paper, but it's still an important exercise.)**
+
 # References

### 2020-07-14 shinyapp/functions.R

@@ -1,0 +1,1 @@
+library(quanteda)

### 2020-07-14 outputs/documents/paper.Rmd
n-gram statistic principal notations

@@
 
-[Statistic notation of n-gram?]
+The probability of a sequence appearing in a corpus follows the chain rules:
 
-[Character-level n-gram?]
+> P(We,had,been,wandering) $=$ P(We)$\cdot$P(had|We)$\cdot$P(been|We,had)$\cdot$P(wandering|We,had,been)
+
+This will soon grow weary if the sentences are long. Instead, the Markov assumption permits us to approximate the probability using only the last $n$ words. More specifically:
+
+> P(We,had,been,wandering)$\approx$ P(We)$\cdot$P(had|We)$\cdot$P(been|had)$\cdot$P(wandering|been)
+
+Simplified and formalized to:
+
+> P($w^{n}_{1}) \approx$ P($w_{1}$) $\cdot$ P($w_{2} | w_{1}$) ... P($w_{n} | w_{n-1}$)
 

### 2020-07-14 outputs/documents/paper.Rmd
n-gram parts revision

@@
 
+To predict a word appearing next, we have to take the sequence of preceeding words into account, which requires knowing the probability of the sequence of words.
 The probability of a sequence appearing in a corpus follows the chain rules:
@@
 
-This will soon grow weary if the sentences are long. Instead, the Markov assumption permits us to approximate the probability using only the last $n$ words. More specifically:
+However, the likelihood that more and more words will occur next to each other in an indentical sequence becomes smaller and smaller, making the prediction difficult.
+Instead, the Markov assumption permits us to approximate the probability using only the last $n$ words **(Keli add citation)**.
+More specifically
 
@@
 
-Simplified and formalized to:
+Because the bigram model looks into one word in the past, the n-gram under Markov assumption can then be generalized into a bigram model with n being any number. The general equation is
 
-> P($w^{n}_{1}) \approx$ P($w_{1}$) $\cdot$ P($w_{2} | w_{1}$) ... P($w_{n} | w_{n-1}$)
+P($w_{n} | w^{n-1}_{1}) \approx$ P($w_{n} | w^{n-1}_{n - N+1})$)
 
-The application of n-gram is versatile. N-gram is widely applied in text prediction, spelling-correction, machine translation...
+The application of n-gram is versatile.
+N-gram is widely applied in text prediction, spelling-correction, machine translation.
+We use n-gram to [add how we use n-gram once the functions in the app are in place]
 

### 2020-07-16 shinyapp/functions.R
trigram word prediction

@@ -1,1 +1,60 @@
-library(quanteda)
+library(tidyverse)
+library(tau)
+library(tm)
+library(hash)
+
+message("load data...")
+message(Sys.time())
+setwd(getSrcDirectory(function(){}))
+
+# Load training text
+blog.sample <- readLines("samples/*.txt")
+
+# Divide the data to 50 batches to speed up n-gram processing later
+nbatch <- 50
+blog.len <- ceiling(length(blog.sample) / nbatch)
+
+# Build functions to preprocess the text by removing non-English characters, twitter handles and urls
+removeURL <- function(x) gsub("http\\S+", "", x)
+removeHash <- function(x) gsub("[@#&]\\S+", "", x)
+removeNumPunct <- function(x) gsub("[^A-z[:space:]']*", "", x)
+
+# Build the n-gram
+h <- hash()
+
+# Loop over the batches
+for (b in 1:nbatch -1) {
+  message(sprintf("Processing the %i-th batch", b))
+  
+  #concatenate text and preprocess
+  blog <- blog.sample[blog.len * b + (1 : blog.len)]
+  trainingText <- blog %>%
+    removeURL() %>% removeHash() %>% removeNumPunct() %>% tolower() %>% stripWhitespace()
+
+  # Get trigrams
+  trigram <- textcnt( trainingText, n=3, split=" ", method = "string", decreasing = TRUE)
+
+  # Create hash table for fast prediction
+  for (i in 1: length(trigram)) {
+    if (trigram[i] < 4) {
+      break
+    } else {
+      gram <- strsplit(names(trigram[i]), split = ' ')[[1]]
+      history <- paste0(gram[1:2], collapse = ' ')
+      candidate <- gram[3]
+      count <- trigram[[i]]
+      if (candidate %in% h[[history]]$candidate) {
+        index <- h[[history]]$candidate == candidate
+        h[[history]]$count[index] <- h[[history]]$count[index] + count
+      } else {
+        h[[history]]$candidate <- c(h[[history]]$candidate, candidate)
+        h[[history]]$count <- c(h[[history]]$count, count)
+      }
+    }
+  }
+  gc()
+}
+if (1) {
+  save(h, file = "hashtable.Rdata")
+}
+

### 2020-07-16 shinyapp/server.R

@@ -1,42 +1,64 @@
-library(shiny)
-library(tidyverse) # tidyverse is needed for hardcoding string replacement, may be removed once we move to next phase
+library(tidyverse)
+library(tau)
+library(tm)
+library(hash)
 source("text-input.R")
-
-function(input, output, session) {
-  # Fucntion that displays input text when get result button is clicked
+load("hashtable.Rdata")
+server <- function(input, output, session) {
   getText <- eventReactive(input$getResult, {
-    # Hardcode text correction by replacing typos manually
-    str_replace_all(input$myText, c("Reeal" = "Reed", "tbere" = "there", "tbat" = "that"))
+    myText <- trimws(input$myText)
+    myText <- gsub('\\s+', ' ', myText)
+    pred.sentence <- paste(myText, "...")
+    words <- strsplit(tolower(myText), split = ' ')[[1]]
+    if (length(words) >= 2) {
+      history <- paste0(tail(words, 2), collapse = ' ')
+      candidates <- h[[history]]$candidate
+      candidates[candidates == 'i'] <- 'I'
+      if (length(candidates)) {
+        maxDisplay <- min(5, length(candidates))
+        for (i in 1:maxDisplay) {
+          pred.sentence <- paste(pred.sentence,
+                                 paste(candidates[i], sep = ' '),
+                                 sep = '\n')
+        }
+      }
+      pred.sentence
+    } else {
+      myText
+    }
   })
+  
   # Displays input text
   output$outputText <- renderText({
     getText()
   })
-
+  
+  
   # Fucntion that displays scores when get result button is clicked
-    getScores <- eventReactive(input$getResult, {
+  getScores <- eventReactive(input$getResult, {
     "Internal consistency score: 0.919 <br> External consistency score: 0.956"
   })
   # Displays scores
   output$outputScores <- renderText({
     getScores()
   })
-
+  
   # Disable get result button and clear button when there is no text data
   observe({
     toggleState(id = "getResult", condition = nchar(input$myText) > 0)
     toggleState(id = "clear", condition = nchar(input$myText) > 0)
   })
-
+  
   # Load demo data when the corresponding button is clicked
   observe({
     input$loadDemo
     updateTextInput(session, "myText", value = demoInput)
   })
-
+  
   # Clear text input area
   observe({
     input$clear
     updateTextInput(session, "myText", value = "")
   })
+  
 }

### 2020-07-16 shinyapp/ui.R

@@ -1,6 +1,13 @@
 library(shiny)
 library(shinythemes)
 library(shinyjs)
+library(shiny)
+library(tidyverse)
+library(tau)
+library(tm)
+library(hash)
+#source("functions.R")
+#load("hashtable.Rdata")
 
 fluidPage(
   useShinyjs(),

### 2020-07-20 outputs/documents/paper.Rmd
Update abstract and introduction.

@@
 title: "Consistency scores in text data"
-thanks: "A Shiny app for interactive and small-scale examples is available: HERE.
-An R package for larger scale applications is available: HERE.
-Our code and datasets are available: HERE.
+thanks: "Thank you to X, Y and Z for helpful comments.
+A Shiny application for interactive and small-scale examples is available: https://kelichiu.shinyapps.io/Arianna/.
+An R package for larger scale applications is available: **HERE**.
+Our code and datasets are available: https://github.com/RohanAlexander/consistency_scores_in_text_datasets.
 Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca."
@@
   affiliation: University of Toronto  
-abstract: "In this paper we introduce a process to clean the text extracted from PDFs using GPT-3.
-Our approach compares originally extracted text with the text generated from GPT-3 using earlier text as stimulus.
-To guide this process, we introduce the notion of a consistency score, which refers to the proporation of text that is unchanged by the model.
+abstract: "In this paper we introduce a process to clean the text extracted from PDFs using various methods from natural language processing.
+Our approach compares originally extracted text with the text generated from, or expected by, these methods using earlier text as stimulus.
+To guide this process, we introduce the notion of a consistency score, which refers to the proportion of text that is unchanged by the model.
 This is used to monitor changes during the cleaning process, and to compare the messiness of different texts.
+The methods that we consider are: n-grams, SOMETHING, GPT-2, and GPT-3.
 We illustrate our process on text from the Canadian Hansard and introduce both a Shiny application and an R package to make our process easier for others to adopt."
-keywords: "text-as-data; GPT-3; quantitative analysis; cleaning."
+keywords: "text-as-data; natural language processing; GPT-3; quantitative analysis."
 date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
 
-The model that we use to forecast is GPT-3 by OpenAI.
-OpenAI is an artificial intelligent research group who claims to be an pursuer of beneficial artificial general intelligence.
-Their works, including GPT-3, have a focus on generative models who have the capacity to generate data tha resemble human creation.
-GPT-3, the Generative Pretrained Transformer 3, is an unsupervised language model that has more than 175 billions of parameters, referring to the weights of the connections that enable the neural network to learn **(Keli add citation)**.
-GPT-3 can learn NLP tasks with few examples and without task-specific dataset **(Keli add citation)**.
-This means GPT-3 can learns like human do, which makes the creation of NLP technologies more accessible.
+We consider various natural language processing methods.
+These include n-grams, **SOMETHING**, as well as GPT2 and GPT-3.
+The n-gram approach involves identifying two, three, or more, words that are commonly found together.
+Here think of a two-gram involving the word 'good' 'good morning'.
+At scale these can identify missing or unusual words, and work quickly, but they lack nuance.
+For instance an equally reasonable two-gram involving the word 'good' is 'good work'.
+For that reason we consider pre-trained generative models.
+These include **SOMETHING**, which works **SOMEHOW**.
+We also consider GPT-2 and GPT-3 from OpenAI.
+GPT-2 and GPT-3 are pre-trained generative unsupervised language models.
+They differ in the number of parameters.
+Generative Pretrained Transformer (GPT)-2 has *X* parameters and GPT-3 has more than 175 billions parameters.
+While GPT-2 is more limited, it can be run on smaller machines, whereas GPT-3 cannot.
 
 We apply our approach to the Canadian Hansard.
-The Canadian Hansard was digitised by **(Rohan add ref)**.
+The Canadian Hansard was digitised by @beelen2017digitization.
 This was a process by which PDFs were scanned, put through optical character recognition (OCR), and then corrected to a reasonable extent.
-The dataset is extensive, fit-for-purpose, and, appropriately, it has been used considerably **(Rohan add refs)**.
+The dataset is extensive, fit-for-purpose, and, appropriately, it has been used considerably, for instance @rheault2020word.
 However, there are many issues with the dataset in terms of reading it.
@@
 **(Ke-Li please find papers that use it, describe them, and add refs - look on arXiv)**.
+These parameters refer to the weights of the connections that enable the neural network to learn **(Keli add citation)**.
+GPT-3 can learn NLP tasks with few examples and without a task-specific dataset **(Keli add citation)**.
 

### 2020-07-21 shinyapp/functions.R
trigram word detection and prediction

@@ -1,60 +1,137 @@
 library(tidyverse)
 library(tau)
 library(tm)
-library(hash)
+library(stringr)
+library(stats)
 
-message("load data...")
-message(Sys.time())
-setwd(getSrcDirectory(function(){}))
+#setwd("~/Desktop/repos/consistency_scores_in_text_datasets/shinyapp")
 
-# Load training text
-blog.sample <- readLines("samples/*.txt")
+demoTraining <- "There was no possibility of taking a walk that day. 
+We had been wandering, indeed, in the leafless shrubbery an hour in the morning; 
+but since dinner (Mrs. Reed, when there was no company, dined early) 
+the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, 
+that further out-door exercise was now out of the question."
 
-# Divide the data to 50 batches to speed up n-gram processing later
-nbatch <- 50
-blog.len <- ceiling(length(blog.sample) / nbatch)
+tokenizer <- function(lines) {
+  lines <- tolower(lines)
+  lines <- gsub("'", "'", lines)
+  #lines <- gsub("[.!?]$|[.!?] |$", " ''split'' ", lines)
+  tokens <- unlist(strsplit(lines, "[^a-z']"))
+  tokens <- tokens[tokens != ""]
+  return(tokens)
+}
+demoTraining <- tokenizer(demoTraining)
+unigrams <- textcnt( demoTraining, n=1, method = "string", decreasing = TRUE)
+bigrams  <- textcnt( demoTraining, n=2, method = "string", decreasing = TRUE)
+trigrams <- textcnt( demoTraining, n=3, method = "string", decreasing = TRUE)
 
-# Build functions to preprocess the text by removing non-English characters, twitter handles and urls
-removeURL <- function(x) gsub("http\\S+", "", x)
-removeHash <- function(x) gsub("[@#&]\\S+", "", x)
-removeNumPunct <- function(x) gsub("[^A-z[:space:]']*", "", x)
+userInput <- "There was no possibility of taking a walk that day. 
+We had been wandering, indeed, in the leafless shrubbery an hour in the morning; 
+but since dinner (Mrs. Reed, when there was no company, dined early) 
+the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, 
+that further out-door exercise was now out of the question."
[…]

### 2020-07-21 shinyapp/server.R

@@ -1,31 +1,14 @@
 library(tidyverse)
 library(tau)
 library(tm)
-library(hash)
 source("text-input.R")
-load("hashtable.Rdata")
+source("functions.R")
+
 server <- function(input, output, session) {
   getText <- eventReactive(input$getResult, {
     myText <- trimws(input$myText)
     myText <- gsub('\\s+', ' ', myText)
-    pred.sentence <- paste(myText, "...")
-    words <- strsplit(tolower(myText), split = ' ')[[1]]
-    if (length(words) >= 2) {
-      history <- paste0(tail(words, 2), collapse = ' ')
-      candidates <- h[[history]]$candidate
-      candidates[candidates == 'i'] <- 'I'
-      if (length(candidates)) {
-        maxDisplay <- min(5, length(candidates))
-        for (i in 1:maxDisplay) {
-          pred.sentence <- paste(pred.sentence,
-                                 paste(candidates[i], sep = ' '),
-                                 sep = '\n')
-        }
-      }
-      pred.sentence
-    } else {
-      myText
-    }
+    detector(myText)
   })
   
   # Displays input text
@@ -60,5 +43,4 @@ server <- function(input, output, session) {
     input$clear
     updateTextInput(session, "myText", value = "")
   })
-  
 }

### 2020-07-21 shinyapp/ui.R

@@ -5,9 +5,6 @@ library(shiny)
 library(tidyverse)
 library(tau)
 library(tm)
-library(hash)
-#source("functions.R")
-#load("hashtable.Rdata")
 
 fluidPage(
   useShinyjs(),

### 2020-07-22 shinyapp/functions.R
code rearrangement

@@ -5,36 +5,37 @@ library(stringr)
 library(stats)
 
 #setwd("~/Desktop/repos/consistency_scores_in_text_datasets/shinyapp")
+path <- getwd()
+demoTraining <- readLines(paste(path,"/samples/demo.txt", sep=""), warn=FALSE)
 
-demoTraining <- "There was no possibility of taking a walk that day. 
-We had been wandering, indeed, in the leafless shrubbery an hour in the morning; 
-but since dinner (Mrs. Reed, when there was no company, dined early) 
-the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, 
-that further out-door exercise was now out of the question."
-
-tokenizer <- function(lines) {
+tokenizer <- function(corpus) {
+  lines <- vector()
+  
+  # Read the corpus line by line
+  for (line in corpus) {
+    # add two start-of-sentence tokens to the beginning of a sentence
+    # line <- paste("<s> <s>", line)
+    lines <-c(lines, line)
+  }
+  
   lines <- tolower(lines)
-  lines <- gsub("'", "'", lines)
-  #lines <- gsub("[.!?]$|[.!?] |$", " ''split'' ", lines)
-  tokens <- unlist(strsplit(lines, "[^a-z']"))
+  lines <- gsub("[.!?]$|[.!?] |$", " </s>", lines) # add end-of-sentence token to the end of a sentence
+  tokens <- unlist(strsplit(lines, "[^a-z<>/]"))
   tokens <- tokens[tokens != ""]
   return(tokens)
 }
-demoTraining <- tokenizer(demoTraining)
-unigrams <- textcnt( demoTraining, n=1, method = "string", decreasing = TRUE)
-bigrams  <- textcnt( demoTraining, n=2, method = "string", decreasing = TRUE)
-trigrams <- textcnt( demoTraining, n=3, method = "string", decreasing = TRUE)
 
-userInput <- "There was no possibility of taking a walk that day. 
-We had been wandering, indeed, in the leafless shrubbery an hour in the morning; 
-but since dinner (Mrs. Reed, when there was no company, dined early) 
-the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, 
-that further out-door exercise was now out of the question."
-userInput <- tokenizer(userInput)
-ui_unigrams <- textcnt( userInput, n=1, method = "string", decreasing = TRUE)
[…]

### 2020-07-23 outputs/documents/paper.Rmd
GPT write-ups in progress

@@
 N-gram is widely applied in text prediction, spelling-correction, machine translation.
-We use n-gram to [add how we use n-gram once the functions in the app are in place]
+In our application, we use a trigram-based model to detect both real-word and non-word errors and correct them with the candidate word that has the highest probability.
+[add more about how we use n-gram once the functions in the app are in place]
 
@@
 
-In recent years, the use of pre-trained language models has played a significant role in advancing natural language processing (NLP) tasks such as reading comprehension, text generation, and **X** **(Ke-Li add ref)**.
+In recent years, the use of pre-trained language models has played a significant role in advancing natural language processing (NLP) tasks such as reading comprehension, text generation, and even creative writing **(Ke-Li add ref)**.
 Pre-trained models remove the need for researchers to create training datasets or devote computational resources to this stage.
@@
 
-OpenAI's GPT-3 is a generative, pre-trained, task-agnostic, language model that was released by OpenAI in **Ke-Li add month** 2020 **(Ke-Li add ref)**.
-OpenAI is **(Ke-Li add background)**.
-GPT-3 is the **Ke-Li add number X** model publicly released by OpenAI.
-It has '175 billion parameters' and OpenAI claim that it as the 'capacity to generate text that ressembles human creation' **(Ke-Li add ref)**.
-Initial research has found...
+OpenAI's GPT-3 is a generative, pre-trained, task-agnostic, language model that was released by OpenAI in June 2020 **(Ke-Li add ref)**.
+OpenAI is an AI research company who openly pursues Artificial General Intelligence—to have machines learn and understand any intellectual tasks as human do **(Ke-Li add ref)**.
+GPT-3 is the third Generative Pre-trained Transformer language model publicly released by OpenAI.
+It has '175 billion parameters' and can learn NLP tasks with just a few examples.
+OpenAI claims that the text generation by GPT-3 is 'difficult for humans to distinguish from authentic human content' **(Ke-Li add ref)**.
+Since the API of GPT-3 at the moment is only available to a few invited users, public applications of GPT-3 are limited.
+On the other hand, its predecessor, GPT-2, has enabled researchers made breakthroughs in the past years.
+Alt, Hubner and Henning extended GPT-2 to relation extraction, which is a task of identifying the relationship between concepts appeared in text.
+They had shown a milestone in predicting larger range of distinct relation types with higher confidence **(Ke-Li add ref)**.
+In the field of speech recognition, data scarcity is often an issue due to the difficulty to collect large amount of data from human for learning **(Ke-Li add ref)**.
+Bird, Faria, Ekart, Premebida and Ayrosa employed GPT-2 as part of the solution to generate augmented data for classification.
+GPT-2 also helped NLP researchers tackling the task of textual paraphrasing **(Ke-Li add ref)**.
+Hedge and Patil used GPT-2 to build an unsupervised paraphrasing model that is not restricted by domain shift within an article, which is a problem faced by the usual supervised models of paraphrasing.
 **(Ke-Li please find papers that use it, describe them, and add refs - look on arXiv)**.
+
+The above examples using GPT-2 were task-agnostic but they were in need of fine-tuning.
 These parameters refer to the weights of the connections that enable the neural network to learn **(Keli add citation)**.
-GPT-3 can learn NLP tasks with few examples and without a task-specific dataset **(Keli add citation)**.
+GPT-3 can learn NLP tasks with few examples and without any task-specific dataset **(Keli add citation)**.
 
 GPT-3 is pre-trained, and takes a sample of text as an input **(Ke-Li add more detail about what it needs and what it outputs and how it works).**. 
-It adapts to the the style of the data that is used an input **(Ke-Li please find the language that they use for this and then we need to make ours consistent with theirs).**
+**(Ke-Li please find the language that they use for this and then we need to make ours consistent with theirs).**
+
+Deep language representations, e.g. those
+learned by the Transformer (Vaswani et al., 2017)
+via language modeling (Radford et al., 2018),
+have been shown to implicitly capture useful semantic and syntactic properties of text solely by
+arXiv:1906.08646v1 [cs.CL] 19 Jun 2019
+unsupervised pre-training (Peters et al., 2018)
 

### 2020-07-23 shinyapp/functions.R

@@ -14,7 +14,7 @@ tokenizer <- function(corpus) {
   # Read the corpus line by line
   for (line in corpus) {
     # add two start-of-sentence tokens to the beginning of a sentence
-    # line <- paste("<s> <s>", line)
+    line <- paste("<s> <s>", line)
     lines <-c(lines, line)
   }
   
@@ -28,8 +28,8 @@ tokenizer <- function(corpus) {
 tokens <- tokenizer(demoTraining)
 
 library(data.table)
-unigrams <- textcnt( tokens, n=1, method = "string", decreasing = TRUE)
-bigrams  <- textcnt( tokens, n=2, method = "string", decreasing = TRUE)
+unigrams <- textcnt( tokens, n=1, method = "string", split = "[[:space:][:digit:]]+", decreasing = TRUE)
+bigrams  <- textcnt( tokens, n=2, method = "string", split = "[[:space:][:digit:]]+", decreasing = TRUE)
 trigrams <- textcnt( tokens, n=3, method = "string", split = "[[:space:][:digit:]]+", decreasing = TRUE)
 
 # Remove trigram that has "</s>" in it, so the trigrams are constructed within a sentence
@@ -90,21 +90,30 @@ trigramsDF <- data.table("FirstWords" = removeLastWord(names(trigrams)),
 
 
 library(dplyr)
-unigramDF <-  unigramDF %>% arrange(desc(Probability))
-bigramsDF <-  bigramsDF %>% arrange(desc(Probability)) %>% filter(Probability > 0.0001)
+unigramDF <- (unigramDF %>% arrange(desc(Probability)))
+bigramsDF <- bigramsDF %>% arrange(desc(Probability)) %>% filter(Probability > 0.0001)
 trigramsDF <- trigramsDF %>% arrange(desc(Probability)) %>% filter(Probability > 0.0001)
 
 detector <- function(input) {
-  input <- input
-  inputTrigrams <- textcnt( input, n=3, method = "string", decreasing = TRUE)
-  correct <- input
+  inputCopy <- unlist(strsplit(input, "\n")) # split the input by sentences
+  tokens <- tokenizer(inputCopy)
+  # Turn input text to trigrams
+  inputTrigrams <- textcnt( tokens, n=3, method = "string", split = "[[:space:][:digit:]]+", decreasing = TRUE)
+  inputTrigrams <- inputTrigrams[!grepl("</s>", names(inputTrigrams))]
+  print(inputTrigrams)
[…]

### 2020-07-23 outputs/documents/paper.Rmd
write-up updates on transformer model and add more referenced literatures

@@
 
-The above examples using GPT-2 were task-agnostic but they were in need of fine-tuning.
-These parameters refer to the weights of the connections that enable the neural network to learn **(Keli add citation)**.
+GPT is a Transformer-based model that has over 175 billions of parameters.
+**(Keli add citation)**.
+Proposed by Vaswani et al., the Transformer is a novel network architecture that outperforms RNN-based models in computational efficiency **(Keli add citation)**.
+The parameters refer to the weights of the connections that enable the neural network to learn **(Keli add citation)**.
+In recent years, the growth of parameters of transformer language models expands rapidly which accelerate the improvements in text synthesis and downstream NLP tasks **(Keli add citation)**.
+
 GPT-3 can learn NLP tasks with few examples and without any task-specific dataset **(Keli add citation)**.
-
+OpenAI refers this ability as few-shot learning.
 GPT-3 is pre-trained, and takes a sample of text as an input **(Ke-Li add more detail about what it needs and what it outputs and how it works).**.
 **(Ke-Li please find the language that they use for this and then we need to make ours consistent with theirs).**
-
-Deep language representations, e.g. those
-learned by the Transformer (Vaswani et al., 2017)
-via language modeling (Radford et al., 2018),
-have been shown to implicitly capture useful semantic and syntactic properties of text solely by
-arXiv:1906.08646v1 [cs.CL] 19 Jun 2019
-unsupervised pre-training (Peters et al., 2018)
 

### 2020-07-23 outputs/documents/paper.Rmd
write-up on few-shot learning

@@
 
-OpenAI's GPT-3 is a generative, pre-trained, task-agnostic, language model that was released by OpenAI in June 2020 **(Ke-Li add ref)**.
+OpenAI's GPT-3 is a generative, pre-trained, task-agnostic, language model that was announced by OpenAI in June 2020 **(Ke-Li add ref)**.
 OpenAI is an AI research company who openly pursues Artificial General Intelligence—to have machines learn and understand any intellectual tasks as human do **(Ke-Li add ref)**.
@@
 
-GPT is a Transformer-based model that has over 175 billions of parameters.
+GPT-3 is a Transformer-based model that has over 175 billions of parameters.
 **(Keli add citation)**.
@@
 The parameters refer to the weights of the connections that enable the neural network to learn **(Keli add citation)**.
-In recent years, the growth of parameters of transformer language models expands rapidly which accelerate the improvements in text synthesis and downstream NLP tasks **(Keli add citation)**.
-
-GPT-3 can learn NLP tasks with few examples and without any task-specific dataset **(Keli add citation)**.
-OpenAI refers this ability as few-shot learning.
-GPT-3 is pre-trained, and takes a sample of text as an input **(Ke-Li add more detail about what it needs and what it outputs and how it works).**.
+In the past few years, the growth of parameters of transformer language models expands rapidly.
+The first generation GPT has 110 million parameters.
+Released in 2018, GPT-2 has 1.6 billion parameters, and within two years, GPT-3 has grown the size to 175 billion, which is over 100 times than its predecessor .
+The more parameters there are the better the model can generalize to new tasks; therefore, the increasing amount of parameters accelerates the improvements in text synthesis and downstream NLP tasks **(Keli add citation)**.
+GPT-3 can learn NLP tasks with just a handful of example; OpenAI refers this ability as few-shot learning, or in-context learning.
+Few-shot learning means that the model  **(Keli add citation)**.
+As a result, we only need to provide GPT-3 a few demonstrations for it to learn a new NLP task.
+**(Ke-Li add more detail about what it needs and what it outputs and how it works).**.
+For example, OpenAI team trained GPT-3 to generate "news articles".
+The prompt input is a plausible first sentence of a news story written by human, and the output is an entire news article generated by the model itself.
+Without the demonstrations, GPT-3 tends to treat the prompts as "tweets" and generate texts that serve as responses or follow-up tweets.
+The team fed a few previous news articles in the model’s context as demonstrations, and the model learned to produce the outputs that adapt to the style and length of news stories **(Keli add citation)**.
+This few-shot learning ability indicates that the model can learn like human do and would democratize the use of language models to broader fields.
 **(Ke-Li please find the language that they use for this and then we need to make ours consistent with theirs).**

### 2020-07-24 outputs/documents/paper.Rmd
Write-up minor revisions

@@
 This is used to monitor changes during the cleaning process, and to compare the messiness of different texts.
-The methods that we consider are: n-grams, SOMETHING, GPT-2, and GPT-3.
+The methods that we consider are: n-grams, [Google’s Word2vec, Standford's GloVe], GPT-2, and GPT-3.
 We illustrate our process on text from the Canadian Hansard and introduce both a Shiny application and an R package to make our process easier for others to adopt."
@@
 We consider various natural language processing methods.
-These include n-grams, **SOMETHING**, as well as GPT2 and GPT-3.
+These include n-grams, [Google’s Word2vec, Standford's GloVe], as well as GPT2 and GPT-3.
 The n-gram approach involves identifying two, three, or more, words that are commonly found together.
@@
 For instance an equally reasonable two-gram involving the word 'good' is 'good work'.
-For that reason we consider pre-trained generative models.
-These include **SOMETHING**, which works **SOMEHOW**.
-We also consider GPT-2 and GPT-3 from OpenAI.
+For that reason we consider pre-trained word embedding models.
+These include [Google’s Word2vec, Standford's GloVe], which works [representing words as dense vectors that capture their semantic relationships with each other]**SOMEHOW**.
+We also consider pre-trained generative models such as GPT-2 and GPT-3 from OpenAI.
 GPT-2 and GPT-3 are pre-trained generative unsupervised language models.
@@
 Without exception, all human languages have some words that co-occur more frequently with others.
-Under this assumption, we can regard the production of English text as a set of conditional probabilities, written as Pr($w_{k}$ | $w^{k-1}_{1}$), where k is the number of words in a sentence, $w_{k}$ is the predicted word, and $w^{k-1}_{1}$ is the history of the word occurring in a sequence (citation).
+Under this assumption, we can regard the production of English text as a set of conditional probabilities, written as Pr($w_{k}$ | $w^{k-1}_{1}$), where k is the number of words in a sentence, $w_{k}$ is the predicted word, and $w^{k-1}_{1}$ is the history of the word occurring in a sequence (@brown1992class).
 In other word, the generation of prediction $w_{k}$ is based on the history $w^{k-1}_{1}$.
@@
 
-"We" is a unigram, "We had" is a bigram, "We had been" is a trigram, and "We had been wandering" is a 4-gram.
-On the other hand, the trigrams of the excerpt are "We had been," and "had been wandering." The overlapping chaining of words reveals the statistical behaviour in human language.
-Moreover, N-gram then enables us to assign a probability to the occurrence of a sequence of words or the likelihood of a word occurring next.
+"We" is a uni-gram, "We had" is a bi-gram, "We had been" is a tri-gram, and "We had been wandering" is a 4-gram.
+On the other hand, the tri-grams of the excerpt are "We had been," and "had been wandering." The overlapping chaining of words reveals the statistical behaviour in human language.
+Moreover, using n-gram models then enable us to assign a probability to the occurrence of a sequence of words or the likelihood of a word occurring next.
 Consider the two sentences: "We had been wandering," and "We had been wangling." The former is likely to be more frequently encountered in a training corpus.
@@
 
-To predict a word appearing next, we have to take the sequence of preceeding words into account, which requires knowing the probability of the sequence of words.
-The probability of a sequence appearing in a corpus follows the chain rules:
+To predict which word appears next, we have to take the sequence of preceeding words into account, which requires knowing the probability of the sequence of words.
+The probability of a sequence appearing in a corpus follows the chain rule:
 
@@
 
-However, the likelihood that more and more words will occur next to each other in an indentical sequence becomes smaller and smaller, making the prediction difficult.
-Instead, the Markov assumption permits us to approximate the probability using only the last $n$ words **(Keli add citation)**.
-More specifically
+However, the likelihood that more and more words will occur next to each other in an identical sequence becomes smaller and smaller, making the prediction difficult.
+Alternatively, we can approximate the probability of a word depending on only the previous word; this is known as Markov assumption.
+Instead, the Markov assumption permits us to approximate the probability using only the last $n$ words (@brown1992class).
+More specifically:
 
@@
 
-Because the bigram model looks into one word in the past, the n-gram under Markov assumption can then be generalized into a bigram model with n being any number. The general equation is
+Because the bi-gram model looks only the immediate preceding word, an n-gram model under Markov assumption can then be reduced into a bi-gram model with n being any number. The general equation is:
 
-P($w_{n} | w^{n-1}_{1}) \approx$ P($w_{n} | w^{n-1}_{n - N+1})$)
+> P($w_{n} | w^{n-1}_{1}) \approx$ P($w_{n} | w^{n-1}_{n - N+1})$)
 
 The application of n-gram is versatile.
-N-gram is widely applied in text prediction, spelling-correction, machine translation.
-In our application, we use a trigram-based model to detect both real-word and non-word errors and correct them with the candidate word that has the highest probability.
+N-gram models are widely applied in text prediction, spelling-correction, machine translation **(Keli add citation)**.
+In our application, we use a tri-gram based model to detect both real-word and non-word errors and correct them with the candidate word that has the highest probability.
 [add more about how we use n-gram once the functions in the app are in place]
@@
 
-### Article: Unsupervised Learning of Sentence Embeddings using Compositional n-Gram Features (2017)
-Embedding of sentences (sequence of words with semantic representations) compositional n-gram features.
-
 ## Pre-trained language models
@@
 In recent years, the use of pre-trained language models has played a significant role in advancing natural language processing (NLP) tasks such as reading comprehension, text generation, and even creative writing **(Ke-Li add ref)**.
-Pre-trained models remove the need for researchers to create training datasets or devote computational resources to this stage.
+Pre-trained models remove the need for researchers to prepare training datasets or allot computational resources to the stage of pre-training.
 Because of this, there has been substantial development in NLP research.
@@
 It has '175 billion parameters' and can learn NLP tasks with just a few examples.
-OpenAI claims that the text generation by GPT-3 is 'difficult for humans to distinguish from authentic human content' **(Ke-Li add ref)**.
+OpenAI claims that the text generation by GPT-3 is 'difficult for humans to distinguish from authentic human content' (@brown2020language).
 Since the API of GPT-3 at the moment is only available to a few invited users, public applications of GPT-3 are limited.
 On the other hand, its predecessor, GPT-2, has enabled researchers made breakthroughs in the past years.
-Alt, Hubner and Henning extended GPT-2 to relation extraction, which is a task of identifying the relationship between concepts appeared in text.
+@alt2019fine extended GPT-2 to relation extraction, which is a task of identifying the relationship between concepts appeared in text.
 They had shown a milestone in predicting larger range of distinct relation types with higher confidence **(Ke-Li add ref)**.
 In the field of speech recognition, data scarcity is often an issue due to the difficulty to collect large amount of data from human for learning **(Ke-Li add ref)**.
-Bird, Faria, Ekart, Premebida and Ayrosa employed GPT-2 as part of the solution to generate augmented data for classification.
+@bird2020lstm employed GPT-2 as part of the solution to generate augmented data for classification.
 GPT-2 also helped NLP researchers tackling the task of textual paraphrasing **(Ke-Li add ref)**.
-Hedge and Patil used GPT-2 to build an unsupervised paraphrasing model that is not restricted by domain shift within an article, which is a problem faced by the usual supervised models of paraphrasing.
+@hegde2020unsupervised used GPT-2 to build an unsupervised paraphrasing model that is not restricted by domain shift within an article, which is a problem faced by the usual supervised models of paraphrasing.
 **(Ke-Li please find papers that use it, describe them, and add refs - look on arXiv)**.
 
-GPT-3 is a Transformer-based model that has over 175 billions of parameters.
-**(Keli add citation)**.
-Proposed by Vaswani et al., the Transformer is a novel network architecture that outperforms RNN-based models in computational efficiency **(Keli add citation)**.
+GPT-3 is a Transformer-based model that has over 175 billions of parameters (@brown2020language).
+Proposed by @vaswani2017attention, the Transformer is a novel network architecture that outperforms RNN-based models in computational efficiency **(Keli add citation)**.
 The parameters refer to the weights of the connections that enable the neural network to learn **(Keli add citation)**.
@@
 Without the demonstrations, GPT-3 tends to treat the prompts as "tweets" and generate texts that serve as responses or follow-up tweets.
-The team fed a few previous news articles in the model’s context as demonstrations, and the model learned to produce the outputs that adapt to the style and length of news stories **(Keli add citation)**.
+The team fed a few previous news articles in the model’s context as demonstrations, and the model learned to produce the outputs that adapt to the style and length of news stories (@brown2020language).
 This few-shot learning ability indicates that the model can learn like human do and would democratize the use of language models to broader fields.

### 2020-07-28 outputs/documents/paper.Rmd
Revisited and add to the section of GPT-2. Add literatures on GloVe, Word2Vec and BERT

@@
 
-### Article: Class-Based n-gram Models of Natural Language (1992)
-Syntax-based (grammartical) and semantic-based (sensible) word classifications in n-gram models.
+## Word Embedding Models: GloVe and Word2Vec
 
-### Article: A Statistical Approach to Automatic OCR Error Correction in Context (1996)
-Context-sensitive correction system to both non-word and real-word errors based on n-gram models applied to OCR postprocessing.
+**Article: GloVe: Global Vectors for Word Representation**
 
-## Pre-trained language models
+GloVe is a pre-trained word embedding model that can capture semantic meanings of words.
+It is an unsupervised learning algorithm developed by Stanford.
+Their disadvantage is that they are context-free and fail to capture higher-level concepts in context.
 
-Draw on and cite: 'Language Models are Few-Shot Learners'
+**Article: word2vec Explained: Deriving Mikolov et al.’s Negative-Sampling Word-Embedding Method**
+
+Developed by Google, Word2Vec is another word embedding model that uses shallow neural network. It's not for deep learning.
+
+## Pre-trained transformer language models
 
@@
 Because of this, there has been substantial development in NLP research.
+OpenAI and Google are two companies that are considered pioneers in artificial intelligence and they both develop pre-trained models that are considered state-of-art; we will discuss the GPT models from OpenAI and the BERT model from Google.
 
-OpenAI's GPT-3 is a generative, pre-trained, task-agnostic, language model that was announced by OpenAI in June 2020 **(Ke-Li add ref)**.
+### GPT
+
+Draw on and cite: 'Language Models are Few-Shot Learners'
+
 OpenAI is an AI research company who openly pursues Artificial General Intelligence—to have machines learn and understand any intellectual tasks as human do **(Ke-Li add ref)**.
-GPT-3 is the third Generative Pre-trained Transformer language model publicly released by OpenAI.
-It has '175 billion parameters' and can learn NLP tasks with just a few examples.
-OpenAI claims that the text generation by GPT-3 is 'difficult for humans to distinguish from authentic human content' (@brown2020language).
+OpenAI's Generative Pre-Trained Transformer (GPT) models are pre-trained language models that are built based on the transformer architecture.
+Proposed by @vaswani2017attention from Google in 2017, the Transformer is a novel network architecture that outperforms RNN-based models in computational efficiency and CNN-based in (something that it doesn't have to look for longer sentences) **(Keli add citation)**.
+[Write about self-attention layer and how it differentiate from RNN and CNN].
+
+As of August 2020, OpenAI has three generations of GPT models.
+GPT was introduced to the public in June, 2018.
+In the year 2019, its successor GPT-2 was released in three stages.
+Not long after, in June 2020, OpenAI announced GPT-3, which is the latest GPT model as of July 2020 (@brown2020language).
 Since the API of GPT-3 at the moment is only available to a few invited users, public applications of GPT-3 are limited.
@@
 @hegde2020unsupervised used GPT-2 to build an unsupervised paraphrasing model that is not restricted by domain shift within an article, which is a problem faced by the usual supervised models of paraphrasing.
-**(Ke-Li please find papers that use it, describe them, and add refs - look on arXiv)**.
 
-GPT-3 is a Transformer-based model that has over 175 billions of parameters (@brown2020language).
-Proposed by @vaswani2017attention, the Transformer is a novel network architecture that outperforms RNN-based models in computational efficiency **(Keli add citation)**.
+GPT-2 has the capacity to adapt to a context and generates custom output with the "fine-tuning" technique **(Ke-Li add ref)**.
+While the pre-training is unsupervised, fine-tuning is conducted by supervised learning or reinforced learning (@ziegler2019fine).
+**(Ke-Li add more about fine-tuning)** We aim to use GPT-2 to clean textual data to **(Ke-Li add more about how we use fine-tuning to train GPT-2 for text cleaning)**
+
+GPT-3 has over 175 billions of parameters (@brown2020language).
 The parameters refer to the weights of the connections that enable the neural network to learn **(Keli add citation)**.
@@
 The first generation GPT has 110 million parameters.
-Released in 2018, GPT-2 has 1.6 billion parameters, and within two years, GPT-3 has grown the size to 175 billion, which is over 100 times than its predecessor .
+Released in 2018, GPT-2 has 1.6 billion parameters, and within two years, GPT-3 has grown the size to 175 billion, which is over 100 times than its predecessor.
 The more parameters there are the better the model can generalize to new tasks; therefore, the increasing amount of parameters accelerates the improvements in text synthesis and downstream NLP tasks **(Keli add citation)**.
 GPT-3 can learn NLP tasks with just a handful of example; OpenAI refers this ability as few-shot learning, or in-context learning.
-Few-shot learning means that the model  **(Keli add citation)**.
+**(Keli add citation)**.
 As a result, we only need to provide GPT-3 a few demonstrations for it to learn a new NLP task.
-**(Ke-Li add more detail about what it needs and what it outputs and how it works).**.
 For example, OpenAI team trained GPT-3 to generate "news articles".
@@
 This few-shot learning ability indicates that the model can learn like human do and would democratize the use of language models to broader fields.
-**(Ke-Li please find the language that they use for this and then we need to make ours consistent with theirs).**
+
+### BERT
+
+**Article: Bert: Pre-training of deep bidirectional transformers for language understanding**
+
+This paper is published by researchers at Google AI Language in 2018.
+The key innovation of BERT is its application of bidirectional training of the Transformer model, which results in a deeper sense of language context and flow.
 

### 2020-07-29 outputs/documents/paper.Rmd
Updated write-ups and added more literatures

@@
 
+## Statistical Language Model
+
+Without exception, all human languages have some words that co-occur more frequently with others.
+Under this assumption, we can regard the production of English text as a set of conditional probabilities, written as Pr($w_{k}$ | $w^{k-1}_{1}$), where k is the number of words in a sentence, $w_{k}$ is the predicted word, and $w^{k-1}_{1}$ is the history of the word occurring in a sequence (@brown1992class).
+In other word, the generation of prediction $w_{k}$ is based on the history $w^{k-1}_{1}$; this is the overarching principle of all language models.
+Essentially, a language model is the probability distribution over sequences of words **(Ke-Li add ref)**.
+The goal of statistical language modeling is to estimate the probability distribution of different linguistic units — words, sentences, and even documents **(Ke-Li add ref)**.
+However, this is fundamentally difficult because language is categorical.
+If we see each word in a set of vocabulary as a category, the dimensions of language is extremely high (@rosenfeld2000two).
+In the course of statistical language model development, [how n-gram, word embedding, transformers tackle this].
+
 ## PDF text extraction and n-grams
@@
 It needs to first be extracted from the PDF, often by optical character recognition (OCR), and then cleaned, prepared and made into some type of tabular dataset.
-
 One issue in this process is that OCR is not perfect.
@@
 This was no longer the case in modern correction techniques as statistical language models (SLMs) and feature-based methods have been used for context-sensitive correction (citation).
-Without exception, all human languages have some words that co-occur more frequently with others.
-Under this assumption, we can regard the production of English text as a set of conditional probabilities, written as Pr($w_{k}$ | $w^{k-1}_{1}$), where k is the number of words in a sentence, $w_{k}$ is the predicted word, and $w^{k-1}_{1}$ is the history of the word occurring in a sequence (@brown1992class).
-In other word, the generation of prediction $w_{k}$ is based on the history $w^{k-1}_{1}$.
-This conditional probability is the foundation of an n-gram language model.
 
+This conditional probability of the language model is the foundation of an n-gram language model.
 An n-gram model is a probabilistic language model that predicts the next word in a sequence of words.
@@
 
-## Word Embedding Models: GloVe and Word2Vec
+The n-gram models work do not take the linguistic structure of language into account.
+As @rosenfeld2000two pointed out, "...
+it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind".
+The prediction of the next word is only from a few words that are immediately preceding; the context is not taking into account.
+For this shortcoming, the preditive power of n-gram is largely limited.
+
+## Word Embedding Models
+Moved from statistical to neural network.
+Word embedding models works by representing words as numerical dense vectors so the neural networks can understand and process.
+Representing words as dense vectors has a long history, but the first neural network language model was proposed in 2013 by @bengio2003neural, and became the pioneer in word embedding models.
+In the work of @bengio2003neural, word embedding works by encoding the words into a feature vector.
+The feature vector represents different aspects of the word and each word is associated with a point in a vector space.
+
+The words that are similar will have similar representations in a vector space.
+The similarity of the words, however, does not depend on their spelling or forms, but by their co-occurrence in given contexts.
+For example, a "cat" is more similar to a "dog" instead of a "car" because we are more likely to talk about cats and dogs in a same conversation.
+Therefore, the co-occurrence of "cat" and "dog" has higher probability than the co-occurrence of "cat" and a "car".
+In other words, the distributed representation is learned by the usage of words.
+This permits words that are used in similar contexts to result in having similar vector representations and their semantic meanings are captured naturally.
+
+**Article: word2vec Explained: Deriving Mikolov et al.’s Negative-Sampling Word-Embedding Method**
+
+### Word2Vec
+
+Developed by Google, Word2Vec is a word embedding model that uses shallow neural network.
+While it is not a deep learning model, it turns words into a vector representations that deep neural networks can understand.
+It takes a corpus as its input and output a set of feature vectors that represent the words in the given corpus.
+@mikolov2013efficient proposed two architectures for learning distributed representations of words within the framework of Word2Vec — Continuous Bag-of-Words Model and Continuous Skip-gram Model.
+
+Example:
+>  vector("King") - vector("Man") + vector("Woman") = vector("Queen")
 
@@
 
-GloVe is a pre-trained word embedding model that can capture semantic meanings of words.
-It is an unsupervised learning algorithm developed by Stanford.
+GloVe stands for global vectors for word representation. It is a pre-trained word embedding model that can capture semantic meanings of words developed by Stanford. 
+It is an unsupervised learning algorithm. 
 Their disadvantage is that they are context-free and fail to capture higher-level concepts in context.
-
-**Article: word2vec Explained: Deriving Mikolov et al.’s Negative-Sampling Word-Embedding Method**
-
-Developed by Google, Word2Vec is another word embedding model that uses shallow neural network. It's not for deep learning.
 
@@
 OpenAI's Generative Pre-Trained Transformer (GPT) models are pre-trained language models that are built based on the transformer architecture.
-Proposed by @vaswani2017attention from Google in 2017, the Transformer is a novel network architecture that outperforms RNN-based models in computational efficiency and CNN-based in (something that it doesn't have to look for longer sentences) **(Keli add citation)**.
+Proposed by @vaswani2017attention from Google in 2017, the Transformer is a novel network architecture for neural network that outperforms RNN-based models in computational efficiency and CNN-based in (something that it doesn't have to look for longer sentences) **(Keli add citation)**.
 [Write about self-attention layer and how it differentiate from RNN and CNN].
+[Paraphrase: The transformer is based solely on attention mechanisms, dispensing with recurrence and convolutions entirely, and has proven extremely effective compared to its predecessor the LSTM].
 
@@
 
-GPT-2 has the capacity to adapt to a context and generates custom output with the "fine-tuning" technique **(Ke-Li add ref)**.
-While the pre-training is unsupervised, fine-tuning is conducted by supervised learning or reinforced learning (@ziegler2019fine).
-**(Ke-Li add more about fine-tuning)** We aim to use GPT-2 to clean textual data to **(Ke-Li add more about how we use fine-tuning to train GPT-2 for text cleaning)**
+Most of prior language models are trained in a single domain by supervised learning to perform a single task.
+OpenAI, on the other hands, train the GPT models to perform many tasks (@radford2019language).
+GPT-2 is pre-trained on WebText, a comprehensive data set that is scrapped from millions of webpages and contains total of 40 GB of text (@radford2019language).
+WebText contains all kinds of textual data: dialogues, comments, translations, questions and answers...etc.
+Such diversity is critical for GPT-2 to be a generalized model that can perform multiple tasks with any domain boundary.
+However, GPT-2 has the capacity to adapt to a context and generates custom output with the "fine-tuning" technique **(Ke-Li add ref)**.
+While the pre-training stage is unsupervised and mandatory, fine-tuning is an optional training stage where GPT can learn to perform new tasks in a supervised or reinforced approach (@ziegler2019fine) **(Ke-Li add examples)**.
+During the pre-training stage, the model is trained to perform text completion task.
+Give it a prompt to start and it will generate texts that continues what has been started.
+In the fine-tuning stage, @radford2019language provided additional training data sets for GPT-2 to generate the texts with specific sentiment, or be descriptive, or even summarize the prompt text.
+
+In our project, we aim to use fine-tune GPT-2 to perform a text cleaning task **(Ke-Li add more about how we use fine-tuning to train GPT-2 for text cleaning)**.
+
+A language model is just a probability:
+> p(output | input)
+
+To model all NLP tasks, we cab represent them with one extra conditioning:
+> p(output | input, task)
+
+OpenAI has made GPT-2 open-sourced through Github with Python and Tensorflow.
+There is also a package in R that wraps the Python code into a R package, created by **(Keli add citation)**.
+However, the R package only permits exploratory experiments with the pre-trained tasks; fine-tuning for new tasks is not available thorough R.
 

### 2020-07-29 outputs/documents/paper.Rmd
Updated write-ups and added more literatures

@@
 ## Word Embedding Models
-Moved from statistical to neural network.
-Word embedding models works by representing words as numerical dense vectors so the neural networks can understand and process.
+[Moved from statistical to neural network] Word embedding models works by representing words as numerical dense vectors so the neural networks can understand and process.
 Representing words as dense vectors has a long history, but the first neural network language model was proposed in 2013 by @bengio2003neural, and became the pioneer in word embedding models.

### 2020-07-30 outputs/documents/paper.Rmd
Language models and Word embedding models

--- 2020-07-29 outputs/documents/paper.Rmd
@@
 ## Word Embedding Models
-Moved from statistical to neural network.
-Word embedding models works by representing words as numerical dense vectors so the neural networks can understand and process.
-Representing words as dense vectors has a long history, but the first neural network language model was proposed in 2013 by @bengio2003neural, and became the pioneer in word embedding models.
-In the work of @bengio2003neural, word embedding works by encoding the words into a feature vector.
+Turning words into numeric representations is an important step for machine to understand and process words.
+The most basic way to turn a word into a vector.
+However, in one-hot encoding, the number of dimensions depends on the size of the vocabulary in a given corpus, and can easily grow to extremely large.
+Word embedding models works by representing words as numerical dense vectors so the dimensions are reduced by representing the words by features **(Keli add citation)**.
+Representing words as dense vectors has a long history.
+but the idea of a dense embedding space for words was first explored in 2003 by @bengio2003neural, and became the prominent driver for word embedding models.
+In the work of @bengio2003neural, word embedding works by encoding the words into feature vectors.
 The feature vector represents different aspects of the word and each word is associated with a point in a vector space.
-
 The words that are similar will have similar representations in a vector space.
@@
 
+### Word2Vec
+
 **Article: word2vec Explained: Deriving Mikolov et al.’s Negative-Sampling Word-Embedding Method**
-
-### Word2Vec
 

### 2020-07-31 outputs/documents/paper.Rmd
Updates to Word2Vec section

--- 2020-07-29 outputs/documents/paper.Rmd
@@
 
-## Statistical Language Model
+## Language Model
 
@@
 ## Word Embedding Models
-Moved from statistical to neural network.
-Word embedding models works by representing words as numerical dense vectors so the neural networks can understand and process.
-Representing words as dense vectors has a long history, but the first neural network language model was proposed in 2013 by @bengio2003neural, and became the pioneer in word embedding models.
-In the work of @bengio2003neural, word embedding works by encoding the words into a feature vector.
-The feature vector represents different aspects of the word and each word is associated with a point in a vector space.
-
+Turning words into numeric representations is an important step for machine to understand and process words.
+The most basic way to turn a word into a vector is one-hot encoding.
+However, with one-hot encoding, the number of dimensions depends on the size of the vocabulary in a given corpus, and can easily grow to extremely large number of dimensions.
+Word embedding models work by representing words as numerical dense vectors so the dimensions are reduced by representing the words by features **(Keli add citation)**.
+Representing words as dense vectors has a long history, but the idea of a dense embedding space for words was first explored in 2003 by @bengio2003neural, and became the prominent driver for word embedding models.
+In the work of @bengio2003neural, words are encoded into feature vectors that represents different aspects of the word, and each word is associated with a point in a vector space.
 The words that are similar will have similar representations in a vector space.
@@
 
-**Article: word2vec Explained: Deriving Mikolov et al.’s Negative-Sampling Word-Embedding Method**
+[The word representations computed using neural networks are very interesting because the learned vectors explicitly encode many linguistic regularities and patterns.
+Somewhat surprisingly, many of these patterns can be represented as linear translations.
+For example, the result of a vector calculation vec("Madrid") - vec("Spain") + vec("France") is closer to vec("Paris") than to any other word vector.]
 
 ### Word2Vec
+
+**Article: word2vec Explained: Deriving Mikolov et al.’s Negative-Sampling Word-Embedding Method**
 
@@
 It takes a corpus as its input and output a set of feature vectors that represent the words in the given corpus.
-@mikolov2013efficient proposed two architectures for learning distributed representations of words within the framework of Word2Vec — Continuous Bag-of-Words Model and Continuous Skip-gram Model.
+@mikolov2013efficient proposed two architectures for learning distributed representations of words within the framework of Word2Vec — Continuous Bag-of-Words Model (CBOW) and Continuous Skip-gram Model.
+CBOW uses several surrounding words, or the context, to predict a target word.
+On the other hand, continuous skip-gram takes a single word as input to predict its neighbouring words.
+For example, consider the sentence "I think therefore I am", a famous quote by Descartes.
+The focus word is we want to learn the embedding is set to "therefore", and the context window is set to two which will give us the two immediate preceding words and two immediate proceeding words of the focus word.
+The input for the CBOW is "I, think, I, am" and the output is "therefore".
+In contrast, the skip-gram model will take "therefore" as the input and out "I, think, I, am" as the prediction.
 
-Example:
+Word2Vec is trained on the Google News dataset that contains 6 billion tokens and the vocabulary size is restricted to one million most common words (@mikolov2013efficient).
+
+Word2Vec use cases:
+1. Find the most similar words (**Keli add example**)
+2. Calculate the similarities between a pair of words (**Keli add example**)
+3. Find the odd word in a list of words (**Keli add example**)
+4. Perform linear algebra to vectorized words. Example:
 >  vector("King") - vector("Man") + vector("Woman") = vector("Queen")

### 2020-08-01 outputs/documents/paper.Rmd
Writing updates. Add one literacture for BERT.

@@
 The prediction of the next word is only from a few words that are immediately preceding; the context is not taking into account.
-For this shortcoming, the preditive power of n-gram is largely limited.
+For this shortcoming, the predictive power of n-gram is largely limited.
 
@@
 
-[The word representations computed using neural networks are very interesting because the learned vectors explicitly encode many linguistic regularities and patterns.
-Somewhat surprisingly, many of these patterns can be represented as linear translations.
-For example, the result of a vector calculation vec("Madrid") - vec("Spain") + vec("France") is closer to vec("Paris") than to any other word vector.]
-
 ### Word2Vec
-
-**Article: word2vec Explained: Deriving Mikolov et al.’s Negative-Sampling Word-Embedding Method**
 
@@
 @mikolov2013efficient proposed two architectures for learning distributed representations of words within the framework of Word2Vec — Continuous Bag-of-Words Model (CBOW) and Continuous Skip-gram Model.
+Both architectures contains three layers: input layer, projection layer and output layer.
+The input layer encodes words within a context window to one-hot vectors.
+Using a projection matrix, the input layer is  projected to a projection layer and the features are.
+Finally, the output layer is a softmax regression classifier that obtains the posterior distribution of words (@rong2014word2vec).
+The input and output vectors differ in the two architecutres.
 CBOW uses several surrounding words, or the context, to predict a target word.
@@
 
-Word2Vec is trained on the Google News dataset that contains 6 billion tokens and the vocabulary size is restricted to one million most common words (@mikolov2013efficient).
+Word2Vec is trained on the Google News data set that contains 6 billion tokens and the vocabulary size is restricted to one million most common words (@mikolov2013efficient).
+The enormous size of data set catches accurate statistical pattern of language and results a comprehensive mapping of word relations to a vector place.
+The positions of the word vectors reflect their semantic relationships between each other.
+Therefore, the Word2Vec model performs especially well in word analogy tasks through answering questions such as "$king$ is to $man$ as $queen$ is to __?" @pennington2014glove.
+The vectorization of words allows the operation of linear algebra for the word vectors in order to get the answer:
 
-Word2Vec use cases:
+>  vector("King") - vector("Man") + vector("Woman") = vector("Queen")
+
+Other Word2Vec use cases:
 1. Find the most similar words (**Keli add example**)
@@
 3. Find the odd word in a list of words (**Keli add example**)
-4. Perform linear algebra to vectorized words. Example:
->  vector("King") - vector("Man") + vector("Woman") = vector("Queen")
 
-**Article: GloVe: Global Vectors for Word Representation**
+### GloVe
 
-GloVe stands for global vectors for word representation. It is a pre-trained word embedding model that can capture semantic meanings of words developed by Stanford. 
-It is an unsupervised learning algorithm. 
-Their disadvantage is that they are context-free and fail to capture higher-level concepts in context.
+GloVe stands for global vectors for word representation.
+It is a pre-trained word embedding model that can capture semantic meanings of words developed by Stanford.
+In the Word2Vec model, the model learn word presentations locally within a context window.
+On the other hand, GloVe considers the properties of the corpus globally.
+The benefit of the global approach is that the model can learn about the global statistics of an entire corpus.
+GloVe can then obtain the co-occurrence probabilities by creating a co-occurrence count matrix for the entire training data set.
+However, GloVe goes beyond calculating distribution of word by the probability of co-occurrence; it reveals properties of words by the ratio of the probability of co-occurrence of words.
+For example, consider the two words we want to obtain the word vectors from "ice" and "steam".
+GloVe reveals their properties by examining their co-occurrence with a variety of probe words, $k$.
+$K$ can be "solid", which is the property of ice; "gas", which is the property of steam; and "water", which is related to both "ice" and "steam".
+The ratio of co-occurrence probability is calculated as the following:
+
+> P(k|ice)/P(k|steam)
+
+The ratio between the co-occurrence probability of "ice, solid" pair and "steam, solid" pair is high because the co-occurrence of "ice" and "solid" is higher than the co-occurrence of "steam" and "solid".
+Reversely, the ratio between the co-occurrence of "ice, gas" pair and "steam, gas" pair is low.
+Other other hand, the ratio of co-occurrence probability of "ice, water" pair and "steam, water" pair are close to 1 because water is relevant to both "ice" and "steam".
+Probability alone does not distinguish well relevant words from irrelevant words.
+For this reason, the researcher of GloVe argued that "the appropriate starting point for word vector learning should be with ratios of co-occurrence probabilities rather than the probabilities themselves (@pennington2014glove)."
 
@@
 
-**Article: Bert: Pre-training of deep bidirectional transformers for language understanding**
+BERT stands for Bidirectional Encoder Representations from Transformers (@devlin2018bert).
+It was released by researchers at Google AI Language in 2018.
+The key innovation of BERT is its application of bidirectional training of the Transformer model, which results in a deeper sense of language context and flow.
+What is bidirectional training?
+In the context of English language, a sentence is constructed from left to right.
+The context on the left is used to determine the next word on the right.
+In this sense, GPT models are unidirectional since the probability of the next word is dependent only on the past.
+It is also possible to reverse the direction and model the sentence from right to left, and BERT incorporates both directions of left-to-right and right-to-left to model bidirectional contexts.
+BERT uses the bidirectional approach to alleviate by "masked language model" (MLM).
+The MLM model masks certain words in the input corpus and predict the original words that are masked.
+The objective is to fuse the representation from the left and right context to enable the pre-training of the deep bidirectional Transformer.
+For example:
 
-This paper is published by researchers at Google AI Language in 2018.
-The key innovation of BERT is its application of bidirectional training of the Transformer model, which results in a deeper sense of language context and flow.
+Because BERT employs the Transformer architecture with the self-attention mechanism, fine-tuning is also available.
 

### 2020-08-03 outputs/documents/paper.Rmd
Update first four pages

@@
 author:
+- name: Ke-Li Chiu
+  affiliation: University of Toronto
 - name: Rohan Alexander
-  affiliation: University of Toronto
-- name: Ke-Li Chiu
   affiliation: University of Toronto
@@
 This is used to monitor changes during the cleaning process, and to compare the messiness of different texts.
-The methods that we consider are: n-grams, [Google’s Word2vec, Standford's GloVe], GPT-2, and GPT-3.
+The methods that we consider are: n-grams, Google's Word2vec, Stanford's GloVe, and OpenAI's GPT-2 and GPT-3.
 We illustrate our process on text from the Canadian Hansard and introduce both a Shiny application and an R package to make our process easier for others to adopt."
-keywords: "text-as-data; natural language processing; GPT-3; quantitative analysis."
+keywords: "text-as-data; natural language processing; quantitative analysis."
 date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
 In this paper we introduce the concept of consistency in a text corpus.
-Consistency refers to the proportion of words that are able to be forecast by a statistical model, based on preceeding words and surrounding context.
+Consistency refers to the proportion of words that are able to be forecast by a statistical model, based on preceding words and surrounding context.
 Further, we define internal consistency as when the model is trained on the corpus itself, and external consistency as when the model is trained on a more general corpus.
@@
 By examining how the consistency scores evolve in response to changes made to the text during the data preparation and cleaning stages we can better understand the effect of the changes.
-Including consistency scores when text corpuses are shared allows researchers to be more transparent about their corpus.
+Including consistency scores and their evolution when text corpuses are shared allows researchers to be more transparent about their corpus.
 And finally, the use of consistency scores allows for an increased level of automation in the cleaning process.
@@
 We consider various natural language processing methods.
-These include n-grams, [Google’s Word2vec, Standford's GloVe], as well as GPT2 and GPT-3.
+These include n-grams, Google's Word2vec, Stanford's GloVe, and OpenAI's GPT-2 and GPT-3.
 The n-gram approach involves identifying two, three, or more, words that are commonly found together.
@@
 For that reason we consider pre-trained word embedding models.
-These include [Google’s Word2vec, Standford's GloVe], which works [representing words as dense vectors that capture their semantic relationships with each other]**SOMEHOW**.
-We also consider pre-trained generative models such as GPT-2 and GPT-3 from OpenAI.
+These include Google's Word2vec and Stanford's GloVe, which place each word in a multi-dimensional space such that distance between words can illustrate their relationship.
+We also consider pre-trained generative models such as OpenAI's GPT-2 and GPT-3.
 GPT-2 and GPT-3 are pre-trained generative unsupervised language models.
 They differ in the number of parameters.
-Generative Pretrained Transformer (GPT)-2 has *X* parameters and GPT-3 has more than 175 billions parameters.
-While GPT-2 is more limited, it can be run on smaller machines, whereas GPT-3 cannot.
+Generative Pretrained Transformer (GPT)-2 has 1.5 billion parameters and GPT-3 has more than 175 billion parameters.
+While GPT-2 is more limited, it can be run on smaller machines, whereas GPT-3 cannot and is accessed via an API.
 
@@
 
-Without exception, all human languages have some words that co-occur more frequently with others.
-Under this assumption, we can regard the production of English text as a set of conditional probabilities, written as Pr($w_{k}$ | $w^{k-1}_{1}$), where k is the number of words in a sentence, $w_{k}$ is the predicted word, and $w^{k-1}_{1}$ is the history of the word occurring in a sequence (@brown1992class).
-In other word, the generation of prediction $w_{k}$ is based on the history $w^{k-1}_{1}$; this is the overarching principle of all language models.
-Essentially, a language model is the probability distribution over sequences of words **(Ke-Li add ref)**.
-The goal of statistical language modeling is to estimate the probability distribution of different linguistic units — words, sentences, and even documents **(Ke-Li add ref)**.
-However, this is fundamentally difficult because language is categorical.
-If we see each word in a set of vocabulary as a category, the dimensions of language is extremely high (@rosenfeld2000two).
-In the course of statistical language model development, [how n-gram, word embedding, transformers tackle this].
+Given the nature of human languages, some combinations of words tend to occur more frequently than others.
+Think of 'good', which is more often followed by 'morning', than 'duck'.
+As such, we could consider English text production as a conditional probability, $\mbox{Pr}(w_{k} | w^{k-1}_{1})$, where $k$ is the number of words in a sentence, $w_{k}$ is the predicted word, and $w^{k-1}_{1}$ is the history of the word occurring in a sequence [@brown1992class].
+In this way, the generation of some prediction, $w_{k}$, is based on the history, $w^{k-1}_{1}$.
+This is the underlying principle of all language models.
+Essentially, a language model is a probability distribution over sequences of words **(Ke-Li add ref)**.
+The goal of statistical language modeling is to estimate probability distributions over different linguistic units — words, sentences, and even documents **(Ke-Li add ref)**.
+However, this is difficult as language is categorical.
+If we consider each word in a vocabulary as a category, then the dimensionality of language becomes large [@rosenfeld2000two].
+The reason that there is scuh a variety of statistical language models is that there are various ways of dealing with this fundamental problem.
 
@@
 
-PDFs contain a lot of information, but typically that information is not able to be analysed directly by statistical models.
-It needs to first be extracted from the PDF, often by optical character recognition (OCR), and then cleaned, prepared and made into some type of tabular dataset.
+The Portable Document Format (PDF) is typically used to render documents in a relatively  fixed way.
+PDFs contain a lot of information, but often that information is not able to be analysed directly by statistical language models.
+The information first needs to be extracted from the PDF, often by optical character recognition (OCR), and then cleaned, prepared, and made into some type of dataset.
 One issue in this process is that OCR is not perfect.
-Hence there is a need to correct the output before it can be analysed.
-Traditionally, word-correction techniques evaluate errors one by one without considering the context of the surrounding words.
-This was no longer the case in modern correction techniques as statistical language models (SLMs) and feature-based methods have been used for context-sensitive correction (citation).
+Hence there is a need to correct the output before it is analysed.
+One way to do this is to correct errors one-by-one and independent of context.
+We all know that 'tbe' is wrong, and should be 'the'.
+But statistical language models have the advantage of correction that can occur within context.
+This allows significant improvements, for instance sometimes 'bat' should be 'hat' and sometimes it really should be 'bat' **(Ke-Li add ref)**.
 
-This conditional probability of the language model is the foundation of an n-gram language model.
-An n-gram model is a probabilistic language model that predicts the next word in a sequence of words.
-The $n$ in n-gram represents the number of words in a sequence.
-Taking an incomplete sentence from the Jane Eyre excerpt as an example:
+The foundation of an n-gram language model is the conditional probability set-up introduced above.
+An n-gram model is a probabilistic language model that predicts the next word in a sequence of words **(Ke-Li add ref)**.
+The $n$ in n-gram referes to the number of words in that sequence.
+Consider the following excerpt from *Jane Eyre*:
 
-> "We had been wandering,"
+> 'We had been wandering'.
 
-"We" is a uni-gram, "We had" is a bi-gram, "We had been" is a tri-gram, and "We had been wandering" is a 4-gram.
-On the other hand, the tri-grams of the excerpt are "We had been," and "had been wandering." The overlapping chaining of words reveals the statistical behaviour in human language.
-Moreover, using n-gram models then enable us to assign a probability to the occurrence of a sequence of words or the likelihood of a word occurring next.
-Consider the two sentences: "We had been wandering," and "We had been wangling." The former is likely to be more frequently encountered in a training corpus.
-Thus, the n-gram would assign a higher probability to "wandering" than to "wangling."
+'We' is a uni-gram, 'We had' is a bi-gram, 'We had been' is a tri-gram, and 'We had been wandering' is a 4-gram.
+Notice that the two tri-grams in this excerpt, 'We had been', and 'had been wandering', overlap.
+The use of n-gram models enables us to assign probabilities to both the next sequence of words and just the next word.
+For instance, consider the two sentences: 'We had been wandering', and 'We had been wangling'.
+The former is likely to be more frequently encountered in a training corpus.
+Thus, an n-gram would assign a higher probability to the next word being 'wandering' than 'wangling', given the sequence 'We had been'.
 
-To predict which word appears next, we have to take the sequence of preceeding words into account, which requires knowing the probability of the sequence of words.
+To predict the next word, we have to take the sequence of preceding words into account, which requires knowing the probability of the sequence of words.
 The probability of a sequence appearing in a corpus follows the chain rule:
 
-> P(We,had,been,wandering) $=$ P(We)$\cdot$P(had|We)$\cdot$P(been|We,had)$\cdot$P(wandering|We,had,been)
+$$\mbox{Pr(We,had,been,wandering)} = \mbox{Pr(We)} \times \mbox{P(had|We)} \times \mbox{P(been|We,had)} \\ 
+\times \mbox{P(wandering|We,had,been)}.$$
 
-However, the likelihood that more and more words will occur next to each other in an identical sequence becomes smaller and smaller, making the prediction difficult.
-Alternatively, we can approximate the probability of a word depending on only the previous word; this is known as Markov assumption.
-Instead, the Markov assumption permits us to approximate the probability using only the last $n$ words (@brown1992class).
-More specifically:
+However, the likelihood that more and more words will occur next to each other in an identical sequence becomes smaller and smaller, making prediction difficult.
+Alternatively, we can approximate the probability of a word depending on only the previous word.
+This is known as the 'Markov assumption' and it allows us to approximate the probability using only the last $n$ words [@brown1992class]:
 
-> P(We,had,been,wandering)$\approx$ P(We)$\cdot$P(had|We)$\cdot$P(been|had)$\cdot$P(wandering|been)
+$$\mbox{Pr(We,had,been,wandering)} \approx \mbox{Pr(We)} \times \mbox{P(had|We)} \times \mbox{P(been|had)} \\ 
+\times \mbox{P(wandering|been)}.$$
 
-Because the bi-gram model looks only the immediate preceding word, an n-gram model under Markov assumption can then be reduced into a bi-gram model with n being any number. The general equation is:
+As a bi-gram model only considers the immediately preceding word, under the Markov assumption, an $n$-gram model can be reduced to a bi-gram model with $n$ being any number:
 
-> P($w_{n} | w^{n-1}_{1}) \approx$ P($w_{n} | w^{n-1}_{n - N+1})$)
+$$\mbox{Pr}(w_{n} | w^{n-1}_{1}) \approx \mbox{P}(w_{n} | w^{n-1}_{n - N+1}))$$
+where $N$ is **(Ke-Li add)**.
 
-The application of n-gram is versatile.
-N-gram models are widely applied in text prediction, spelling-correction, machine translation **(Keli add citation)**.
+Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation **(Ke-Li add citation)**.
 In our application, we use a tri-gram based model to detect both real-word and non-word errors and correct them with the candidate word that has the highest probability.
-[add more about how we use n-gram once the functions in the app are in place]
+**(TODO: Add more about how we use n-gram once the functions in the app are in place.)**
 
-The n-gram models work do not take the linguistic structure of language into account.
-As @rosenfeld2000two pointed out, "...
-it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind".
-The prediction of the next word is only from a few words that are immediately preceding; the context is not taking into account.
-For this shortcoming, the predictive power of n-gram is largely limited.
+However, n-gram models do not take the linguistic structure of language into account.
+For instance, @rosenfeld2000two [p.
+X]  **(Ke-Li add)** discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'.
+The prediction of the next word is based on only a few preceding words and broader context is not taken into account.
+Hence next-word prediction using n-gram based language models can be limited.
 
@@
 Finally, the output layer is a softmax regression classifier that obtains the posterior distribution of words (@rong2014word2vec).
-The input and output vectors differ in the two architecutres.
+The input and output vectors differ in the two architectures.
 CBOW uses several surrounding words, or the context, to predict a target word.
@@
 For example, consider the sentence "I think therefore I am", a famous quote by Descartes.
-The focus word is we want to learn the embedding is set to "therefore", and the context window is set to two which will give us the two immediate preceding words and two immediate proceeding words of the focus word.
+The focus word is we want to learn the embedding is set to "therefore", and the context window is set to two which will give us the two immediately preceding words and two immediate proceeding words of the focus word.
 The input for the CBOW is "I, think, I, am" and the output is "therefore".
@@
 Reversely, the ratio between the co-occurrence of "ice, gas" pair and "steam, gas" pair is low.
-Other other hand, the ratio of co-occurrence probability of "ice, water" pair and "steam, water" pair are close to 1 because water is relevant to both "ice" and "steam".
+On the other hand, the ratio of co-occurrence probability of "ice, water" pair and "steam, water" pair are close to 1 because water is relevant to both "ice" and "steam".
 Probability alone does not distinguish well relevant words from irrelevant words.
@@
 
-OpenAI has made GPT-2 open-sourced through Github with Python and Tensorflow.
+OpenAI has made GPT-2 open-sourced through GitHub with Python and TensorFlow.
 There is also a package in R that wraps the Python code into a R package, created by **(Keli add citation)**.

### 2020-08-03 scripts/00-have_a_go_at_gpt2.R

@@ -1,0 +1,21 @@
+# Based on: https://blogs.rstudio.com/ai/posts/2019-10-23-gpt-2/
+
+
+
+remotes::install_github("r-tensorflow/gpt2")
+
+gpt2::install_gpt2(envname = "r-gpt2")
+
+library(gpt2)
+
+install.packages("reticulate")
+library(reticulate)
+use_virtualenv("r-tensorflow")
+install.packages("keras")
+library(keras)
+import("scipy")
+
+install_gpt2()
+
+gpt2(model = "124M")
+

### 2020-08-03 outputs/documents/paper.Rmd
recover previous changes in citations in later paragraphs

@@
 This is the underlying principle of all language models.
-Essentially, a language model is a probability distribution over sequences of words **(Ke-Li add ref)**.
-The goal of statistical language modeling is to estimate probability distributions over different linguistic units — words, sentences, and even documents **(Ke-Li add ref)**.
+Essentially, a language model is a probability distribution over sequences of words.
+The goal of statistical language modeling is to estimate probability distributions over different linguistic units — words, sentences, and even documents (@bengio2003neural).
 However, this is difficult as language is categorical.
@@
 
-Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation **(Ke-Li add citation)**.
+Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation  (@brown1992class).
 In our application, we use a tri-gram based model to detect both real-word and non-word errors and correct them with the candidate word that has the highest probability.
@@
 ## Word Embedding Models
+
 Turning words into numeric representations is an important step for machine to understand and process words.
@@
 However, with one-hot encoding, the number of dimensions depends on the size of the vocabulary in a given corpus, and can easily grow to extremely large number of dimensions.
-Word embedding models work by representing words as numerical dense vectors so the dimensions are reduced by representing the words by features **(Keli add citation)**.
+Word embedding models work by representing words as numerical dense vectors so the dimensions are reduced by representing the words by features (@bengio2003neural).
 Representing words as dense vectors has a long history, but the idea of a dense embedding space for words was first explored in 2003 by @bengio2003neural, and became the prominent driver for word embedding models.
@@
 This permits words that are used in similar contexts to result in having similar vector representations and their semantic meanings are captured naturally.
+As a result, word embedding models are capable of finding the most similar words, calculate the similarities between a pair of words, and find the odd word in a group of words @mikolov2013efficient.
+The most commonly used word embedding models are Word2Vec and GloVe, which will be described in the following sections.
 
@@
 
-Other Word2Vec use cases:
-1. Find the most similar words (**Keli add example**)
-2. Calculate the similarities between a pair of words (**Keli add example**)
-3. Find the odd word in a list of words (**Keli add example**)
-
 ### GloVe
@@
 
-In recent years, the use of pre-trained language models has played a significant role in advancing natural language processing (NLP) tasks such as reading comprehension, text generation, and even creative writing **(Ke-Li add ref)**.
+In recent years, the use of pre-trained language models has played a significant role in advancing natural language processing (NLP) tasks such as reading comprehension, text generation, and even creative writing (@brown2020language).
 Pre-trained models remove the need for researchers to prepare training datasets or allot computational resources to the stage of pre-training.
@@
 
-Draw on and cite: 'Language Models are Few-Shot Learners'
-
-OpenAI is an AI research company who openly pursues Artificial General Intelligence—to have machines learn and understand any intellectual tasks as human do **(Ke-Li add ref)**.
+OpenAI is an AI research company who openly pursues Artificial General Intelligence — to have machines learn and understand any intellectual tasks as human do **(Cite OpenAI blog)**.
 OpenAI's Generative Pre-Trained Transformer (GPT) models are pre-trained language models that are built based on the transformer architecture.
-Proposed by @vaswani2017attention from Google in 2017, the Transformer is a novel network architecture for neural network that outperforms RNN-based models in computational efficiency and CNN-based in (something that it doesn't have to look for longer sentences) **(Keli add citation)**.
-[Write about self-attention layer and how it differentiate from RNN and CNN].
-[Paraphrase: The transformer is based solely on attention mechanisms, dispensing with recurrence and convolutions entirely, and has proven extremely effective compared to its predecessor the LSTM].
+Proposed by @vaswani2017attention from Google in 2017, the Transformer is a novel network architecture for neural network that outperforms RNN-based and CNN-based models in computational efficiency (@vaswani2017attention).
+Since the emergence of the Transformer model, most of the representative pre-trained models are built based on this architecture (@hanretty2018comparing).
+Therefore, we can say that Transformer marks a new era of language models.
 
@@
 @alt2019fine extended GPT-2 to relation extraction, which is a task of identifying the relationship between concepts appeared in text.
-They had shown a milestone in predicting larger range of distinct relation types with higher confidence **(Ke-Li add ref)**.
-In the field of speech recognition, data scarcity is often an issue due to the difficulty to collect large amount of data from human for learning **(Ke-Li add ref)**.
+They had shown a milestone in predicting larger range of distinct relation types with higher confidence (@alt2019fine).
+In the field of speech recognition, data scarcity is often an issue due to the difficulty to collect large amount of data from human for learning (@bird2020lst).
 @bird2020lstm employed GPT-2 as part of the solution to generate augmented data for classification.
-GPT-2 also helped NLP researchers tackling the task of textual paraphrasing **(Ke-Li add ref)**.
+GPT-2 also helped NLP researchers tackling the task of textual paraphrasing.
 @hegde2020unsupervised used GPT-2 to build an unsupervised paraphrasing model that is not restricted by domain shift within an article, which is a problem faced by the usual supervised models of paraphrasing.
@@
 Such diversity is critical for GPT-2 to be a generalized model that can perform multiple tasks with any domain boundary.
-However, GPT-2 has the capacity to adapt to a context and generates custom output with the "fine-tuning" technique **(Ke-Li add ref)**.
-While the pre-training stage is unsupervised and mandatory, fine-tuning is an optional training stage where GPT can learn to perform new tasks in a supervised or reinforced approach (@ziegler2019fine) **(Ke-Li add examples)**.
+However, GPT-2 has the capacity to adapt to a context and generates custom output with the "fine-tuning" technique (@ziegler2019fine).
+While the pre-training stage is unsupervised and mandatory, fine-tuning is an optional training stage where GPT can learn to perform new tasks in a supervised or reinforced approach such as stylistic continuation and summarization (@ziegler2019fine).
 During the pre-training stage, the model is trained to perform text completion task.
@@
 
-In our project, we aim to use fine-tune GPT-2 to perform a text cleaning task **(Ke-Li add more about how we use fine-tuning to train GPT-2 for text cleaning)**.
-
-A language model is just a probability:
-> p(output | input)
-
-To model all NLP tasks, we cab represent them with one extra conditioning:
-> p(output | input, task)
-
-OpenAI has made GPT-2 open-sourced through GitHub with Python and TensorFlow.
-There is also a package in R that wraps the Python code into a R package, created by **(Keli add citation)**.
+OpenAI has made GPT-2 open-sourced through Github with Python and Tensorflow.
+There is also a package in R that wraps the Python code into a R package, created by @keydanaluraschi2019gpt2.
 However, the R package only permits exploratory experiments with the pre-trained tasks; fine-tuning for new tasks is not available thorough R.
+In our project, we aim to use fine-tune GPT-2 to perform a text cleaning task **(Ke-Li add more about how we use fine-tuning to train GPT-2 for text cleaning, or try with other task if text cleaning is not achieved)**.
 
 GPT-3 has over 175 billions of parameters (@brown2020language).
-The parameters refer to the weights of the connections that enable the neural network to learn **(Keli add citation)**.
 In the past few years, the growth of parameters of transformer language models expands rapidly.
@@
 Released in 2018, GPT-2 has 1.6 billion parameters, and within two years, GPT-3 has grown the size to 175 billion, which is over 100 times than its predecessor.
-The more parameters there are the better the model can generalize to new tasks; therefore, the increasing amount of parameters accelerates the improvements in text synthesis and downstream NLP tasks **(Keli add citation)**.
-GPT-3 can learn NLP tasks with just a handful of example; OpenAI refers this ability as few-shot learning, or in-context learning.
-**(Keli add citation)**.
+The more parameters there are the better the model can generalize to new tasks; therefore, the increasing amount of parameters accelerates the improvements in text synthesis and downstream NLP tasks (@brown2020language).
+GPT-3 can learn NLP tasks with just a handful of example; OpenAI refers this ability as few-shot learning, or in-context learning (@brown2020language).
 As a result, we only need to provide GPT-3 a few demonstrations for it to learn a new NLP task.
@@
 
-Because BERT employs the Transformer architecture with the self-attention mechanism, fine-tuning is also available.
+> **(Keli adds example)**
+
+Because BERT employs the Transformer architecture with the self-attention mechanism, fine-tuning is also available.[To be continued...]
 

### 2020-08-03 outputs/documents/paper.Rmd
Update through to the pre-trained transformer language models section

@@
 title: "Consistency scores in text data"
-thanks: "Thank you to X, Y and Z for helpful comments.
+thanks: "Thank you to **X, Y and Z** for helpful comments.
 A Shiny application for interactive and small-scale examples is available: https://kelichiu.shinyapps.io/Arianna/.
@@
 This is used to monitor changes during the cleaning process, and to compare the messiness of different texts.
-The methods that we consider are: n-grams, Google's Word2vec, Stanford's GloVe, and OpenAI's GPT-2 and GPT-3.
+The methods that we consider are: n-grams, word2vec, GloVe, and the GPT models.
 We illustrate our process on text from the Canadian Hansard and introduce both a Shiny application and an R package to make our process easier for others to adopt."
@@
 We consider various natural language processing methods.
-These include n-grams, Google's Word2vec, Stanford's GloVe, and OpenAI's GPT-2 and GPT-3.
+These include n-grams, word2vec, GloVe, and the GPT models.
 The n-gram approach involves identifying two, three, or more, words that are commonly found together.
@@
 At scale these can identify missing or unusual words, and work quickly, but they lack nuance.
-For instance an equally reasonable two-gram involving the word 'good' is 'good work'.
-For that reason we consider pre-trained word embedding models.
-These include Google's Word2vec and Stanford's GloVe, which place each word in a multi-dimensional space such that distance between words can illustrate their relationship.
-We also consider pre-trained generative models such as OpenAI's GPT-2 and GPT-3.
+For instance, an equally reasonable two-gram involving the word 'good' is 'good work'.
+For that reason, we consider pre-trained word embedding models.
+These include word2vec and GloVe, which place each word in a multi-dimensional space such that distance between words can illustrate their relationship.
+We also consider pre-trained generative models such as GPT-2 and GPT-3.
 GPT-2 and GPT-3 are pre-trained generative unsupervised language models.
 They differ in the number of parameters.
-Generative Pretrained Transformer (GPT)-2 has 1.5 billion parameters and GPT-3 has more than 175 billion parameters.
+GPT-2 has 1.5 billion parameters and GPT-3 has more than 175 billion parameters.
 While GPT-2 is more limited, it can be run on smaller machines, whereas GPT-3 cannot and is accessed via an API.
@@
 Essentially, a language model is a probability distribution over sequences of words.
-The goal of statistical language modeling is to estimate probability distributions over different linguistic units — words, sentences, and even documents (@bengio2003neural).
+The goal of statistical language modeling is to estimate probability distributions over different linguistic units — words, sentences, and even documents [@bengio2003neural].
 However, this is difficult as language is categorical.
 If we consider each word in a vocabulary as a category, then the dimensionality of language becomes large [@rosenfeld2000two].
-The reason that there is scuh a variety of statistical language models is that there are various ways of dealing with this fundamental problem.
+The reason that there is such a variety of statistical language models is that there are various ways of dealing with this fundamental problem.
 
@@
 One way to do this is to correct errors one-by-one and independent of context.
-We all know that 'tbe' is wrong, and should be 'the'.
+We all know that 'tbe' is wrong and should be 'the'.
 But statistical language models have the advantage of correction that can occur within context.
@@
 An n-gram model is a probabilistic language model that predicts the next word in a sequence of words **(Ke-Li add ref)**.
-The $n$ in n-gram referes to the number of words in that sequence.
-Consider the following excerpt from *Jane Eyre*:
-
-> 'We had been wandering'.
-
+The $n$ in n-gram refers to the number of words in that sequence.
+Consider the following excerpt from *Jane Eyre*: 'We had been wandering'.
 'We' is a uni-gram, 'We had' is a bi-gram, 'We had been' is a tri-gram, and 'We had been wandering' is a 4-gram.
@@
 
-$$\mbox{Pr(We,had,been,wandering)} = \mbox{Pr(We)} \times \mbox{P(had|We)} \times \mbox{P(been|We,had)} \\ 
-\times \mbox{P(wandering|We,had,been)}.$$
+$$\mbox{Pr(We,had,been,wandering)} = \mbox{Pr(We)} \times \mbox{Pr(had|We)} \times \mbox{Pr(been|We,had)} \\ 
+\times \mbox{Pr(wandering|We,had,been)}.$$
 
@@
 
-$$\mbox{Pr(We,had,been,wandering)} \approx \mbox{Pr(We)} \times \mbox{P(had|We)} \times \mbox{P(been|had)} \\ 
-\times \mbox{P(wandering|been)}.$$
+$$\mbox{Pr(We,had,been,wandering)} \approx \mbox{Pr(We)} \times \mbox{Pr(had|We)} \times \mbox{Pr(been|had)} \\ 
+\times \mbox{Pr(wandering|been)}.$$
 
@@
 For instance, @rosenfeld2000two [p.
-X]  **(Ke-Li add)** discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'.
+X]  **(Ke-Li add page please)** discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'.
 The prediction of the next word is based on only a few preceding words and broader context is not taken into account.
@@
 
-## Word Embedding Models
+**(Ke-Li add brief write-up of implementation in R please. Probably quanteda unless there's a better option.)**
 
-Turning words into numeric representations is an important step for machine to understand and process words.
-The most basic way to turn a word into a vector is one-hot encoding.
-However, with one-hot encoding, the number of dimensions depends on the size of the vocabulary in a given corpus, and can easily grow to extremely large number of dimensions.
-Word embedding models work by representing words as numerical dense vectors so the dimensions are reduced by representing the words by features (@bengio2003neural).
-Representing words as dense vectors has a long history, but the idea of a dense embedding space for words was first explored in 2003 by @bengio2003neural, and became the prominent driver for word embedding models.
-In the work of @bengio2003neural, words are encoded into feature vectors that represents different aspects of the word, and each word is associated with a point in a vector space.
-The words that are similar will have similar representations in a vector space.
-The similarity of the words, however, does not depend on their spelling or forms, but by their co-occurrence in given contexts.
-For example, a "cat" is more similar to a "dog" instead of a "car" because we are more likely to talk about cats and dogs in a same conversation.
-Therefore, the co-occurrence of "cat" and "dog" has higher probability than the co-occurrence of "cat" and a "car".
-In other words, the distributed representation is learned by the usage of words.
-This permits words that are used in similar contexts to result in having similar vector representations and their semantic meanings are captured naturally.
-As a result, word embedding models are capable of finding the most similar words, calculate the similarities between a pair of words, and find the odd word in a group of words @mikolov2013efficient.
-The most commonly used word embedding models are Word2Vec and GloVe, which will be described in the following sections.
+## Word embedding models
 
-### Word2Vec
+Although, from an implementation perspective, there is an inherent numerical representation of language in n-gram models, it is not overly tractable.
+The numeric representation typically relies on each term being represented in a vector, which can quickly result in quite high-dimensional objects.
+In contrast, word embedding models represent terms by vectors of real numbers where each vector is a 'feature', hence the final object has considerably lower dimensionality [@bengio2003neural].
 
-Developed by Google, Word2Vec is a word embedding model that uses shallow neural network.
-While it is not a deep learning model, it turns words into a vector representations that deep neural networks can understand.
-It takes a corpus as its input and output a set of feature vectors that represent the words in the given corpus.
-@mikolov2013efficient proposed two architectures for learning distributed representations of words within the framework of Word2Vec — Continuous Bag-of-Words Model (CBOW) and Continuous Skip-gram Model.
-Both architectures contains three layers: input layer, projection layer and output layer.
-The input layer encodes words within a context window to one-hot vectors.
+This approach has a long history, but the implementation of @bengio2003neural has become the foundation for much subsequent work.
+Here, terms are encoded into feature vectors that represent different aspects of the term, and so each term is represented by a position in the vector space.
+Terms that are used in similar ways will have similar representations in the vector space.
+This will be due to their co-occurrence in usage.
+For example, 'cat' will be closer to 'dog' than to 'car', because we are more likely to talk about cats and dogs, than cats and cars.
+See how they might be close when considered along a 'pets' dimension, but far when considered along a 'related-to-tigers' dimension.
+
+Terms that are often used in similar contexts will have similar vector representations.
+Hence, word embedding models can identify similar terms, calculate the similarity between a pair of terms, and find the odd term in a group of terms [@mikolov2013efficient].
+There are a variety of ways to achieve this dimensionality reduction, including word2vec and GloVe.
+
+### Word2vec
+
+Word2vec is a word embedding model that uses a neural network to learn word associations [@mikolov2013efficient] [@mikolov2013distributed].
+It takes a corpus as an input and outputs a set of feature vectors that allow the terms in that corpus to be represented.
+@mikolov2013efficient proposed two architectures for determining this distributed representation of terms: continuous bag-of-words (CBOW), and continuous skip-gram.
+Both have an input layer, a projection layer, and an output layer.
+But CBOW uses several surrounding words, or the context, to predict a target word.
+On the other hand, continuous skip-gram takes a single word as input to predict its neighbouring words.
+
+To illustrate the difference, consider the quote from Descartes, 'I think therefore I am'.
+The focus word for which we want to learn the embedding is 'therefore', and the context window is set to two.
+This means that we consider the two words immediately preceding and proceeding 'therefore'.
+The CBOW input is 'I, think, I, am', and the output is 'therefore'.
+In contrast, the continuous skip-gram input is 'therefore', while 'I, think, I, am' is the output.
+The input layer encodes terms within a context window to one-hot vectors.
 Using a projection matrix, the input layer is  projected to a projection layer and the features are.
-Finally, the output layer is a softmax regression classifier that obtains the posterior distribution of words (@rong2014word2vec).
-The input and output vectors differ in the two architectures.
-CBOW uses several surrounding words, or the context, to predict a target word.
-On the other hand, continuous skip-gram takes a single word as input to predict its neighbouring words.
-For example, consider the sentence "I think therefore I am", a famous quote by Descartes.
-The focus word is we want to learn the embedding is set to "therefore", and the context window is set to two which will give us the two immediately preceding words and two immediate proceeding words of the focus word.
-The input for the CBOW is "I, think, I, am" and the output is "therefore".
-In contrast, the skip-gram model will take "therefore" as the input and out "I, think, I, am" as the prediction.
+**(Ke-Li: Please fix this sentence.)** Finally, the output layer is a softmax regression classifier that obtains the posterior distribution of words (@rong2014word2vec).
+It is the input and output vectors that differ between the two architectures.
 
-Word2Vec is trained on the Google News data set that contains 6 billion tokens and the vocabulary size is restricted to one million most common words (@mikolov2013efficient).
-The enormous size of data set catches accurate statistical pattern of language and results a comprehensive mapping of word relations to a vector place.
-The positions of the word vectors reflect their semantic relationships between each other.
-Therefore, the Word2Vec model performs especially well in word analogy tasks through answering questions such as "$king$ is to $man$ as $queen$ is to __?" @pennington2014glove.
-The vectorization of words allows the operation of linear algebra for the word vectors in order to get the answer:
+Word2vec is trained on the Google News dataset **(Ke-Li add cite to the dataset please.)**.
+This contains six billion tokens and the vocabulary size is restricted to the one million most common words (@mikolov2013efficient).
+As this is a large dataset, it provides an accurate statistical pattern of language, as it is used in this dataset, and a comprehensive mapping between terms and vector positions.
 
->  vector("King") - vector("Man") + vector("Woman") = vector("Queen")
+To the extent that the relative positions of terms reflects their semantic relationship, word2vec will perform well for word analogies.
+However, terms that have context-specific meaning may not perform as well.
+
+**(Ke-Li add brief write-up of implementation in R please.)**
 
@@
 
-GloVe stands for global vectors for word representation.
-It is a pre-trained word embedding model that can capture semantic meanings of words developed by Stanford.
-In the Word2Vec model, the model learn word presentations locally within a context window.
+The global vectors (GloVe) model is a pre-trained word embedding model that emphasises the semantic meaning of words [@pennington2014glove].
+In the word2vec model, word representations are learnt locally within the context window.
 On the other hand, GloVe considers the properties of the corpus globally.
-The benefit of the global approach is that the model can learn about the global statistics of an entire corpus.
-GloVe can then obtain the co-occurrence probabilities by creating a co-occurrence count matrix for the entire training data set.
-However, GloVe goes beyond calculating distribution of word by the probability of co-occurrence; it reveals properties of words by the ratio of the probability of co-occurrence of words.
-For example, consider the two words we want to obtain the word vectors from "ice" and "steam".
-GloVe reveals their properties by examining their co-occurrence with a variety of probe words, $k$.
-$K$ can be "solid", which is the property of ice; "gas", which is the property of steam; and "water", which is related to both "ice" and "steam".
-The ratio of co-occurrence probability is calculated as the following:
+As the model considers the context of an entire corpus the co-occurrence probabilities are based on counts for the entire training dataset.
 
-> P(k|ice)/P(k|steam)
+GloVe additionally considers the relative probability of word co-occurrence.
+For instance, consider the two words 'ice' and 'steam'.
+GloVe considers their co-occurrence with a variety of probe words, $k$.
+For instance $k$ could be 'solid', which is a property of ice; 'gas', which is a property of steam; and 'water', which is related to both.The ratio of co-occurrence probability is:
 
-The ratio between the co-occurrence probability of "ice, solid" pair and "steam, solid" pair is high because the co-occurrence of "ice" and "solid" is higher than the co-occurrence of "steam" and "solid".
-Reversely, the ratio between the co-occurrence of "ice, gas" pair and "steam, gas" pair is low.
-On the other hand, the ratio of co-occurrence probability of "ice, water" pair and "steam, water" pair are close to 1 because water is relevant to both "ice" and "steam".
-Probability alone does not distinguish well relevant words from irrelevant words.
-For this reason, the researcher of GloVe argued that "the appropriate starting point for word vector learning should be with ratios of co-occurrence probabilities rather than the probabilities themselves (@pennington2014glove)."
+$$\frac{\mbox{Pr}(k|\mbox{ice})}{\mbox{Pr}(k|\mbox{steam})}.$$
+
+The ratio between the co-occurrence probability of the 'ice, solid' pair and the 'steam, solid' pair is high because the co-occurrence of 'ice' and 'solid' is higher than the co-occurrence of 'steam' and 'solid'.
+Conversely, the ratio between the co-occurrence of the 'ice, gas' pair and the 'steam, gas' pair is low.
+On the other hand, the ratio of the co-occurrence probability of the 'ice, water' pair and the 'steam, water' pair us close to one because water is relevant to both 'ice' and 'steam'.
+
+**(Ke-Li add brief paragraph about negatives of this approach.)**
+
+**(Ke-Li add brief write-up of implementation in R please.)**
 
@@
 
-In recent years, the use of pre-trained language models has played a significant role in advancing natural language processing (NLP) tasks such as reading comprehension, text generation, and even creative writing (@brown2020language).
+In recent years, the use of pre-trained language models has played a significant role in advancing natural language processing (NLP) tasks such as reading comprehension, text generation, and even creative writing [@brown2020language].
 Pre-trained models remove the need for researchers to prepare training datasets or allot computational resources to the stage of pre-training.
 Because of this, there has been substantial development in NLP research.
-OpenAI and Google are two companies that are considered pioneers in artificial intelligence and they both develop pre-trained models that are considered state-of-art; we will discuss the GPT models from OpenAI and the BERT model from Google.
+In this section we discuss the GPT models from OpenAI and the BERT model from Google.
 

### 2020-08-03 outputs/documents/paper.Rmd
notes on to-dos

@@
 For instance, @rosenfeld2000two [p.
-X]  **(Ke-Li add page please)** discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'.
+1]  **(Ke-Li add page please)** discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'.
 The prediction of the next word is based on only a few preceding words and broader context is not taken into account.
@@
 **(Ke-Li add brief write-up of implementation in R please. Probably quanteda unless there's a better option.)**
+The data set we used for training the n-gram model is [**(Ke-Li add a data set once settled the choice)**].
+For generating n-grams, we used the Text Analysis Utilities (tau) R package to tokenize the training corpus into tri-grams.
 
@@
 **(Ke-Li add brief paragraph about negatives of this approach.)**
+#### Notes not content ####
+- Inability to handle unknown or out-of-vocabulary (OOV) words.
+- No shared representations at sub-word levels
+- Word presentations are not contextual. Representation of apple is the same in context of fruit and tech company
+- [Although these pre-trained embeddings can capture semantic meanings of words, they are context-free and fail to capture higher-level concepts in context, such as polysemous disambiguation, syntactic structures, semantic roles, anaphora.] (@hanretty2018comparing).
 
 **(Ke-Li add brief write-up of implementation in R please.)** 
+GloVe word embedding is available to operate in R with the text2vec package developed by Dmitriy Selivanov **(Ke-Li add package citation)**.
 

### 2020-08-04 outputs/documents/paper.Rmd
n-gram implementation write up

@@
 In our application, we use a tri-gram based model to detect both real-word and non-word errors and correct them with the candidate word that has the highest probability.
-**(TODO: Add more about how we use n-gram once the functions in the app are in place.)**
 
@@
 For instance, @rosenfeld2000two [p.
-1]  **(Ke-Li add page please)** discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'.
+1] discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'.
 The prediction of the next word is based on only a few preceding words and broader context is not taken into account.
@@
 **(Ke-Li add brief write-up of implementation in R please. Probably quanteda unless there's a better option.)**
-The data set we used for training the n-gram model is [**(Ke-Li add a data set once settled the choice)**].
-For generating n-grams, we used the Text Analysis Utilities (tau) R package to tokenize the training corpus into tri-grams.
+To implement the n-gram model, we used the Text Analysis Utilities (tau) R package to tokenize the training corpus (citation('tau')).
+To ensure that the three-word sequences are split within a sentence, we split the corpus into sentences by period, exponential mark, or question marks.
+We then manually added an end of sentence token $<s>$ to the end of each sentence.
+We used the textcnt() function turn the training corpus into tri-grams, and the function outputs a textcnt object that resembles a table object that has contains the tri-grams and their counts.
+We then removed all the tri-grams that contains an end-of-sentence token, so the tri-grams that include a word from a new sentence will be removed.
+For example, the tri-gram "day </s> we" will be removed because it has words from two different sentences.
+The following is the sample output that show the first 4 tokens:
+
+```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
+tabl <- " 
+| there was no  | a rain so | a walk that | an hour in  | 
+|---------------|-----------|-------------|-------------|
+| 2             | 1         | 1           | 1           |
+"
+cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
+```
+We can then access the tri-grams by calling the names of the object and use their counts to proceed to probability calculation.
+In calculating the probability of the tri-grams, we applied Kneser–Ney smoothing, the smoothing technique recommended by @chen1999empirical to improve the accuracy of n-gram models **(This function is written manually, but it's not my own code.
+I am reading about the Kneser–Ney smoothing to understand the math, will replace the function afterwards)**.
+Finally, we create a data table that contains the first two words and the last of each tri-gram in separate column and the smoothed probability in another column.
+The example row of the table is the following:
+
+```{r table2, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
+tabl <- " 
+| First Words   | Last Word | Probability |
+|---------------|-----------|-------------|
+| there was     | no        | 0.6278947   | 
+"
+cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
+```
 
@@
 The input layer encodes terms within a context window to one-hot vectors.
-Using a projection matrix, the input layer is  projected to a projection layer and the features are.
-**(Ke-Li: Please fix this sentence.)** Finally, the output layer is a softmax regression classifier that obtains the posterior distribution of words (@rong2014word2vec).
+Using a projection matrix, the projection layer then project the discrete one-hot vectors to lower-dimension continuous vectors.
+Finally, the output layer is a softmax regression classifier that obtains the posterior distribution of words (@rong2014word2vec).
 It is the input and output vectors that differ between the two architectures.
@@
 **(Ke-Li add brief write-up of implementation in R please.)**
+#### Notes ####
+CBOW might be better for text cleaning because it uses the context to predict the center word.
 

### 2020-08-04 shinyapp/functions.R

@@ -4,6 +4,9 @@ library(tm)
 library(stringr)
 library(stats)
 
+library(reshape2)
+library(RWeka)
+
 #setwd("~/Desktop/repos/consistency_scores_in_text_datasets/shinyapp")
 path <- getwd()
 demoTraining <- readLines(paste(path,"/samples/demo.txt", sep=""), warn=FALSE)
@@ -14,7 +17,7 @@ tokenizer <- function(corpus) {
   # Read the corpus line by line
   for (line in corpus) {
     # add two start-of-sentence tokens to the beginning of a sentence
-    line <- paste("<s> <s>", line)
+    #line <- paste("<s> <s>", line)
     lines <-c(lines, line)
   }
   
@@ -42,6 +45,7 @@ getLastWords <- function(string, words) {
   pattern <- paste("[a-z']+( [a-z']+){", words - 1, "}$", sep="")
   return(substring(string, str_locate(string, pattern)[,1]))
 }
+
 removeLastWord <- function(string) {
   sub(" [a-z']+$", "", string)
 }
@@ -73,6 +77,7 @@ kneserNay <- function(ngrams, d) {
   return(probabilities)
 }
 
+# Get the probabilities after smoothing
 unigramProbs <- kneserNay(unigrams, 0.75)
 bigramProbs <- kneserNay(bigrams, 0.75)
 trigramProbs <- kneserNay(trigrams, 0.75)
@@ -81,11 +86,11 @@ library(data.table)
 
 unigramDF <- data.table("Words" = (names(unigrams)), 
                         "Probability" = as.vector(unigramProbs), stringsAsFactors=F)
-bigramsDF <- data.table("FirstWords" = removeLastWord(names(bigrams)), 
-                        "LastWord" = getLastWords(names(bigrams), 1), 
+bigramsDF <- data.table("First Words" = removeLastWord(names(bigrams)), 
+                        "Las tWord" = getLastWords(names(bigrams), 1), 
                         "Probability" = as.vector(bigramProbs), stringsAsFactors=F)
-trigramsDF <- data.table("FirstWords" = removeLastWord(names(trigrams)), 
-                         "LastWord" = getLastWords(names(trigrams), 1), 
+trigramsDF <- data.table("Firs tWords" = removeLastWord(names(trigrams)), 
+                         "Last Word" = getLastWords(names(trigrams), 1), 
                          "Probability" = as.vector(trigramProbs), stringsAsFactors=F)
 
 

### 2020-08-05 outputs/documents/paper.Rmd
implementations on Word2Vec and GloVe, WIP

@@
 **(Ke-Li add brief write-up of implementation in R please.)**
-#### Notes ####
-CBOW might be better for text cleaning because it uses the context to predict the center word.
+There are a few packages for implementing a Word2Vec model in R.
+The packages are "rword2vec" by Mukul Chaware and "word2vec" by Jan Wijffels, BNOSAC, and Max Fomichev.
+Both packages uses the same algorithm that Word2Vec uses to produce word representations, but they do not have an affiliation with Google.
+Another way to implement a Word2Vec model in R is through Keras package.
+Keras is an API for deep learning written in Python (@chollet2015keras) and the Keras package provides an R interface to Keras **(Ke-Li add citation for the package)**.
+RStudio published a tutorial for implementing the skip-gram model of Word2Vec (@falbel2017word); we intend to follow this tutorial as the approach to implement a Word2Vec model.
 
@@
 **(Ke-Li add brief paragraph about negatives of this approach.)**
-#### Notes not content ####
-- Inability to handle unknown or out-of-vocabulary (OOV) words.
-- No shared representations at sub-word levels
-- Word presentations are not contextual. Representation of apple is the same in context of fruit and tech company
-- [Although these pre-trained embeddings can capture semantic meanings of words, they are context-free and fail to capture higher-level concepts in context, such as polysemous disambiguation, syntactic structures, semantic roles, anaphora.] (@hanretty2018comparing).
+**(This limitation paragraph applies to both Word2Vec and GloVe)**
+As @hanretty2018comparing point out, although word embedding models can capture semantic means of words, "they are context-free and fail to capture higher-level concepts in context, such as polysemous disambiguation, syntactic structures, semantic roles, anaphora".
+Word embedding models do not represent a homonyms that has difference meanings in different vectors depending on the context.
+For example, "squash" can be a type of hard shelled vegetable or can be referring to a racket and ball sport, but there is only one series of vectors to represent the word "squash", instead of dynamically changing the vectors depending on the context.
+Moreover, from our observation, the word embedding models are unable to handle novel words that are not included in the vocabulary of the training corpus; for this reason, we conclude that the generalizability of word embedding models is sub-optimal.
 
-**(Ke-Li add brief write-up of implementation in R please.)** 
-GloVe word embedding is available to operate in R with the text2vec package developed by Dmitriy Selivanov **(Ke-Li add package citation)**.
+There are more than one way to implement GloVe word embedding model in R.
+One way is to load GloVe embeddings into a frozen Keras Embedding layer **(Keli add citation to the blog post by Keras)** using the Keras R package (**(Keli add citation to package)**).
+Another way, as mentioned and briefly demonstrated by Rstudio in their blog post, is to use the "text2vec" package developed by Dmitriy Selivanov **(Ke-Li add package citation)**.
 

### 2020-08-05 outputs/documents/paper.Rmd
implementations on Word2Vec and GloVe

@@
 Keras is an API for deep learning written in Python (@chollet2015keras) and the Keras package provides an R interface to Keras **(Ke-Li add citation for the package)**.
-RStudio published a tutorial for implementing the skip-gram model of Word2Vec (@falbel2017word); we intend to follow this tutorial as the approach to implement a Word2Vec model.
+RStudio published a tutorial for implementing the skip-gram model of Word2Vec (@falbel2017word) through Keras.
+We intend to follow this tutorial as the approach to implement a Word2Vec model.
+Following the demonstration, functions are defined to mimic the skip-gram algorithm of Word2Vec.
+The embedding dimensions, the size of the context window and the number of samples are configurable.
+As a result, a embedding matrix is obtained and serves as the look-up table of the word vectors.
 
@@
 Another way, as mentioned and briefly demonstrated by Rstudio in their blog post, is to use the "text2vec" package developed by Dmitriy Selivanov **(Ke-Li add package citation)**.
+Keras published a blog post on pre-trained word embeddings through GloVe in R.
+In this project, we intend to follow the blog post for implementing GloVe through the Keras R package **(Keli add citation to the blog post by Keras)**.
+The blog post is aimed to create a word embedding model for text classification.
+However, we will only use the guide create a embedding matrix and skip the steps of text classification.
+Similar to the Word2Vec embedding matrix, the GloVe embedding matrix is obtained and serves as a look-up table.
 
@@
 However, the R package only permits exploratory experiments with the pre-trained tasks; fine-tuning for new tasks is not available thorough R.
-In our project, we aim to use fine-tune GPT-2 to perform a text cleaning task **(Ke-Li add more about how we use fine-tuning to train GPT-2 for text cleaning, or try with other task if text cleaning is not achieved)**.
+In our project, we aim to use fine-tune GPT-2 to perform a text cleaning task.
 
@@
 Because BERT employs the Transformer architecture with the self-attention mechanism, fine-tuning is also available.[To be continued...]
+
+**(Keli adds implementation of BERT in R)**
+BERT implementation in R is also possible through Keras.
+Zhao HG (**Replace by citation**) created a python module called Keras-bert and the module can be called by R through reticulate package, which enables interoperability between Python and R.
 

### 2020-08-05 outputs/documents/paper.Rmd
BERT write up

@@
 It is also possible to reverse the direction and model the sentence from right to left, and BERT incorporates both directions of left-to-right and right-to-left to model bidirectional contexts.
-BERT uses the bidirectional approach to alleviate by "masked language model" (MLM).
-The MLM model masks certain words in the input corpus and predict the original words that are masked.
-The objective is to fuse the representation from the left and right context to enable the pre-training of the deep bidirectional Transformer.
-For example:
 
-> **(Keli adds example)**
+BERT is pre-trained by using two unsupervised tasks.
+The first task is Masked Language Model and and second task is Next Sentence Prediction.
+The bidirectional approach of BERT is alleviated by the "Masked Language Model" (@devlin2018bert).
+The Masked Language Model masks certain words in the input corpus which forces the model to predict the original words by "filling in the blank".
+This training process blends the representation from the left and right context and therefore increases the model's context sensitivity (@nozza2020mask).
+In the Next Sentence Prediction process, the model receives two sentences in sequence as input.
+50% of the sentences are paired in the right sequence and the ones from the other half have sentences that are paired randomly (@devlin2018bert).
+The goal of this process is to have BERT evaluate if the second sentence is the next sentence of the first sentence in the original text and sharpen its capacity to predict the likelihood of the next sentence.
 
-Because BERT employs the Transformer architecture with the self-attention mechanism, fine-tuning is also available.[To be continued...]
+Because BERT employs the Transformer architecture with the self-attention mechanism, fine-tuning is also available with one additional output layer and a task-specific input (@devlin2018bert).
+Google AI researchers have fine-tuned BERT for 11 NLP tasks including question-answering and common-sense inference.
 
-**(Keli adds implementation of BERT in R)**
 BERT implementation in R is also possible through Keras.
-Zhao HG (**Replace by citation**) created a python module called Keras-bert and the module can be called by R through reticulate package, which enables interoperability between Python and R.
+Zhao HG (**Replace by citation**) created a python module called "keras-bert" and the module can be called by R through "reticulate" package (@reticulateR), which enables the operation of Python modules in R.
+To start, the "keras-bert" module is installed in Python environment and is called in R through the 'reticulate' package.
 

### 2020-08-05 outputs/documents/paper.Rmd
BERT implementation WIP

@@
 There is also a package in R that wraps the Python code into a R package, created by @keydanaluraschi2019gpt2.
-However, the R package only permits exploratory experiments with the pre-trained tasks; fine-tuning for new tasks is not available thorough R.
-In our project, we aim to use fine-tune GPT-2 to perform a text cleaning task.
+There are three choices of models in terms of parameter sizes — "124M", "355M" and "774M".
+The R package only permits exploratory experiments with the pre-trained tasks; fine-tuning for new tasks is not available thorough R.
+To proceed with the implementation of GPT-2 in R, we will fine-tune the model through Python and create a wrapper code for R (**Keli to check if this is possible**).
 
@@
 
+BERT was pre-trained on BooksCorpus data set with 800 million words and English Wikipedia with 2,500 million words (@devlin2018bert).
 Because BERT employs the Transformer architecture with the self-attention mechanism, fine-tuning is also available with one additional output layer and a task-specific input (@devlin2018bert).

### 2020-08-06 outputs/documents/paper.Rmd
BERT implementation WIP
BERT Word2VEc Glove implementations revisions

@@
 
-**(Ke-Li add brief write-up of implementation in R please.)**
 There are a few packages for implementing a Word2Vec model in R.
@@
 We intend to follow this tutorial as the approach to implement a Word2Vec model.
-Following the demonstration, functions are defined to mimic the skip-gram algorithm of Word2Vec.
+There are three basic steps within the Keras deep learning framework: first is to prepare the data, second is to define the model and third is to train and evaluate the model.
+Following the steps, we first downloaded a the Amazon Fine Foods Reviews dataset from Stanford **(Ke-Li add citation for data set)** and preprocess the text into tokens.
+Then we defined a model that mimics the skip-gram algorithm of Word2Vec.
 The embedding dimensions, the size of the context window and the number of samples are configurable.
-As a result, a embedding matrix is obtained and serves as the look-up table of the word vectors.
+Finally, we used the fit_generator() function of Keras to train the model.
+**What's supposed to happen:** As a result, a embedding matrix is obtained and serves as the look-up table of the word vectors.
+**What really happened:** However, the training process was stuck in the operation of fit_generator() function.
 
@@
 **(This limitation paragraph applies to both Word2Vec and GloVe)**
-As @hanretty2018comparing point out, although word embedding models can capture semantic means of words, "they are context-free and fail to capture higher-level concepts in context, such as polysemous disambiguation, syntactic structures, semantic roles, anaphora".
+As @hanretty2018comparing point out, although word embedding models can capture semantic meanings of words, "they are context-free and fail to capture higher-level concepts in context, such as polysemous disambiguation, syntactic structures, semantic roles, anaphora".
 Word embedding models do not represent a homonyms that has difference meanings in different vectors depending on the context.
@@
 There are more than one way to implement GloVe word embedding model in R.
-One way is to load GloVe embeddings into a frozen Keras Embedding layer **(Keli add citation to the blog post by Keras)** using the Keras R package (**(Keli add citation to package)**).
+One way is to load GloVe embeddings into a frozen Keras Embedding layer **(Keli add citation to the blog post by Keras)** using the Keras R package **(Keli add citation to package)**.
 Another way, as mentioned and briefly demonstrated by Rstudio in their blog post, is to use the "text2vec" package developed by Dmitriy Selivanov **(Ke-Li add package citation)**.
@@
 However, we will only use the guide create a embedding matrix and skip the steps of text classification.
-Similar to the Word2Vec embedding matrix, the GloVe embedding matrix is obtained and serves as a look-up table.
+Since we are employing Keras again, the process remains similar.
+We first retrieved the training data set from Stanford and preprocessed the text in to tokens.
+We then defined the model by configuring the embedding layer.
+Finally the model is trained and evaluated, and the GloVe embedding matrix is obtained.
 
@@
 To start, the "keras-bert" module is installed in Python environment and is called in R through the 'reticulate' package.
+We can then load, configure and train the BERT model.
+To train the BERT model, we load a task-specific dataset for it to learn.
+[(**Keli Add more details**)]
 
@@
 # References
+
+--
+There are a variety of ways to implement an n-gram model within R **[cite R]** including X, Y and Z.
+We used the Text Analysis Utilities, `tau`, R package **[cite tau]** because of X, Y, and Z.
+To illustrate the use of `tau`, we will first install and load the library, and create a small corpus.
+```{r, eval = FALSE}
+install.packages(“tau”)
+library(tau)
+corpus <- “There was no possibility of taking a walk that day.
+We had been wandering, indeed, in the leafless shrubbery an hour in the morning; but since dinner (Mrs.
+Reed, when there was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, that further out-door exercise was now out of the question.”
+```
+Our purpose is to construct tri-grams and so we first split the corpus into sentences by period, exponential mark, or question marks.
+We then manually added an end of sentence token $<s>$ to the end of each sentence.
+```{r}
+#Add simple code for however this is done in tua?
+```
+Next we remove all punctuation and other non-letter characters, and convert all text to lower case.
+```{r}
+#however this is done in tua?
+```
+We will use the `textcnt()` function to turn the training corpus into tri-grams.
+The most important argument of this function is the corpus, as well as X, Y and Z.
+The function outputs a `textcnt` object, which is essentially a table with each tri-gram and their counts.
+```{r}
+#code to do this in tau
+```
+At this point it is possible to take other measures around fine-tuning the tri-grams.
+For instance we may like to remove the tri-grams that contain an end-of-sentence token.
+This would mean that tri-grams that include a word from a new sentence will be removed.
+For example, the tri-gram ‘day </s> we’ will be removed because it has words from two different sentences.
+Table **X** shows the the first four tri-grams from the small sample corpus after this process.
+```{r table3, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
+tabl <- "
+| there was no  | a rain so | a walk that | an hour in  |
+|---------------|-----------|-------------|-------------|
+| 2             | 1         | 1           | 1           |
+"
+cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
+```
+The tri-grams can be accessed by calling the names of this table object and the counts can be used to calculate probabilities.
+There are a variety of options around constructing these probabilities.
+One is Kneser–Ney smoothing, which is the smoothing technique recommended by @chen1999empirical to improve the accuracy of n-gram models **(This function is written manually, but it’s not my own code.
+I am reading about the Kneser–Ney smoothing to understand the math, will replace the function afterwards)**.
+Finally, it is helpful to create a data table that contains the first two words and the last of each tri-gram in separate column and the smoothed probability in another column.
+For instance, Table **X** provides and example of this set-up.
+```{r table4, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
+tabl <- "
+| First Words   | Last Word | Probability |
+|---------------|-----------|-------------|
+| there was     | no        | 0.6278947   |
+"
+cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
+```

### 2020-08-06 outputs/documents/paper.Rmd
Expand on Tau example

@@
 
-**(Ke-Li add brief write-up of implementation in R please. Probably quanteda unless there's a better option.)**
-To implement the n-gram model, we used the Text Analysis Utilities (tau) R package to tokenize the training corpus (citation('tau')).
-To ensure that the three-word sequences are split within a sentence, we split the corpus into sentences by period, exponential mark, or question marks.
+There are a variety of ways to implement an n-gram model within R @citeR including **X, Y and Z**.
+We used the Text Analysis Utilities, `tau`, R package @citetau because of **X, Y, and Z**.
+To illustrate the use of `tau`, we will first install and load the library, and create a small corpus.
+
+```{r, eval = FALSE}
+install.packages("tau")
+```
+
+```{r}
+library(tau)
+corpus <- "There was no possibility of taking a walk that day.
+We had been wandering, indeed, in the leafless shrubbery an hour in the morning; but since dinner (Mrs.
+Reed, when there was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, that further out-door exercise was now out of the question."
+```
+
+Our purpose is to construct tri-grams and so we first split the corpus into sentences by period, exponential mark, or question marks.
 We then manually added an end of sentence token $<s>$ to the end of each sentence.
-We used the textcnt() function turn the training corpus into tri-grams, and the function outputs a textcnt object that resembles a table object that has contains the tri-grams and their counts.
-We then removed all the tri-grams that contains an end-of-sentence token, so the tri-grams that include a word from a new sentence will be removed.
-For example, the tri-gram "day </s> we" will be removed because it has words from two different sentences.
-The following is the sample output that show the first 4 tokens:
+
+```{r}
+# Add simple code for however this is done in tua?
+```
+
+Next we remove all punctuation and other non-letter characters, and convert all text to lower case.
+
+```{r}
+# Add however this is done in tua?
+```
+
+We will use the `textcnt()` function to turn the training corpus into tri-grams.
+The most important argument of this function is the corpus, as well as **X, Y and Z**.
+The function outputs a `textcnt` object, which is essentially a table with each tri-gram and their counts.
+
+```{r}
+# Add code to do this in tau
+```
+
+At this point it is possible to take other measures around fine-tuning the tri-grams.
+For instance we may like to remove the tri-grams that contain an end-of-sentence token.
+This would mean that tri-grams that include a word from a new sentence will be removed.
+For example, the tri-gram 'day </s> we' will be removed because it has words from two different sentences.
+Table **X** shows the the first four tri-grams from the small sample corpus after this process.
 
@@
 ```
-We can then access the tri-grams by calling the names of the object and use their counts to proceed to probability calculation.
-In calculating the probability of the tri-grams, we applied Kneser–Ney smoothing, the smoothing technique recommended by @chen1999empirical to improve the accuracy of n-gram models **(This function is written manually, but it's not my own code.
+The tri-grams can be accessed by calling the names of this table object and the counts can be used to calculate probabilities.
+There are a variety of options around constructing these probabilities.
+One is Kneser–Ney smoothing, which is the smoothing technique recommended by @chen1999empirical to improve the accuracy of n-gram models **(This function is written manually, but it’s not my own code.
 I am reading about the Kneser–Ney smoothing to understand the math, will replace the function afterwards)**.
-Finally, we create a data table that contains the first two words and the last of each tri-gram in separate column and the smoothed probability in another column.
-The example row of the table is the following:
+
+Finally, it is helpful to create a data table that contains the first two words and the last of each tri-gram in separate column and the smoothed probability in another column.
+For instance, Table **X** provides and example of this set-up.
 

### 2020-08-06 outputs/documents/paper.Rmd
BERT implementation write-up

@@
 There are three basic steps within the Keras deep learning framework: first is to prepare the data, second is to define the model and third is to train and evaluate the model.
-Following the steps, we first downloaded a the Amazon Fine Foods Reviews dataset from Stanford **(Ke-Li add citation for data set)** and preprocess the text into tokens.
-Then we defined a model that mimics the skip-gram algorithm of Word2Vec.
-The embedding dimensions, the size of the context window and the number of samples are configurable.
+Following the steps, we first downloaded a the [Amazon Fine Foods Reviews dataset from Stanford] **(Ke-Li add citation for data set)** and preprocess the text into tokens.
+Then we defined a model that mimics the skip-gram algorithm of Word2Vec in a function called skipgrams_generator().
+The embedding dimensions, the size of the context window and the number of samples are configurable within this function.
 Finally, we used the fit_generator() function of Keras to train the model.
@@
 **What really happened:** However, the training process was stuck in the operation of fit_generator() function.
+According to Stack Flow this might be an issue of incompatible versions between Keras and Tensorflow.
+More investigation is needed.
 
@@
 The blog post is aimed to create a word embedding model for text classification.
-However, we will only use the guide create a embedding matrix and skip the steps of text classification.
-Since we are employing Keras again, the process remains similar.
-We first retrieved the training data set from Stanford and preprocessed the text in to tokens.
-We then defined the model by configuring the embedding layer.
-Finally the model is trained and evaluated, and the GloVe embedding matrix is obtained.
+However, we will only use the guide create a embedding index and skip the steps of text classification.
+Instead of building a model for word embedding from scratch as we did for the skip-gram model of Word2Vec, here we are merely loading the Glove model and its pretrained word vectors as an index.
+The word embedding index is then saved to an R environment.
 
@@
 Zhao HG (**Replace by citation**) created a python module called "keras-bert" and the module can be called by R through "reticulate" package (@reticulateR), which enables the operation of Python modules in R.
-To start, the "keras-bert" module is installed in Python environment and is called in R through the 'reticulate' package.
-We can then load, configure and train the BERT model.
-To train the BERT model, we load a task-specific dataset for it to learn.
-[(**Keli Add more details**)]
+To start, we install the keras-bert model in Python on the machine and load BERT model into R through "reticulate".
+Once keras-bert is loaded, we can use it to fine-tune BERT on a task-specific training dataset.
+Fine-tuning is started by tokenizing the training data.
+The output of the tokenization is then fed to BERT as the input.
+The next step is to define the model by configuring its parameters such as batch size, number of epochs and the learning rate.
+Finally, the model is compiled and we can begin the fine-tuning.
 

### 2020-08-10 outputs/documents/paper.Rmd
write-up updates, literactures added, model starters

@@
 This is used to monitor changes during the cleaning process, and to compare the messiness of different texts.
-The methods that we consider are: n-grams, word2vec, GloVe, and the GPT models.
+The methods that we consider are: n-grams, Word2Vec, GloVe, and the GPT models.
 We illustrate our process on text from the Canadian Hansard and introduce both a Shiny application and an R package to make our process easier for others to adopt."
@@
 We consider various natural language processing methods.
-These include n-grams, word2vec, GloVe, and the GPT models.
+These include n-grams, Word2Cec, GloVe, GPT models and BERT.
 The n-gram approach involves identifying two, three, or more, words that are commonly found together.
@@
 For that reason, we consider pre-trained word embedding models.
-These include word2vec and GloVe, which place each word in a multi-dimensional space such that distance between words can illustrate their relationship.
-We also consider pre-trained generative models such as GPT-2 and GPT-3.
-GPT-2 and GPT-3 are pre-trained generative unsupervised language models.
-They differ in the number of parameters.
+These include Word2Vec and GloVe, which place each word in a multi-dimensional space such that distance between words can illustrate their relationship.
+We also consider pre-trained generative models such as GPT-2 and GPT-3 and BERT.
+GPT-2, GPT-3 and BERT are pre-trained generative unsupervised language models.
+GPT-2 and GPT-3 are the two generations of the same model that differ in the number of parameters.
 GPT-2 has 1.5 billion parameters and GPT-3 has more than 175 billion parameters.
 While GPT-2 is more limited, it can be run on smaller machines, whereas GPT-3 cannot and is accessed via an API.
+BERT has 340 millions parameters which is smaller than both GPT-2 and GPT-3.
+BERT also differs from GPT models because its encoding process is bidirectional instead of unidirectional.
 
@@
 We all know that 'tbe' is wrong and should be 'the'.
-But statistical language models have the advantage of correction that can occur within context.
-This allows significant improvements, for instance sometimes 'bat' should be 'hat' and sometimes it really should be 'bat' **(Ke-Li add ref)**.
+But statistical language models have the advantage of correction that can occur within context (@rosenfeld2000two).
+This allows significant improvements, for instance sometimes 'bat' should be 'hat' and sometimes it really should be 'bat'.
 
 The foundation of an n-gram language model is the conditional probability set-up introduced above.
-An n-gram model is a probabilistic language model that predicts the next word in a sequence of words **(Ke-Li add ref)**.
+An n-gram model is a probabilistic language model that predicts the next word in a sequence of words (@bengio2003neural).
 The $n$ in n-gram refers to the number of words in that sequence.
@@
 
-$$\mbox{Pr}(w_{n} | w^{n-1}_{1}) \approx \mbox{P}(w_{n} | w^{n-1}_{n - N+1}))$$
-where $N$ is **(Ke-Li add)**.
+$$\mbox{Pr}(w_{n} | w^{n-1}_{1}) \approx \mbox{Pr}(w_{n} | w_{n -1})$$
 
@@
 
-There are a variety of ways to implement an n-gram model within R @citeR including **X, Y and Z**.
-We used the Text Analysis Utilities, `tau`, R package @citetau because of **X, Y, and Z**.
+There are a variety of ways to implement an n-gram model within R @citeR including using R packages such as `Quanteda` (@citequanteda), `tidyText` (@citetidytext) and `tm` (@citetm).
+We used the Text Analysis Utilities, `tau`, R package @citetau because of we can separate text-cleaning and n-gram making in two different steps.
+We want to have control to manually clean the text because we want to add an end-of-sentence token to each sentence int he corpus.
 To illustrate the use of `tau`, we will first install and load the library, and create a small corpus.
@@
 
-```{r}
+```{r, eval=FALSE}
 library(tau)
-corpus <- "There was no possibility of taking a walk that day.
-We had been wandering, indeed, in the leafless shrubbery an hour in the morning; but since dinner (Mrs.
-Reed, when there was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, that further out-door exercise was now out of the question."
+corpus <- "There was no possibility of taking a walk that day. We had been 
+wandering, indeed, in the leafless shrubbery an hour in the morning; but 
+since dinner (Mrs. Reed, when there was no company, dined early) the cold 
+winter wind had brought with it clouds so sombre, and a rain so penetrating, 
+that further out-door exercise was now out of the question."
 ```
@@
 Our purpose is to construct tri-grams and so we first split the corpus into sentences by period, exponential mark, or question marks.
-We then manually added an end of sentence token $<s>$ to the end of each sentence.
 
-```{r}
-# Add simple code for however this is done in tua?
-```
+First, we remove all punctuation and other non-letter characters, and convert all text to lower case.
 
-Next we remove all punctuation and other non-letter characters, and convert all text to lower case.
-
-```{r}
-# Add however this is done in tua?
+```{r, eval=FALSE}
+tokenizer <- function(corpus) {
+  lines <- vector()
+  # Read the corpus line by line
+  for (line in corpus) {
+    # add two start-of-sentence tokens to the beginning of a sentence
+    lines <-c(lines, line)
+  }
+  lines <- tolower(lines)
+  # add end-of-sentence token to the end of a sentence
+  lines <- gsub("[.!?]$|[.!?] |$", " </s>", lines) 
+  tokens <- unlist(strsplit(lines, "[^a-z<>/]"))
+  tokens <- tokens[tokens != ""]
+  return(tokens)
+}
 ```
@@
 We will use the `textcnt()` function to turn the training corpus into tri-grams.
-The most important argument of this function is the corpus, as well as **X, Y and Z**.
-The function outputs a `textcnt` object, which is essentially a table with each tri-gram and their counts.
+The most important argument of this function is the _corpus_, as well as the _n_ to set the number of words in a sequence, _method_ to determine which type of counts to compute, and _split_ to designate the regular expression pattern for splitting.
+The function outputs a `textcnt` object, which is essentially a table with each tri-gram and their counts (@citetau).
 
-```{r}
-# Add code to do this in tau
+```{r, eval=FALSE}
+trigrams <- textcnt( tokens, n=3, method = "string", 
+                     split = "[[:space:][:digit:]]+", decreasing = TRUE)
+```
+
+We then manually added an end of sentence token </s> to the end of each sentence.
+
+```{r, eval=FALSE}
+trigrams <- trigrams[!grepl("</s>", names(trigrams))]
 ```
@@
 For example, the tri-gram 'day </s> we' will be removed because it has words from two different sentences.
-Table **X** shows the the first four tri-grams from the small sample corpus after this process.
+Table 1.
+shows four sample tri-grams from the small sample corpus after this process.
 
@@
 | 2             | 1         | 1           | 1           |
+
+Table 1. Four sample tri-grams generated from the corpus
 "
@@
 ```
+
 The tri-grams can be accessed by calling the names of this table object and the counts can be used to calculate probabilities.
 There are a variety of options around constructing these probabilities.
-One is Kneser–Ney smoothing, which is the smoothing technique recommended by @chen1999empirical to improve the accuracy of n-gram models **(This function is written manually, but it’s not my own code.
-I am reading about the Kneser–Ney smoothing to understand the math, will replace the function afterwards)**.
+One is Kneser–Ney smoothing, which is the smoothing technique recommended by @chen1999empirical to improve the accuracy of n-gram models.
 
 Finally, it is helpful to create a data table that contains the first two words and the last of each tri-gram in separate column and the smoothed probability in another column.
-For instance, Table **X** provides and example of this set-up.
+For instance, Table 2.
+provides and example of this set-up.
 
@@
 | there was     | no        | 0.6278947   |
+
+Table2. The smoothed probability table
 "
@@
 Hence, word embedding models can identify similar terms, calculate the similarity between a pair of terms, and find the odd term in a group of terms [@mikolov2013efficient].
-There are a variety of ways to achieve this dimensionality reduction, including word2vec and GloVe.
+There are a variety of ways to achieve this dimensionality reduction, including Word2Vec and GloVe.
 
@@
 
-Word2vec is trained on the Google News dataset **(Ke-Li add cite to the dataset please.)**.
+Word2Vec is trained on the Google News dataset (@citegooglenews).
 This contains six billion tokens and the vocabulary size is restricted to the one million most common words (@mikolov2013efficient).
@@
 
-To the extent that the relative positions of terms reflects their semantic relationship, word2vec will perform well for word analogies.
+To the extent that the relative positions of terms reflects their semantic relationship, Word2Vec will perform well for word analogies.
 However, terms that have context-specific meaning may not perform as well.
@@
 There are a few packages for implementing a Word2Vec model in R.
-The packages are "rword2vec" by Mukul Chaware and "word2vec" by Jan Wijffels, BNOSAC, and Max Fomichev.
-Both packages uses the same algorithm that Word2Vec uses to produce word representations, but they do not have an affiliation with Google.
+The packages are `rword2vec` by @rword2vec and `word2vec` by @word2vecR.
+Both packages use the same algorithm that Word2Vec uses to produce word representations, but they do not have an affiliation with Google.
 Another way to implement a Word2Vec model in R is through Keras package.
-Keras is an API for deep learning written in Python (@chollet2015keras) and the Keras package provides an R interface to Keras **(Ke-Li add citation for the package)**.
+Keras is an API for deep learning written in Python (@chollet2015keras) and the Keras package provides an R interface to Keras (@kerasR2020).
 RStudio published a tutorial for implementing the skip-gram model of Word2Vec (@falbel2017word) through Keras.
@@
 There are three basic steps within the Keras deep learning framework: first is to prepare the data, second is to define the model and third is to train and evaluate the model.
-Following the steps, we first downloaded a the [Amazon Fine Foods Reviews dataset from Stanford] **(Ke-Li add citation for data set)** and preprocess the text into tokens.
-Then we defined a model that mimics the skip-gram algorithm of Word2Vec in a function called skipgrams_generator().
+Following the steps, we first downloaded a the [Amazon Fine Foods Reviews dataset from Stanford] (@amzfinefooddata) and preprocess the text into tokens.
+Then we defined a model that mimics the skip-gram algorithm of Word2Vec in a function called `skipgrams_generator()`.
 The embedding dimensions, the size of the context window and the number of samples are configurable within this function.
-Finally, we used the fit_generator() function of Keras to train the model.
+Finally, we used the `fit_generator()` function of Keras to train the model.
 **What's supposed to happen:** As a result, a embedding matrix is obtained and serves as the look-up table of the word vectors.
-**What really happened:** However, the training process was stuck in the operation of fit_generator() function.
+**What really happened:** However, the training process was stuck in the operation of `fit_generator()` function.
 According to Stack Flow this might be an issue of incompatible versions between Keras and Tensorflow.
@@
 
-The global vectors (GloVe) model is a pre-trained word embedding model that emphasises the semantic meaning of words [@pennington2014glove].
-In the word2vec model, word representations are learnt locally within the context window.
+The Global Vectors (GloVe) model is a pre-trained word embedding model that emphasises the semantic meaning of words [@pennington2014glove].
+In the Word2Vec model, word representations are learnt locally within the context window.
 On the other hand, GloVe considers the properties of the corpus globally.
@@
 
-**(Ke-Li add brief paragraph about negatives of this approach.)**
-**(This limitation paragraph applies to both Word2Vec and GloVe)**
 As @hanretty2018comparing point out, although word embedding models can capture semantic meanings of words, "they are context-free and fail to capture higher-level concepts in context, such as polysemous disambiguation, syntactic structures, semantic roles, anaphora".
@@
 There are more than one way to implement GloVe word embedding model in R.
-One way is to load GloVe embeddings into a frozen Keras Embedding layer **(Keli add citation to the blog post by Keras)** using the Keras R package **(Keli add citation to package)**.
-Another way, as mentioned and briefly demonstrated by Rstudio in their blog post, is to use the "text2vec" package developed by Dmitriy Selivanov **(Ke-Li add package citation)**.
+One way is to load GloVe embeddings into a frozen Keras Embedding layer (@kerasglove) using the `Keras` R package (@kerasR2020).
+Another way, as mentioned and briefly demonstrated by Rstudio in their blog post, is to use the `text2vec` package developed by @text2vec.
 Keras published a blog post on pre-trained word embeddings through GloVe in R.
-In this project, we intend to follow the blog post for implementing GloVe through the Keras R package **(Keli add citation to the blog post by Keras)**.
+In this project, we intend to follow the blog post for implementing GloVe through the Keras R package (@kerasglove).
 The blog post is aimed to create a word embedding model for text classification.
 However, we will only use the guide create a embedding index and skip the steps of text classification.
-Instead of building a model for word embedding from scratch as we did for the skip-gram model of Word2Vec, here we are merely loading the Glove model and its pretrained word vectors as an index.
+Instead of building a model for word embedding from scratch as we did for the skip-gram model of Word2Vec, here we are merely loading the Glove model and its pre-trained word vectors as an index.
 The word embedding index is then saved to an R environment.
@@
 
-OpenAI is an AI research company who openly pursues Artificial General Intelligence — to have machines learn and understand any intellectual tasks as human do **(Cite OpenAI blog)**.
+OpenAI is an AI research company who openly pursues Artificial General Intelligence — to have machines learn and understand any intellectual tasks as human do (@brundage2018malicious).
 OpenAI's Generative Pre-Trained Transformer (GPT) models are pre-trained language models that are built based on the transformer architecture.
@@
 They had shown a milestone in predicting larger range of distinct relation types with higher confidence (@alt2019fine).
-In the field of speech recognition, data scarcity is often an issue due to the difficulty to collect large amount of data from human for learning (@bird2020lst).
+In the field of speech recognition, data scarcity is often an issue due to the difficulty to collect large amount of data from human for learning (@bird2020lstm).
 @bird2020lstm employed GPT-2 as part of the solution to generate augmented data for classification.
@@
 OpenAI, on the other hands, train the GPT models to perform many tasks (@radford2019language).
-GPT-2 is pre-trained on WebText, a comprehensive data set that is scrapped from millions of webpages and contains total of 40 GB of text (@radford2019language).
+GPT-2 is pre-trained on WebText, a comprehensive dataset that is scrapped from millions of webpages and contains total of 40 GB of text (@radford2019language).
 WebText contains all kinds of textual data: dialogues, comments, translations, questions and answers...etc.
@@
 Give it a prompt to start and it will generate texts that continues what has been started.
-In the fine-tuning stage, @radford2019language provided additional training data sets for GPT-2 to generate the texts with specific sentiment, or be descriptive, or even summarize the prompt text.
+In the fine-tuning stage, @radford2019language provided additional training datasets for GPT-2 to generate the texts with specific sentiment, or be descriptive, or even summarize the prompt text.
 
-OpenAI has made GPT-2 open-sourced through Github with Python and Tensorflow.
+OpenAI has made GPT-2 open-sourced through Github with Python and Tensorflow (@tensorflow2015-whitepaper).
 There is also a package in R that wraps the Python code into a R package, created by @keydanaluraschi2019gpt2.
@@
 The R package only permits exploratory experiments with the pre-trained tasks; fine-tuning for new tasks is not available thorough R.
-To proceed with the implementation of GPT-2 in R, we will fine-tune the model through Python and create a wrapper code for R (**Keli to check if this is possible**).
+To proceed with the implementation of GPT-2 in R, theoretically, we will fine-tune the model through Python and create a wrapper code for R.
+This could possibility be achieved by employing the `reticulate` R package (@reticulateR) that allows operation of Python code in R.
+The practicality of this approach is to be further investigated.
 
@@
 
-BERT was pre-trained on BooksCorpus data set with 800 million words and English Wikipedia with 2,500 million words (@devlin2018bert).
+BERT was pre-trained on BooksCorpus dataset with 800 million words and English Wikipedia with 2,500 million words (@devlin2018bert).
 Because BERT employs the Transformer architecture with the self-attention mechanism, fine-tuning is also available with one additional output layer and a task-specific input (@devlin2018bert).
@@
 BERT implementation in R is also possible through Keras.
-Zhao HG (**Replace by citation**) created a python module called "keras-bert" and the module can be called by R through "reticulate" package (@reticulateR), which enables the operation of Python modules in R.
-To start, we install the keras-bert model in Python on the machine and load BERT model into R through "reticulate".
-Once keras-bert is loaded, we can use it to fine-tune BERT on a task-specific training dataset.
+@kerasbert created a python module called `keras-bert` and the module can be called by R through `reticulate` package (@reticulateR), which enables the operation of Python modules in R.
+To start, we install the keras-bert model in Python on the machine and load BERT model into R through `reticulate`.
+Once `keras-bert` is loaded, we can use it to fine-tune BERT on a task-specific training dataset.
 Fine-tuning is started by tokenizing the training data.
@@
 # References
-
---
-There are a variety of ways to implement an n-gram model within R **[cite R]** including X, Y and Z.
-We used the Text Analysis Utilities, `tau`, R package **[cite tau]** because of X, Y, and Z.
-To illustrate the use of `tau`, we will first install and load the library, and create a small corpus.
-```{r, eval = FALSE}
-install.packages(“tau”)
-library(tau)
-corpus <- “There was no possibility of taking a walk that day.
-We had been wandering, indeed, in the leafless shrubbery an hour in the morning; but since dinner (Mrs.
-Reed, when there was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, that further out-door exercise was now out of the question.”
-```
-Our purpose is to construct tri-grams and so we first split the corpus into sentences by period, exponential mark, or question marks.
-We then manually added an end of sentence token $<s>$ to the end of each sentence.
-```{r}
-#Add simple code for however this is done in tua?
-```
-Next we remove all punctuation and other non-letter characters, and convert all text to lower case.
-```{r}
-#however this is done in tua?
-```
-We will use the `textcnt()` function to turn the training corpus into tri-grams.
-The most important argument of this function is the corpus, as well as X, Y and Z.
-The function outputs a `textcnt` object, which is essentially a table with each tri-gram and their counts.
-```{r}
-#code to do this in tau
-```
-At this point it is possible to take other measures around fine-tuning the tri-grams.
-For instance we may like to remove the tri-grams that contain an end-of-sentence token.
-This would mean that tri-grams that include a word from a new sentence will be removed.
-For example, the tri-gram ‘day </s> we’ will be removed because it has words from two different sentences.
-Table **X** shows the the first four tri-grams from the small sample corpus after this process.
-```{r table3, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
-tabl <- "
-| there was no  | a rain so | a walk that | an hour in  |
-|---------------|-----------|-------------|-------------|
-| 2             | 1         | 1           | 1           |
-"
-cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
-```
-The tri-grams can be accessed by calling the names of this table object and the counts can be used to calculate probabilities.
-There are a variety of options around constructing these probabilities.
-One is Kneser–Ney smoothing, which is the smoothing technique recommended by @chen1999empirical to improve the accuracy of n-gram models **(This function is written manually, but it’s not my own code.
-I am reading about the Kneser–Ney smoothing to understand the math, will replace the function afterwards)**.
-Finally, it is helpful to create a data table that contains the first two words and the last of each tri-gram in separate column and the smoothed probability in another column.
-For instance, Table **X** provides and example of this set-up.
-```{r table4, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
-tabl <- "
-| First Words   | Last Word | Probability |
-|---------------|-----------|-------------|
-| there was     | no        | 0.6278947   |
-"
-cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
-```

### 2020-08-10 model starters/BERT.R

@@ -1,0 +1,87 @@
+ ## Source tutorial: https://blogs.rstudio.com/ai/posts/2019-09-30-bert-r/
+
+Sys.setenv(TF_KERAS=1)
+
+pretrained_path = '/change/this/path/uncased_L-12_H-768_A-12'
+
+config_path = file.path(pretrained_path, 'bert_config.json')
+checkpoint_path = file.path(pretrained_path, 'bert_model.ckpt')
+vocab_path = file.path(pretrained_path, 'vocab.txt')
+
+library(reticulate)
+k_bert = import('keras_bert')
+token_dict = k_bert$load_vocabulary(vocab_path)
+tokenizer = k_bert$Tokenizer(token_dict)
+
+seq_length = 50L
+bch_size = 70
+epochs = 1
+learning_rate = 1e-4
+
+DATA_COLUMN = 'comment_text'
+LABEL_COLUMN = 'target'
+
+model = k_bert$load_trained_model_from_checkpoint(
+  config_path,
+  checkpoint_path,
+  training=T,
+  trainable=T,
+  seq_len=seq_length)
+
+# tokenize text
+tokenize_fun = function(dataset) {
+  c(indices, target, segments) %<-% list(list(),list(),list())
+  for ( i in 1:nrow(dataset)) {
+    c(indices_tok, segments_tok) %<-% tokenizer$encode(dataset[[DATA_COLUMN]][i], 
+                                                       max_len=seq_length)
+    indices = indices %>% append(list(as.matrix(indices_tok)))
+    target = target %>% append(dataset[[LABEL_COLUMN]][i])
+    segments = segments %>% append(list(as.matrix(segments_tok)))
+  }
+  return(list(indices,segments, target))
+}
+# read data
+dt_data = function(dir, rows_to_read){
+  data = data.table::fread(dir, nrows=rows_to_read)
+  c(x_train, x_segment, y_train) %<-% tokenize_fun(data)
+  return(list(x_train, x_segment, y_train))
+}
+
+c(x_train,x_segment, y_train) %<-% dt_data('/Users/kelichiu/Desktop/train.csv',2000)
+
+train = do.call(cbind,x_train) %>% t()
+segments = do.call(cbind,x_segment) %>% t()
+targets = do.call(cbind,y_train) %>% t()
+
+concat = c(list(train ),list(segments))
+
+c(decay_steps, warmup_steps) %<-% k_bert$calc_train_steps(
+  targets %>% length(),
+  batch_size=bch_size,
+  epochs=epochs
+)
+
[…]

### 2020-08-10 model starters/GloVe.R

@@ -1,0 +1,38 @@
+# Source tutorial: https://keras.rstudio.com/articles/examples/pretrained_word_embeddings.html
+library(keras)
+
+GLOVE_DIR <- 'glove.6B'
+TEXT_DATA_DIR <- '20_newsgroup'
+MAX_SEQUENCE_LENGTH <- 1000
+MAX_NUM_WORDS <- 20000
+EMBEDDING_DIM <- 100
+VALIDATION_SPLIT <- 0.2
+
+# download data if necessary
+download_data <- function(data_dir, url_path, data_file) {
+  if (!dir.exists(data_dir)) {
+    download.file(paste0(url_path, data_file), data_file, mode = "wb")
+    if (tools::file_ext(data_file) == "zip")
+      unzip(data_file, exdir = tools::file_path_sans_ext(data_file))
+    else
+      untar(data_file)
+    unlink(data_file)
+  }
+}
+download_data(GLOVE_DIR, 'http://nlp.stanford.edu/data/', 'glove.6B.zip')
+
+# first, build index mapping words in the embeddings set
+# to their embedding vector
+
+cat('Indexing word vectors.\n')
+
+embeddings_index <- new.env(parent = emptyenv())
+lines <- readLines(file.path(GLOVE_DIR, 'glove.6B.100d.txt'))
+for (line in lines) {
+  values <- strsplit(line, ' ', fixed = TRUE)[[1]]
+  word <- values[[1]]
+  coefs <- as.numeric(values[-1])
+  embeddings_index[[word]] <- coefs
+}
+
+cat(sprintf('Found %s word vectors.\n', length(embeddings_index)))

### 2020-08-10 shinyapp/functions.R

@@ -3,7 +3,6 @@ library(tau)
 library(tm)
 library(stringr)
 library(stats)
-
 library(reshape2)
 library(RWeka)
 
@@ -13,14 +12,12 @@ demoTraining <- readLines(paste(path,"/samples/demo.txt", sep=""), warn=FALSE)
 
 tokenizer <- function(corpus) {
   lines <- vector()
-  
   # Read the corpus line by line
   for (line in corpus) {
     # add two start-of-sentence tokens to the beginning of a sentence
     #line <- paste("<s> <s>", line)
     lines <-c(lines, line)
   }
-  
   lines <- tolower(lines)
   lines <- gsub("[.!?]$|[.!?] |$", " </s>", lines) # add end-of-sentence token to the end of a sentence
   tokens <- unlist(strsplit(lines, "[^a-z<>/]"))
@@ -92,7 +89,6 @@ bigramsDF <- data.table("First Words" = removeLastWord(names(bigrams)),
 trigramsDF <- data.table("Firs tWords" = removeLastWord(names(trigrams)), 
                          "Last Word" = getLastWords(names(trigrams), 1), 
                          "Probability" = as.vector(trigramProbs), stringsAsFactors=F)
-
 
 library(dplyr)
 unigramDF <- (unigramDF %>% arrange(desc(Probability)))

### 2020-08-10 outputs/documents/paper.Rmd
brackets correction

@@
 Using a projection matrix, the projection layer then project the discrete one-hot vectors to lower-dimension continuous vectors.
-Finally, the output layer is a softmax regression classifier that obtains the posterior distribution of words (@rong2014word2vec).
+Finally, the output layer is a softmax regression classifier that obtains the posterior distribution of words [@rong2014word2vec].
 It is the input and output vectors that differ between the two architectures.
 
-Word2Vec is trained on the Google News dataset (@citegooglenews).
-This contains six billion tokens and the vocabulary size is restricted to the one million most common words (@mikolov2013efficient).
+Word2Vec is trained on the Google News dataset [@citegooglenews].
+This contains six billion tokens and the vocabulary size is restricted to the one million most common words [@mikolov2013efficient].
 As this is a large dataset, it provides an accurate statistical pattern of language, as it is used in this dataset, and a comprehensive mapping between terms and vector positions.
@@
 Another way to implement a Word2Vec model in R is through Keras package.
-Keras is an API for deep learning written in Python (@chollet2015keras) and the Keras package provides an R interface to Keras (@kerasR2020).
-RStudio published a tutorial for implementing the skip-gram model of Word2Vec (@falbel2017word) through Keras.
+Keras is an API for deep learning written in Python [@chollet2015keras] and the Keras package provides an R interface to Keras [@kerasR2020].
+RStudio published a tutorial for implementing the skip-gram model of Word2Vec [@falbel2017word] through Keras.
 We intend to follow this tutorial as the approach to implement a Word2Vec model.
 There are three basic steps within the Keras deep learning framework: first is to prepare the data, second is to define the model and third is to train and evaluate the model.
-Following the steps, we first downloaded a the [Amazon Fine Foods Reviews dataset from Stanford] (@amzfinefooddata) and preprocess the text into tokens.
+Following the steps, we first downloaded a the [Amazon Fine Foods Reviews dataset from Stanford] [@amzfinefooddata] and preprocess the text into tokens.
 Then we defined a model that mimics the skip-gram algorithm of Word2Vec in a function called `skipgrams_generator()`.
@@
 There are more than one way to implement GloVe word embedding model in R.
-One way is to load GloVe embeddings into a frozen Keras Embedding layer (@kerasglove) using the `Keras` R package (@kerasR2020).
+One way is to load GloVe embeddings into a frozen Keras Embedding layer [@kerasglove] using the `Keras` R package [@kerasR2020].
 Another way, as mentioned and briefly demonstrated by Rstudio in their blog post, is to use the `text2vec` package developed by @text2vec.
 Keras published a blog post on pre-trained word embeddings through GloVe in R.
-In this project, we intend to follow the blog post for implementing GloVe through the Keras R package (@kerasglove).
+In this project, we intend to follow the blog post for implementing GloVe through the Keras R package [@kerasglove].
 The blog post is aimed to create a word embedding model for text classification.
@@
 
-OpenAI is an AI research company who openly pursues Artificial General Intelligence — to have machines learn and understand any intellectual tasks as human do (@brundage2018malicious).
+OpenAI is an AI research company who openly pursues Artificial General Intelligence — to have machines learn and understand any intellectual tasks as human do [@brundage2018malicious].
 OpenAI's Generative Pre-Trained Transformer (GPT) models are pre-trained language models that are built based on the transformer architecture.
-Proposed by @vaswani2017attention from Google in 2017, the Transformer is a novel network architecture for neural network that outperforms RNN-based and CNN-based models in computational efficiency (@vaswani2017attention).
-Since the emergence of the Transformer model, most of the representative pre-trained models are built based on this architecture (@hanretty2018comparing).
+Proposed by @vaswani2017attention from Google in 2017, the Transformer is a novel network architecture for neural network that outperforms RNN-based and CNN-based models in computational efficiency [@vaswani2017attention].
+Since the emergence of the Transformer model, most of the representative pre-trained models are built based on this architecture [@hanretty2018comparing].
 Therefore, we can say that Transformer marks a new era of language models.
@@
 In the year 2019, its successor GPT-2 was released in three stages.
-Not long after, in June 2020, OpenAI announced GPT-3, which is the latest GPT model as of July 2020 (@brown2020language).
+Not long after, in June 2020, OpenAI announced GPT-3, which is the latest GPT model as of July 2020 [@brown2020language].
 Since the API of GPT-3 at the moment is only available to a few invited users, public applications of GPT-3 are limited.
@@
 @alt2019fine extended GPT-2 to relation extraction, which is a task of identifying the relationship between concepts appeared in text.
-They had shown a milestone in predicting larger range of distinct relation types with higher confidence (@alt2019fine).
-In the field of speech recognition, data scarcity is often an issue due to the difficulty to collect large amount of data from human for learning (@bird2020lstm).
+They had shown a milestone in predicting larger range of distinct relation types with higher confidence [@alt2019fine].
+In the field of speech recognition, data scarcity is often an issue due to the difficulty to collect large amount of data from human for learning [@bird2020lstm].
 @bird2020lstm employed GPT-2 as part of the solution to generate augmented data for classification.
@@
 Most of prior language models are trained in a single domain by supervised learning to perform a single task.
-OpenAI, on the other hands, train the GPT models to perform many tasks (@radford2019language).
-GPT-2 is pre-trained on WebText, a comprehensive dataset that is scrapped from millions of webpages and contains total of 40 GB of text (@radford2019language).
+OpenAI, on the other hands, train the GPT models to perform many tasks [@radford2019language].
+GPT-2 is pre-trained on WebText, a comprehensive dataset that is scrapped from millions of webpages and contains total of 40 GB of text [@radford2019language].
 WebText contains all kinds of textual data: dialogues, comments, translations, questions and answers...etc.
 Such diversity is critical for GPT-2 to be a generalized model that can perform multiple tasks with any domain boundary.
-However, GPT-2 has the capacity to adapt to a context and generates custom output with the "fine-tuning" technique (@ziegler2019fine).
-While the pre-training stage is unsupervised and mandatory, fine-tuning is an optional training stage where GPT can learn to perform new tasks in a supervised or reinforced approach such as stylistic continuation and summarization (@ziegler2019fine).
+However, GPT-2 has the capacity to adapt to a context and generates custom output with the "fine-tuning" technique [@ziegler2019fine].
+While the pre-training stage is unsupervised and mandatory, fine-tuning is an optional training stage where GPT can learn to perform new tasks in a supervised or reinforced approach such as stylistic continuation and summarization [@ziegler2019fine].
 During the pre-training stage, the model is trained to perform text completion task.
@@
 
-OpenAI has made GPT-2 open-sourced through Github with Python and Tensorflow (@tensorflow2015-whitepaper).
+OpenAI has made GPT-2 open-sourced through Github with Python and Tensorflow [@tensorflow2015-whitepaper].
 There is also a package in R that wraps the Python code into a R package, created by @keydanaluraschi2019gpt2.
@@
 To proceed with the implementation of GPT-2 in R, theoretically, we will fine-tune the model through Python and create a wrapper code for R.
-This could possibility be achieved by employing the `reticulate` R package (@reticulateR) that allows operation of Python code in R.
+This could possibility be achieved by employing the `reticulate` R package [@reticulateR] that allows operation of Python code in R.
 The practicality of this approach is to be further investigated.
 
-GPT-3 has over 175 billions of parameters (@brown2020language).
+GPT-3 has over 175 billions of parameters [@brown2020language].
 In the past few years, the growth of parameters of transformer language models expands rapidly.
@@
 Released in 2018, GPT-2 has 1.6 billion parameters, and within two years, GPT-3 has grown the size to 175 billion, which is over 100 times than its predecessor.
-The more parameters there are the better the model can generalize to new tasks; therefore, the increasing amount of parameters accelerates the improvements in text synthesis and downstream NLP tasks (@brown2020language).
-GPT-3 can learn NLP tasks with just a handful of example; OpenAI refers this ability as few-shot learning, or in-context learning (@brown2020language).
+The more parameters there are the better the model can generalize to new tasks; therefore, the increasing amount of parameters accelerates the improvements in text synthesis and downstream NLP tasks [@brown2020language].
+GPT-3 can learn NLP tasks with just a handful of example; OpenAI refers this ability as few-shot learning, or in-context learning [@brown2020language].
 As a result, we only need to provide GPT-3 a few demonstrations for it to learn a new NLP task.
@@
 Without the demonstrations, GPT-3 tends to treat the prompts as "tweets" and generate texts that serve as responses or follow-up tweets.
-The team fed a few previous news articles in the model’s context as demonstrations, and the model learned to produce the outputs that adapt to the style and length of news stories (@brown2020language).
+The team fed a few previous news articles in the model’s context as demonstrations, and the model learned to produce the outputs that adapt to the style and length of news stories [@brown2020language].
 This few-shot learning ability indicates that the model can learn like human do and would democratize the use of language models to broader fields.
@@
 
-BERT stands for Bidirectional Encoder Representations from Transformers (@devlin2018bert).
+BERT stands for Bidirectional Encoder Representations from Transformers [@devlin2018bert].
 It was released by researchers at Google AI Language in 2018.
@@
 The first task is Masked Language Model and and second task is Next Sentence Prediction.
-The bidirectional approach of BERT is alleviated by the "Masked Language Model" (@devlin2018bert).
+The bidirectional approach of BERT is alleviated by the "Masked Language Model" [@devlin2018bert].
 The Masked Language Model masks certain words in the input corpus which forces the model to predict the original words by "filling in the blank".
-This training process blends the representation from the left and right context and therefore increases the model's context sensitivity (@nozza2020mask).
+This training process blends the representation from the left and right context and therefore increases the model's context sensitivity [@nozza2020mask].
 In the Next Sentence Prediction process, the model receives two sentences in sequence as input.
-50% of the sentences are paired in the right sequence and the ones from the other half have sentences that are paired randomly (@devlin2018bert).
+50% of the sentences are paired in the right sequence and the ones from the other half have sentences that are paired randomly [@devlin2018bert].
 The goal of this process is to have BERT evaluate if the second sentence is the next sentence of the first sentence in the original text and sharpen its capacity to predict the likelihood of the next sentence.
 
-BERT was pre-trained on BooksCorpus dataset with 800 million words and English Wikipedia with 2,500 million words (@devlin2018bert).
-Because BERT employs the Transformer architecture with the self-attention mechanism, fine-tuning is also available with one additional output layer and a task-specific input (@devlin2018bert).
+BERT was pre-trained on BooksCorpus dataset with 800 million words and English Wikipedia with 2,500 million words [@devlin2018bert].
+Because BERT employs the Transformer architecture with the self-attention mechanism, fine-tuning is also available with one additional output layer and a task-specific input [@devlin2018bert].
 Google AI researchers have fine-tuned BERT for 11 NLP tasks including question-answering and common-sense inference.
@@
 BERT implementation in R is also possible through Keras.
-@kerasbert created a python module called `keras-bert` and the module can be called by R through `reticulate` package (@reticulateR), which enables the operation of Python modules in R.
+@kerasbert created a python module called `keras-bert` and the module can be called by R through `reticulate` package [@reticulateR], which enables the operation of Python modules in R.
 To start, we install the keras-bert model in Python on the machine and load BERT model into R through `reticulate`.

### 2020-08-10 model starters/BERT.R
added comment for BERT file download source

@@ -1,7 +1,8 @@
- ## Source tutorial: https://blogs.rstudio.com/ai/posts/2019-09-30-bert-r/
+# Source tutorial: https://blogs.rstudio.com/ai/posts/2019-09-30-bert-r/
 
 Sys.setenv(TF_KERAS=1)
 
+# Go to the tutorial link in the above comment to download the files
 pretrained_path = '/change/this/path/uncased_L-12_H-768_A-12'
 
 config_path = file.path(pretrained_path, 'bert_config.json')

### 2020-08-11 outputs/documents/paper.Rmd
Minor updates to the paper.

@@
 We will use the `textcnt()` function to turn the training corpus into tri-grams.
-The most important argument of this function is the _corpus_, as well as the _n_ to set the number of words in a sequence, _method_ to determine which type of counts to compute, and _split_ to designate the regular expression pattern for splitting.
+The most important argument of this function is _corpus_, as well as _n_ to set the number of words in a sequence, _method_ to determine which type of counts to compute, and _split_ to designate the regular expression pattern for splitting.
 The function outputs a `textcnt` object, which is essentially a table with each tri-gram and their counts (@citetau).
@@
 
-We then manually added an end of sentence token </s> to the end of each sentence.
+We then manually added an end of sentence token, `</s>`, to the end of each sentence.
 
@@
 
-There are a few packages for implementing a Word2Vec model in R.
-The packages are `rword2vec` by @rword2vec and `word2vec` by @word2vecR.
-Both packages use the same algorithm that Word2Vec uses to produce word representations, but they do not have an affiliation with Google.
-Another way to implement a Word2Vec model in R is through Keras package.
-Keras is an API for deep learning written in Python [@chollet2015keras] and the Keras package provides an R interface to Keras [@kerasR2020].
-RStudio published a tutorial for implementing the skip-gram model of Word2Vec [@falbel2017word] through Keras.
-We intend to follow this tutorial as the approach to implement a Word2Vec model.
-There are three basic steps within the Keras deep learning framework: first is to prepare the data, second is to define the model and third is to train and evaluate the model.
-Following the steps, we first downloaded a the [Amazon Fine Foods Reviews dataset from Stanford] [@amzfinefooddata] and preprocess the text into tokens.
-Then we defined a model that mimics the skip-gram algorithm of Word2Vec in a function called `skipgrams_generator()`.
+There are a few packages for implementing a word2vec model in R, including `rword2vec` [@citerwordvecchaware] and `word2vec` [@citewordvecwijffels].
+Both packages use the same algorithm.
+Another way to implement a word2vec model in R is through Keras which is an API for deep learning written in Python [@chollet2015keras] for which the `keras` package provides an R interface [@citekeras].
+
+Here, we follow @falbel2017word to implement a skip-gram model of word2vec through Keras.
+There are three steps in the Keras deep learning framework: 1) prepare the data; 2) define the model; and 3) train and evaluate the model.
+**(Ke-Li: Can we change this to be a different dataset please?)** Following these steps, we first downloaded the **(Ke-Li: Is there a reason this is in brackets?)** [Amazon Fine Foods Reviews dataset from Stanford] [@amzfinefooddata] and preprocess the text into tokens.
+
+**(Ke-Li: Add code.)**
+
+We then define a model that mimics the skip-gram algorithm of word2vec using the function `skipgrams_generator()`.
+
+**(Ke-Li: Add code.)**
+
 The embedding dimensions, the size of the context window and the number of samples are configurable within this function.
-Finally, we used the `fit_generator()` function of Keras to train the model.
+
+Finally, we use the `fit_generator()` function to train the model.
+
+**(Ke-Li: Add code.)**
+
 **What's supposed to happen:** As a result, a embedding matrix is obtained and serves as the look-up table of the word vectors.
@@
 
-There are more than one way to implement GloVe word embedding model in R.
-One way is to load GloVe embeddings into a frozen Keras Embedding layer [@kerasglove] using the `Keras` R package [@kerasR2020].
-Another way, as mentioned and briefly demonstrated by Rstudio in their blog post, is to use the `text2vec` package developed by @text2vec.
-Keras published a blog post on pre-trained word embeddings through GloVe in R.
-In this project, we intend to follow the blog post for implementing GloVe through the Keras R package [@kerasglove].
+There are a variety of ways to implement GloVe word embedding models in R.
+One is to load GloVe embeddings into a frozen Keras Embedding layer [@kerasglove] using the `Keras` R package [@kerasR2020].
+This means **(Ke-Li: Add what it means please.)** Another way, is to use the `text2vec` package [@text2vec].
+
+Here we follow **CITE** to implement GloVe using the `keras` R package [@kerasglove].
 The blog post is aimed to create a word embedding model for text classification.
@@
 The word embedding index is then saved to an R environment.
+
+**[As before, we need code and examples here]**
 

### 2020-08-16 model starters/DocumentEmbedding.Rmd
word embedding for document similarity and classification tryouts

---
title: "Document embedding trial"
author: "Keli Chiu"
date: "15/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(textdata)
library(tidyverse)
library(widyr)
library(tidytext)
library(SnowballC)
```

## Can we use word embedding for document similarity and document classification?

Yes! I learned that similar to represent word in vectors, we can represent documents in vectors, and it's referred as document embedding. Once we have the document vectors, we can then find similar documents and also do classification.

The most mentioned way to do document embedding is Paragraph Vector (widely referred as Doc2Vec)), which is based on Word2Vec and proposed by the same researchers. (The related article is here: https://arxiv.org/abs/1405.4053) "the paragraph vector is concatenated or averaged with local context word vectors", and then an addtional vector is added for the document(paragraph) ID. In the end, the vectors of the document can be used as the features of the document. The features can be fed to machine learning algorithms, including classification.

Another way that's also frequently proposed is to aggregate all the word vectors in a document as the representation of the document. After obtaining hte aggregation of the word embeddings (usually the sum or the average), we can then calculate the cosine distance between the document vectors in order to detect similarity.

## What are other ways to do document embedding?

- **Topic modeling with LDA**, the documents are represented by topic distributions
- **Sentence-BERT (SBERT)**
- **TF-IDF**

## Example of document embedding

Doc2Vec is available in Python but doesn't seem to be available in R. The following example is the aggregation of pretrained GloVe word vectors, following the tutorial of Julia Silge: https://smltar.com/embeddings.html# In the tutorial, Julia sums GloVe word embeddings in a document as the representation vectors of that document. The training documents are replaced with paper abstracts downloaded from Kaggle: https://www.kaggle.com/nikhilmittal/research-paper-abstracts

```{r,warning=FALSE}
# Tokenize the abstract
abstracts <- read_csv("abstracts_101.csv")
tidy_abstracts <- abstracts %>%
  select(id, abstract) %>%
  unnest_tokens(word, abstract) %>%
  add_count(word) %>%
  filter(n >= 50) %>%
  select(-n)
```

```{r}
# Download the glove word embeddings
glove6b <- embedding_glove6b(dimensions = 100)
glove6b
```

```{r}
# Change the word embedding format to be more tidy
tidy_glove <- glove6b %>%
  pivot_longer(contains("d"),
               names_to = "dimension"
  ) %>%
  rename(item1 = token)
```

```{r}
# Change the word embedding format to be more tidy
tidy_glove <- glove6b %>%
  pivot_longer(contains("d"),
               names_to = "dimension"
  ) %>%
  rename(item1 = token)
```

```{r}
# Change the word embedding format to be more tidy
tidy_glove <- glove6b %>%
  pivot_longer(contains("d"),
               names_to = "dimension"
  ) %>%
  rename(item1 = token)
```

```{r}
# Document embedding
## GloVe embedding doens't handle unseen words, so we have to match 
## the abstracts and GloVe embedding to have the same set of vocabulary
word_matrix <- tidy_abstracts %>%
  inner_join(tidy_glove %>%
               distinct(item1) %>%
               rename(word = item1)) %>%
  count(id, word)
doc_ids <- distinct(word_matrix, id) # Retrieve the document ID
word_matrix_cast <-cast_sparse(word_matrix, id, word, n)

glove_matrix <- tidy_glove %>%
  inner_join(tidy_abstracts %>%
               distinct(word) %>%
               rename(item1 = word)) %>%
  cast_sparse(item1, dimension, value)

## Make the document matrix that contains document vectors
doc_matrix <- word_matrix_cast %*% glove_matrix
doc_matrix <- as.matrix(doc_matrix)
doc_matrix <- as.data.frame(doc_matrix)
doc_matrix <- cbind(doc_ids,doc_matrix ) # Put the document IDs back
doc_matrix
```
Up to this step, a document matrix with the corresponding vectors that represents the documentation is obtained. We can then perform matrix algera, or use the vectors to feed in machine learning algorithms for similarity detection or classfication. Further investigation is needed to perform the tasks!

### 2020-08-17 model starters/BERT.R
BERT OCR correctiong Python example

@@ -1,9 +1,29 @@
 # Source tutorial: https://blogs.rstudio.com/ai/posts/2019-09-30-bert-r/
+setwd("~/Desktop/repos/consistency_scores_in_text_datasets/model starters")
+library(tesseract)
+eng <- tesseract("eng")
+text_original <- tesseract::ocr("sample6.png", engine = eng)
+cat(text_original)
+
+library(textclean)
+testtxt <- c("His Majesty’s Government wish to add that they have no im@@@l@fon of requesting the establishment of military bases in peace time within the area of Palestine now united to the Kis@@®m of Jordan.")
+testtxt <- strip(testtxt, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
+      lower.case = FALSE)
+
+library(hunspell)
+incorrectwords <- hunspell(testtxt)
+print(incorrectwords)
+
+for (w in incorrectwords) {
+  testtxt <- str_replace_all(testtxt, w, "[MASK]")
+}
+
+
+
 
 Sys.setenv(TF_KERAS=1)
-
 # Go to the tutorial link in the above comment to download the files
-pretrained_path = '/change/this/path/uncased_L-12_H-768_A-12'
+pretrained_path = 'uncased_L-12_H-768_A-12'
 
 config_path = file.path(pretrained_path, 'bert_config.json')
 checkpoint_path = file.path(pretrained_path, 'bert_model.ckpt')

### 2020-08-19 model starters/BERT2.R
BERT implementation in R for word prediction

@@ -4,46 +4,44 @@ Sys.setenv(TF_KERAS=1)
 reticulate::py_config()
 setwd("~/Desktop/repos/consistency_scores_in_text_datasets/model starters")
 
-# 1. Read text from image
+# Read text from image
 library(tesseract)
 eng <- tesseract("eng")
 text_original <- tesseract::ocr("BERT Python/image1.png", engine = eng)
 cat(text_original)
 
-# 2. Collect incorrect words
+# Collect incorrect words
 library(textclean)
 text_stripped <- strip(text_original, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
-                       lower.case = FALSE)
+      lower.case = FALSE)
 library(hunspell)
 incorrectwords <- hunspell(text_stripped)
 
-# 3. Configure paths to the BERT pre-trained files variables
-### The zip file is available through the tutorial link at the first line
+# Configure paths to the pre-trained files variables
 pretrained_path = 'uncased_L-12_H-768_A-12'
 config_path = file.path(pretrained_path, 'bert_config.json')
 checkpoint_path = file.path(pretrained_path, 'bert_model.ckpt')
 vocab_path = file.path(pretrained_path, 'vocab.txt')
 seq_length = NULL
 
-# 4. Load BERT and tokenize the text
+# Load BERT and tokenize the text
 library(reticulate)
 k_bert = import('keras_bert')
 token_dict = k_bert$load_vocabulary(vocab_path)
 tokenizer = k_bert$Tokenizer(token_dict)
+encodes <- tokenizer$encode(text_original)
 model = k_bert$load_trained_model_from_checkpoint(
   config_path,
   checkpoint_path,
   seq_len=seq_length,
   training=T)
 
-# 5. Tokenize the text and get the encoding
 tokens <- tokenizer$tokenize(text_original)
-encodes <- tokenizer$encode(text_original)
+length(tokens)
 
-# 6. Get encoded inputs for the model: train, segments, targets
+# Get encoded inputs for the model: train, segments, targets
 indices = encodes[1]
 segments = encodes[2]
-## Manually flag word 13 and word 33 as masked words to be predicted 
 masks = list(list(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0,
[…]

### 2020-08-19 scripts/00-try_ngrams.R
Get example n-grams working

@@ -1,0 +1,98 @@
+#### Preamble ####
+# Contact: Rohan Alexander
+# Email: rohan.alexander@utoronto.ca
+# Date: 19 August 2020
+# Purpose: This script attempts to use n-grams to identify words that may be incorrect and 
+# need to be corrected.
+
+
+#### Workspace setup ####
+library(quanteda)
+library(tidyverse)
+
+
+#### Generate comparison dataset ####
+# First we need a collection of n-grams that we will search for, so we are creating
+# that in this section. We will just look at 3-grams, so once we have this list 
+# we can search for the first two tokens and then compare the third with what we
+# expect.
+example_paragraph <- "I should like to bury something precious in every place 
+where I've been happy and then, when I'm old and ugly and miserable, I could 
+come back and dig it up and remember. I should like to also ensure that my children
+and grandchildren knew where they were. For when I am old and ugly, I should like 
+that they too dig it up and imagine me as I am now - happy and content. I should lIke 
+also to be to be happy and content then, but really, who can promise that?"
+
+# Create tokens
+tokens_from_example <- quanteda::tokens(example_paragraph, remove_punct = TRUE)
+
+# Create ngrams from the tokens
+toks_ngram <- quanteda::tokens_ngrams(tokens_from_example, n = 3)
+head(toks_ngram[[1]])
+
+# Convert to tibble so we can use our familiar verbs
+all_tokens <- tibble(tokens = toks_ngram[[1]])
+
+# We only want the common ones, not every one.
+all_tokens <- 
+  all_tokens %>% 
+  group_by(tokens) %>% 
+  count() %>% 
+  filter(n > 1) %>% 
+  ungroup()
+
+# Create a tibble that has the first two words in one column then the third
+all_tokens <- 
+  all_tokens %>% 
+  mutate(tokens = str_replace_all(tokens, "_", " "),
+         first_words = word(tokens, start = 1, end = 2),
+         last_word = word(tokens, -1),
+         tokens = str_replace_all(tokens, " ", "_"),
+         first_words = str_replace_all(first_words, " ", "_")
+         ) %>% 
+  rename(last_word_expected = last_word) %>% 
+  select(-n, -tokens)
+
+
+#### Apply to the dataset to create a consistency score ####
+# Now that we have our collection of n-grams (this will be internal consistency
[…]

### 2020-08-21 outputs/documents/paper.Rmd
externalu consistency ngram

@@
 
-Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation (@brown1992class).
+Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation [@brown1992class].
 In our application, we use a tri-gram based model to detect both real-word and non-word errors and correct them with the candidate word that has the highest probability.
@@
 
-There are a variety of ways to implement an n-gram model within R @citeR including using R packages such as `Quanteda` (@citequanteda), `tidyText` (@citetidytext) and `tm` (@citetm).
+There are a variety of ways to implement an n-gram model within R @citeR including using R packages such as `Quanteda` [@citequanteda], `tidyText` [@citetidytext] and `tm` [@citetm].
 We used the Text Analysis Utilities, `tau`, R package @citetau because of we can separate text-cleaning and n-gram making in two different steps.
@@
 The most important argument of this function is _corpus_, as well as _n_ to set the number of words in a sequence, _method_ to determine which type of counts to compute, and _split_ to designate the regular expression pattern for splitting.
-The function outputs a `textcnt` object, which is essentially a table with each tri-gram and their counts (@citetau).
+The function outputs a `textcnt` object, which is essentially a table with each tri-gram and their counts [@citetau].
 
@@
 
-Here, we follow @falbel2017word to implement a skip-gram model of word2vec through Keras.
+Here, we follow @falbel2017word to implement a skip-gram model of Word2Vec through Keras.
 There are three steps in the Keras deep learning framework: 1) prepare the data; 2) define the model; and 3) train and evaluate the model.
-**(Ke-Li: Can we change this to be a different dataset please?)** Following these steps, we first downloaded the **(Ke-Li: Is there a reason this is in brackets?)** [Amazon Fine Foods Reviews dataset from Stanford] [@amzfinefooddata] and preprocess the text into tokens.
+**(Ke-Li: Can we change this to be a different dataset please?)** Following these steps, we first downloaded the Google News dataset [@citegooglenews] and preprocess the text into tokens.
 
-**(Ke-Li: Add code.)**
+```{r, eval=FALSE}
+tokenizer %>% fit_text_tokenizer(training_corpus)
+```
 
@@
 
-**(Ke-Li: Add code.)**
+```{r, eval=FALSE}
+skipgrams_generator <- function(text, tokenizer, window_size, negative_samples) {
+  gen <- texts_to_sequences_generator(tokenizer, sample(text))
+  function() {
+    skip <- generator_next(gen) %>%
+      skipgrams(
+        vocabulary_size = tokenizer$num_words, 
+        window_size = window_size, 
+        negative_samples = 1
+      )
+    x <- transpose(skip$couples) %>% map(. %>% unlist %>% as.matrix(ncol = 1))
+    y <- skip$labels %>% as.matrix(ncol = 1)
+    list(x, y)
+  }
+}
+```
 
@@
 
-**(Ke-Li: Add code.)**
+```{r, eval=FALSE}
+model %>%
+  fit_generator(
+    skipgrams_generator(training_corpus, tokenizer, skip_window, negative_samples), 
+    steps_per_epoch = 10, epochs = 1
+  )
+```
 
@@
 One is to load GloVe embeddings into a frozen Keras Embedding layer [@kerasglove] using the `Keras` R package [@kerasR2020].
-This means **(Ke-Li: Add what it means please.)** Another way, is to use the `text2vec` package [@text2vec].
+This means the embedding vectors will not be updated during the training process.
+Another way, is to use the `text2vec` package [@text2vec].
 
@@
 The bidirectional approach of BERT is alleviated by the "Masked Language Model" [@devlin2018bert].
-The Masked Language Model masks certain words in the input corpus which forces the model to predict the original words by "filling in the blank".
+The Masked Language Model masks random words in the input corpus which forces the model to predict the original words by "filling in the blank".
 This training process blends the representation from the left and right context and therefore increases the model's context sensitivity [@nozza2020mask].
@@
 Finally, the model is compiled and we can begin the fine-tuning.
+
+For the task of text cleaning, fine-tuning is not required.
+The process of text cleaning involves detecting and correcting mis-spelled words.
+Correcting the incorrect words can be seen as predicting the correct words.
+The Masked Language Model of BERT can be employed to do such prediction.
+In the pre-training process, the Masked Language Model would mask random words to do the prediction.
+However, we can tell the model what we want to "mask" by assigning the words to be encoded as `[MASK]`.
+Consider the erroneous sentence that has ten words:
+
+> "There was no possibility of taking a wlak that day."
+
+Using a spell-checker, "wlak" is identified as an error and is substituted by `[MASK]`:
+
+> "There was no possibility of taking a [MASK] that day."
+
+The implementation of text correction with BERT in R with the `keras-bert` Python module is not yet completed.
+For the purpose of demonstrating how BERT predicts the missing words, we will show the implementation process in Python using the `pytorch_pretrained_bert` library.
+The reason to use `pytorch_pretrained_bert`(**Keli add citation**) instead of `keras-bert`is that `pytorch_pretrained_bert` has relatively comprehensive functions that makes the prediction of missing words more straightforward.
+
+After we retrieved the new text data with the error replaced by `[MASK]`, we use the `tokenizer()` function in `pytorch_pretrained_bert` to tokenize the text:
+
+```{python, eval=FALSE}
+tokenized_text = tokenizer.tokenize("There was no possibility of taking a [MASK] that day.")
+tokenized_text
+>>> ['there', 'was', 'no', 'possibility', 'of', 'taking', 'a', '[MASK]', 'that', 'day', '.']
+```
+
+We then get the index of `[MASK]` within the sentence and save it to a list. Since we only have one `[MASK]` token, the list has only one element:
+
+```{python, eval=FALSE}
+mask_index = 7
+```
+
+`pytorch_pretrained_bert` has a dictionary of tokens generated from the pre-training stage.
+The tokens include the vocabulary, puncuations and special symbols.
+Each token in the dictionary is given an index, represented as `{token: index}`.
+The function `convert_tokens_to_ids()` encodes the tokens we got previously to their positional indices in the tokens dictionary and collects them into a list:
+
+```{python, eval=FALSE}
+indexed_tokens = tokenizer.convert_tokens_to_ids(tokenized_text)
+indexed_tokens
+>>> [2045, 2001, 2053, 6061, 1997, 2635, 1037, 103, 2008, 2154, 1012]
+```
+
+The next step is to encode the text by its segments.
+The model encodes a sentence of representing each word in the sentence with the positional index of the sentence in the provided corpus.
+For example, the example text has only one sentence "There was no possibility of taking a [MASK] that day." Therefore, the position of the sentence is "0", and the encoding for each word in the sentence, including the period is a series of "0" to the length of the sentence:
+
+```{python, eval=FALSE}
+segments_tokens = [0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
+```
+
+The previous two steps give us two set of encoding — `indexed_tokens` that contains the encoding for each token, and `segments_tokens` that contains the encoding for each sentence.
+These two encoding layers that we feed to BERT Masked Language Model.
+The model takes tensor objects as inputs [@paszke2017automatic], therefore we transformed the two lists into tensor objects before hands.
+These two tensor objects has two dimensions.
+The first dimension is the container, and the second dimension is the length of the token lists.
+
+```{python, eval=FALSE}
+tokens_tensor = torch.tensor([indexed_tokens])
+tokens_tensor
+
+segments_tensor = torch.tensor([segments_encodes])
+segments_tensor
+>>> [[0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]]
+```
+
+The encoding layers can be represented as the following:
+
+```{r table3, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
+tabl <- "
+|        |'there' |'was' |'no' |'possibility'|'of'|'taking'|'a' |'[MASK]'|'that'|'day'| '.'|
+|--------|--------|------|-----|-------------|----|--------|----|--------|------|-----|----|
+|token   |2045    |2001  |2053 |6061         |1997|2635    |1037|103     |2008  |2154 |1012|
+|segment |0       |0     |0    |0            | 0  |0       |0   |0       |0     |0    |0   |
+
+Table2. The smoothed probability table
+"
+cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
+```
+
+We then load the pre-trained BertForMaskedLM and feed the two tensor objects to it.
+The model then returns a tensor object that has three dimensions.
+The `predictions` tensor objects contains the prediction score of all the words in the vocabulary.
+
+```{python, eval=FALSE}
+model = BertForMaskedLM.from_pretrained('bert-base-uncased')
+predictions = model(tokens_tensor, segments_tensor)
+predictions
+```
+
+To get the prediction, we will get the element with the highest prediction score and retrieve its index.
+The retrieved index is then used to find the corresponding token in the tokens dictionary.
+The predicted word is "walk", which matches with the original sentence in **Jane Eyre**.
+
+```{python, eval=FALSE}
+predicted_word_index = torch.argmax(predictions[0, mask_index]).item()
+predicted_word_index
+>>> 3328
+predicted_word_token = tokenizer.convert_ids_to_tokens([predicted_index])[0]
+predicted_word_token
+>>> walk
+```
 

### 2020-08-21 scripts/00-try_document_embedding.Rmd
externalu consistency ngram

--- 2020-08-16 model starters/DocumentEmbedding.Rmd
@@
 # Tokenize the abstract
-abstracts <- read_csv("abstracts_101.csv")
+abstracts <- read_csv("input-abstracts.csv")
 tidy_abstracts <- abstracts %>%
@@
 ```
+
 Up to this step, a document matrix with the corresponding vectors that represents the documentation is obtained.

### 2020-08-21 model_starters/Word2Vec.R

@@ -62,7 +62,7 @@ summary(model)
 #model <- keras::train_on_batch(model)
 
 model %>%
-  fit(
+  fit_generator(
     skipgrams_generator(training_corpus, tokenizer, skip_window, negative_samples), 
     steps_per_epoch = 10, epochs = 1
   )

### 2020-08-21 scripts/00-try_external_consistency.R

@@ -1,0 +1,140 @@
+library(tidyverse)
+library(tau)
+library(tm)
+library(stringr)
+library(stats)
+library(reshape2)
+library(RWeka)
+
+#setwd("~/Desktop/repos/consistency_scores_in_text_datasets/scripts/")
+path <- getwd()
+demoTraining <- readLines(paste(path,"/input-sample_en_US.blogs.txt", sep=""), warn=FALSE)
+
+#### OCR Read text from image ####
+# To-do: make an UI for this to upload PDF
+library(tesseract)
+eng <- tesseract("eng")
+input_original <- tesseract::ocr("input-image.png", engine = eng)
+cat(input_original)
+
+#### Collect incorrect words from the OCR ####
+library(textclean)
+input_stripped <- strip(input_original, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
+                       lower.case = FALSE)
+# Tokenizer function
+tokenizer <- function(corpus) {
+  lines <- vector()
+  # Read the corpus line by line
+  for (line in corpus) {
+    lines <-c(lines, line)
+  }
+  lines <- tolower(lines)
+  lines <- gsub("[.!?]$|[.!?] |$", " </s>", lines) # add end-of-sentence token to the end of a sentence
+  tokens <- unlist(strsplit(lines, "[^a-z<>/]"))
+  tokens <- tokens[tokens != ""]
+  return(tokens)
+}
+
+library(hunspell)
+# Use hunspell spell checker to get text errors
+incorrectwords <- hunspell(input_stripped)
+incorrectwords <- lapply(incorrectwords, tolower)
+# Split the input text
+input_tokens <- tokenizer(input_stripped)
+# Get the indices of the incorrect words in the tokens
+incorrectwords_indices <- match(as.vector(incorrectwords[[1]]),input_tokens)
+# Get the previous two words of the text error and put them in the list of bigrams
+bigrams_for_prediction <- vector()
+for(i in incorrectwords_indices) {
+  previous_two_words <- paste(input_tokens[i-2], input_tokens[i-1])
+  bigrams_for_prediction <- append(bigrams_for_prediction, previous_two_words)
+}
+
+#### Training text tokenization ####
+tokens <- tokenizer(demoTraining)
+
+#### Make data table for the ngrams ####
+library(data.table)
[…]

### 2020-08-21 scripts/00-try_ngrams.R

@@ -96,3 +96,12 @@ all_tokens_with_errors %>%
 all_tokens_with_errors %>% 
   mutate(as_expected = last_word == last_word_expected) %>% 
   filter(as_expected == FALSE)
+
+### To-Dos
+# Text pre-processing: to lowercase and remove stopwords
+# Code clean up and tighten up
+# Function 1: Create ngrams for internal source of text
+# Function 2: Create ngrams for input text
+# Output 1: Internal consistency score
+# Output 2: Summary tables of expected words
+# Package website

### 2020-08-21 shinyapp/external_consistency_func.R

@@ -8,8 +8,15 @@ library(RWeka)
 
 #setwd("~/Desktop/repos/consistency_scores_in_text_datasets/shinyapp")
 path <- getwd()
-demoTraining <- readLines(paste(path,"/samples/demo.txt", sep=""), warn=FALSE)
+demoTraining <- readLines(paste(path,"/samples/sample_en_US.blogs.txt", sep=""), warn=FALSE)
 
+######## OCR Read text from image ########
+library(tesseract)
+eng <- tesseract("eng")
+text_original <- tesseract::ocr("BERT Python/image1.png", engine = eng)
+cat(text_original)
+
+######## Training Text Tokenization ######## 
 tokenizer <- function(corpus) {
   lines <- vector()
   # Read the corpus line by line
@@ -37,7 +44,22 @@ unigrams <- unigrams[!grepl("</s>", names(unigrams))]
 bigrams <-   bigrams[!grepl("</s>", names(bigrams))]
 trigrams <- trigrams[!grepl("</s>", names(trigrams))]
 
-### ngram probability
+######## Collect incorrect words ######## 
+text_original <- "There was na possibility of taking a wlak that day"
+library(textclean)
+text_stripped <- strip(text_original, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
+                       lower.case = FALSE)
+library(hunspell)
+incorrectwords <- hunspell(text_stripped)
+text_tokens <- tokenizer(text_stripped)
+indices <- match(as.vector(incorrectwords[[1]]),text_tokens)
+bigrams <- vector()
+for(i in indices) {
+  bigram <- paste(text_tokens[i-2], text_tokens[i-1])
+  bigrams <- append(bigrams, bigram)
+}
+
+########  ngram probability ######## 
 getLastWords <- function(string, words) {
   pattern <- paste("[a-z']+( [a-z']+){", words - 1, "}$", sep="")
   return(substring(string, str_locate(string, pattern)[,1]))
@@ -47,7 +69,7 @@ removeLastWord <- function(string) {
   sub(" [a-z']+$", "", string)
 }
 
-# Kneser-Ney Smoothing
+######## Kneser-Ney Smoothing ########
 kneserNay <- function(ngrams, d) {
   n <- length(strsplit(names(ngrams[1]), " ")[[1]])
   # Special case for unigrams
@@ -83,11 +105,11 @@ library(data.table)
 
 unigramDF <- data.table("Words" = (names(unigrams)), 
[…]

### 2020-08-21 shinyapp/internal_consistency_func.R

@@ -1,0 +1,87 @@
+#### Preamble ####
+# Contact: Rohan Alexander
+# Email: rohan.alexander@utoronto.ca
+# Date: 19 August 2020
+# Purpose: This script attempts to use n-grams to identify words that may be incorrect and 
+# need to be corrected.
+
+#### Workspace setup ####
+library(quanteda)
+library(tidyverse)
+
+#### Generate comparison dataset ####
+# First we need a collection of n-grams that we will search for, so we are creating
+# that in this section. We will just look at 3-grams, so once we have this list 
+# we can search for the first two tokens and then compare the third with what we
+# expect.
+
+create_ngrams_tibble <- function(corpus, is_training) {
+  # Create tokens
+  tokens <- quanteda::tokens(corpus, remove_punct = TRUE)
+  # Create ngrams from the tokens
+  toks_ngram <- quanteda::tokens_ngrams(tokens, n = 3)
+  # Convert to tibble so we can use our familiar verbs
+  all_tokens <- tibble(tokens = toks_ngram[[1]])
+  
+  # If it's a training corpus, we only want the common ones, not every one.
+  if (is_training) {
+    all_tokens <- 
+      all_tokens %>% 
+      group_by(tokens) %>% 
+      count() %>% 
+      filter(n > 1) %>% 
+      ungroup()
+  }
+  
+  # Create a tibble that has the first two words in one column then the third
+  all_tokens <- 
+    all_tokens %>% 
+    mutate(tokens = str_replace_all(tokens, "_", " "),
+           first_words = word(tokens, start = 1, end = 2),
+           last_word = word(tokens, -1),
+           tokens = str_replace_all(tokens, " ", "_"),
+           first_words = str_replace_all(first_words, " ", "_")
+    ) 
+  
+  # If it's a training corpus, drop n column and tokens column
+  if (is_training) {
+    all_tokens <- all_tokens %>% 
+      rename(last_word_expected = last_word) %>% 
+      select(-n, -tokens)
+  }
+  
+  return(all_tokens)
+}
+
+# Get training n-gram tibble
+#training_tokens<- create_ngrams_tibble(example_paragraph, is_training = TRUE)
+# Get example n-gram tibble
+#input_tokens<- create_ngrams_tibble(example_sentence_with_errors, is_training = FALSE)
+
+#### Apply to the dataset to create a consistency score ####
+# Now that we have our collection of n-grams (this will be internal consistency
[…]

### 2020-08-21 shinyapp/server.R

@@ -1,46 +1,57 @@
 library(tidyverse)
 library(tau)
 library(tm)
-source("text-input.R")
-source("functions.R")
+source("training_data.R")
+source("internal_consistency_func.R")
 
 server <- function(input, output, session) {
-  getText <- eventReactive(input$getResult, {
-    myText <- trimws(input$myText)
-    myText <- gsub('\\s+', ' ', myText)
-    detector(myText)
+  getResults <- eventReactive(input$getResult, {
+    internalTrainingData <- trimws(input$internalTrainingData)
+    myTextData <- trimws(input$myTextData)
+    myTextData <- gsub('\\s+', ' ', myTextData)
+    get_results(create_ngrams_tibble(internalTrainingData, is_training = TRUE), create_ngrams_tibble(myTextData, is_training = FALSE))
   })
   
   # Displays input text
-  output$outputText <- renderText({
-    getText()
+  output$unexpected_words <- renderTable({
+    getResults()[2]
   })
   
-  
-  # Fucntion that displays scores when get result button is clicked
-  getScores <- eventReactive(input$getResult, {
-    "Internal consistency score: 0.919 <br> External consistency score: 0.956"
-  })
   # Displays scores
-  output$outputScores <- renderText({
-    getScores()
+  output$internal_consistency_score <- renderTable({
+    getResults()[1]
   })
   
   # Disable get result button and clear button when there is no text data
   observe({
-    toggleState(id = "getResult", condition = nchar(input$myText) > 0)
-    toggleState(id = "clear", condition = nchar(input$myText) > 0)
+    toggleState(id = "getResult", condition = nchar(input$myTextData) > 0)
+    toggleState(id = "clearText", condition = nchar(input$myTextData) > 0)
+    toggleState(id = "clearTraining", condition = nchar(input$internalTrainingData) > 0)
+  })
+  
+  observeEvent(input$getResult, {
+    shinyjs::show("results_row")
   })
   
   # Load demo data when the corresponding button is clicked
   observe({
     input$loadDemo
-    updateTextInput(session, "myText", value = demoInput)
+    updateTextInput(session, "myTextData", value = demoTextData)
+  })
+  
+  observe({
+    input$loadTrainingDemo
+    updateTextInput(session, "internalTrainingData", value = demoInternalTrainingData)
   })
   
   # Clear text input area
   observe({
-    input$clear
-    updateTextInput(session, "myText", value = "")
+    input$clearTraining
[…]

### 2020-08-21 shinyapp/text-input.R

@@ -1,1 +1,0 @@
-demoInput <- "There was no possibility of taking a walk that day. We had been wandering, indeed, in tbe leafless shrubbery an hour in tbe morning; but since dinner (Mrs. Reeal, when tbere was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, tbat further out-door exercise was now out of the question."

### 2020-08-21 shinyapp/training_data.R

@@ -1,0 +1,8 @@
+demoTextData <- "I should lIke also to be to be happy and content 
+then, but really, who can promise that?"
+demoInternalTrainingData <- "I should like to bury something precious in every place 
+where I've been happy and then, when I'm old and ugly and miserable, I could 
+come back and dig it up and remember. I should like to also ensure that my children
+and grandchildren knew where they were. For when I am old and ugly, I should like 
+that they too dig it up and imagine me as I am now - happy and content. I should lIke 
+also to be to be happy and content then, but really, who can promise that?"

### 2020-08-21 shinyapp/ui.R

@@ -10,21 +10,40 @@ fluidPage(
   useShinyjs(),
   theme = shinytheme("cosmo"),
   titlePanel("Text Consistency Score"),
-
+  # Input row
   fluidRow(
+    # Input Panel: Internal Training Data
     column(6, wellPanel(
-      h4("Place your data here"),
-      actionButton("loadDemo", "Load Demo Data", class = "btn-sm btn-primary"),
-      textAreaInput("myText", "", value = "", height = "500px"),
-      actionButton("getResult", "Get Result", class = "btn-primary"),
-      actionButton("clear", "Clear", class = "btn-outline-secondary"),
-
+      h4("Internal Training Data"),
+      actionButton("loadTrainingDemo", "Load Demo Data", class = "btn-sm btn-primary"),
+      textAreaInput("internalTrainingData", "", value = "", height = "300px"),
+      actionButton("clearTraining", "Clear", class = "btn-sm btn-outline-secondary"),
     )),
+    # Input Panel
     column(6, wellPanel(
-      h4("See your result here"),
-      tableOutput("outputText"),
-      hr(),
-      tableOutput("outputScores"),
-    ))
+      h4("Your Text Data"),
+      actionButton("loadDemo", "Load Demo Data", class = "btn-sm btn-primary"),
+      textAreaInput("myTextData", "", value = "", height = "300px"),
+      actionButton("clearText", "Clear", class = "btn-sm btn-outline-secondary"),
+    )),
+  ),
+  # Action button row
+  fluidRow(
+    column(12, 
+      actionButton("getResult", "Get Result", class = "btn-primary"),
+    )
+  ),
+  # Result row
+  shinyjs::hidden(
+    fluidRow( id = "results_row",
+      column(8, 
+        h5("Unexpected Words"),
+        tableOutput("unexpected_words"),
+      ),
+      column(4, 
+        h5("Internal Consistency Score"),
+        tableOutput("internal_consistency_score"),
+      )
+    )
   )
 )

### 2020-08-21 outputs/documents/paper.Rmd
glove code examples

@@
 There are three steps in the Keras deep learning framework: 1) prepare the data; 2) define the model; and 3) train and evaluate the model.
-**(Ke-Li: Can we change this to be a different dataset please?)** Following these steps, we first downloaded the Google News dataset [@citegooglenews] and preprocess the text into tokens.
+**(Ke-Li: Can we change this to be a different dataset please?
+Yes, I changed it to Google News for now, let's change it to an official data set that we will decide on later!)** Following these steps, we first downloaded the Google News dataset [@citegooglenews] and preprocess the text into tokens.
 
@@
 
-Here we follow **CITE** to implement GloVe using the `keras` R package [@kerasglove].
+Here we follow the blog post published by Keras [@kerasglove] to implement GloVe using the `keras` R package [@kerasglove].
 The blog post is aimed to create a word embedding model for text classification.
@@
 
-**[As before, we need code and examples here]**
+To use GloVe in R, we first download the pre-trained GloVe vectors. The file comes in zip format, which needs to be unzipped:
+
+```{r, eval=FALSE}
+download.file('http://nlp.stanford.edu/data/glove.6B.zip',destfile = 'glove.6B.zip')
+unzip('glove.6B.zip')
+```
+
+The next step is to load the GloVe vectors into R:
+
+```{r, eval=FALSE}
+vectors = data.table::fread('glove.6B.300d.txt', data.table = F,  encoding = 'UTF-8')
+colnames(vectors) = c('word',paste('dim',1:300,sep = '_'))
+```
+
+The GloVe vectors are now available in R. We restructure the word vectors into a tibble data frame for easy access. The tibble dataset has 400,000 words and 301 dimensions to represent each word:
+
+```{r, eval=FALSE}
+as_tibble(vectors)
+# A tibble: 400,000 x 301
+```
 

### 2020-08-21 model_starters/GloVe.R

@@ -1,38 +1,19 @@
 # Source tutorial: https://keras.rstudio.com/articles/examples/pretrained_word_embeddings.html
+# This example shows how one can quickly load glove vectors
+# and train a Keras model in R
+
 library(keras)
+library(dplyr)
 
-GLOVE_DIR <- 'glove.6B'
-TEXT_DATA_DIR <- '20_newsgroup'
-MAX_SEQUENCE_LENGTH <- 1000
-MAX_NUM_WORDS <- 20000
-EMBEDDING_DIM <- 100
-VALIDATION_SPLIT <- 0.2
-
-# download data if necessary
-download_data <- function(data_dir, url_path, data_file) {
-  if (!dir.exists(data_dir)) {
-    download.file(paste0(url_path, data_file), data_file, mode = "wb")
-    if (tools::file_ext(data_file) == "zip")
-      unzip(data_file, exdir = tools::file_path_sans_ext(data_file))
-    else
-      untar(data_file)
-    unlink(data_file)
-  }
-}
-download_data(GLOVE_DIR, 'http://nlp.stanford.edu/data/', 'glove.6B.zip')
-
-# first, build index mapping words in the embeddings set
-# to their embedding vector
-
-cat('Indexing word vectors.\n')
-
-embeddings_index <- new.env(parent = emptyenv())
-lines <- readLines(file.path(GLOVE_DIR, 'glove.6B.100d.txt'))
-for (line in lines) {
-  values <- strsplit(line, ' ', fixed = TRUE)[[1]]
-  word <- values[[1]]
-  coefs <- as.numeric(values[-1])
-  embeddings_index[[word]] <- coefs
+# Download Glove vectors if necessary
+if (!file.exists('glove.6B.zip')) {
+  download.file('http://nlp.stanford.edu/data/glove.6B.zip',destfile = 'glove.6B.zip')
+  unzip('glove.6B.zip')
 }
 
-cat(sprintf('Found %s word vectors.\n', length(embeddings_index)))
+# load glove vectors into R
+vectors = data.table::fread('glove.6B.300d.txt', data.table = F,  encoding = 'UTF-8')
+colnames(vectors) = c('word',paste('dim',1:300,sep = '_'))
+
+# structure of the vectors
+as_tibble(vectors)

### 2020-08-24 shinyapp/OCR_text_func.R
OCR text error suggestion

@@ -1,0 +1,7 @@
+#### OCR Read text from image ####
+library(tesseract)
+eng <- tesseract("eng")
+get_OCR_text<- function(file_path){
+  input_original <- tesseract::ocr(file_path, engine = eng)
+  return(input_original)
+}

### 2020-08-24 shinyapp/dropbox_auth.R

@@ -1,0 +1,10 @@
+library(rdrop2)
+drop_auth()
+token <- drop_auth()
+saveRDS(token, file = "token.rds")
+drop_dir() %>% 
+  filter(.tag == "Arianna")
+
+trigramsDF <- drop_read_csv("Arianna/trigramsDF.csv")
+
+runExample("09_upload")

### 2020-08-24 shinyapp/external_consistency_func.R

@@ -1,28 +1,16 @@
+#### Create bigrams from text errros' previous two words ####
+library(textclean)
+library(hunspell)
 library(tidyverse)
-library(tau)
-library(tm)
-library(stringr)
-library(stats)
-library(reshape2)
-library(RWeka)
+library(rdrop2)
+#trigramsDF <- drop_read_csv("Arianna/trigramsDF.csv")
+trigramsDF <- read_csv("trigramsDF.csv")
 
-#setwd("~/Desktop/repos/consistency_scores_in_text_datasets/shinyapp")
-path <- getwd()
-demoTraining <- readLines(paste(path,"/samples/sample_en_US.blogs.txt", sep=""), warn=FALSE)
-
-######## OCR Read text from image ########
-library(tesseract)
-eng <- tesseract("eng")
-text_original <- tesseract::ocr("BERT Python/image1.png", engine = eng)
-cat(text_original)
-
-######## Training Text Tokenization ######## 
+# Tokenizer function
 tokenizer <- function(corpus) {
   lines <- vector()
   # Read the corpus line by line
   for (line in corpus) {
-    # add two start-of-sentence tokens to the beginning of a sentence
-    #line <- paste("<s> <s>", line)
     lines <-c(lines, line)
   }
   lines <- tolower(lines)
@@ -32,96 +20,42 @@ tokenizer <- function(corpus) {
   return(tokens)
 }
 
-tokens <- tokenizer(demoTraining)
-
-library(data.table)
-unigrams <- textcnt( tokens, n=1, method = "string", split = "[[:space:][:digit:]]+", decreasing = TRUE)
-bigrams  <- textcnt( tokens, n=2, method = "string", split = "[[:space:][:digit:]]+", decreasing = TRUE)
-trigrams <- textcnt( tokens, n=3, method = "string", split = "[[:space:][:digit:]]+", decreasing = TRUE)
-
-# Remove trigram that has "</s>" in it, so the trigrams are constructed within a sentence
-unigrams <- unigrams[!grepl("</s>", names(unigrams))]
-bigrams <-   bigrams[!grepl("</s>", names(bigrams))]
-trigrams <- trigrams[!grepl("</s>", names(trigrams))]
-
-######## Collect incorrect words ######## 
-text_original <- "There was na possibility of taking a wlak that day"
-library(textclean)
-text_stripped <- strip(text_original, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
[…]

### 2020-08-24 shinyapp/generate_trigramsDF_csv.R

@@ -1,0 +1,123 @@
+library(tidyverse)
+library(tau)
+library(tm)
+library(stringr)
+library(stats)
+library(reshape2)
+
+path <- getwd()
+external_training_data <- readLines(paste(path,"/samples/input-sample_en_US.blogs.txt", sep=""), warn=FALSE)
+
+library(textclean)
+library(hunspell)
+
+#### Tokenizer function ####
+tokenizer <- function(corpus) {
+  lines <- vector()
+  # Read the corpus line by line
+  for (line in corpus) {
+    lines <-c(lines, line)
+  }
+  lines <- tolower(lines)
+  lines <- gsub("[.!?]$|[.!?] |$", " </s> ", lines) # add end-of-sentence token to the end of a sentence
+  tokens <- unlist(strsplit(lines, "[^a-z<>/]"))
+  tokens <- tokens[tokens != ""]
+  return(tokens)
+}
+
+#### Create bigrams from text errros' previous two words ####
+create_bigrams_for_predictions <- function(input_original) {
+  
+  input_stripped <- strip(input_original, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
+                          lower.case = FALSE)
+  # Use hunspell spell checker to get text errors
+  incorrectwords <- hunspell(input_stripped)
+  incorrectwords <- lapply(incorrectwords, tolower)
+  # Split the input text
+  input_tokens <- tokenizer(input_stripped)
+  # Get the indices of the incorrect words in the tokens
+  incorrectwords_indices <- match(as.vector(incorrectwords[[1]]),input_tokens)
+  # Get the previous two words of the text error and put them in the list of bigrams
+  bigrams_for_prediction <- vector()
+  for(i in incorrectwords_indices) {
+    previous_two_words <- paste(input_tokens[i-2], input_tokens[i-1])
+    bigrams_for_prediction <- append(bigrams_for_prediction, previous_two_words)
+  }
+  split_trigrams <- data.frame(bigrams_for_prediction, "incorrectwords" = incorrectwords[[1]])
+  
+  return(split_trigrams)
+}
+
+#### Training text tokenization ####
+tokens <- tokenizer(external_training_data)
+
+#### Make data table for the ngrams ####
+library(data.table)
[…]

### 2020-08-24 shinyapp/server.R

@@ -3,9 +3,21 @@ library(tau)
 library(tm)
 source("training_data.R")
 source("internal_consistency_func.R")
+source("external_consistency_func.R")
+source("OCR_text_func.R")
 
+#### Dorpbox authentication ####
+library(rdrop2)
+drop_auth()
+token <- drop_auth()
+saveRDS(token, file = "token.rds")
+drop_dir() %>% 
+  filter(.tag == "Arianna")
+#trigramsDF <- drop_read_csv("Arianna/trigramsDF.csv")
+trigramsDF <- read_csv("trigramsDF.csv")
 server <- function(input, output, session) {
-  getResults <- eventReactive(input$getResult, {
+#### Internal consistency score ####
+  getInternalResults <- eventReactive(input$getInternalResult, {
     internalTrainingData <- trimws(input$internalTrainingData)
     myTextData <- trimws(input$myTextData)
     myTextData <- gsub('\\s+', ' ', myTextData)
@@ -14,29 +26,59 @@ server <- function(input, output, session) {
   
   # Displays input text
   output$unexpected_words <- renderTable({
-    getResults()[2]
+    getInternalResults()[2]
   })
   
   # Displays scores
   output$internal_consistency_score <- renderTable({
-    getResults()[1]
+    getInternalResults()[1]
   })
+  
+#### Exteranl consistency score ####
+  getExternalResults <- eventReactive(input$getExternalResult, {
+    # Get the DF of the split trigrams
+    myTextData_e <- trimws(input$myTextData_e)
+    split_trigrams <- create_bigrams_for_predictions(myTextData_e)
+    return(predict_words(split_trigrams))
+  })
+  
+  # Displays input text
+  output$unexpected_words_e <- renderTable({
+    getExternalResults()
+  })
+  
   
   # Disable get result button and clear button when there is no text data
   observe({
-    toggleState(id = "getResult", condition = nchar(input$myTextData) > 0)
+    toggleState(id = "getInternalResult", condition = nchar(input$myTextData) > 0)
+    toggleState(id = "getExternalResult", condition = nchar(input$myTextData_e) > 0)
     toggleState(id = "clearText", condition = nchar(input$myTextData) > 0)
+    toggleState(id = "clearText_e", condition = nchar(input$myTextData_e) > 0)
     toggleState(id = "clearTraining", condition = nchar(input$internalTrainingData) > 0)
   })
   
-  observeEvent(input$getResult, {
-    shinyjs::show("results_row")
-  })
+  observeEvent(
[…]

### 2020-08-24 shinyapp/ui.R

@@ -6,44 +6,81 @@ library(tidyverse)
 library(tau)
 library(tm)
 
-fluidPage(
-  useShinyjs(),
-  theme = shinytheme("cosmo"),
-  titlePanel("Text Consistency Score"),
-  # Input row
-  fluidRow(
-    # Input Panel: Internal Training Data
-    column(6, wellPanel(
-      h4("Internal Training Data"),
-      actionButton("loadTrainingDemo", "Load Demo Data", class = "btn-sm btn-primary"),
-      textAreaInput("internalTrainingData", "", value = "", height = "300px"),
-      actionButton("clearTraining", "Clear", class = "btn-sm btn-outline-secondary"),
-    )),
-    # Input Panel
-    column(6, wellPanel(
-      h4("Your Text Data"),
-      actionButton("loadDemo", "Load Demo Data", class = "btn-sm btn-primary"),
-      textAreaInput("myTextData", "", value = "", height = "300px"),
-      actionButton("clearText", "Clear", class = "btn-sm btn-outline-secondary"),
-    )),
-  ),
-  # Action button row
-  fluidRow(
-    column(12, 
-      actionButton("getResult", "Get Result", class = "btn-primary"),
-    )
-  ),
-  # Result row
-  shinyjs::hidden(
-    fluidRow( id = "results_row",
-      column(8, 
-        h5("Unexpected Words"),
-        tableOutput("unexpected_words"),
+navbarPage("Arianna",
+   theme = shinytheme("cosmo"),
+   tabPanel("Internal Consistency",
+      useShinyjs(),
+      fluidRow(
+        # Input Panel: Internal Training Data
+        column(6, wellPanel(
+          h4("Internal Training Data"),
+          actionButton("loadTrainingDemo", "Load Demo Data", class = "btn-sm btn-primary"),
+          textAreaInput("internalTrainingData", "", value = "", height = "300px"),
+          actionButton("clearTraining", "Clear", class = "btn-sm btn-outline-secondary"),
+        )),
+        # Input Panel
+        column(6, wellPanel(
+          h4("Your Text Data"),
+          actionButton("loadDemo", "Load Demo Data", class = "btn-sm btn-primary"),
+          textAreaInput("myTextData", "", value = "", height = "300px"),
+          actionButton("clearText", "Clear", class = "btn-sm btn-outline-secondary"),
+        )),
       ),
-      column(4, 
-        h5("Internal Consistency Score"),
[…]

### 2020-08-26 model_starters/BERT.R
BERT implementation in R updated

@@ -2,30 +2,18 @@
 Sys.setenv(TF_KERAS=1) 
 # to see python version
 reticulate::py_config()
-setwd("~/Desktop/repos/consistency_scores_in_text_datasets/model starters")
+setwd("~/Desktop/repos/consistency_scores_in_text_datasets/model_starters")
 
-# 1. Read text from image
-library(tesseract)
-eng <- tesseract("eng")
-text_original <- tesseract::ocr("BERT Python/image1.png", engine = eng)
-cat(text_original)
 
-# 2. Collect incorrect words
-library(textclean)
-text_stripped <- strip(text_original, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
-                       lower.case = FALSE)
-library(hunspell)
-incorrectwords <- hunspell(text_stripped)
-
-# 3. Configure paths to the BERT pre-trained files variables
+# 1. Configure paths to the BERT pre-trained files variables
 ### The zip file is available through the tutorial link at the first line
-pretrained_path = 'uncased_L-12_H-768_A-12'
+pretrained_path = '../../../uncased_L-12_H-768_A-12'
 config_path = file.path(pretrained_path, 'bert_config.json')
 checkpoint_path = file.path(pretrained_path, 'bert_model.ckpt')
 vocab_path = file.path(pretrained_path, 'vocab.txt')
 seq_length = NULL
 
-# 4. Load BERT and tokenize the text
+# 2. Load BERT and tokenize the text
 library(reticulate)
 k_bert = import('keras_bert')
 token_dict = k_bert$load_vocabulary(vocab_path)
@@ -35,26 +23,3 @@ model = k_bert$load_trained_model_from_checkpoint(
   checkpoint_path,
   seq_len=seq_length,
   training=T)
-
-# 5. Tokenize the text and get the encoding
-tokens <- tokenizer$tokenize(text_original)
-encodes <- tokenizer$encode(text_original)
-
-# 6. Get encoded inputs for the model: train, segments, targets
-indices = encodes[1]
-segments = encodes[2]
-## Manually flag word 13 and word 33 as masked words to be predicted 
-masks = list(list(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0,
-                  0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
[…]

### 2020-08-26 scripts/00-try-BERT-text-prediction.R

@@ -2,53 +2,70 @@
 Sys.setenv(TF_KERAS=1) 
 # to see python version
 reticulate::py_config()
-setwd("~/Desktop/repos/consistency_scores_in_text_datasets/model starters")
+setwd("~/Desktop/repos/consistency_scores_in_text_datasets/model_starters")
 
-# Read text from image
+#### 1. Read text from image
 library(tesseract)
 eng <- tesseract("eng")
-text_original <- tesseract::ocr("BERT Python/image1.png", engine = eng)
+text_original <- tesseract::ocr("BERT_Python/image1.png", engine = eng)
 cat(text_original)
 
-# Collect incorrect words
+#### 2. Collect incorrect words
 library(textclean)
 text_stripped <- strip(text_original, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
-      lower.case = FALSE)
+                       lower.case = FALSE)
+
 library(hunspell)
 incorrectwords <- hunspell(text_stripped)
 
-# Configure paths to the pre-trained files variables
-pretrained_path = 'uncased_L-12_H-768_A-12'
+tokens <- unlist(strsplit(text_stripped, "[^aA-zZ<>/]"))
+mask_indices <- match(c("imlfon", "Kism"),tokens)
+tokens[12] = tokens[33] = "[MASK]"
+toString(tokens)
+#### 3. Configure paths to the BERT pre-trained files variables
+# The zip file is available through the tutorial link at the first line
+pretrained_path = '../../../uncased_L-12_H-768_A-12'
 config_path = file.path(pretrained_path, 'bert_config.json')
 checkpoint_path = file.path(pretrained_path, 'bert_model.ckpt')
 vocab_path = file.path(pretrained_path, 'vocab.txt')
 seq_length = NULL
 
-# Load BERT and tokenize the text
+#### 4. Load BERT and tokenize the text
 library(reticulate)
 k_bert = import('keras_bert')
+np = import('numpy')
 token_dict = k_bert$load_vocabulary(vocab_path)
 tokenizer = k_bert$Tokenizer(token_dict)
-encodes <- tokenizer$encode(text_original)
 model = k_bert$load_trained_model_from_checkpoint(
   config_path,
   checkpoint_path,
   seq_len=seq_length,
   training=T)
 
[…]

### 2020-08-27 shinyapp/server.R
shiny app updated

@@ -1,53 +1,54 @@
-library(tidyverse)
-library(tau)
-library(tm)
+library(aRianna)
+#library(tau)
+#library(tm)
 source("training_data.R")
-source("internal_consistency_func.R")
-source("external_consistency_func.R")
+#source("internal_consistency_func.R")
+#source("external_consistency_func.R")
 source("OCR_text_func.R")
 
-#### Dorpbox authentication ####
-library(rdrop2)
-drop_auth()
-token <- drop_auth()
-saveRDS(token, file = "token.rds")
-drop_dir() %>% 
-  filter(.tag == "Arianna")
-trigramsDF <- drop_read_csv("Arianna/trigramsDF.csv")
-#trigramsDF <- read_csv("trigramsDF.csv")
 server <- function(input, output, session) {
 #### Internal consistency score ####
   getInternalResults <- eventReactive(input$getInternalResult, {
-    internalTrainingData <- trimws(input$internalTrainingData)
-    myTextData <- trimws(input$myTextData)
-    myTextData <- gsub('\\s+', ' ', myTextData)
-    get_results(create_ngrams_tibble(internalTrainingData, is_training = TRUE), create_ngrams_tibble(myTextData, is_training = FALSE))
+    internalTrainingData <- aRianna::make_internal_consistency_dataset(input$internalTrainingData)
+    aRianna::generate_internal_consistency_score(input$myTextData, internalTrainingData)
   })
-  
+
   # Displays input text
   output$unexpected_words <- renderTable({
-    getInternalResults()[2]
+    unexpected_words <- getInternalResults()[[1]]
+    colnames(unexpected_words) <- c("As Expected", "Count", "Consistency Score")
+    unexpected_words
   })
-  
+
   # Displays scores
   output$internal_consistency_score <- renderTable({
-    getInternalResults()[1]
+    internal_consistency_score <- getInternalResults()[[2]]
+    colnames(internal_consistency_score) <- c("First Words", "Last Word", "Last Word Expected", "Is Expected")
+    internal_consistency_score
   })
-  
+
 #### Exteranl consistency score ####
   getExternalResults <- eventReactive(input$getExternalResult, {
     # Get the DF of the split trigrams
     myTextData_e <- trimws(input$myTextData_e)
-    split_trigrams <- create_bigrams_for_predictions(myTextData_e)
-    return(predict_words(split_trigrams))
+    aRianna::generate_external_consistency_score(myTextData_e)
   })
-  
+
   # Displays input text
   output$unexpected_words_e <- renderTable({
[…]

### 2020-08-27 shinyapp/training_data.R

@@ -1,8 +1,2 @@
-demoTextData <- "I should lIke also to be to be happy and content 
-then, but really, who can promise that?"
-demoInternalTrainingData <- "I should like to bury something precious in every place 
-where I've been happy and then, when I'm old and ugly and miserable, I could 
-come back and dig it up and remember. I should like to also ensure that my children
-and grandchildren knew where they were. For when I am old and ugly, I should like 
-that they too dig it up and imagine me as I am now - happy and content. I should lIke 
-also to be to be happy and content then, but really, who can promise that?"
+demoTextData <- "we had no idea that you had no idae"
+demoInternalTrainingData <- "we had no idea that you had no idea that they had no idea"

### 2020-08-27 shinyapp/ui.R

@@ -28,19 +28,19 @@ navbarPage("Arianna",
       ),
       # Action button row
       fluidRow(
-        column(12, 
+        column(12,
                actionButton("getInternalResult", "Get Result", class = "btn-primary"),
         )
       ),
       # Result row
       shinyjs::hidden(
         fluidRow( id = "internal_results_row",
-          column(8, 
-                 h5("Unexpected Words"),
+          column(6,
+                 h5("Internal Consistency Score"),
                  tableOutput("unexpected_words"),
           ),
-          column(4, 
-                 h5("Internal Consistency Score"),
+          column(6,
+                 h5("Unexpected Words"),
                  tableOutput("internal_consistency_score"),
           )
         )
@@ -57,25 +57,25 @@ navbarPage("Arianna",
           textAreaInput("myTextData_e", "", value = "", height = "300px"),
           actionButton("clearText_e", "Clear", class = "btn-sm btn-outline-secondary"),
         )),
-        column(6, wellPanel(
+        column(6,
           imageOutput("OCR_image")
-        ))
+        )
       ),
       # Action button row
       fluidRow(
-        column(12, 
+        column(12,
                actionButton("getExternalResult", "Get Result", class = "btn-primary"),
         )
       ),
       # Result row
       shinyjs::hidden(
         fluidRow( id = "external_results_row",
-          column(8, 
-                 h5("Unexpected Words"),
+          column(6,
+                 h5("External Consistency Score"),
                  tableOutput("unexpected_words_e"),
           ),
-          column(4, 
-                 h5("External Consistency Score"),
+          column(6,
+                 h5("Unexpected Words"),
                  tableOutput("external_consistency_score"),
           )
         )

### 2020-08-27 shinyapp/server.R
added an about page to shiny app

@@ -1,9 +1,5 @@
 library(aRianna)
-#library(tau)
-#library(tm)
 source("training_data.R")
-#source("internal_consistency_func.R")
-#source("external_consistency_func.R")
 source("OCR_text_func.R")
 
 server <- function(input, output, session) {

### 2020-08-27 shinyapp/ui.R

@@ -2,12 +2,37 @@ library(shiny)
 library(shinythemes)
 library(shinyjs)
 library(shiny)
-library(tidyverse)
-library(tau)
-library(tm)
 
-navbarPage("Arianna",
+introduction <- c("Alphabets Really In Absolute Need of New Alphabet-ing or `aRianna` is a package that helps identify potential issues in text datasets by comparing the actual text with that expected by various natural language models. The package also provides a 'consistency score' that can be used to monitor how the text dataset changes during the cleaning and preparation process.")
+package_repo <- c("Github repository is available at: ")
+more_information <- c("More information is available at: ")
+installation_code <- c("devtools::install_github('RohanAlexander/arianna')")
+github_arianna <- c("https://github.com/RohanAlexander/aRianna")
+github_consistency <- c("https://github.com/RohanAlexander/consistency_scores_in_text_datasets")
+author1 <- c("Rohan Alexander: \t rohan.alexander@utoronto.ca")
+author2 <- c("Ke-Li Chiu: \t keli.chiu@mail.utoronto.ca")
+
+navbarPage("aRianna",
    theme = shinytheme("cosmo"),
+   tabPanel("About",
+      fluidRow(
+        # Input Panel: Internal Training Data
+        column(12, wellPanel(
+          h1("Introduction to aRianna"),
+          p(introduction),
+          p(package_repo),
+          a(github_arianna),
+          p(more_information),
+          a(github_consistency),
+          h3("Installation"),
+          code(installation_code),
+          h3("Authors"),
+          p(author1),
+          p(author2),
+
+        ))
+      ),
+    ),
    tabPanel("Internal Consistency",
       useShinyjs(),
       fluidRow(

### 2020-08-28 outputs/documents/paper.Rmd
aRianna bug documentation

@@
 We consider various natural language processing methods.
-These include n-grams, Word2Cec, GloVe, GPT models and BERT.
+These include n-grams, Word2Vec, GloVe, GPT models and BERT.
 The n-gram approach involves identifying two, three, or more, words that are commonly found together.
@@
 For the purpose of demonstrating how BERT predicts the missing words, we will show the implementation process in Python using the `pytorch_pretrained_bert` library.
-The reason to use `pytorch_pretrained_bert`(**Keli add citation**) instead of `keras-bert`is that `pytorch_pretrained_bert` has relatively comprehensive functions that makes the prediction of missing words more straightforward.
+The reason to use `pytorch_pretrained_bert` [@Wolf2019HuggingFacesTS] instead of `keras-bert`is that `pytorch_pretrained_bert` has relatively comprehensive functions that makes the prediction of missing words more straightforward.
 

### 2020-08-28 scripts/01-bug_documentation_consistency_score.R

@@ -1,0 +1,71 @@
+#### Bug documentation: get_internal_consistency_score function in aRianna ####
+# get_internal_consistency_score in this script does not
+# return consistency score accurately. For example:
+
+## text_to_check <- c("apple is red")
+## internal_corpus <- c("apple is red, apple is green, apple is yellow,apple is red, apple is green, apple is yellow,")
+## generate_internal_consistency_score(text_to_check, internal_corpus)
+
+# The output will give an internal consistency socre of 0.33, instead of 1.
+# This is because although `apple_is + red` is in the dataset, `red` is not in `apple_is + green` or `apple_is + yellow`
+# The error is generated form the code chunk between line 52 to line 58
+
+generate_internal_consistency_score <- function(text_to_check, consistency_dataset) {
+  #### Apply to the dataset to create a consistency score ####
+  # Now that we have our collection of n-grams (this will be internal consistency
+  # because that collection was based on the data itself) we want to work out a measure
+  # of consistency.
+
+  # Create tokens with errors
+  tokens_from_example_with_errors <- quanteda::tokens(text_to_check, remove_punct = TRUE)
+  tokens_from_example_with_errors <- quanteda::tokens_tolower(tokens_from_example_with_errors)
+
+  # Create ngrams from the tokens with errors
+  toks_ngram_with_errors <- quanteda::tokens_ngrams(tokens_from_example_with_errors, n = 3)
+
+  all_tokens_with_errors <- tibble::tibble(tokens = toks_ngram_with_errors[[1]])
+
+  all_tokens_with_errors <-
+    all_tokens_with_errors %>%
+    dplyr::mutate(tokens = stringr::str_replace_all(tokens, "_", " "),
+                  first_words = stringr::word(tokens, start = 1, end = 2),
+                  last_word = stringr::word(tokens, -1),
+                  tokens = stringr::str_replace_all(tokens, " ", "_"),
+                  first_words = stringr::str_replace_all(first_words, " ", "_")
+    )
+
+  # Now we combine them so last_word will be what we have and last_word_expected will
+  # be what we expect.
+  all_tokens_with_errors <-
+    all_tokens_with_errors %>%
+    dplyr::left_join(dplyr::select(consistency_dataset, -tokens), by = c("first_words"))
+
+  all_tokens_with_errors_only <- all_tokens_with_errors
+
[…]

### 2020-08-31 outputs/documents/paper_ngrams.Rmd
paper focusing on ngrams

--- 2020-08-21 outputs/documents/paper.Rmd
@@
   affiliation: University of Toronto
-abstract: "In this paper we introduce a process to clean the text extracted from PDFs using various methods from natural language processing.
+abstract: "In this paper we introduce a process to clean the text extracted from PDFs using n-gram models that is widely used in statistical natural language processing.
 Our approach compares originally extracted text with the text generated from, or expected by, these methods using earlier text as stimulus.
@@
 This is used to monitor changes during the cleaning process, and to compare the messiness of different texts.
-The methods that we consider are: n-grams, Word2Vec, GloVe, and the GPT models.
-We illustrate our process on text from the Canadian Hansard and introduce both a Shiny application and an R package to make our process easier for others to adopt."
+We illustrate our process on text from the book Jane Eyre and introduce both a Shiny application and an R package to make our process easier for others to adopt."
 keywords: "text-as-data; natural language processing; quantitative analysis."
@@
 
-We consider various natural language processing methods.
-These include n-grams, Word2Cec, GloVe, GPT models and BERT.
+We employ n-gram models to calculate consistency scores and generate word candidates for text corrections.
 The n-gram approach involves identifying two, three, or more, words that are commonly found together.
-Here think of a two-gram involving the word 'good' 'good morning'.
-At scale these can identify missing or unusual words, and work quickly, but they lack nuance.
-For instance, an equally reasonable two-gram involving the word 'good' is 'good work'.
-For that reason, we consider pre-trained word embedding models.
-These include Word2Vec and GloVe, which place each word in a multi-dimensional space such that distance between words can illustrate their relationship.
-We also consider pre-trained generative models such as GPT-2 and GPT-3 and BERT.
-GPT-2, GPT-3 and BERT are pre-trained generative unsupervised language models.
-GPT-2 and GPT-3 are the two generations of the same model that differ in the number of parameters.
-GPT-2 has 1.5 billion parameters and GPT-3 has more than 175 billion parameters.
-While GPT-2 is more limited, it can be run on smaller machines, whereas GPT-3 cannot and is accessed via an API.
-BERT has 340 millions parameters which is smaller than both GPT-2 and GPT-3.
-BERT also differs from GPT models because its encoding process is bidirectional instead of unidirectional.
+**(Keli adds descriptions on how n-gram can be used to calculate consistency scores and generate word candidates)**
 
-We apply our approach to the Canadian Hansard.
-The Canadian Hansard was digitised by @beelen2017digitization.
-This was a process by which PDFs were scanned, put through optical character recognition (OCR), and then corrected to a reasonable extent.
-The dataset is extensive, fit-for-purpose, and, appropriately, it has been used considerably, for instance @rheault2020word.
-However, there are many issues with the dataset in terms of reading it.
-We focus on one particular year - **(Rohan pick one)** - and show that our approach can be used to relatively quickly improve the quality of the dataset in a reproducible and consistent manner.
-
-The remainder of our paper is structured as follows: **(Rohan add structure)**.
-Additionally, we construct a Shiny app available at: **(Rohan add link)**.
-That app computes internal and external consistency scores for corpus excerpts, and have developed an R Package that allows our approach to be used on larger datasets, which is available at: **(Rohan add link)**.
+The remainder of our paper is structured as follows: **(Keli add structure)**.
+Additionally, we construct a Shiny app available at: https://kelichiu.shinyapps.io/aRianna/.
+That app computes internal and external consistency scores for corpus excerpts, and have developed an R Package, `aRianna`, that allows our approach to be used on larger datasets, which is available at: https://github.com/RohanAlexander/aRianna.
 
@@
 
-## PDF text extraction and n-grams
-
-The Portable Document Format (PDF) is typically used to render documents in a relatively fixed way.
-PDFs contain a lot of information, but often that information is not able to be analysed directly by statistical language models.
-The information first needs to be extracted from the PDF, often by optical character recognition (OCR), and then cleaned, prepared, and made into some type of dataset.
-One issue in this process is that OCR is not perfect.
-Hence there is a need to correct the output before it is analysed.
-One way to do this is to correct errors one-by-one and independent of context.
-We all know that 'tbe' is wrong and should be 'the'.
-But statistical language models have the advantage of correction that can occur within context (@rosenfeld2000two).
-This allows significant improvements, for instance sometimes 'bat' should be 'hat' and sometimes it really should be 'bat'.
+## N-gram Models
 
@@
 
-However, n-gram models do not take the linguistic structure of language into account.
-For instance, @rosenfeld2000two [p.
-1] discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'.
-The prediction of the next word is based on only a few preceding words and broader context is not taken into account.
-Hence next-word prediction using n-gram based language models can be limited.
+### Package Dependencies
 
 There are a variety of ways to implement an n-gram model within R @citeR including using R packages such as `Quanteda` [@citequanteda], `tidyText` [@citetidytext] and `tm` [@citetm].
-We used the Text Analysis Utilities, `tau`, R package @citetau because of we can separate text-cleaning and n-gram making in two different steps.
-We want to have control to manually clean the text because we want to add an end-of-sentence token to each sentence int he corpus.
-To illustrate the use of `tau`, we will first install and load the library, and create a small corpus.
+In our package, we used the `Quanteda` R package because it has a comprehensive set of functions for conducting text analysis.
+We also employed `dplyr` (**Keli adds citation**) for data frame manipulation and `tibble` (**Keli adds citation**) for transforming data frames to the tibble format:
 
 ```{r, eval = FALSE}
-install.packages("tau")
+install.packages("quanteda")
+install.packages("dplyr")
+install.packages("tibble")
 ```
 
+### make_internal_consistency_dataset Function
+
+To obtain the internal consistency score, we will construct a internal consistency dataset from a body of text that is the source of the text being evaluated.
+The function `make_internal_consistency_dataset` is created for this purpose.
+The function turns the body_of_text into an internal consistency dataset which the text being evaluated is compared with in the following steps.
+To do so, the function first transforms the body of text into word tokens with the punctuation removed and the letters turned to lowercase:
+
 ```{r, eval=FALSE}
-library(tau)
-corpus <- "There was no possibility of taking a walk that day. We had been 
+tokens_from_example <- quanteda::tokens(body_of_text, remove_punct = TRUE, )
+tokens_from_example <- quanteda::tokens_tolower(tokens_from_example)
+```
+
+Next, the `make_internal_consistency_dataset` function turns the individual tokens into tri-grams, and keeps only the tri-grams that appears in the data for more than once.
+The common tri-grams are preserved in a tibble:
+
+```{r, eval=FALSE}
+# Create ngrams from the tokens
+toks_ngram <- quanteda::tokens_ngrams(tokens_from_example, n = 3)
+
+# Convert to tibble so we can use our familiar verbs
+all_tokens <- tibble::tibble(tokens = toks_ngram[[1]])
+
+# We only want the common ones, not every one.
+all_tokens <-
+  all_tokens %>%
+  dplyr::group_by(tokens) %>%
+  dplyr::count() %>%
+  dplyr::filter(n > 1) %>%
+  dplyr::ungroup()
+```
+
+Finally, the `make_internal_consistency_dataset` function splits the tri-grams are to two parts: the previous two words and the last word:
+
+```{r, eval=FALSE}
+# Create a tibble that has the first two words in one column then the third
+  all_tokens <-
+    all_tokens %>%
+    dplyr::mutate(tokens = stringr::str_replace_all(tokens, "_", " "),
+                  first_words = stringr::word(tokens, start = 1, end = 2),
+                  last_word = stringr::word(tokens, -1),
+                  tokens = stringr::str_replace_all(tokens, " ", "_"),
+                  first_words = stringr::str_replace_all(first_words, " ", "_")
+                  ) %>%
+    dplyr::rename(last_word_expected = last_word) %>%
+    dplyr::select(-n)
+```
+
+### generate_internal_consistency_score Function
+
+After we have the internal consistency dataset, we need a function to compare the text being evaluated and the internal dataset in order to retrieve the internal consistency score — the `generate_internal_consistency_score` function.
+Similar to the `make_internal_consistency_dataset` function, the function first transforms the text being evaluated into tokens, then split tri-grams and preserve them into a tibble:
+
+```{r, eval=FALSE}
+# Create tokens with errors
+tokens_from_example_with_errors <- quanteda::tokens(text_to_check, remove_punct = TRUE)
+tokens_from_example_with_errors <- quanteda::tokens_tolower(tokens_from_example_with_errors)
+
+# Create ngrams from the tokens with errors
+toks_ngram_with_errors <- quanteda::tokens_ngrams(tokens_from_example_with_errors, n = 3)
+all_tokens_with_errors <- tibble::tibble(tokens = toks_ngram_with_errors[[1]])
+
+all_tokens_with_errors <-
+  all_tokens_with_errors %>%
+  dplyr::mutate(tokens = stringr::str_replace_all(tokens, "_", " "),
+                first_words = stringr::word(tokens, start = 1, end = 2),
+                last_word = stringr::word(tokens, -1),
+                tokens = stringr::str_replace_all(tokens, " ", "_"),
+                first_words = stringr::str_replace_all(first_words, " ", "_")
+)
+```
+
+Next, the function combines the text tokens tibble and the internal consistency tibble by the left_join function of `dplyr`.
+The left_join function returns all the rows from the text tokens tibble and all the columns from both tibbles.
+By combining the two tibbles, the function generates two addtional columns — the `last_word` column are the last words of all tri-grams, and the `last_word_expected` contains the words that are in the internal consistency dataset.
+By comparing the words in `last_word` and `last_word_expected`.
+If a last word is in `last_word` but not `last_word_expected`, it is regarded as an unexpected word and the corresponding word in `last_word_expected` serves as a replacement candidate:
+
+```{r, eval=FALSE}
+all_tokens_with_errors <-
+    all_tokens_with_errors %>%
+    dplyr::left_join(dplyr::select(consistency_dataset, -tokens), by = c("first_words"))
+```
+
+The consistency score is calculated by the number of words that can be predicted by the model divided by the total number of words in the data.
+The function identifies the counts of words that are predicted by the model and divided by the total number of words in the internal consistency dataset:
+
+```{r, eval=FALSE}
+# Calculate the internal consistency score:
+internal_consistency <-
+  internal_consistency %>%
+  dplyr::ungroup()%>%
+  dplyr::filter(!is.na(as_expected)) %>%
+  dplyr::count(as_expected) %>%
+  dplyr::mutate(consistency = true_count/sum(n))
+```
+
+The `generate_internal_consistency_score` function also lists the identified text errors and generates replacement candidate to the text errors:
+
+```{r, eval=FALSE}
+# Identify which words were unexpected
+unexpected <-
+  all_tokens_with_errors_only %>%
+  dplyr::mutate(as_expected = last_word == last_word_expected) %>%
+  dplyr::filter(as_expected == FALSE) %>%
+  dplyr::select(-tokens)
+```
+
+### generate_internal_consistency_score Function
+
+The `generate_external_consistency_score` function works identically with the `generate_internal_consistency_score`.
+The only difference is that it compares the text being evaluated with a external consistency dataset that is larger and more general.
+The external consistency dataset is from **(Keli adds source of the external dataset)**.
+
+### Demonstration
+
+For demonstration, we use the first paragraph of Jane Eyre as the internal text data.
+A sentence within the paragraph is modified with the intention to contain an error and serves as the text to be evaluated:
+
+```{r, eval=FALSE}
+body_of_text <- "There was no possibility of taking a walk that day. We had been 
 wandering, indeed, in the leafless shrubbery an hour in the morning; but 
@@
 that further out-door exercise was now out of the question."
+
+text_to_evaluate <- "when there was na company"
 ```
 
-Our purpose is to construct tri-grams and so we first split the corpus into sentences by period, exponential mark, or question marks.
-
-First, we remove all punctuation and other non-letter characters, and convert all text to lower case.
+The first step is to turn the body of text into the internal consistency dataset.
+The generated internal consistency dataset is a tibble that contains the tri-grams that appear in the internal text data more than once.
+Since only `there_was_no` has more than one occurrence, it is the only reference tri-gram in the internal consistency dataset:
 
 ```{r, eval=FALSE}
-tokenizer <- function(corpus) {
-  lines <- vector()
-  # Read the corpus line by line
-  for (line in corpus) {
-    # add two start-of-sentence tokens to the beginning of a sentence
-    lines <-c(lines, line)
-  }
-  lines <- tolower(lines)
-  # add end-of-sentence token to the end of a sentence
-  lines <- gsub("[.!?]$|[.!?] |$", " </s>", lines) 
-  tokens <- unlist(strsplit(lines, "[^a-z<>/]"))
-  tokens <- tokens[tokens != ""]
-  return(tokens)
-}
+internal_consistency_dataset <- aRianna::make_internal_consistency_dataset(body_of_text)
+internal_consistency_dataset
+## # A tibble: 1 x 3
+##   tokens       first_words last_word_expected
+##   <chr>        <chr>       <chr>             
+## 1 there_was_no there_was   no                
 ```
 
-We will use the `textcnt()` function to turn the training corpus into tri-grams.
-The most important argument of this function is _corpus_, as well as _n_ to set the number of words in a sequence, _method_ to determine which type of counts to compute, and _split_ to designate the regular expression pattern for splitting.
-The function outputs a `textcnt` object, which is essentially a table with each tri-gram and their counts [@citetau].
+The next step is to compare the text to be evaluated with the internal consistency dataset that we created in the previous step.
+The function `generate_internal_consistency_score` takes in two arguments — the text to evaluate and the internal consistency dataset.
+The function identifies the word "na" as an unexpected word, and generates the internal consistency score as zero.
+The function also lists "no" as the replacement of "na":
 
 ```{r, eval=FALSE}
-trigrams <- textcnt( tokens, n=3, method = "string", 
-                     split = "[[:space:][:digit:]]+", decreasing = TRUE)
+aRianna::generate_internal_consistency_score(text_to_evaluate, internal_consistency_dataset)
+
+## $`internal consistency`
+## # A tibble: 1 x 3
+##   as_expected     n consistency
+##   <lgl>       <int>       <dbl>
+## 1 FALSE           1           0
+
+## $`unexpected words`
+## # A tibble: 1 x 4
+##   first_words last_word last_word_expected as_expected
+##   <chr>       <chr>     <chr>              <lgl>      
+## 1 there_was   na        no                 FALSE
+
 ```
 
-We then manually added an end of sentence token, `</s>`, to the end of each sentence.
+To get the external consistency score, we will employ the `generate_external_consistency_score` function.
+The function `generate_internal_consistency_score` takes in only one argument — the text to evaluate, and compares it with the external consistency dataset.
+The function identifies the word "na" as an unexpected word, and generates the external consistency score as zero.
+Because the external dataset is larger, there are several words generated as the replacement candidates of "na":
 
 ```{r, eval=FALSE}
-trigrams <- trigrams[!grepl("</s>", names(trigrams))]
-```
+aRianna::generate_external_consistency_score(text_to_evaluate)
 
-At this point it is possible to take other measures around fine-tuning the tri-grams.
-For instance we may like to remove the tri-grams that contain an end-of-sentence token.
-This would mean that tri-grams that include a word from a new sentence will be removed.
-For example, the tri-gram 'day </s> we' will be removed because it has words from two different sentences.
-Table 1.
-shows four sample tri-grams from the small sample corpus after this process.
+## aRianna::generate_external_consistency_score(text_to_evaluate)
+## $`external consistency`
+## # A tibble: 1 x 3
+##   as_expected     n consistency
+##   <lgl>       <int>       <dbl>
+## 1 FALSE           2           0
 
-```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
-tabl <- "
-| there was no  | a rain so | a walk that | an hour in  |
-|---------------|-----------|-------------|-------------|
-| 2             | 1         | 1           | 1           |
+## $`unexpected words`
+## # A tibble: 13 x 4
+##    first_words last_word last_word_expected as_expected
+##    <chr>       <chr>     <chr>              <lgl>      
+##  1 there_was   na        a                  FALSE      
+##  2 there_was   na        an                 FALSE      
+##  3 there_was   na        no                 FALSE      
+##  4 there_was   na        not                FALSE      
+##  5 there_was   na        nothing            FALSE      
+##  6 there_was   na        of                 FALSE      
+##  7 there_was   na        one                FALSE      
+##  8 there_was   na        only               FALSE      
+##  9 there_was   na        plenty             FALSE      
+## 10 there_was   na        someone            FALSE      
+## 11 there_was   na        something          FALSE
 
-Table 1. Four sample tri-grams generated from the corpus
-"
-cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
-```
-
-The tri-grams can be accessed by calling the names of this table object and the counts can be used to calculate probabilities.
-There are a variety of options around constructing these probabilities.
-One is Kneser–Ney smoothing, which is the smoothing technique recommended by @chen1999empirical to improve the accuracy of n-gram models.
-
-Finally, it is helpful to create a data table that contains the first two words and the last of each tri-gram in separate column and the smoothed probability in another column.
-For instance, Table 2.
-provides and example of this set-up.
-
-```{r table2, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
-tabl <- "
-| First Words   | Last Word | Probability |
-|---------------|-----------|-------------|
-| there was     | no        | 0.6278947   |
-
-Table2. The smoothed probability table
-"
-cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
-```
-
-## Word embedding models
-
-Although, from an implementation perspective, there is an inherent numerical representation of language in n-gram models, it is not overly tractable.
-The numeric representation typically relies on each term being represented in a vector, which can quickly result in quite high-dimensional objects.
-In contrast, word embedding models represent terms by vectors of real numbers where each vector is a 'feature', hence the final object has considerably lower dimensionality [@bengio2003neural].
-
-This approach has a long history, but the implementation of @bengio2003neural has become the foundation for much subsequent work.
-Here, terms are encoded into feature vectors that represent different aspects of the term, and so each term is represented by a position in the vector space.
-Terms that are used in similar ways will have similar representations in the vector space.
-This will be due to their co-occurrence in usage.
-For example, 'cat' will be closer to 'dog' than to 'car', because we are more likely to talk about cats and dogs, than cats and cars.
-See how they might be close when considered along a 'pets' dimension, but far when considered along a 'related-to-tigers' dimension.
-
-Terms that are often used in similar contexts will have similar vector representations.
-Hence, word embedding models can identify similar terms, calculate the similarity between a pair of terms, and find the odd term in a group of terms [@mikolov2013efficient].
-There are a variety of ways to achieve this dimensionality reduction, including Word2Vec and GloVe.
-
-### Word2vec
-
-Word2vec is a word embedding model that uses a neural network to learn word associations [@mikolov2013efficient] [@mikolov2013distributed].
-It takes a corpus as an input and outputs a set of feature vectors that allow the terms in that corpus to be represented.
-@mikolov2013efficient proposed two architectures for determining this distributed representation of terms: continuous bag-of-words (CBOW), and continuous skip-gram.
-Both have an input layer, a projection layer, and an output layer.
-But CBOW uses several surrounding words, or the context, to predict a target word.
-On the other hand, continuous skip-gram takes a single word as input to predict its neighbouring words.
-
-To illustrate the difference, consider the quote from Descartes, 'I think therefore I am'.
-The focus word for which we want to learn the embedding is 'therefore', and the context window is set to two.
-This means that we consider the two words immediately preceding and proceeding 'therefore'.
-The CBOW input is 'I, think, I, am', and the output is 'therefore'.
-In contrast, the continuous skip-gram input is 'therefore', while 'I, think, I, am' is the output.
-The input layer encodes terms within a context window to one-hot vectors.
-Using a projection matrix, the projection layer then project the discrete one-hot vectors to lower-dimension continuous vectors.
-Finally, the output layer is a softmax regression classifier that obtains the posterior distribution of words [@rong2014word2vec].
-It is the input and output vectors that differ between the two architectures.
-
-Word2Vec is trained on the Google News dataset [@citegooglenews].
-This contains six billion tokens and the vocabulary size is restricted to the one million most common words [@mikolov2013efficient].
-As this is a large dataset, it provides an accurate statistical pattern of language, as it is used in this dataset, and a comprehensive mapping between terms and vector positions.
-
-To the extent that the relative positions of terms reflects their semantic relationship, Word2Vec will perform well for word analogies.
-However, terms that have context-specific meaning may not perform as well.
-
-There are a few packages for implementing a word2vec model in R, including `rword2vec` [@citerwordvecchaware] and `word2vec` [@citewordvecwijffels].
-Both packages use the same algorithm.
-Another way to implement a word2vec model in R is through Keras which is an API for deep learning written in Python [@chollet2015keras] for which the `keras` package provides an R interface [@citekeras].
-
-Here, we follow @falbel2017word to implement a skip-gram model of Word2Vec through Keras.
-There are three steps in the Keras deep learning framework: 1) prepare the data; 2) define the model; and 3) train and evaluate the model.
-**(Ke-Li: Can we change this to be a different dataset please?)** Following these steps, we first downloaded the Google News dataset [@citegooglenews] and preprocess the text into tokens.
-
-```{r, eval=FALSE}
-tokenizer %>% fit_text_tokenizer(training_corpus)
-```
-
-We then define a model that mimics the skip-gram algorithm of word2vec using the function `skipgrams_generator()`.
-
-```{r, eval=FALSE}
-skipgrams_generator <- function(text, tokenizer, window_size, negative_samples) {
-  gen <- texts_to_sequences_generator(tokenizer, sample(text))
-  function() {
-    skip <- generator_next(gen) %>%
-      skipgrams(
-        vocabulary_size = tokenizer$num_words, 
-        window_size = window_size, 
-        negative_samples = 1
-      )
-    x <- transpose(skip$couples) %>% map(. %>% unlist %>% as.matrix(ncol = 1))
-    y <- skip$labels %>% as.matrix(ncol = 1)
-    list(x, y)
-  }
-}
-```
-
-The embedding dimensions, the size of the context window and the number of samples are configurable within this function.
-
-Finally, we use the `fit_generator()` function to train the model.
-
-```{r, eval=FALSE}
-model %>%
-  fit_generator(
-    skipgrams_generator(training_corpus, tokenizer, skip_window, negative_samples), 
-    steps_per_epoch = 10, epochs = 1
-  )
-```
-
-**What's supposed to happen:** As a result, a embedding matrix is obtained and serves as the look-up table of the word vectors.
-**What really happened:** However, the training process was stuck in the operation of `fit_generator()` function.
-According to Stack Flow this might be an issue of incompatible versions between Keras and Tensorflow.
-More investigation is needed.
-
-### GloVe
-
-The Global Vectors (GloVe) model is a pre-trained word embedding model that emphasises the semantic meaning of words [@pennington2014glove].
-In the Word2Vec model, word representations are learnt locally within the context window.
-On the other hand, GloVe considers the properties of the corpus globally.
-As the model considers the context of an entire corpus the co-occurrence probabilities are based on counts for the entire training dataset.
-
-GloVe additionally considers the relative probability of word co-occurrence.
-For instance, consider the two words 'ice' and 'steam'.
-GloVe considers their co-occurrence with a variety of probe words, $k$.
-For instance $k$ could be 'solid', which is a property of ice; 'gas', which is a property of steam; and 'water', which is related to both.The ratio of co-occurrence probability is:
-
-$$\frac{\mbox{Pr}(k|\mbox{ice})}{\mbox{Pr}(k|\mbox{steam})}.$$
-
-The ratio between the co-occurrence probability of the 'ice, solid' pair and the 'steam, solid' pair is high because the co-occurrence of 'ice' and 'solid' is higher than the co-occurrence of 'steam' and 'solid'.
-Conversely, the ratio between the co-occurrence of the 'ice, gas' pair and the 'steam, gas' pair is low.
-On the other hand, the ratio of the co-occurrence probability of the 'ice, water' pair and the 'steam, water' pair us close to one because water is relevant to both 'ice' and 'steam'.
-
-As @hanretty2018comparing point out, although word embedding models can capture semantic meanings of words, "they are context-free and fail to capture higher-level concepts in context, such as polysemous disambiguation, syntactic structures, semantic roles, anaphora".
-Word embedding models do not represent a homonyms that has difference meanings in different vectors depending on the context.
-For example, "squash" can be a type of hard shelled vegetable or can be referring to a racket and ball sport, but there is only one series of vectors to represent the word "squash", instead of dynamically changing the vectors depending on the context.
-Moreover, from our observation, the word embedding models are unable to handle novel words that are not included in the vocabulary of the training corpus; for this reason, we conclude that the generalizability of word embedding models is sub-optimal.
-
-There are a variety of ways to implement GloVe word embedding models in R.
-One is to load GloVe embeddings into a frozen Keras Embedding layer [@kerasglove] using the `Keras` R package [@kerasR2020].
-This means the embedding vectors will not be updated during the training process.
-Another way, is to use the `text2vec` package [@text2vec].
-
-Here we follow **CITE** to implement GloVe using the `keras` R package [@kerasglove].
-The blog post is aimed to create a word embedding model for text classification.
-However, we will only use the guide create a embedding index and skip the steps of text classification.
-Instead of building a model for word embedding from scratch as we did for the skip-gram model of Word2Vec, here we are merely loading the Glove model and its pre-trained word vectors as an index.
-The word embedding index is then saved to an R environment.
-
-**[As before, we need code and examples here]**
-
-## Pre-trained transformer language models
-
-In recent years, the use of pre-trained language models has played a significant role in advancing natural language processing (NLP) tasks such as reading comprehension, text generation, and even creative writing [@brown2020language].
-Pre-trained models remove the need for researchers to prepare training datasets or allot computational resources to the stage of pre-training.
-Because of this, there has been substantial development in NLP research.
-In this section we discuss the GPT models from OpenAI and the BERT model from Google.
-
-### GPT
-
-OpenAI is an AI research company who openly pursues Artificial General Intelligence — to have machines learn and understand any intellectual tasks as human do [@brundage2018malicious].
-OpenAI's Generative Pre-Trained Transformer (GPT) models are pre-trained language models that are built based on the transformer architecture.
-Proposed by @vaswani2017attention from Google in 2017, the Transformer is a novel network architecture for neural network that outperforms RNN-based and CNN-based models in computational efficiency [@vaswani2017attention].
-Since the emergence of the Transformer model, most of the representative pre-trained models are built based on this architecture [@hanretty2018comparing].
-Therefore, we can say that Transformer marks a new era of language models.
-
-As of August 2020, OpenAI has three generations of GPT models.
-GPT was introduced to the public in June, 2018.
-In the year 2019, its successor GPT-2 was released in three stages.
-Not long after, in June 2020, OpenAI announced GPT-3, which is the latest GPT model as of July 2020 [@brown2020language].
-Since the API of GPT-3 at the moment is only available to a few invited users, public applications of GPT-3 are limited.
-On the other hand, its predecessor, GPT-2, has enabled researchers made breakthroughs in the past years.
-@alt2019fine extended GPT-2 to relation extraction, which is a task of identifying the relationship between concepts appeared in text.
-They had shown a milestone in predicting larger range of distinct relation types with higher confidence [@alt2019fine].
-In the field of speech recognition, data scarcity is often an issue due to the difficulty to collect large amount of data from human for learning [@bird2020lstm].
-@bird2020lstm employed GPT-2 as part of the solution to generate augmented data for classification.
-GPT-2 also helped NLP researchers tackling the task of textual paraphrasing.
-@hegde2020unsupervised used GPT-2 to build an unsupervised paraphrasing model that is not restricted by domain shift within an article, which is a problem faced by the usual supervised models of paraphrasing.
-
-Most of prior language models are trained in a single domain by supervised learning to perform a single task.
-OpenAI, on the other hands, train the GPT models to perform many tasks [@radford2019language].
-GPT-2 is pre-trained on WebText, a comprehensive dataset that is scrapped from millions of webpages and contains total of 40 GB of text [@radford2019language].
-WebText contains all kinds of textual data: dialogues, comments, translations, questions and answers...etc.
-Such diversity is critical for GPT-2 to be a generalized model that can perform multiple tasks with any domain boundary.
-However, GPT-2 has the capacity to adapt to a context and generates custom output with the "fine-tuning" technique [@ziegler2019fine].
-While the pre-training stage is unsupervised and mandatory, fine-tuning is an optional training stage where GPT can learn to perform new tasks in a supervised or reinforced approach such as stylistic continuation and summarization [@ziegler2019fine].
-During the pre-training stage, the model is trained to perform text completion task.
-Give it a prompt to start and it will generate texts that continues what has been started.
-In the fine-tuning stage, @radford2019language provided additional training datasets for GPT-2 to generate the texts with specific sentiment, or be descriptive, or even summarize the prompt text.
-
-OpenAI has made GPT-2 open-sourced through Github with Python and Tensorflow [@tensorflow2015-whitepaper].
-There is also a package in R that wraps the Python code into a R package, created by @keydanaluraschi2019gpt2.
-There are three choices of models in terms of parameter sizes — "124M", "355M" and "774M".
-The R package only permits exploratory experiments with the pre-trained tasks; fine-tuning for new tasks is not available thorough R.
-To proceed with the implementation of GPT-2 in R, theoretically, we will fine-tune the model through Python and create a wrapper code for R.
-This could possibility be achieved by employing the `reticulate` R package [@reticulateR] that allows operation of Python code in R.
-The practicality of this approach is to be further investigated.
-
-GPT-3 has over 175 billions of parameters [@brown2020language].
-In the past few years, the growth of parameters of transformer language models expands rapidly.
-The first generation GPT has 110 million parameters.
-Released in 2018, GPT-2 has 1.6 billion parameters, and within two years, GPT-3 has grown the size to 175 billion, which is over 100 times than its predecessor.
-The more parameters there are the better the model can generalize to new tasks; therefore, the increasing amount of parameters accelerates the improvements in text synthesis and downstream NLP tasks [@brown2020language].
-GPT-3 can learn NLP tasks with just a handful of example; OpenAI refers this ability as few-shot learning, or in-context learning [@brown2020language].
-As a result, we only need to provide GPT-3 a few demonstrations for it to learn a new NLP task.
-For example, OpenAI team trained GPT-3 to generate "news articles".
-The prompt input is a plausible first sentence of a news story written by human, and the output is an entire news article generated by the model itself.
-Without the demonstrations, GPT-3 tends to treat the prompts as "tweets" and generate texts that serve as responses or follow-up tweets.
-The team fed a few previous news articles in the model’s context as demonstrations, and the model learned to produce the outputs that adapt to the style and length of news stories [@brown2020language].
-This few-shot learning ability indicates that the model can learn like human do and would democratize the use of language models to broader fields.
-
-### BERT
-
-BERT stands for Bidirectional Encoder Representations from Transformers [@devlin2018bert].
-It was released by researchers at Google AI Language in 2018.
-The key innovation of BERT is its application of bidirectional training of the Transformer model, which results in a deeper sense of language context and flow.
-What is bidirectional training?
-In the context of English language, a sentence is constructed from left to right.
-The context on the left is used to determine the next word on the right.
-In this sense, GPT models are unidirectional since the probability of the next word is dependent only on the past.
-It is also possible to reverse the direction and model the sentence from right to left, and BERT incorporates both directions of left-to-right and right-to-left to model bidirectional contexts.
-
-BERT is pre-trained by using two unsupervised tasks.
-The first task is Masked Language Model and and second task is Next Sentence Prediction.
-The bidirectional approach of BERT is alleviated by the "Masked Language Model" [@devlin2018bert].
-The Masked Language Model masks random words in the input corpus which forces the model to predict the original words by "filling in the blank".
-This training process blends the representation from the left and right context and therefore increases the model's context sensitivity [@nozza2020mask].
-In the Next Sentence Prediction process, the model receives two sentences in sequence as input.
-50% of the sentences are paired in the right sequence and the ones from the other half have sentences that are paired randomly [@devlin2018bert].
-The goal of this process is to have BERT evaluate if the second sentence is the next sentence of the first sentence in the original text and sharpen its capacity to predict the likelihood of the next sentence.
-
-BERT was pre-trained on BooksCorpus dataset with 800 million words and English Wikipedia with 2,500 million words [@devlin2018bert].
-Because BERT employs the Transformer architecture with the self-attention mechanism, fine-tuning is also available with one additional output layer and a task-specific input [@devlin2018bert].
-Google AI researchers have fine-tuned BERT for 11 NLP tasks including question-answering and common-sense inference.
-
-BERT implementation in R is also possible through Keras.
-@kerasbert created a python module called `keras-bert` and the module can be called by R through `reticulate` package [@reticulateR], which enables the operation of Python modules in R.
-To start, we install the keras-bert model in Python on the machine and load BERT model into R through `reticulate`.
-Once `keras-bert` is loaded, we can use it to fine-tune BERT on a task-specific training dataset.
-Fine-tuning is started by tokenizing the training data.
-The output of the tokenization is then fed to BERT as the input.
-The next step is to define the model by configuring its parameters such as batch size, number of epochs and the learning rate.
-Finally, the model is compiled and we can begin the fine-tuning.
-
-For the task of text cleaning, fine-tuning is not required.
-The process of text cleaning involves detecting and correcting mis-spelled words.
-Correcting the incorrect words can be seen as predicting the correct words.
-The Masked Language Model of BERT can be employed to do such prediction.
-In the pre-training process, the Masked Language Model would mask random words to do the prediction.
-However, we can tell the model what we want to "mask" by assigning the words to be encoded as `[MASK]`.
-Consider the erroneous sentence that has ten words:
-
-> "There was no possibility of taking a wlak that day."
-
-Using a spell-checker, "wlak" is identified as an error and is substituted by `[MASK]`:
-
-> "There was no possibility of taking a [MASK] that day."
-
-The implementation of text correction with BERT in R with the `keras-bert` Python module is not yet completed.
-For the purpose of demonstrating how BERT predicts the missing words, we will show the implementation process in Python using the `pytorch_pretrained_bert` library.
-The reason to use `pytorch_pretrained_bert`(**Keli add citation**) instead of `keras-bert`is that `pytorch_pretrained_bert` has relatively comprehensive functions that makes the prediction of missing words more straightforward.
-
-After we retrieved the new text data with the error replaced by `[MASK]`, we use the `tokenizer()` function in `pytorch_pretrained_bert` to tokenize the text:
-
-```{python, eval=FALSE}
-tokenized_text = tokenizer.tokenize("There was no possibility of taking a [MASK] that day.")
-tokenized_text
->>> ['there', 'was', 'no', 'possibility', 'of', 'taking', 'a', '[MASK]', 'that', 'day', '.']
-```
-
-We then get the index of `[MASK]` within the sentence and save it to a list. Since we only have one `[MASK]` token, the list has only one element:
-
-```{python, eval=FALSE}
-mask_index = 7
-```
-
-`pytorch_pretrained_bert` has a dictionary of tokens generated from the pre-training stage.
-The tokens include the vocabulary, puncuations and special symbols.
-Each token in the dictionary is given an index, represented as `{token: index}`.
-The function `convert_tokens_to_ids()` encodes the tokens we got previously to their positional indices in the tokens dictionary and collects them into a list:
-
-```{python, eval=FALSE}
-indexed_tokens = tokenizer.convert_tokens_to_ids(tokenized_text)
-indexed_tokens
->>> [2045, 2001, 2053, 6061, 1997, 2635, 1037, 103, 2008, 2154, 1012]
-```
-
-The next step is to encode the text by its segments.
-The model encodes a sentence of representing each word in the sentence with the positional index of the sentence in the provided corpus.
-For example, the example text has only one sentence "There was no possibility of taking a [MASK] that day." Therefore, the position of the sentence is "0", and the encoding for each word in the sentence, including the period is a series of "0" to the length of the sentence:
-
-```{python, eval=FALSE}
-segments_tokens = [0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
-```
-
-The previous two steps give us two set of encoding — `indexed_tokens` that contains the encoding for each token, and `segments_tokens` that contains the encoding for each sentence.
-These two encoding layers that we feed to BERT Masked Language Model.
-The model takes tensor objects as inputs [@paszke2017automatic], therefore we transformed the two lists into tensor objects before hands.
-These two tensor objects has two dimensions.
-The first dimension is the container, and the second dimension is the length of the token lists.
-
-```{python, eval=FALSE}
-tokens_tensor = torch.tensor([indexed_tokens])
-tokens_tensor
-
-segments_tensor = torch.tensor([segments_encodes])
-segments_tensor
->>> [[0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]]
-```
-
-The encoding layers can be represented as the following:
-
-```{r table3, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
-tabl <- "
-|        |'there' |'was' |'no' |'possibility'|'of'|'taking'|'a' |'[MASK]'|'that'|'day'| '.'|
-|--------|--------|------|-----|-------------|----|--------|----|--------|------|-----|----|
-|token   |2045    |2001  |2053 |6061         |1997|2635    |1037|103     |2008  |2154 |1012|
-|segment |0       |0     |0    |0            | 0  |0       |0   |0       |0     |0    |0   |
-
-Table2. The smoothed probability table
-"
-cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
-```
-
-We then load the pre-trained BertForMaskedLM and feed the two tensor objects to it.
-The model then returns a tensor object that has three dimensions.
-The `predictions` tensor objects contains the prediction score of all the words in the vocabulary.
-
-```{python, eval=FALSE}
-model = BertForMaskedLM.from_pretrained('bert-base-uncased')
-predictions = model(tokens_tensor, segments_tensor)
-predictions
-```
-
-To get the prediction, we will get the element with the highest prediction score and retrieve its index.
-The retrieved index is then used to find the corresponding token in the tokens dictionary.
-The predicted word is "walk", which matches with the original sentence in **Jane Eyre**.
-
-```{python, eval=FALSE}
-predicted_word_index = torch.argmax(predictions[0, mask_index]).item()
-predicted_word_index
->>> 3328
-predicted_word_token = tokenizer.convert_ids_to_tokens([predicted_index])[0]
-predicted_word_token
->>> walk
 ```
@@
 # Stylized example
-
-**(TODO: This stylized example need to be changed and updated to actually use GPT-3 at some point.)**
-
-In our application, we deploy OpenAI GPT-3 to generate text and ...
 
@@
 
-# Data
-
-Various data can be used....
-
 # Application
 
+**(Keli adds application in OCR tasks)**
+
+[Internal consistency: ground truth of the text and the text that are aligned with ground truth]
+
 # Discussion
+
+**(Keli organizes these text to illustrate limitations and future works)**
 
@@
 
-# Appendix
+However, n-gram models do not take the linguistic structure of language into account.
+For instance, @rosenfeld2000two [p.
+1] discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'.
+The prediction of the next word is based on only a few preceding words and broader context is not taken into account.
+Hence next-word prediction using n-gram based language models can be limited.
 
-## Literature summaries
-
-(These are just here for now - we'll remove later.)
-
-### Article: Improving Language Understanding by Generative Pre-Training (2018)
-Introduction of a semi-supervised approach for language understanding machine learning tasks. The approach is the combinaiton of unsupervised pre-training and supervised fine-tuning.
-
-### Article: Language Models are Few-Shot Learners (2020)
-Official paper from OpenAI to introduce OpenAI GPT-3. Emphasizes on the "fewer-shots" and "task-agnostic" aspects of the latest langual model compared to its predecessor GPT-2.
-
-## Discussion of how GPT-3 works
-
-**(Ke-Li - it would be good for us to go through how it actually works in as much detail as we can manage (in our own words).
-This is mostly just to force us to learn how it works and may be taken out of the final version of the paper, but it's still an important exercise.)**
+Here think of a two-gram involving the word 'good' 'good morning'.
+At scale these can identify missing or unusual words, and work quickly, but they lack nuance.
+For instance, an equally reasonable two-gram involving the word 'good' is 'good work'.
+For that reason, we consider pre-trained word embedding models.
+These include Word2Vec and GloVe, which place each word in a multi-dimensional space such that distance between words can illustrate their relationship.
+We also consider pre-trained generative models such as GPT-2 and GPT-3 and BERT.
+GPT-2, GPT-3 and BERT are pre-trained generative unsupervised language models.
+GPT-2 and GPT-3 are the two generations of the same model that differ in the number of parameters.
+GPT-2 has 1.5 billion parameters and GPT-3 has more than 175 billion parameters.
+While GPT-2 is more limited, it can be run on smaller machines, whereas GPT-3 cannot and is accessed via an API.
+BERT has 340 millions parameters which is smaller than both GPT-2 and GPT-3.
+BERT also differs from GPT models because its encoding process is bidirectional instead of unidirectional.
 

### 2020-09-01 outputs/documents/paper_ngrams.Rmd
ngram paper draft ready for revie

@@
 biblio-style: apalike
-title: "Consistency scores in text data"
+title: "aRianna: R package for consistency scores in text data"
 thanks: "Thank you to **X, Y and Z** for helpful comments.
 A Shiny application for interactive and small-scale examples is available: https://kelichiu.shinyapps.io/Arianna/.
-An R package for larger scale applications is available: **HERE**.
+An R package for larger scale applications is available: https://github.com/RohanAlexander/arianna.
 Our code and datasets are available: https://github.com/RohanAlexander/consistency_scores_in_text_datasets.
@@
   affiliation: University of Toronto
-abstract: "In this paper we introduce a process to clean the text extracted from PDFs using n-gram models that is widely used in statistical natural language processing.
+abstract: "In this paper, we introduce a process to clean the text extracted from PDFs using n-gram models that is widely used in statistical natural language processing.
 Our approach compares originally extracted text with the text generated from, or expected by, these methods using earlier text as stimulus.
@@
 
-# Introduction
+# 1. Introduction
 
@@
 The n-gram approach involves identifying two, three, or more, words that are commonly found together.
-**(Keli adds descriptions on how n-gram can be used to calculate consistency scores and generate word candidates)**
+In our application, we use the sequence of three words, or tri-grams, to calculate consistency scores and generate word candidates for text corrections.
+By using tri-grams, we identify the first two words in the sequence and evaluate if the last word in the sequence is "as expected" by comparing it to the corresponding tri-gram in a comparison dataset.
+The comparison dataset can be internal, meaning that the dataset is the source where the text being evaluated is extracted from.
+On the other hand, a comparison dataset is external when the dataset is from a different source than where the text being evaluated is from, and is a larger and more universal dataset that holds generalizability.
+The external dataset we use in the application is constructed using a subset of text data sourced from Open WebText Corpus by @Gokaslan2019OpenWeb.
+Since the content of Open WebText Corpus is from the web, the variety of the text data is promising for meeting the generalizability and universality an external dataset should possess.
 
-The remainder of our paper is structured as follows: **(Keli add structure)**.
-Additionally, we construct a Shiny app available at: https://kelichiu.shinyapps.io/aRianna/.
+In this paper, we introduce the `aRianna` R Package with the following structure: In Section 2, we provide an overview of language model and n-grams.
+In section we deconstruct the package functions and provide demonstration example of the package usage.
+Finally, we discuss the limitation and the future directions of aRianna in Section 4.
+Beyond this paper, we construct a Shiny app available at: https://kelichiu.shinyapps.io/aRianna/.
 That app computes internal and external consistency scores for corpus excerpts, and have developed an R Package, `aRianna`, that allows our approach to be used on larger datasets, which is available at: https://github.com/RohanAlexander/aRianna.
 
-# Background
+# 2. Background
 
@@
 
-Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation [@brown1992class].
-In our application, we use a tri-gram based model to detect both real-word and non-word errors and correct them with the candidate word that has the highest probability.
+# 3. The aRianna R Package
 
-### Package Dependencies
+## Package Dependencies
 
@@
 In our package, we used the `Quanteda` R package because it has a comprehensive set of functions for conducting text analysis.
-We also employed `dplyr` (**Keli adds citation**) for data frame manipulation and `tibble` (**Keli adds citation**) for transforming data frames to the tibble format:
+We also employed `dplyr` [@2018dplyr] for data frame manipulation and `tibble` [@citeTibble] for transforming data frames to the tibble format:
 
@@
 
-### make_internal_consistency_dataset Function
+## Function to Make the Internal DataSet
 
@@
 The function `make_internal_consistency_dataset` is created for this purpose.
-The function turns the body_of_text into an internal consistency dataset which the text being evaluated is compared with in the following steps.
+The function turns the body_of_text into an internal consistency dataset which the text being evaluated is compared with.
 To do so, the function first transforms the body of text into word tokens with the punctuation removed and the letters turned to lowercase:
@@
 ```{r, eval=FALSE}
-tokens_from_example <- quanteda::tokens(body_of_text, remove_punct = TRUE, )
-tokens_from_example <- quanteda::tokens_tolower(tokens_from_example)
+tokens_from_example <- 
+  quanteda::tokens(body_of_text, remove_punct = TRUE, )
+tokens_from_example <- 
+  quanteda::tokens_tolower(tokens_from_example)
 ```
@@
 # Create ngrams from the tokens
-toks_ngram <- quanteda::tokens_ngrams(tokens_from_example, n = 3)
+toks_ngram <- 
+  quanteda::tokens_ngrams(tokens_from_example, n = 3)
 
 # Convert to tibble so we can use our familiar verbs
-all_tokens <- tibble::tibble(tokens = toks_ngram[[1]])
+all_tokens <- 
+  tibble::tibble(tokens = toks_ngram[[1]])
 
@@
 
-Finally, the `make_internal_consistency_dataset` function splits the tri-grams are to two parts: the previous two words and the last word:
+Finally, the `make_internal_consistency_dataset` function splits the tri-grams to two parts: the previous two words and the last word.
+In the end, a tibble is created with each tri-gram containing its original three-word sequence, the first two words, and the last word:
 
@@
 
-### generate_internal_consistency_score Function
+## Function to Generate Internal Consistency Score
 
 After we have the internal consistency dataset, we need a function to compare the text being evaluated and the internal dataset in order to retrieve the internal consistency score — the `generate_internal_consistency_score` function.
-Similar to the `make_internal_consistency_dataset` function, the function first transforms the text being evaluated into tokens, then split tri-grams and preserve them into a tibble:
+Similar to the `make_internal_consistency_dataset` function, the function first transforms the text being evaluated into tri-gram tokens, then splits the tri-grams and preserves them into a tibble:
 
@@
 # Create tokens with errors
-tokens_from_example_with_errors <- quanteda::tokens(text_to_check, remove_punct = TRUE)
-tokens_from_example_with_errors <- quanteda::tokens_tolower(tokens_from_example_with_errors)
+tokens_from_example_with_errors <- 
+  quanteda::tokens(text_to_check, remove_punct = TRUE)
+tokens_from_example_with_errors <- 
+  quanteda::tokens_tolower(tokens_from_example_with_errors)
 
 # Create ngrams from the tokens with errors
-toks_ngram_with_errors <- quanteda::tokens_ngrams(tokens_from_example_with_errors, n = 3)
-all_tokens_with_errors <- tibble::tibble(tokens = toks_ngram_with_errors[[1]])
+toks_ngram_with_errors <- 
+  quanteda::tokens_ngrams(tokens_from_example_with_errors, n = 3)
+all_tokens_with_errors <- 
+  tibble::tibble(tokens = toks_ngram_with_errors[[1]])
 
@@
 
-Next, the function combines the text tokens tibble and the internal consistency tibble by the left_join function of `dplyr`.
-The left_join function returns all the rows from the text tokens tibble and all the columns from both tibbles.
-By combining the two tibbles, the function generates two addtional columns — the `last_word` column are the last words of all tri-grams, and the `last_word_expected` contains the words that are in the internal consistency dataset.
-By comparing the words in `last_word` and `last_word_expected`.
+Next, the function combines the text tokens tibble and the internal consistency tibble by the `left_join()` function of `dplyr`.
+The `left_join()` function returns all the rows from the text tokens tibble and all the columns from both tibbles.
+By combining the two tibbles, the function generates two additional columns — the `last_word` column are the last words of all tri-grams, and the `last_word_expected` contains the words that are in the internal consistency dataset.
+By comparing the words in `last_word` and `last_word_expected`, we can identify the "unexpected" words.
 If a last word is in `last_word` but not `last_word_expected`, it is regarded as an unexpected word and the corresponding word in `last_word_expected` serves as a replacement candidate:
@@
   all_tokens_with_errors %>%
-    dplyr::left_join(dplyr::select(consistency_dataset, -tokens), by = c("first_words"))
+  dplyr::left_join(dplyr::select(consistency_dataset, -tokens), 
+                   by = c("first_words"))
 ```
@@
 
-### generate_internal_consistency_score Function
+## Function to Generate External Consistency Score
 
@@
 The only difference is that it compares the text being evaluated with a external consistency dataset that is larger and more general.
-The external consistency dataset is from **(Keli adds source of the external dataset)**.
+The external consistency dataset we employed in the application is constructed using a subset text data from Open WebText Corpus by @Gokaslan2019OpenWeb.
 
-### Demonstration
+## Demonstration
+
+To install the package and load the library:
+
+```{r, eval=FALSE}
+devtools::install_github("RohanAlexander/arianna")
+library(aRianna)
+```
 
@@
 ```{r, eval=FALSE}
-body_of_text <- "There was no possibility of taking a walk that day. We had been 
-wandering, indeed, in the leafless shrubbery an hour in the morning; but 
-since dinner (Mrs. Reed, when there was no company, dined early) the cold 
-winter wind had brought with it clouds so sombre, and a rain so penetrating, 
-that further out-door exercise was now out of the question."
+body_of_text <- "There was no possibility of taking a walk that day. 
+We had been wandering, indeed, in the leafless shrubbery an hour in 
+the morning; but since dinner (Mrs. Reed, when there was no company, 
+dined early) the cold winter wind had brought with it clouds so sombre, 
+and a rain so penetrating, that further out-door exercise was now out 
+of the question."
 
@@
 ```{r, eval=FALSE}
-internal_consistency_dataset <- aRianna::make_internal_consistency_dataset(body_of_text)
+internal_consistency_dataset <- 
+  aRianna::make_internal_consistency_dataset(body_of_text)
 internal_consistency_dataset
@@
 ```{r, eval=FALSE}
-aRianna::generate_internal_consistency_score(text_to_evaluate, internal_consistency_dataset)
+aRianna::generate_internal_consistency_score(
+  text_to_evaluate, internal_consistency_dataset)
 
@@
 ## $`unexpected words`
-## # A tibble: 13 x 4
+## # A tibble: 11 x 4
 ##    first_words last_word last_word_expected as_expected
@@
 
-# Stylized example
+# 4. Discussion
 
-As a stylized example, let's consider the following actual paragraph from Jane Eyre, by Charlotte Bronte:
-
-> There was no possibility of taking a walk that day.
-We had been wandering, indeed, in the leafless shrubbery an hour in the morning; but since dinner (Mrs.
-Reed, when there was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, that further out-door exercise was now out of the question.
-
-Let's pretend that this text had been created from optical character recognition and that it had the following errors: some 'h' were replaced with 'b'; and some 'd' have been replaced with 'al':
-
-> There was no possibility of taking a walk that day.
-We had been wandering, indeed, in tbe leafless shrubbery an hour in tbe morning; but since dinner (Mrs.
-Reeal, when tbere was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, tbat further out-door exercise was now out of the question.
-
-Assume a model that is trained to perfectly forecast the next word in Jane Eyre.
-For this fragment there are 62 words, comprising 5 errors and 57 correct words.
-So the internal consistency score of this fragment would be: 57/62 = 0.919.
-When we recognise and correct the errors, this consistency score would increase to 1.
-
-Similarly, assume a model that is trained on an external data source.
-This means that it will recognise the the cases where some 'h' were replaced with 'b', but not recognise that 'Reed' has become 'Reeal'.
-Hence, the external consistency score would be 58/62 = 0.935.
-
-# Application
-
-**(Keli adds application in OCR tasks)**
-
-[Internal consistency: ground truth of the text and the text that are aligned with ground truth]
-
-# Discussion
-
-**(Keli organizes these text to illustrate limitations and future works)**
-
-Internal validity vs external- one is their own words the other is a general set of words.
-
-In the same way that precision and recall provide important measures…
-
+Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation [@brown1992class].
+As demonstrated in our application, we use a tri-gram based model to calculate internal and external consistency scores and generate a list of words as text correction candidates.
 However, n-gram models do not take the linguistic structure of language into account.
@@
 Hence next-word prediction using n-gram based language models can be limited.
-
 Here think of a two-gram involving the word 'good' 'good morning'.
@@
 For instance, an equally reasonable two-gram involving the word 'good' is 'good work'.
-For that reason, we consider pre-trained word embedding models.
-These include Word2Vec and GloVe, which place each word in a multi-dimensional space such that distance between words can illustrate their relationship.
-We also consider pre-trained generative models such as GPT-2 and GPT-3 and BERT.
-GPT-2, GPT-3 and BERT are pre-trained generative unsupervised language models.
-GPT-2 and GPT-3 are the two generations of the same model that differ in the number of parameters.
-GPT-2 has 1.5 billion parameters and GPT-3 has more than 175 billion parameters.
-While GPT-2 is more limited, it can be run on smaller machines, whereas GPT-3 cannot and is accessed via an API.
-BERT has 340 millions parameters which is smaller than both GPT-2 and GPT-3.
-BERT also differs from GPT models because its encoding process is bidirectional instead of unidirectional.
+However, because n-gram models does not take the surrounding text into account, it has not capacity of judging if 'morining' or 'work' is to be suggested as the next word after 'good'.
+
+In future works, we intend to investigate in more advanced language models such as word embeddings and Transformers.
+We consider pre-trained word embedding models including Word2Vec [@mikolov2013efficient] [@mikolov2013distributed] and GloVe [@pennington2014glove], which place each word in a multi-dimensional space such that distance between words can illustrate their relationship [@bengio2003neural].
+This approach of representing words in vectors has a long history, but the implementation of @bengio2003neural has become the foundation for much subsequent work.
+Here, terms are encoded into feature vectors that represent different aspects of the term, and so each term is represented by a position in the vector space.
+
+We also consider incorporating generative pre-trained transformer models such as GPT-2 [@radford2019language] and GPT-3 [@brown2020language] and BERT [@devlin2018bert].
+GPT-2, GPT-3 and BERT are pre-trained language models that are built based on the Transformer architecture [@vaswani2017attention].
+Proposed by @vaswani2017attention from Google in 2017, the Transformer is a novel network architecture for neural network that outperforms RNN-based and CNN-based models in computational efficiency [@vaswani2017attention].
+Since the emergence of the Transformer model, most of the representative pre-trained models are built based on this architecture [@hanretty2018comparing].
+Therefore, we can say that Transformer marks a new era of language models.
+In our future works, it will be important for us to implement different language models compare their effectiveness in the tasks of generating consistency score and providing suggested text corrections and improve the `aRianna` package.
+
+\newpage
 

### 2020-09-02 outputs/documents/paper_ngrams.Rmd
Update abstract and introduction

@@
 biblio-style: apalike
-title: "aRianna: R package for consistency scores in text data"
-thanks: "Thank you to **X, Y and Z** for helpful comments.
+title: "On consistency scores in text data with an implementation in R"
+thanks: "Thank you to Monica Howlett for helpful comments.
 A Shiny application for interactive and small-scale examples is available: https://kelichiu.shinyapps.io/Arianna/.
@@
   affiliation: University of Toronto
-abstract: "In this paper, we introduce a process to clean the text extracted from PDFs using n-gram models that is widely used in statistical natural language processing.
-Our approach compares originally extracted text with the text generated from, or expected by, these methods using earlier text as stimulus.
-To guide this process, we introduce the notion of a consistency score, which refers to the proportion of text that is unchanged by the model.
-This is used to monitor changes during the cleaning process, and to compare the messiness of different texts.
+abstract: "In this paper, we introduce a reproducible cleaning process for the text extracted from PDFs using n-gram models.
+Our approach compares the originally extracted text with the text generated from, or expected by, these models using earlier text as stimulus.
+To guide this process, we introduce the notion of a consistency score, which refers to the proportion of text that is expected by the model.
+This is used to monitor changes during the cleaning process, and across different corpuses.
 We illustrate our process on text from the book Jane Eyre and introduce both a Shiny application and an R package to make our process easier for others to adopt."
-keywords: "text-as-data; natural language processing; quantitative analysis."
+keywords: "text-as-data; natural language processing; quantitative analysis; optical character recognition"
 date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
 fontfamily: mathpazo
-fontsize: 12pt
+fontsize: 11pt
 # spacing: double
@@
 Further, we define internal consistency as when the model is trained on the corpus itself, and external consistency as when the model is trained on a more general corpus.
-Together, these concepts provide a guide to the cleanliness and consistency of a text dataset.
+Together, these concepts provide a guide to the cleanliness and reasonableness of a text dataset.
 This can be important when deciding whether a dataset is fit for purpose; when carrying out data cleaning and preparation tasks; and as a comparison between datasets.
@@
 After we correct this error, the consistency score would likely increase.
-By examining how the consistency scores evolve in response to changes made to the text during the data preparation and cleaning stages we can better understand the effect of the changes.
+By examining how consistency scores evolve in response to changes made to the text during the data preparation and cleaning stages we can better understand the effect of the changes.
 Including consistency scores and their evolution when text corpuses are shared allows researchers to be more transparent about their corpus.
@@
 We employ n-gram models to calculate consistency scores and generate word candidates for text corrections.
-The n-gram approach involves identifying two, three, or more, words that are commonly found together.
-In our application, we use the sequence of three words, or tri-grams, to calculate consistency scores and generate word candidates for text corrections.
-By using tri-grams, we identify the first two words in the sequence and evaluate if the last word in the sequence is "as expected" by comparing it to the corresponding tri-gram in a comparison dataset.
+The n-gram approach involves identifying words that are commonly found together.
+In our application, we use these sequences to calculate consistency scores and generate word candidates for text corrections.
+For instance, if using a sequence of three words, or tri-grams, then we identify the first two words in the sequence and evaluate if the last word in the sequence is 'as expected' by comparing it to the corresponding tri-gram in a comparison dataset.
+
 The comparison dataset can be internal, meaning that the dataset is the source where the text being evaluated is extracted from.
-On the other hand, a comparison dataset is external when the dataset is from a different source than where the text being evaluated is from, and is a larger and more universal dataset that holds generalizability.
-The external dataset we use in the application is constructed using a subset of text data sourced from Open WebText Corpus by @Gokaslan2019OpenWeb.
-Since the content of Open WebText Corpus is from the web, the variety of the text data is promising for meeting the generalizability and universality an external dataset should possess.
+On the other hand, a comparison dataset can be external, which means the dataset is from a different source; typically a larger and more universal dataset.
+The external dataset we use in the application is constructed using a subset of text data sourced from Open WebText Corpus [@Gokaslan2019OpenWeb].
+Since the content of Open WebText Corpus is from the internet, the variety of the text data is promising for meeting the generalizability and universality that an external dataset should possess.
 
-In this paper, we introduce the `aRianna` R Package with the following structure: In Section 2, we provide an overview of language model and n-grams.
-In section we deconstruct the package functions and provide demonstration example of the package usage.
-Finally, we discuss the limitation and the future directions of aRianna in Section 4.
-Beyond this paper, we construct a Shiny app available at: https://kelichiu.shinyapps.io/aRianna/.
-That app computes internal and external consistency scores for corpus excerpts, and have developed an R Package, `aRianna`, that allows our approach to be used on larger datasets, which is available at: https://github.com/RohanAlexander/aRianna.
+The resources that we have developed to support this paper include a Shiny app available at: https://kelichiu.shinyapps.io/aRianna/.
+That app computes internal and external consistency scores for corpus excerpts.
+We have also developed an R Package, `aRianna`, that allows our approach to be used on larger datasets, which is available at: https://github.com/RohanAlexander/aRianna.
+
+In this paper, we introduce consistency scores and an R package that implements them.
+In Section 2, we provide an overview of language models and n-grams in particular.
+In Section 3 we deconstruct the package functions and provide demonstration example of the package usage.
+Finally, in Section 4 we discuss limitations and future directions.
 
@@
 
-## Language Model
+## Language model
 
@@
 
-## N-gram Models
+## N-gram models
 
@@
 
-# 3. The aRianna R Package
+# 3. The aRianna R package
 
-## Package Dependencies
+## Package dependencies
 
@@
 
-## Function to Make the Internal DataSet
+## Function to make the internal dataset
 
@@
 
-## Function to Generate Internal Consistency Score
+## Function to generate internal consistency
 
@@
 
-## Function to Generate External Consistency Score
+## Function to generate external consistency
 
 The `generate_external_consistency_score` function works identically with the `generate_internal_consistency_score`.
-The only difference is that it compares the text being evaluated with a external consistency dataset that is larger and more general.
+The only difference is that it compares the text being evaluated with an external consistency dataset that is larger and more general.
 The external consistency dataset we employed in the application is constructed using a subset text data from Open WebText Corpus by @Gokaslan2019OpenWeb.

### 2020-09-02 outputs/documents/paper_ngrams.Rmd
ngram paper updated

@@
 For instance, if using a sequence of three words, or tri-grams, then we identify the first two words in the sequence and evaluate if the last word in the sequence is 'as expected' by comparing it to the corresponding tri-gram in a comparison dataset.
+To calculate consistency scores and generate word candidates for text corrections, we use mixed sequences of five, four and three consecutive words also known as 5-grams, 4-grams and tri-grams.
 
@@
 
+In our application, we use a model that has a mixture of 5-grams, 4-grams and tri-grams to provide word replacements to a text error.
+Longer n-grams such as 5-grams would capture more context of the sequence and potentially generate more accurate predictions.
+However, there is less probability for a longer sequence of words to fit a specific combination.
+Therefore, shorter n-grams should still be included to preserve generalizability of the predictions.
+
+Consider the common English idiom "elephant in the room".
+If by mistake, the idiom is written as "elephant in the roon" and correction is needed for the last word, we can apply both a tri-gram model or a 4-gram model to predict the candidate word:
+
+> Using tri-gram model: "in the `<prediction>`"
+
+> Using 4-gram model: "elephant in the `<prediction>`"
+
+In the 4-gram model, one extra more word "elephant" is included and thus provides more context for the prediction, and the word "room" is predicted by the 4-gram model.
+However, there are times people would simply describe the location of an elephant, for example "elephant in the pond".
+This combination of this sequence as a 4-gram might not be in a comparison dataset.
+Therefore, we have to also rely on the tri-gram models to provide other word candidates including "in the pond".
+
 # 3. The aRianna R package
@@
 
-Next, the `make_internal_consistency_dataset` function turns the individual tokens into tri-grams, and keeps only the tri-grams that appears in the data for more than once.
-The common tri-grams are preserved in a tibble:
+Next, the `make_internal_consistency_dataset` function turns the individual tokens into 5-grams, 4-grams and tri-grams, and keeps only the n-grams that appears in the data for more than once.
+The common n-grams are preserved in a tibble:
 
@@
 toks_ngram <- 
-  quanteda::tokens_ngrams(tokens_from_example, n = 3)
+  toks_ngram <- quanteda::tokens_ngrams(tokens_from_example, n = 5:3)
 
@@
 
-Finally, the `make_internal_consistency_dataset` function splits the tri-grams to two parts: the previous two words and the last word.
-In the end, a tibble is created with each tri-gram containing its original three-word sequence, the first two words, and the last word:
+Finally, the `make_internal_consistency_dataset` function splits the n-grams to two parts: the previous $n-1$ words and the last word.
+In the end, a tibble is created with each n-gram containing its original sequence, the first words, and the last word:
 
@@
   dplyr::mutate(tokens = stringr::str_replace_all(tokens, "_", " "),
-                first_words = stringr::word(tokens, start = 1, end = 2),
+                first_words = stringr::word(tokens, start = 1, end = -2),
                 last_word = stringr::word(tokens, -1),
@@
 After we have the internal consistency dataset, we need a function to compare the text being evaluated and the internal dataset in order to retrieve the internal consistency score — the `generate_internal_consistency_score` function.
-Similar to the `make_internal_consistency_dataset` function, the function first transforms the text being evaluated into tri-gram tokens, then splits the tri-grams and preserves them into a tibble:
+Similar to the `make_internal_consistency_dataset` function, the function first transforms the text being evaluated into 5-grams, 4-grams and tri-grams tokens, then splits the n-grams and preserves them into a tibble:
 
@@
 toks_ngram_with_errors <- 
-  quanteda::tokens_ngrams(tokens_from_example_with_errors, n = 3)
+  quanteda::tokens_ngrams(tokens_from_example_with_errors, n = 5:3)
+
 all_tokens_with_errors <- 
@@
   dplyr::mutate(tokens = stringr::str_replace_all(tokens, "_", " "),
-                first_words = stringr::word(tokens, start = 1, end = 2),
+                first_words = stringr::word(tokens, start = 1, end = -2),
                 last_word = stringr::word(tokens, -1),
                 tokens = stringr::str_replace_all(tokens, " ", "_"),
-                first_words = stringr::str_replace_all(first_words, " ", "_")
-  )
+                first_words = stringr::str_replace_all(first_words, " ", "_"))
 ```
@@
 
-The `generate_internal_consistency_score` function also lists the identified text errors and generates replacement candidate to the text errors:
+The `generate_internal_consistency_score` function also lists the identified text errors and generates replacement candidates to the text errors:
 
@@
 The first step is to turn the body of text into the internal consistency dataset.
-The generated internal consistency dataset is a tibble that contains the tri-grams that appear in the internal text data more than once.
+The generated internal consistency dataset is a tibble that contains the 5-grams, 4-grams and tri-grams that appear in the internal text data more than once.
 Since only `there_was_no` has more than one occurrence, it is the only reference tri-gram in the internal consistency dataset:
@@
 ## $`unexpected words`
-## # A tibble: 1 x 4
-##   first_words last_word last_word_expected as_expected
-##   <chr>       <chr>     <chr>              <lgl>      
-## 1 there_was   na        no                 FALSE
+## # A tibble: 1 x 3
+##   first_words last_word last_word_expected 
+##   <chr>       <chr>     <chr>                 
+## 1 there_was   na        no
 
@@
 The function `generate_internal_consistency_score` takes in only one argument — the text to evaluate, and compares it with the external consistency dataset.
-The function identifies the word "na" as an unexpected word, and generates the external consistency score as zero.
-Because the external dataset is larger, there are several words generated as the replacement candidates of "na":
+Here, the text to evaluate is "there was no possibiliti", with possibility being wrongly spelled.
+The function identifies the word "possibiliti" as an unexpected word, and generates the external consistency score as 0.833.
+Because the external dataset is larger, there are several words generated as the replacement candidates of "possibiliti".
+The replacements based on 4-grams are displayed first followed by results based on tri-grams:
 
@@
 ##   <lgl>       <int>       <dbl>
-## 1 FALSE           2           0
+## 1 FALSE           2       0.833
+## 2 TRUE           10       0.833
 
 ## $`unexpected words`
-## # A tibble: 11 x 4
-##    first_words last_word last_word_expected as_expected
-##    <chr>       <chr>     <chr>              <lgl>      
-##  1 there_was   na        a                  FALSE      
-##  2 there_was   na        an                 FALSE      
-##  3 there_was   na        no                 FALSE      
-##  4 there_was   na        not                FALSE      
-##  5 there_was   na        nothing            FALSE      
-##  6 there_was   na        of                 FALSE      
-##  7 there_was   na        one                FALSE      
-##  8 there_was   na        only               FALSE      
-##  9 there_was   na        plenty             FALSE      
-## 10 there_was   na        someone            FALSE      
-## 11 there_was   na        something          FALSE
-
+## # A tibble: 11 x 3
+##    first_words  last_word    last_word_expected 
+##    <chr>        <chr>        <chr>                    
+##  1 there_was_no possibiliti  evidence          
+##  2 there_was_no possibiliti  immediate         
+##  3 there_was_no possibiliti  infrastructure    
+##  4 there_was_no possibiliti  one               
+##  5 there_was_no possibiliti  possibility       
+##  6 there_was_no possibiliti  sound             
+##  7 there_was_no possibiliti  way               
+##  8 was_no       possibiliti  good              
+##  9 was_no       possibiliti  longer            
+## 10 was_no       possibiliti  more              
+## 11 was_no       possibiliti  wonder 
 ```
@@
 For instance, an equally reasonable two-gram involving the word 'good' is 'good work'.
-However, because n-gram models does not take the surrounding text into account, it has not capacity of judging if 'morining' or 'work' is to be suggested as the next word after 'good'.
+However, because n-gram models does not take the surrounding text into account, it has not capacity of judging if 'morning' or 'work' is to be suggested as the next word after 'good'.
 

### 2020-09-03 outputs/documents/paper_ngrams.Rmd
paper minor updates

@@
 For instance, if using a sequence of three words, or tri-grams, then we identify the first two words in the sequence and evaluate if the last word in the sequence is 'as expected' by comparing it to the corresponding tri-gram in a comparison dataset.
-To calculate consistency scores and generate word candidates for text corrections, we use mixed sequences of five, four and three consecutive words also known as 5-grams, 4-grams and tri-grams.
+The sequences we employed are a mixture of five, four and three consecutive words also known as 5-grams, 4-grams and tri-grams.
 
@@
 In Section 2, we provide an overview of language models and n-grams in particular.
-In Section 3 we deconstruct the package functions and provide demonstration example of the package usage.
-Finally, in Section 4 we discuss limitations and future directions.
+In Section 3, we deconstruct the package functions and provide demonstration example of the package usage.
+Finally, in Section 4, we discuss limitations and future directions.
 
@@
 The reason that there is such a variety of statistical language models is that there are various ways of dealing with this fundamental problem.
+N-gram is one of the statistical language models that works by taking the co-occurrence of words in sequence into account.
 
@@
 In our application, we use a model that has a mixture of 5-grams, 4-grams and tri-grams to provide word replacements to a text error.
-Longer n-grams such as 5-grams would capture more context of the sequence and potentially generate more accurate predictions.
-However, there is less probability for a longer sequence of words to fit a specific combination.
+Longer n-grams such as 5-grams capture more context of word sequences and potentially generate more accurate predictions.
+However, there is less probability for a longer sequence of words to fit a specific combination and leads to data sparsity.
 Therefore, shorter n-grams should still be included to preserve generalizability of the predictions.
@@
 
-In the 4-gram model, one extra more word "elephant" is included and thus provides more context for the prediction, and the word "room" is predicted by the 4-gram model.
-However, there are times people would simply describe the location of an elephant, for example "elephant in the pond".
-This combination of this sequence as a 4-gram might not be in a comparison dataset.
+In the 4-gram model, one extra more word "elephant" is included and thus provides more context for the word "room" to be accurately predicted.
+However, there are times people would simply describe the location of an elephant; for example, "elephant in the pond".
+The combination of this sequence as a 4-gram token might not be in a comparison dataset and no prediction would be available if only the 4-gram models are used.
 Therefore, we have to also rely on the tri-gram models to provide other word candidates including "in the pond".
@@
 
-To obtain the internal consistency score, we will construct a internal consistency dataset from a body of text that is the source of the text being evaluated.
+To obtain the internal consistency score, we will construct an internal consistency dataset from a body of text that is the source of the text being evaluated.
 The function `make_internal_consistency_dataset` is created for this purpose.
-The function turns the body_of_text into an internal consistency dataset which the text being evaluated is compared with.
+The function turns the body_of_text into an internal consistency dataset which the text being evaluated will be compared with.
 To do so, the function first transforms the body of text into word tokens with the punctuation removed and the letters turned to lowercase:
@@
 Finally, the `make_internal_consistency_dataset` function splits the n-grams to two parts: the previous $n-1$ words and the last word.
-In the end, a tibble is created with each n-gram containing its original sequence, the first words, and the last word:
+As a result, a tibble is created with each n-gram containing its original sequence, the first words, and the last word:
 
@@
 After we have the internal consistency dataset, we need a function to compare the text being evaluated and the internal dataset in order to retrieve the internal consistency score — the `generate_internal_consistency_score` function.
-Similar to the `make_internal_consistency_dataset` function, the function first transforms the text being evaluated into 5-grams, 4-grams and tri-grams tokens, then splits the n-grams and preserves them into a tibble:
+Similar to the `make_internal_consistency_dataset` function, the function first transforms the text being evaluated into 5-grams, 4-grams and tri-grams tokens, then splits the n-grams to their first words and last words:
 
@@
 Next, the function combines the text tokens tibble and the internal consistency tibble by the `left_join()` function of `dplyr`.
-The `left_join()` function returns all the rows from the text tokens tibble and all the columns from both tibbles.
-By combining the two tibbles, the function generates two additional columns — the `last_word` column are the last words of all tri-grams, and the `last_word_expected` contains the words that are in the internal consistency dataset.
+The `left_join()` function returns all the rows from the text tokens tibble and all the columns from both tibbles [@2018dplyr].
+By combining the two tibbles, the function generates two additional columns — the `last_word` column are the last words of all n-grams, and the `last_word_expected` contains the last words that are in the internal consistency dataset.
 By comparing the words in `last_word` and `last_word_expected`, we can identify the "unexpected" words.
@@
 
-The `generate_external_consistency_score` function works identically with the `generate_internal_consistency_score`.
-The only difference is that it compares the text being evaluated with an external consistency dataset that is larger and more general.
+The `generate_external_consistency_score` function works identically as the `generate_internal_consistency_score` does.
+The only difference lies in that it compares the text being evaluated with an external consistency dataset that is larger and more general.
 The external consistency dataset we employed in the application is constructed using a subset text data from Open WebText Corpus by @Gokaslan2019OpenWeb.
@@
 The generated internal consistency dataset is a tibble that contains the 5-grams, 4-grams and tri-grams that appear in the internal text data more than once.
-Since only `there_was_no` has more than one occurrence, it is the only reference tri-gram in the internal consistency dataset:
+Since only `there_was_no` has more than one occurrence, it is the only reference n-gram in the internal consistency dataset:
 
@@
 The function `generate_internal_consistency_score` takes in only one argument — the text to evaluate, and compares it with the external consistency dataset.
-Here, the text to evaluate is "there was no possibiliti", with possibility being wrongly spelled.
-The function identifies the word "possibiliti" as an unexpected word, and generates the external consistency score as 0.833.
-Because the external dataset is larger, there are several words generated as the replacement candidates of "possibiliti".
-The replacements based on 4-grams are displayed first followed by results based on tri-grams:
+Here, the text to evaluate is "there was no possibiliti", with the word "possibility" being wrongly spelled.
+The function identifies the word "possibiliti" as an unexpected word, and generates the external consistency score 0.833.
+Because the external dataset is larger, it is capable of providing more replacement candidates for "possibiliti".
+The replacements based on 4-grams are displayed first followed by the replacements based on tri-grams:
 
@@
 Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation [@brown1992class].
-As demonstrated in our application, we use a tri-gram based model to calculate internal and external consistency scores and generate a list of words as text correction candidates.
+As demonstrated in our application, we use a n-gram based model to calculate internal and external consistency scores and generate a list of words as text correction candidates.
 However, n-gram models do not take the linguistic structure of language into account.
@@
 
-We also consider incorporating generative pre-trained transformer models such as GPT-2 [@radford2019language] and GPT-3 [@brown2020language] and BERT [@devlin2018bert].
+We also consider incorporating generative pre-trained transformer models such as GPT-2 [@radford2019language], GPT-3 [@brown2020language], and BERT [@devlin2018bert].
 GPT-2, GPT-3 and BERT are pre-trained language models that are built based on the Transformer architecture [@vaswani2017attention].
@@
 Therefore, we can say that Transformer marks a new era of language models.
-In our future works, it will be important for us to implement different language models compare their effectiveness in the tasks of generating consistency score and providing suggested text corrections and improve the `aRianna` package.
+In our future works, it will be important for us to implement different language models compare their effectiveness in the tasks of generating consistency score and providing suggested text corrections.
+We intend to use our future findings to continuously improve the `aRianna` package.
 

### 2020-09-03 scripts/00-try-BERT-text-prediction.R
BERT in R code organization

@@ -1,5 +1,5 @@
 # Source tutorial: https://blogs.rstudio.com/ai/posts/2019-09-30-bert-r/
-Sys.setenv(TF_KERAS=1) 
+Sys.setenv(TF_KERAS=1)
 # to see python version
 reticulate::py_config()
 setwd("~/Desktop/repos/consistency_scores_in_text_datasets/model_starters")
@@ -19,9 +19,12 @@ library(hunspell)
 incorrectwords <- hunspell(text_stripped)
 
 tokens <- unlist(strsplit(text_stripped, "[^aA-zZ<>/]"))
-mask_indices <- match(c("imlfon", "Kism"),tokens)
-tokens[12] = tokens[33] = "[MASK]"
-toString(tokens)
+mask_indices <- match(incorrectwords[[1]],tokens)
+for(i in mask_indices){
+  tokens[i] = "MASK"
+}
+str <- toString(tokens)
+str <- strip(str)
 #### 3. Configure paths to the BERT pre-trained files variables
 # The zip file is available through the tutorial link at the first line
 pretrained_path = '../../../uncased_L-12_H-768_A-12'
@@ -44,21 +47,29 @@ model = k_bert$load_trained_model_from_checkpoint(
 
 library(zeallot)
 #### 5. Tokenize the text and get the encoding
-tokens <- tokenizer$tokenize(toString(tokens))
-encodes <- tokenizer$encode(toString(tokens))
+tokens <- tokenizer$tokenize(str)
+encodes <- tokenizer$encode(str)
+
+for(i in mask_indices){
+  encodes[[1]][i+1] = 103
+}
 
 #### 6. Get encoded inputs for the model: train, segments, targets
 indices <- encodes[1]
 indices <- as.matrix(indices)
 segments <- encodes[2]
 segments <- as.matrix(segments)
-## Manually flag word 13 and word 33 as masked words to be predicted 
-masks <- as.matrix(list(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0,
-                        0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
-                        0, 0, 0, 0, 0, 0))
+
+## Manually flag word 13 and word 33 as masked words to be predicted
+masks <- as.matrix(list(0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
[…]

### 2020-10-03 outputs/documents/notes/literacture_review.Rmd
lit review first article

---
title: "Literature_review"
author: "Keli Chiu"
date: "03/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## MaskGAN: Better text generation via filling in the ________

Published date: March 1, 2018

Authors: William Fedus, Ian Goodfellow and Andrew M. Dai

Affiliation: Google Brain

URL: https://arxiv.org/pdf/1801.07736.pdf

### Background

At the time where the paper was published, text generation techniques with the best results were commonly built on Recurrent Neural Networks (RNNs). In this paper, the authors proposed a new framework that involves the usage of Generative Adversarial Networks (GANs) (Goodfellow et al., 2014). GAN is a framework for training generative models in an adversarial setup that consists of a discriminator and a generator, where the model learns from the competition between the discriminator and the generator. GAN is widely used in image generation. In this paper, the authors applied the GAN model in the context of natural language to achieve text generation.

### How it works
According to the paper, GANs commonly suffer from issues such as training instability and mode dropping, both are exacerbated in a textual setting due to the discrete nature of text. The authors reduce the impact of these problems by training the model on a text in-filing task. In this task, portions of a body of text are deleted or redacted, and the model aims to infill the missing portions of text to be indistinguishable from the original data in an autoregressive manner.

The network is trained by the task of text-infilling using the sequence-to-sequence (Seq2Seq) architecture (Sutskever et al., 2014) for both the generator and the discriminator. During the training process, the encoder of the generator produces the context representation using the masked sequence. The discriminator then conditions on the generator’s output and the masked sequence to compute the probability of a word in the generator’s output as being fake or real. If one wishes to do text-generation instead of text-infilling, it can be done by setting all inputs to be blanks (masked).

### Limitations

* According to one reviewer on OpenReview: “because the model aims to fill-in blanks from the text around (up to that time), generated texts are generally locally valid but not always valid globally.”

* Another reviewer raised the concern of the model performance.

* As pointed out by the paper Enabling text infilling by language models: “While their approach is effective at infilling individual words, it is somewhat redundant as the model must “predict” the unmasked text in $\tilde{x}$. Additionally, a model is not guaranteed to exactly reproduce the unmasked text.”

### Other notes

* One of the authors, Goodfellow, is the person who co-invented GAN.

### 2020-10-03 outputs/documents/notes/literacture_review.Rmd
lit reviews all four articles

@@
 * One of the authors, Goodfellow, is the person who co-invented GAN.
+
+## Text Infilling
+
+Published date: January 18, 2019
+
+Authors: Wanrong Zhu, Zhiting Hu, Eric P. Xing
+
+Affiliation: Peking University, Carnegie Mellon University, Petuum Inc.
+
+URL: https://arxiv.org/pdf/1901.00158.pdf
+
+### Background
+
+This paper proposes an approach to achieve text-infilling tasks based on several representative solutions, including RNN with sequence-to-sequence (Seq2Seq) (Bahdanau et al., 2014), a GAN-based model (Goodfellow et al., 2014), and a self-attention method (Vaswani et al., 2017).
+The authors concluded that the self-attention method has the best performance for the text-infilling task.
+Compared to MaskGAN, the main breakthrough of the approach is its ability to generate unconstraining length of words in one blank gap.
+The proposed approach aims to fill the missing portions based on the global and surrounding context, to make the text complete and meaningful.
+
+See the example given by the paper, where there are six words inserted in the two blanks:
+
+![Text infilling example](figures/TIF01.png)
+
+### How it works
+
+In the training process, the authors randomly sample sentences from the training corpus and blank out words with different strategies (taking out prepositions and articles, keeping only achor words...etc).
+For each input, the number of blanks and their positions are known, but the number of tokens to be filled in each blank is not known.
+A model must decide by itself how many tokens to generate for a blank.
+With this approach, the model learns to produce in-filling results that does not restrict to a single word.
+
+### Limitations
+
+* According to a reviewer on Open Review: “the size of the data set is far from sufficient.
+Unless the authors are trying to show that the transformer is more data-efficient (which is doubtful), the dataset needs to be much larger than the 1M token it appears to be now.
+The size of the vocabularies is also far from being representative of any real-world setting.”
+
+* As pointed out by the paper Enable text-infilling by language models”: “this approach relies on specialized architecture and hence cannot easily leverage large-scale pre-trained models.”
+
+### Other notes
+
+* The table provides the comparison of different approaches. Golden is the original data. The approach that uses self-attention has the best meaningful result.
+
+* Self-attention mechanism is what Transformer architecture is based on.
+
+## XL-Editor: Post-editing Sentences with XLNet
+
+Published date: October 19, 2019
+
+Authors: Yong-Siang Shih, Wei-Cheng Chang, Yiming Yang
+
+Affiliation: Carnegie Mellon University
+
+URL: https://arxiv.org/pdf/1910.10479.pdf
+
+### Background
+
+This paper proposed a novel training framework for XLNet (Yang et al.
+2019) to execute operation of post-editing of text data.
+The extension of XLNet called XLEditor, was then created to perform text post-editing operations such as insertion, deletion, and replacement.
+While the paper aims to perform text editing, text in-filling is one of the tasks the model was trained to do.
+Created in 2019, XLNet is a new pre-trained language model based on the Transformer architecture with recurrence.
+According to the authors, XLNet outperforms other Transformer-based language models including GPT-2 from OpenAI and BERT from Google.
+The authors examined the capability of XLEditor by three tasks: including Locate (which position to insert), Text Infilling (what tokens to be inserted), and Text Deletion (what tokens to be deleted).
+The author also claimed that XLEditor outperforms XLNet in the above tasks.
+
+### How it works
+
+According to the authors, the heart of the XL-Editor model is to estimate the variable length insertion probability for the post-editing process.
+Still working to understand the meaning of this claim.
+The training method for the three tasks for post-editing are as follows:
+
+Locate task: In the training sentences for locating the need of text insertion, a randomly selected subsequence is deleted.
+The model is given 5 positions and is asked to decide which position to perform text insertion.
+
+Text infilling task: Text infilling task is built upon the work of (Zhu, Hu, and Xing 2019; Liu et al.
+2019).
+The improvement provided by the authors was to provide more context in the training to restrict the possibilities of plausible insertions.
+
+Text deletion: In each training example for text deletion task, there are five consecutive sentences sampled from the testing articles.
+One of the three sentences in the middle is sampled from a different article and is to be removed for text coherence.
+The model is asked to decide which sentence is not coherent with the rest.
+
+### Limitations
+
+Limitations are not discussed in the paper, not found in other papers, and public reviews on open reviewers are not found.
+
+### Other notes
+
+Now made as a desktop application: https://www.xldevelopment.net/editor.php
+
+## Enabling Language Models to Fill in the Blanks
+
+Published date: September 10, 2020
+
+Authors: Chris Donahue, Mina Lee, Percy Liang
+
+Affiliation: Stanford University
+
+URL: https://arxiv.org/pdf/2005.05339.pdf
+
+### Background
+
+In this paper, the authors present a simple framework that enables language models to infill variable-length spans called infilling by language modeling (ILM).
+The authors claimed that the framework is model-agnostic, which means it can be applied to any existing language models.
+Moreover, the framework performs better with pre-trained language models instead of training from scratch.
+Their best result presented in the paper was from fine-tuning GPT-2.
+
+### How it works
+
+There are four different training strategies they employed for the experiment.
+LM considers only the left context while LM-Rev only considers the right.
+LM-All considers context on both sides but uses long sequence lengths.
+ILM, which is the proposed model, considers context from both directions with fewer tokens.
+
+What’s valuable about the paper is that it demonstrates how to prepare training data for fine-tuning the language model (GPT-2 in their case), which is a knowledge we didn’t have in the summer.
+To produce an infilling example, a number of random spans in a sentence were replaced with [blank] tokens.
+The masked sentence, the separation token ([sep]), and the answers are then concatenated to be one example.
+The following screenshot is the training examples for different strategies they employed:
+
+![Training examples for different strategies](figures/ILM01.png)
+
+\newpage
+
+After the model is trained, the model takes text input with blanks, predicts the answers, and then merges them as one sentence to produce the output. For example:
+
+  > She ate [blank] for [blank] + leftover pasta [answer] lunch [answer] =
+  > She ate leftover pasta for lunch
+
+### Limitations
+
+Limitations are not discussed in the paper, not found in other papers, and public reviews on open reviewers are not found.
+
+### Other notes
+
+None

### 2020-10-04 outputs/documents/notes/paper_code_documentation.Rmd
code trial for text infilling paper

---
title: "paper_code_documentation"
author: "Keli Chiu"
date: "03/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Paper: Text Infilling

Published date: January 18, 2019

Authors: Wanrong Zhu, Zhiting Hu, Eric P. Xing

Affiliation: Peking University, Carnegie Mellon University, Petuum Inc.

Code URL: https://github.com/VegB/Text_Infilling

### Result

Not able to follow through the code tutorial.

### Steps:

1. Download the code and install Texar.

``` {python, eval = FALSE}
git clone https://github.com/VegB/Text_Infilling
cd Text_Infilling
pip install --user -e .    
```

2. Navigate to the infilling demo directory and download the dataset. The instruction provided on Github didn't work, I followed the `README.MD` in the `text_infilling` directory from this point. Downloaded data is stored in `yelp_data/`.

``` {python, eval = FALSE}
cd text_infilling
python data_utils.py
```

You might have to run extra code to install required packages if you don't have them already. In my case, I had to get `tensorflow` and `matplotlib`.

``` {python, eval = FALSE}
pip install "tensorflow>=1.15,<2.0" 
pip install matplotlib
```

3. Code template to execute is provided as follow:

``` {python, eval = FALSE}
python [MODEL].py --mask_rate [MASK_RATE] --blank_num [BLANK_NUM] 
  --filename_prefix 'pos.' --data_dir './yelp_data/pos/'
```

Here is what I tried, set mask rate to 30%, number of blank in an example is 1, and chose the self-attention mode.

``` {python, eval = FALSE}
python self_attn.py --mask_rate 30 --blank_num 1 --filename_prefix 'pos.' 
  --data_dir './yelp_data/pos/'
```

An Error was raised from the above code. The error message:

``` {python, eval = FALSE}
Traceback (most recent call last):
  File "self_attn.py", line 335, in <module>
    tf.app.run(main=_main)
  File "/Users/kelichiu/Desktop/project_tif/lib/python3.7/site-packages/tensorflow_core/python/platform/app.py", 
  line 40, in run
    _run(main=main, argv=argv, flags_parser=_parse_flags_tolerate_undef)
  File "/Users/kelichiu/Desktop/project_tif/lib/python3.7/site-packages/absl/app.py", 
  line 300, in run
    _run_main(main, args)
  File "/Users/kelichiu/Desktop/project_tif/lib/python3.7/site-packages/absl/app.py", 
  line 251, in _run_main
    sys.exit(main(argv))
  File "self_attn.py", line 64, in _main
    hparams=args.word_embedding_hparams)
  File "/Users/kelichiu/Desktop/repos/Text_Infilling/texar/modules/embedders/embedders.py", 
  line 53, in __init__
    self._hparams)
  File "/Users/kelichiu/Desktop/repos/Text_Infilling/texar/modules/embedders/embedder_base.py", 
  line 33, in _init_parameterized_embedding
    hparams, init_value, num_embeds, self.variable_scope)
  File "/Users/kelichiu/Desktop/repos/Text_Infilling/texar/modules/embedders/embedder_utils.py", 
  line 191, in get_embedding
    initializer = layers.get_initializer(hparams["initializer"])
  File "/Users/kelichiu/Desktop/repos/Text_Infilling/texar/core/layers.py", 
  line 333, in get_initializer
    initializer = utils.get_instance(hparams["type"], kwargs, modules)
  File "/Users/kelichiu/Desktop/repos/Text_Infilling/texar/utils/utils.py", 
  line 209, in get_instance
    (class_.__module__, class_.__name__, key, class_args))
ValueError: Invalid argument for class 
  tensorflow.python.ops.init_ops.RandomNormal: mean, valid args:set()
```

I have change the --mask_rate 30 and --blank_num 1 with variations between int, float and percentage, but no value could successfully run the code. Googled solutions but no relevant answers were found.

## Paper: Text Infilling

Extra step
pip install nltk

### 2020-10-04 outputs/documents/notes/paper_code_documentation.Rmd
paper code trials

@@
 
-## Paper: Text Infilling
+## Paper: Enabling Language Models to Fill in the Blanks
 
-Extra step
+Published date: September 10, 2020
+
+Authors: Chris Donahue, Mina Lee, Percy Liang
+
+Affiliation: Stanford University
+
+Code URL: https://github.com/chrisdonahue/ilm
+
+### Result
+
+Not able to follow through the code tutorial from Github. Google Collab code generates demo successfully.
+
+## Steps for Github repo
+
+1. Download the code and initialize
+
+``` {python, eval = FALSE}
+git clone git@github.com:chrisdonahue/ilm.git
+cd ilm
+pip install -r requirements.txt
+python -c "import nltk; nltk.download('punkt')"
+pip install -e .
+```
+
+You might have to run extra code to install required packages if you don't have them already. In my case, I had to get `nltk`.
+
+``` {python, eval = FALSE}
 pip install nltk
+```
+
+2. The following script was provided to download training dataset and create training examples:
+
+``` {sh, eval = FALSE}
+DATASET=arxiv_cs_abstracts
+
+pushd data
+./get_${DATASET}.sh
+popd
+
+for SPLIT in train valid
+do
+	python create_ilm_examples.py \
+		${SPLIT} \
+		data/char_masks/${DATASET} \
+		--seed 0 \
+		--data_name ${DATASET} \
+		--data_split ${SPLIT}
+done
+```
+
+The above code was saved as a sh script and execute by the following command in the terminal:
+
+``` {python, eval = FALSE}
+./script.sh
+```
+
+And the above execution raised an error:
+
+``` {python, eval = FALSE}
+~/desktop/repos/ilm/data ~/desktop/repos/ilm
+~/desktop/repos/ilm/data/raw_data/arxiv_cs_abstracts ~/desktop/repos/ilm/data ~/desktop/repos/ilm
+./get_arxiv_cs_abstracts.sh: line 5: wget: command not found
+sed: illegal option -- r
+usage: sed script [-Ealn] [-i extension] [file ...]
+       sed [-Ealn] [-i extension] [-e script] ... [-f script_file] ... [file ...]
+./get_arxiv_cs_abstracts.sh: line 5: wget: command not found
+gunzip: can't stat: arxiv_cs_abstracts.txt.gz (arxiv_cs_abstracts.txt.gz.gz): No such file or directory
+./get_arxiv_cs_abstracts.sh: line 7: sha256sum: command not found
+~/desktop/repos/ilm/data ~/desktop/repos/ilm
+~/desktop/repos/ilm
+Traceback (most recent call last):
+  File "create_ilm_examples.py", line 158, in <module>
+    limit=args.max_num_documents)
+  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 24, in get_dataset
+    d = arxiv_cs_abstracts(split, *args, data_dir=data_dir, **kwargs)
+  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 63, in arxiv_cs_abstracts
+    with open(os.path.join(data_dir, 'arxiv_cs_abstracts.txt'), 'r') as f:
+FileNotFoundError: [Errno 2] No such file or directory: '/Users/kelichiu/Desktop/repos/ilm/data/raw_data/arxiv_cs_abstracts/arxiv_cs_abstracts.txt'
+Traceback (most recent call last):
+  File "create_ilm_examples.py", line 158, in <module>
+    limit=args.max_num_documents)
+  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 24, in get_dataset
+    d = arxiv_cs_abstracts(split, *args, data_dir=data_dir, **kwargs)
+  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 63, in arxiv_cs_abstracts
+    with open(os.path.join(data_dir, 'arxiv_cs_abstracts.txt'), 'r') as f:
+FileNotFoundError: [Errno 2] No such file or directory: '/Users/kelichiu/Desktop/repos/ilm/data/raw_data/arxiv_cs_abstracts/arxiv_cs_abstracts.txt'
+(project_ilm) (base) Kelis-MacBook-Pro:ilm kelichiu$ ./script.sh
+~/desktop/repos/ilm/data ~/desktop/repos/ilm
+./script.sh: line 4: data/get_arxiv_cs_abstracts.sh: No such file or directory
+~/desktop/repos/ilm
+Traceback (most recent call last):
+  File "create_ilm_examples.py", line 158, in <module>
+    limit=args.max_num_documents)
+  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 24, in get_dataset
+    d = arxiv_cs_abstracts(split, *args, data_dir=data_dir, **kwargs)
+  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 63, in arxiv_cs_abstracts
+    with open(os.path.join(data_dir, 'arxiv_cs_abstracts.txt'), 'r') as f:
+FileNotFoundError: [Errno 2] No such file or directory: '/Users/kelichiu/Desktop/repos/ilm/data/raw_data/arxiv_cs_abstracts/arxiv_cs_abstracts.txt'
+Traceback (most recent call last):
+  File "create_ilm_examples.py", line 158, in <module>
+    limit=args.max_num_documents)
+  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 24, in get_dataset
+    d = arxiv_cs_abstracts(split, *args, data_dir=data_dir, **kwargs)
+  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 63, in arxiv_cs_abstracts
+    with open(os.path.join(data_dir, 'arxiv_cs_abstracts.txt'), 'r') as f:
+FileNotFoundError: [Errno 2] No such file or directory: '/Users/kelichiu/Desktop/repos/ilm/data/raw_data/arxiv_cs_abstracts/arxiv_cs_abstracts.txt'
+(project_ilm) (base) Kelis-MacBook-Pro:ilm kelichiu$ ./script.sh
+~/desktop/repos/ilm/data ~/desktop/repos/ilm
+./script.sh: line 4: data/get_arxiv_cs_abstracts.sh: No such file or directory
+~/desktop/repos/ilm
+Traceback (most recent call last):
+  File "create_ilm_examples.py", line 158, in <module>
+    limit=args.max_num_documents)
+  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 24, in get_dataset
+    d = arxiv_cs_abstracts(split, *args, data_dir=data_dir, **kwargs)
+  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 63, in arxiv_cs_abstracts
+    with open(os.path.join(data_dir, 'arxiv_cs_abstracts.txt'), 'r') as f:
+FileNotFoundError: [Errno 2] No such file or directory: '/Users/kelichiu/Desktop/repos/ilm/data/raw_data/arxiv_cs_abstracts/arxiv_cs_abstracts.txt'
+Traceback (most recent call last):
+  File "create_ilm_examples.py", line 158, in <module>
+    limit=args.max_num_documents)
+  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 24, in get_dataset
+    d = arxiv_cs_abstracts(split, *args, data_dir=data_dir, **kwargs)
+  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 63, in arxiv_cs_abstracts
+    with open(os.path.join(data_dir, 'arxiv_cs_abstracts.txt'), 'r') as f:
+FileNotFoundError: [Errno 2] No such file or directory: '/Users/kelichiu/Desktop/repos/ilm/data/raw_data/arxiv_cs_abstracts/arxiv_cs_abstracts.txt'
+
+```
+
+## Google Colab
+
+Link: https://colab.research.google.com/drive/1So95M0hHefyNm_eELglCna_ZayoDX6KV?usp=sharing
+
+### Input
+
+``` {txt, eval = FALSE}
+English Class
+
+Christie was good at doing _. _ _ _ She ended up failing the test.
+```
+
+### Encode
+
+``` {txt, eval = FALSE}
+English Class
+
+Christie was good at doing<|infill_word|>.<|infill_paragraph|><|infill_sentence|>
+  <|infill_sentence|> She ended up failing the test.
+```
+
+### Output
+
+``` {txt, eval = FALSE}
+English Class
+
+Christie was good at doing things. One day at the library she had to take a math test. 
+She made a terrible grade. Christie felt terrible and did it anyways. 
+She ended up failing the test.
+
+English Class
+
+Christie was good at doing homework. She would stay up late to study. 
+When she was home for class, she would go into and out of class. 
+But her teacher couldn't find her anywhere. She ended up failing the test.
+```

### 2020-10-10 outputs/documents/notes/literacture_review.Rmd
lit review for XLNet

@@
 None
+
+## XLNet: Generalized Autoregressive Pretrainingfor Language Understanding
+
+Published date: January 2, 2020
+
+Authors: Zhilin Yang, Zihang Dai, Yiming Yang, Jaime Carbonell, Ruslan Salakhutdinov, Quoc V. Le
+
+Affiliation: Carnegie Mellon University, Google AI Brain Team
+
+URL: https://arxiv.org/pdf/1906.08237.pdf
+
+### Background
+
+According to the author, there are two common pre-training objectives in the real of unsupervised representation learning for NLP — Autoregressive and autoencoding.
+Autoregressive models such as GPT-2 seek to estimate the probability distribution of a corpus by factorizing the likelihood with either a forward direction (from left to right) or a backward direction (from right to left).
+This training method is not effective for the models to take contexts on both direction into account.
+On the other hand, autoencoding-based model, BERT, is not restricted to the uni-directional density estimation and is able to learn in a bi-directional way.
+However, BERT is not able to model the joint probability and assumes that predicted blanks are independent from each other.
+The strategy of XLNet is to bring the best of both autoregressive and autoencoding models and leave out their weakness.
+
+### How it works
+
+The strategy of XLNet is based on permutation language modeling, which captures bidirectional context by training an autoregressive model on all possible permutation of tokens in a sentence.
+For example, in a sequence of three tokens [1, 2, 3], the permutations are 3!
+= {[1,2,3], [1,3,2], [2,1,3], [2,3,1], [3,1,2], [3,2,1]}.
+The model is auto-regressive over all the permutations and calculates the probability of token $x_i$ given the preceding context $x<i$ from any permutation of Z.
+Taking all three positions and all 3!
+permutations into account, the permutation operation makes each position learns to use contextual information from all positions, including bidirectional context.
+
+Another strategy XLNet utilizes is two-stream self-attention.
+For each token position $i$, there are two associated vectors at each self-attention layer.
+One layer is content stream attention (denoted as $h$), which is a standard hidden states in Transformer.
+The other layer, called the query representation stream, is an additional self-attention layer unique to XLNet, created to preserve the contexts information.
+For example, a token $x_i$ can serve as two roles.
+When it is used as content to predict other tokens, the content representation is used.
+On the other hand, if $x_i$ is the token to be predicted, the query representation is used to preserve context information ($x<i$) and the position information of $x_i$.
+
+### Limitations
+
+* According to Ajit Rajasekharan's post on Quora dated July 6, 2019, XLNet requires more compute power and memory (GPU/TPU memory) in comparison to BERT.
+
+### Other notes
+
+* XLNet integrates ideas from Transformer-XL, the state-of-the-art autoregressive model

### 2020-10-11 outputs/documents/notes/paper_code_documentation.Rmd
paper code implementation documentation

@@
 
-2. The following script was provided to download training dataset and create training examples:
+2.
+The following script was provided to download training dataset and create training example.
+It requires us saving the script as an sh file in order to execute it.
+I saved it in the repo root directory as script.sh.
 
@@
 
-The above code was saved as a sh script and execute by the following command in the terminal:
+3. Execute the script by `./script.sh`. If permission error is displayed, try  `chmod +x script.sh` then execute the script again:
 
-``` {python, eval = FALSE}
+``` {sh, eval = FALSE}
+# chmod +x script.sh
 ./script.sh
@@
 
-And the above execution raised an error:
+3.1.
+If the above execution raises an error that tells you that `arxiv_cs_abstracts.txt` is not found, go to this link to download the missing file: https://docs.google.com/uc?export=download&id=1N3MbvpgZAmNgiZgnpXAQFzHrU7Tt3Blb
 
-``` {python, eval = FALSE}
-~/desktop/repos/ilm/data ~/desktop/repos/ilm
-~/desktop/repos/ilm/data/raw_data/arxiv_cs_abstracts ~/desktop/repos/ilm/data ~/desktop/repos/ilm
-./get_arxiv_cs_abstracts.sh: line 5: wget: command not found
-sed: illegal option -- r
-usage: sed script [-Ealn] [-i extension] [file ...]
-       sed [-Ealn] [-i extension] [-e script] ... [-f script_file] ... [file ...]
-./get_arxiv_cs_abstracts.sh: line 5: wget: command not found
-gunzip: can't stat: arxiv_cs_abstracts.txt.gz (arxiv_cs_abstracts.txt.gz.gz): No such file or directory
-./get_arxiv_cs_abstracts.sh: line 7: sha256sum: command not found
-~/desktop/repos/ilm/data ~/desktop/repos/ilm
-~/desktop/repos/ilm
-Traceback (most recent call last):
-  File "create_ilm_examples.py", line 158, in <module>
-    limit=args.max_num_documents)
-  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 24, in get_dataset
-    d = arxiv_cs_abstracts(split, *args, data_dir=data_dir, **kwargs)
-  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 63, in arxiv_cs_abstracts
-    with open(os.path.join(data_dir, 'arxiv_cs_abstracts.txt'), 'r') as f:
-FileNotFoundError: [Errno 2] No such file or directory: '/Users/kelichiu/Desktop/repos/ilm/data/raw_data/arxiv_cs_abstracts/arxiv_cs_abstracts.txt'
-Traceback (most recent call last):
-  File "create_ilm_examples.py", line 158, in <module>
-    limit=args.max_num_documents)
-  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 24, in get_dataset
-    d = arxiv_cs_abstracts(split, *args, data_dir=data_dir, **kwargs)
-  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 63, in arxiv_cs_abstracts
-    with open(os.path.join(data_dir, 'arxiv_cs_abstracts.txt'), 'r') as f:
-FileNotFoundError: [Errno 2] No such file or directory: '/Users/kelichiu/Desktop/repos/ilm/data/raw_data/arxiv_cs_abstracts/arxiv_cs_abstracts.txt'
-(project_ilm) (base) Kelis-MacBook-Pro:ilm kelichiu$ ./script.sh
-~/desktop/repos/ilm/data ~/desktop/repos/ilm
-./script.sh: line 4: data/get_arxiv_cs_abstracts.sh: No such file or directory
-~/desktop/repos/ilm
-Traceback (most recent call last):
-  File "create_ilm_examples.py", line 158, in <module>
-    limit=args.max_num_documents)
-  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 24, in get_dataset
-    d = arxiv_cs_abstracts(split, *args, data_dir=data_dir, **kwargs)
-  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 63, in arxiv_cs_abstracts
-    with open(os.path.join(data_dir, 'arxiv_cs_abstracts.txt'), 'r') as f:
-FileNotFoundError: [Errno 2] No such file or directory: '/Users/kelichiu/Desktop/repos/ilm/data/raw_data/arxiv_cs_abstracts/arxiv_cs_abstracts.txt'
-Traceback (most recent call last):
-  File "create_ilm_examples.py", line 158, in <module>
-    limit=args.max_num_documents)
-  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 24, in get_dataset
-    d = arxiv_cs_abstracts(split, *args, data_dir=data_dir, **kwargs)
-  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 63, in arxiv_cs_abstracts
-    with open(os.path.join(data_dir, 'arxiv_cs_abstracts.txt'), 'r') as f:
-FileNotFoundError: [Errno 2] No such file or directory: '/Users/kelichiu/Desktop/repos/ilm/data/raw_data/arxiv_cs_abstracts/arxiv_cs_abstracts.txt'
-(project_ilm) (base) Kelis-MacBook-Pro:ilm kelichiu$ ./script.sh
-~/desktop/repos/ilm/data ~/desktop/repos/ilm
-./script.sh: line 4: data/get_arxiv_cs_abstracts.sh: No such file or directory
-~/desktop/repos/ilm
-Traceback (most recent call last):
-  File "create_ilm_examples.py", line 158, in <module>
-    limit=args.max_num_documents)
-  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 24, in get_dataset
-    d = arxiv_cs_abstracts(split, *args, data_dir=data_dir, **kwargs)
-  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 63, in arxiv_cs_abstracts
-    with open(os.path.join(data_dir, 'arxiv_cs_abstracts.txt'), 'r') as f:
-FileNotFoundError: [Errno 2] No such file or directory: '/Users/kelichiu/Desktop/repos/ilm/data/raw_data/arxiv_cs_abstracts/arxiv_cs_abstracts.txt'
-Traceback (most recent call last):
-  File "create_ilm_examples.py", line 158, in <module>
-    limit=args.max_num_documents)
-  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 24, in get_dataset
-    d = arxiv_cs_abstracts(split, *args, data_dir=data_dir, **kwargs)
-  File "/Users/kelichiu/Desktop/repos/ilm/ilm/datasets.py", line 63, in arxiv_cs_abstracts
-    with open(os.path.join(data_dir, 'arxiv_cs_abstracts.txt'), 'r') as f:
-FileNotFoundError: [Errno 2] No such file or directory: '/Users/kelichiu/Desktop/repos/ilm/data/raw_data/arxiv_cs_abstracts/arxiv_cs_abstracts.txt'
+3.2. Unzip the file and rename it to `arxiv_cs_abstracts.txt`, move it to `repo/data/raw_data/arxiv_cs_abstracts/`
 
+4.
+The following script was provided to train the model (fine-tune GPT-2) by the training examples.
+It requires us saving the script as an sh file in order to execute it.
+I saved it in the repo root directory as training.sh.
+Execute it by `./training.sh`:
+
+``` {sh, eval = FALSE}
+DATASET=arxiv_cs_abstracts
+TRAIN_DIR=train
+EXAMPLES_DIR=data/char_masks/${DATASET}
+python train_ilm.py \
+	experiment_${DATASET} \
+	${TRAIN_DIR} \
+	${EXAMPLES_DIR} \
+	--seed 0 \
+	--train_examples_tag train \
+	--eval_examples_tag valid \
+	--eval_max_num_examples 512
 ```
+
+The fine-tuning will start. According to the paper, it will take 1 day to complete with GPU. For a computer that doesn't have a GPU, it will take longer.
 

### 2020-10-11 outputs/documents/notes/model_comparison.Rmd
model comparison notes

---
title: "Model comparison"
author: "Keli Chiu"
date: "05/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## BERT

### Pros

* The Masked Language Model of BERT does text-infilling by its own mechanism, no fine-tuning is needed.

* In contrast to autoregressive models such as GPT-2, BERT uses bidirectional contexts when predicting tokens, leading to improved performance.

### Cons

* The use of bidirectional attention limits its infilling capability to fixed-length spans. (Chris Donahue)

* The [MASK] symbol used by BERT during pre-training are not present in real data at fine-tuning time, resulting in a pretrain-finetune discrepancy.

* BERT is not able to model the joint probability and assumes the predicted tokens are independent of each other.

## GPT-2

### Pros

* GPT-2 does not rely on data corruption and therefore does not suffer from the pretrain-finetune discrepancy like BERT does

### Cons

* The inference time is relatively long.

* According to Caleb Kaiser on Too big to deploy: How GPT-2 is breaking servers, GPT-2 is both compute hungry and memory hungry: "GPT-2 is compute hungry. In order to serve a single prediction, GPT-2 can occupy a CPU at 100% utilization for several minutes. Even with a GPU, a single prediction can still take seconds. ... GPT-2 is memory hungry. Beyond its considerable disk space and compute requirements, GPT-2 also needs large amounts of memory to run without crashing."

* Needs to be fine-tuned for text-infilling task. Fine-tuning would typically take 1 GPU a day.

## XLNet

### Pros

* Bidirectional contexts are taken into account when predicting tokens by calculating all possible permutations of the factorization order.

* XLNet does not rely on data corruption and therefore does not suffer from the pretrain-finetune discrepancy like BERT does.

### Cons

* According to Ajit Rajasekharan's post on Quora dated July 6, 2019, XLNet requires more compute power and memory (GPU/TPU memory) in comparison to BERT.

* Needs to be fine-tuned for text-infilling task. Fine-tuning would typically take 1 GPU a day.

### 2020-10-11 outputs/documents/notes/paper_code_documentation.Rmd
model comparison notes
updated documentation

@@
 
-Not able to follow through the code tutorial from Github. Google Collab code generates demo successfully.
+Able to follow through the code tutorial from Github to the point to fine-tuning. Google Colab code generates demo successfully.
 
@@
 
+The Google Colab demostrated text-infilling by the model after fine-tuning.
 Link: https://colab.research.google.com/drive/1So95M0hHefyNm_eELglCna_ZayoDX6KV?usp=sharing

### 2020-10-18 outputs/documents/paper_bert.Rmd
Pytorch Bert text-infilling

--- 2020-08-21 outputs/documents/paper.Rmd
@@
 
-We consider various natural language processing methods.
-These include n-grams, Word2Cec, GloVe, GPT models and BERT.
-The n-gram approach involves identifying two, three, or more, words that are commonly found together.
-Here think of a two-gram involving the word 'good' 'good morning'.
-At scale these can identify missing or unusual words, and work quickly, but they lack nuance.
-For instance, an equally reasonable two-gram involving the word 'good' is 'good work'.
-For that reason, we consider pre-trained word embedding models.
-These include Word2Vec and GloVe, which place each word in a multi-dimensional space such that distance between words can illustrate their relationship.
-We also consider pre-trained generative models such as GPT-2 and GPT-3 and BERT.
-GPT-2, GPT-3 and BERT are pre-trained generative unsupervised language models.
-GPT-2 and GPT-3 are the two generations of the same model that differ in the number of parameters.
-GPT-2 has 1.5 billion parameters and GPT-3 has more than 175 billion parameters.
-While GPT-2 is more limited, it can be run on smaller machines, whereas GPT-3 cannot and is accessed via an API.
-BERT has 340 millions parameters which is smaller than both GPT-2 and GPT-3.
-BERT also differs from GPT models because its encoding process is bidirectional instead of unidirectional.
+We consider various Transformer language models.
+These include BERT [@devlin2018bert], GPT-2 [@radford2019language] and XLNet [@yang2019xlnet].
+These three models are pre-trained generative unsupervised language models based on the Transformer architecture [@vaswani2017attention].
+We started our experiment with BERT.
+Released by researchers at Google AI Language in 2018, the key innovation of BERT is its application of bidirectional training of the Transformer model, which enables BERT to perform text-infilling without fine-tuning.
 
@@
 The reason that there is such a variety of statistical language models is that there are various ways of dealing with this fundamental problem.
-
-## PDF text extraction and n-grams
-
-The Portable Document Format (PDF) is typically used to render documents in a relatively fixed way.
-PDFs contain a lot of information, but often that information is not able to be analysed directly by statistical language models.
-The information first needs to be extracted from the PDF, often by optical character recognition (OCR), and then cleaned, prepared, and made into some type of dataset.
-One issue in this process is that OCR is not perfect.
-Hence there is a need to correct the output before it is analysed.
-One way to do this is to correct errors one-by-one and independent of context.
-We all know that 'tbe' is wrong and should be 'the'.
-But statistical language models have the advantage of correction that can occur within context (@rosenfeld2000two).
-This allows significant improvements, for instance sometimes 'bat' should be 'hat' and sometimes it really should be 'bat'.
-
-The foundation of an n-gram language model is the conditional probability set-up introduced above.
-An n-gram model is a probabilistic language model that predicts the next word in a sequence of words (@bengio2003neural).
-The $n$ in n-gram refers to the number of words in that sequence.
-Consider the following excerpt from *Jane Eyre*: 'We had been wandering'.
-'We' is a uni-gram, 'We had' is a bi-gram, 'We had been' is a tri-gram, and 'We had been wandering' is a 4-gram.
-Notice that the two tri-grams in this excerpt, 'We had been', and 'had been wandering', overlap.
-The use of n-gram models enables us to assign probabilities to both the next sequence of words and just the next word.
-For instance, consider the two sentences: 'We had been wandering', and 'We had been wangling'.
-The former is likely to be more frequently encountered in a training corpus.
-Thus, an n-gram would assign a higher probability to the next word being 'wandering' than 'wangling', given the sequence 'We had been'.
-
-To predict the next word, we have to take the sequence of preceding words into account, which requires knowing the probability of the sequence of words.
-The probability of a sequence appearing in a corpus follows the chain rule:
-
-$$\mbox{Pr(We,had,been,wandering)} = \mbox{Pr(We)} \times \mbox{Pr(had|We)} \times \mbox{Pr(been|We,had)} \\  
-\times \mbox{Pr(wandering|We,had,been)}.$$
-
-However, the likelihood that more and more words will occur next to each other in an identical sequence becomes smaller and smaller, making prediction difficult.
-Alternatively, we can approximate the probability of a word depending on only the previous word.
-This is known as the 'Markov assumption' and it allows us to approximate the probability using only the last $n$ words [@brown1992class]:
-
-$$\mbox{Pr(We,had,been,wandering)} \approx \mbox{Pr(We)} \times \mbox{Pr(had|We)} \times \mbox{Pr(been|had)} \\ 
-\times \mbox{Pr(wandering|been)}.$$
-
-As a bi-gram model only considers the immediately preceding word, under the Markov assumption, an $n$-gram model can be reduced to a bi-gram model with $n$ being any number:
-
-$$\mbox{Pr}(w_{n} | w^{n-1}_{1}) \approx \mbox{Pr}(w_{n} | w_{n -1})$$
-
-Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation [@brown1992class].
-In our application, we use a tri-gram based model to detect both real-word and non-word errors and correct them with the candidate word that has the highest probability.
-
-However, n-gram models do not take the linguistic structure of language into account.
-For instance, @rosenfeld2000two [p.
-1] discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'.
-The prediction of the next word is based on only a few preceding words and broader context is not taken into account.
-Hence next-word prediction using n-gram based language models can be limited.
-
-There are a variety of ways to implement an n-gram model within R @citeR including using R packages such as `Quanteda` [@citequanteda], `tidyText` [@citetidytext] and `tm` [@citetm].
-We used the Text Analysis Utilities, `tau`, R package @citetau because of we can separate text-cleaning and n-gram making in two different steps.
-We want to have control to manually clean the text because we want to add an end-of-sentence token to each sentence int he corpus.
-To illustrate the use of `tau`, we will first install and load the library, and create a small corpus.
-
-```{r, eval = FALSE}
-install.packages("tau")
-```
-
-```{r, eval=FALSE}
-library(tau)
-corpus <- "There was no possibility of taking a walk that day. We had been 
-wandering, indeed, in the leafless shrubbery an hour in the morning; but 
-since dinner (Mrs. Reed, when there was no company, dined early) the cold 
-winter wind had brought with it clouds so sombre, and a rain so penetrating, 
-that further out-door exercise was now out of the question."
-```
-
-Our purpose is to construct tri-grams and so we first split the corpus into sentences by period, exponential mark, or question marks.
-
-First, we remove all punctuation and other non-letter characters, and convert all text to lower case.
-
-```{r, eval=FALSE}
-tokenizer <- function(corpus) {
-  lines <- vector()
-  # Read the corpus line by line
-  for (line in corpus) {
-    # add two start-of-sentence tokens to the beginning of a sentence
-    lines <-c(lines, line)
-  }
-  lines <- tolower(lines)
-  # add end-of-sentence token to the end of a sentence
-  lines <- gsub("[.!?]$|[.!?] |$", " </s>", lines) 
-  tokens <- unlist(strsplit(lines, "[^a-z<>/]"))
-  tokens <- tokens[tokens != ""]
-  return(tokens)
-}
-```
-
-We will use the `textcnt()` function to turn the training corpus into tri-grams.
-The most important argument of this function is _corpus_, as well as _n_ to set the number of words in a sequence, _method_ to determine which type of counts to compute, and _split_ to designate the regular expression pattern for splitting.
-The function outputs a `textcnt` object, which is essentially a table with each tri-gram and their counts [@citetau].
-
-```{r, eval=FALSE}
-trigrams <- textcnt( tokens, n=3, method = "string", 
-                     split = "[[:space:][:digit:]]+", decreasing = TRUE)
-```
-
-We then manually added an end of sentence token, `</s>`, to the end of each sentence.
-
-```{r, eval=FALSE}
-trigrams <- trigrams[!grepl("</s>", names(trigrams))]
-```
-
-At this point it is possible to take other measures around fine-tuning the tri-grams.
-For instance we may like to remove the tri-grams that contain an end-of-sentence token.
-This would mean that tri-grams that include a word from a new sentence will be removed.
-For example, the tri-gram 'day </s> we' will be removed because it has words from two different sentences.
-Table 1.
-shows four sample tri-grams from the small sample corpus after this process.
-
-```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
-tabl <- "
-| there was no  | a rain so | a walk that | an hour in  |
-|---------------|-----------|-------------|-------------|
-| 2             | 1         | 1           | 1           |
-
-Table 1. Four sample tri-grams generated from the corpus
-"
-cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
-```
-
-The tri-grams can be accessed by calling the names of this table object and the counts can be used to calculate probabilities.
-There are a variety of options around constructing these probabilities.
-One is Kneser–Ney smoothing, which is the smoothing technique recommended by @chen1999empirical to improve the accuracy of n-gram models.
-
-Finally, it is helpful to create a data table that contains the first two words and the last of each tri-gram in separate column and the smoothed probability in another column.
-For instance, Table 2.
-provides and example of this set-up.
-
-```{r table2, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
-tabl <- "
-| First Words   | Last Word | Probability |
-|---------------|-----------|-------------|
-| there was     | no        | 0.6278947   |
-
-Table2. The smoothed probability table
-"
-cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
-```
-
-## Word embedding models
-
-Although, from an implementation perspective, there is an inherent numerical representation of language in n-gram models, it is not overly tractable.
-The numeric representation typically relies on each term being represented in a vector, which can quickly result in quite high-dimensional objects.
-In contrast, word embedding models represent terms by vectors of real numbers where each vector is a 'feature', hence the final object has considerably lower dimensionality [@bengio2003neural].
-
-This approach has a long history, but the implementation of @bengio2003neural has become the foundation for much subsequent work.
-Here, terms are encoded into feature vectors that represent different aspects of the term, and so each term is represented by a position in the vector space.
-Terms that are used in similar ways will have similar representations in the vector space.
-This will be due to their co-occurrence in usage.
-For example, 'cat' will be closer to 'dog' than to 'car', because we are more likely to talk about cats and dogs, than cats and cars.
-See how they might be close when considered along a 'pets' dimension, but far when considered along a 'related-to-tigers' dimension.
-
-Terms that are often used in similar contexts will have similar vector representations.
-Hence, word embedding models can identify similar terms, calculate the similarity between a pair of terms, and find the odd term in a group of terms [@mikolov2013efficient].
-There are a variety of ways to achieve this dimensionality reduction, including Word2Vec and GloVe.
-
-### Word2vec
-
-Word2vec is a word embedding model that uses a neural network to learn word associations [@mikolov2013efficient] [@mikolov2013distributed].
-It takes a corpus as an input and outputs a set of feature vectors that allow the terms in that corpus to be represented.
-@mikolov2013efficient proposed two architectures for determining this distributed representation of terms: continuous bag-of-words (CBOW), and continuous skip-gram.
-Both have an input layer, a projection layer, and an output layer.
-But CBOW uses several surrounding words, or the context, to predict a target word.
-On the other hand, continuous skip-gram takes a single word as input to predict its neighbouring words.
-
-To illustrate the difference, consider the quote from Descartes, 'I think therefore I am'.
-The focus word for which we want to learn the embedding is 'therefore', and the context window is set to two.
-This means that we consider the two words immediately preceding and proceeding 'therefore'.
-The CBOW input is 'I, think, I, am', and the output is 'therefore'.
-In contrast, the continuous skip-gram input is 'therefore', while 'I, think, I, am' is the output.
-The input layer encodes terms within a context window to one-hot vectors.
-Using a projection matrix, the projection layer then project the discrete one-hot vectors to lower-dimension continuous vectors.
-Finally, the output layer is a softmax regression classifier that obtains the posterior distribution of words [@rong2014word2vec].
-It is the input and output vectors that differ between the two architectures.
-
-Word2Vec is trained on the Google News dataset [@citegooglenews].
-This contains six billion tokens and the vocabulary size is restricted to the one million most common words [@mikolov2013efficient].
-As this is a large dataset, it provides an accurate statistical pattern of language, as it is used in this dataset, and a comprehensive mapping between terms and vector positions.
-
-To the extent that the relative positions of terms reflects their semantic relationship, Word2Vec will perform well for word analogies.
-However, terms that have context-specific meaning may not perform as well.
-
-There are a few packages for implementing a word2vec model in R, including `rword2vec` [@citerwordvecchaware] and `word2vec` [@citewordvecwijffels].
-Both packages use the same algorithm.
-Another way to implement a word2vec model in R is through Keras which is an API for deep learning written in Python [@chollet2015keras] for which the `keras` package provides an R interface [@citekeras].
-
-Here, we follow @falbel2017word to implement a skip-gram model of Word2Vec through Keras.
-There are three steps in the Keras deep learning framework: 1) prepare the data; 2) define the model; and 3) train and evaluate the model.
-**(Ke-Li: Can we change this to be a different dataset please?)** Following these steps, we first downloaded the Google News dataset [@citegooglenews] and preprocess the text into tokens.
-
-```{r, eval=FALSE}
-tokenizer %>% fit_text_tokenizer(training_corpus)
-```
-
-We then define a model that mimics the skip-gram algorithm of word2vec using the function `skipgrams_generator()`.
-
-```{r, eval=FALSE}
-skipgrams_generator <- function(text, tokenizer, window_size, negative_samples) {
-  gen <- texts_to_sequences_generator(tokenizer, sample(text))
-  function() {
-    skip <- generator_next(gen) %>%
-      skipgrams(
-        vocabulary_size = tokenizer$num_words, 
-        window_size = window_size, 
-        negative_samples = 1
-      )
-    x <- transpose(skip$couples) %>% map(. %>% unlist %>% as.matrix(ncol = 1))
-    y <- skip$labels %>% as.matrix(ncol = 1)
-    list(x, y)
-  }
-}
-```
-
-The embedding dimensions, the size of the context window and the number of samples are configurable within this function.
-
-Finally, we use the `fit_generator()` function to train the model.
-
-```{r, eval=FALSE}
-model %>%
-  fit_generator(
-    skipgrams_generator(training_corpus, tokenizer, skip_window, negative_samples), 
-    steps_per_epoch = 10, epochs = 1
-  )
-```
-
-**What's supposed to happen:** As a result, a embedding matrix is obtained and serves as the look-up table of the word vectors.
-**What really happened:** However, the training process was stuck in the operation of `fit_generator()` function.
-According to Stack Flow this might be an issue of incompatible versions between Keras and Tensorflow.
-More investigation is needed.
-
-### GloVe
-
-The Global Vectors (GloVe) model is a pre-trained word embedding model that emphasises the semantic meaning of words [@pennington2014glove].
-In the Word2Vec model, word representations are learnt locally within the context window.
-On the other hand, GloVe considers the properties of the corpus globally.
-As the model considers the context of an entire corpus the co-occurrence probabilities are based on counts for the entire training dataset.
-
-GloVe additionally considers the relative probability of word co-occurrence.
-For instance, consider the two words 'ice' and 'steam'.
-GloVe considers their co-occurrence with a variety of probe words, $k$.
-For instance $k$ could be 'solid', which is a property of ice; 'gas', which is a property of steam; and 'water', which is related to both.The ratio of co-occurrence probability is:
-
-$$\frac{\mbox{Pr}(k|\mbox{ice})}{\mbox{Pr}(k|\mbox{steam})}.$$
-
-The ratio between the co-occurrence probability of the 'ice, solid' pair and the 'steam, solid' pair is high because the co-occurrence of 'ice' and 'solid' is higher than the co-occurrence of 'steam' and 'solid'.
-Conversely, the ratio between the co-occurrence of the 'ice, gas' pair and the 'steam, gas' pair is low.
-On the other hand, the ratio of the co-occurrence probability of the 'ice, water' pair and the 'steam, water' pair us close to one because water is relevant to both 'ice' and 'steam'.
-
-As @hanretty2018comparing point out, although word embedding models can capture semantic meanings of words, "they are context-free and fail to capture higher-level concepts in context, such as polysemous disambiguation, syntactic structures, semantic roles, anaphora".
-Word embedding models do not represent a homonyms that has difference meanings in different vectors depending on the context.
-For example, "squash" can be a type of hard shelled vegetable or can be referring to a racket and ball sport, but there is only one series of vectors to represent the word "squash", instead of dynamically changing the vectors depending on the context.
-Moreover, from our observation, the word embedding models are unable to handle novel words that are not included in the vocabulary of the training corpus; for this reason, we conclude that the generalizability of word embedding models is sub-optimal.
-
-There are a variety of ways to implement GloVe word embedding models in R.
-One is to load GloVe embeddings into a frozen Keras Embedding layer [@kerasglove] using the `Keras` R package [@kerasR2020].
-This means the embedding vectors will not be updated during the training process.
-Another way, is to use the `text2vec` package [@text2vec].
-
-Here we follow **CITE** to implement GloVe using the `keras` R package [@kerasglove].
-The blog post is aimed to create a word embedding model for text classification.
-However, we will only use the guide create a embedding index and skip the steps of text classification.
-Instead of building a model for word embedding from scratch as we did for the skip-gram model of Word2Vec, here we are merely loading the Glove model and its pre-trained word vectors as an index.
-The word embedding index is then saved to an R environment.
-
-**[As before, we need code and examples here]**
 
@@
 
-# Stylized example
-
-**(TODO: This stylized example need to be changed and updated to actually use GPT-3 at some point.)**
-
-In our application, we deploy OpenAI GPT-3 to generate text and ...
-
-As a stylized example, let's consider the following actual paragraph from Jane Eyre, by Charlotte Bronte:
-
-> There was no possibility of taking a walk that day.
-We had been wandering, indeed, in the leafless shrubbery an hour in the morning; but since dinner (Mrs.
-Reed, when there was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, that further out-door exercise was now out of the question.
-
-Let's pretend that this text had been created from optical character recognition and that it had the following errors: some 'h' were replaced with 'b'; and some 'd' have been replaced with 'al':
-
-> There was no possibility of taking a walk that day.
-We had been wandering, indeed, in tbe leafless shrubbery an hour in tbe morning; but since dinner (Mrs.
-Reeal, when tbere was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, tbat further out-door exercise was now out of the question.
-
-Assume a model that is trained to perfectly forecast the next word in Jane Eyre.
-For this fragment there are 62 words, comprising 5 errors and 57 correct words.
-So the internal consistency score of this fragment would be: 57/62 = 0.919.
-When we recognise and correct the errors, this consistency score would increase to 1.
-
-Similarly, assume a model that is trained on an external data source.
-This means that it will recognise the the cases where some 'h' were replaced with 'b', but not recognise that 'Reed' has become 'Reeal'.
-Hence, the external consistency score would be 58/62 = 0.935.
-
-# Data
-
-Various data can be used....
-
-# Application
-
-# Discussion
-
-Internal validity vs external- one is their own words the other is a general set of words.
-
-In the same way that precision and recall provide important measures…
-
 # Appendix
 
-## Literature summaries
-
-(These are just here for now - we'll remove later.)
-
-### Article: Improving Language Understanding by Generative Pre-Training (2018)
-Introduction of a semi-supervised approach for language understanding machine learning tasks. The approach is the combinaiton of unsupervised pre-training and supervised fine-tuning.
-
-### Article: Language Models are Few-Shot Learners (2020)
-Official paper from OpenAI to introduce OpenAI GPT-3. Emphasizes on the "fewer-shots" and "task-agnostic" aspects of the latest langual model compared to its predecessor GPT-2.
-
-## Discussion of how GPT-3 works
-
-**(Ke-Li - it would be good for us to go through how it actually works in as much detail as we can manage (in our own words).
-This is mostly just to force us to learn how it works and may be taken out of the final version of the paper, but it's still an important exercise.)**
-
 # References

### 2020-10-18 scripts/model_starters/R/Pytorch-Bert-Text-Infilling.Rmd
Pytorch Bert text-infilling

---
title: "Pytorch-Bert-Text-Infilling"
author: "Keli Chiu"
date: "17/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Calling Python in R by `reticulate`

The pretrained BERT model is from Pytorch, we are loading the necessary modules through `reticulate` library which permites operation of Python modules in R, and `rTorch`, an R interface to Pytorch. It's recommended to create a conda python environment for rTorch. This part is done in command line. `my-torch` is the name of the environment that we are going to tell R later, you can change it to something else.

``` {sh, eval = FALSE}
### Create a conda environment
conda create -n my-torch python=3.7 -y

### Activate the new environment with 
conda activate my-torch

### Inside the new environment, install PyTorch and related packages with:

conda install python=3.6 pytorch torchvision matplotlib pandas -c pytorch
```

Back to R, we load the `reticulate` library and tell it to use `my-torch` as the python environment. After that, we will import all the necessary modules and models.

``` {r}
library(reticulate)
use_condaenv("my-torch", required = TRUE) # Specify the python environment

pytorch_pretrained_bert <- import("pytorch_pretrained_bert")
BertModel <- pytorch_pretrained_bert$BertModel
BertForMaskedLM <- pytorch_pretrained_bert$BertForMaskedLM
BertTokenizer <- pytorch_pretrained_bert$BertTokenizer
tokenizer <- BertTokenizer$from_pretrained('bert-base-uncased')
model <- BertForMaskedLM$from_pretrained('bert-base-uncased')
```

## Load and clean text (to be included in the function)

The ground truth text is:

> "His Majesty Government wish to add that they have no intention of requesting the establishment of military bases in peace time within the area of Palestine now united to the Kingdom of Jordan."

The words "intention" and "Kingdom" are turned into errors:

> "His Majesty Government wish to add that they have no im@@@l@fon of requesting the establishment of military bases in peace time within the area of Palestine now united to the Kis@@®m of Jordan."

We load the text with errors and clean it by removing digits and special characters. "?", ".", "!" are kept as they help BERT to do the segmentation of sentences.

``` {r}
### Get text
text_original <- c("His Majesty Government wish to add that they have no im@@@l@fon of 
                   requesting the establishment of military bases in peace time within 
                   the area of Palestine now united to the Kis@@®m of Jordan.")

### Clean text
library(textclean)
text_stripped <- strip(text_original, char.keep = c("?", ".", "!"), digit.remove = TRUE, 
                       apostrophe.remove = FALSE, lower.case = FALSE)

text_stripped
```

## Identify errors and encode them (to be included in the function)

Right now, the error detection is done through a spellchecker as a start, real-word errors shall be tackled in the future. Once we have a list of errors, we replace them with a `[MASK]` token, which BERT recognizes as the token to be predicted.

``` {r}
### Collect text errors
library(hunspell)
errors <- hunspell(text_stripped)

### Replace errors to [MASK]
for(e in errors){
  pats <- paste(e, collapse = '|')
}
library(stringr)
text <- str_replace_all(text_stripped, pats, '[MASK]')

text
```

## Tokenize the text and encode them (to be included in the function)

We then tokenize the text and encode them. There are two layers of encoding:

* `segments_ids` is the encoding of the sentences

* `indexed_tokens` is the encoding of each word in the sentences

We then turn the above two into tensors as BERT's input by the function `torch$tensor`, namely `segments_tensors` and `tokens_tensor`.

``` {r}
### Tokenize the text
tokenized_text <- tokenizer$tokenize(text)

### Encode the text and create segments tensors
indexed_tokens <- tokenizer$convert_tokens_to_ids(tokenized_text)
SEGS <- which(tokenized_text == ".")
segments_ids <- rep(as.integer(0), times = SEGS)

### Get the position of the [MASK] tokens
MASKIDS <- which(tokenized_text == "[MASK]")

### Make the encoding into tensors as BERT inputs
library(rTorch)
segments_tensors <- torch$tensor(list(segments_ids))
tokens_tensor <- torch$tensor(list(indexed_tokens))

segments_tensors
tokens_tensor
```

## Model preduciton (to be included in the function)

We feed the `tokens_tensor` and `segments_tensors` in the model to produce prediction tensor.

``` {r}
### Generate prediction
py <- torch$no_grad()
with(py, {
  predictions <- model(tokens_tensor, segments_tensors)
})

predictions
```

## Prediction function

Here is a half-ly constructed function to produce the text prediction, the above scripts are to included to make the function complete. Here, we replace `[MASK]` token in the text to the predicted words, which are "intention" and "area".

``` {r, warning = FALSE}
### Prediction function
predict_tokens <- function(text, predictions, MASKIDS) {
  predicted_token <- list()
  for(i in 1:length(MASKIDS)){
    preds <- torch$topk(predictions[[0]][MASKIDS[i]], k=as.integer(50))
    indices <- preds$indices$tolist()
    list <- tokenizer$convert_ids_to_tokens(indices)
    predicted_token <- append(predicted_token, list(list[1]))
  }
  predicted_token.i <- 1

  for (i in 1:length(text)) {
    while (grepl(pattern = "\\Q[MASK]\\E", text[i])) {
      text[i] <- sub(pattern = "\\Q[MASK]\\E", replacement = predicted_token[predicted_token.i], x = text[i])
      predicted_token.i <- predicted_token.i + 1
    }
  }

  return(text)
}

predict_tokens(text, predictions, MASKIDS)
```

### 2020-10-18 scripts/00-try-BERT-text-prediction.R

@@ -2,7 +2,7 @@
 Sys.setenv(TF_KERAS=1)
 # to see python version
 reticulate::py_config()
-setwd("~/Desktop/repos/consistency_scores_in_text_datasets/model_starters")
+setwd("~/Desktop/repos/consistency_scores_in_text_datasets/scripts/model_starters")
 
 #### 1. Read text from image
 library(tesseract)

### 2020-10-18 scripts/model_starters/R/.Rprofile.R

@@ -1,0 +1,8 @@
+transformers <- import("transformers")
+DistilBertForMaskedLM <- transformers$DistilBertForMaskedLM
+DistilBertModel <- transformers$DistilBertModel
+DistilBertTokenizer <- transformers$DistilBertTokenizer
+tokenizer <- DistilBertTokenizer$from_pretrained('distilbert-base-uncased')
+
+# pip install "transformers==3.1.0"
+model <- DistilBertModel$from_pretrained('distilbert-base-uncased', return_dict=TRUE)

### 2020-10-18 scripts/model_starters/R/BERT_pytorch.R

@@ -1,0 +1,75 @@
+library(reticulate)
+use_condaenv("my-torch", required = TRUE)
+
+pytorch_pretrained_bert <- import("pytorch_pretrained_bert")
+BertModel <- pytorch_pretrained_bert$BertModel
+BertForMaskedLM <- pytorch_pretrained_bert$BertForMaskedLM
+BertTokenizer <- pytorch_pretrained_bert$BertTokenizer
+tokenizer <- BertTokenizer$from_pretrained('bert-base-uncased')
+
+# pip install "transformers==3.1.0"
+model <- BertForMaskedLM$from_pretrained('bert-base-uncased')
+
+#### 1. Get text
+text_original <- c("His Majesty Government wish to add that they have no im@@@l@fon of requesting the establishment of military bases in peace time within the area of Palestine now united to the Kis@@®m of Jordan.")
+
+#### 2. Clean text
+library(textclean)
+text_stripped <- strip(text_original, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
+                      lower.case = FALSE)
+
+#### 3. Collect text errors
+library(hunspell)
+errors <- hunspell(text_stripped)
+
+### 4. Replace errors to [MASK]
+for(e in errors){
+  pats <- paste(e, collapse = '|')
+}
+library(stringr)
+text <- str_replace_all(text_stripped, pats, '[MASK]')
+
+### 5. Tokenize the text
+tokenized_text <- tokenizer$tokenize(text)
+indexed_tokens <- tokenizer$convert_tokens_to_ids(tokenized_text)
+
+### 6. Encode the text and create segments tensors
+MASKIDS <- which(tokenized_text == "[MASK]")
+SEGS <- which(tokenized_text == ".")
+
+segments_ids <- rep(as.integer(0), times = SEGS)
+
+library(rTorch)
+segments_tensors <- torch$tensor(list(segments_ids))
+tokens_tensor <- torch$tensor(list(indexed_tokens))
+
+### 7. Generate prediction
+py <- torch$no_grad()
+with(py, {
+  predictions <- model(tokens_tensor, segments_tensors)
+})
+
+### 8. Prediction function
+predict_tokens <- function(text, predictions, MASKIDS) {
+  predicted_token <- list()
+  for(i in 1:length(MASKIDS)){
[…]

### 2020-10-26 scripts/model_starters/R/Pytorch-Bert-Text-Infilling.Rmd
update Pytorch-Bert-Text-Infilling

@@
 ```
+
+## Updated prediction unction
+
+```{r, warning=FALSE}
+##### Prediction function
+predict_tokens <- function(text_original) {
+  ### Clean text
+  text_stripped <- strip(text_original, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
+                         lower.case = FALSE)
+  ### Collect text errors
+  errors <- hunspell(text_stripped)
+
+  ### Replace errors to [MASK]
+  for(e in errors){
+    pats <- paste(e, collapse = '|')
+  }
+  text <- str_replace_all(text_stripped, pats, '[MASK]')
+
+  ### Tokenize the text
+  tokenized_text <- tokenizer$tokenize(text)
+  indexed_tokens <- tokenizer$convert_tokens_to_ids(tokenized_text)
+
+  ### Encode the text and create segments tensors
+  MASKIDS <- which(tokenized_text == "[MASK]")
+  SEGS <- which(tokenized_text == ".")
+
+  segments_ids <- rep(as.integer(0), times = SEGS)
+
+  segments_tensors <- torch$tensor(list(segments_ids))
+  tokens_tensor <- torch$tensor(list(indexed_tokens))
+
+  ### Generate prediction
+  py <- torch$no_grad()
+  with(py, {
+    predictions <- model(tokens_tensor, segments_tensors)
+  })
+
+  predicted_token <- list()
+  for(i in 1:length(MASKIDS)){
+    preds <- torch$topk(predictions[[0]][MASKIDS[i]], k=as.integer(50))
+    indices <- preds$indices$tolist()
+    list <- tokenizer$convert_ids_to_tokens(indices)
+    predicted_token <- append(predicted_token, list(list[1]))
+  }
+  predicted_token.i <- 1
+
+  for (i in 1:length(text)) {
+    while (grepl(pattern = "\\Q[MASK]\\E", text[i])) {
+      text[i] <- sub(pattern = "\\Q[MASK]\\E", replacement = predicted_token[predicted_token.i], x = text[i])
+      predicted_token.i <- predicted_token.i + 1
+    }
+  }
+
+  return(text)
+}
+
+predict_tokens(text_original)
+```

### 2020-10-26 scripts/model_starters/R/Pytorch-Bert-Text-Infilling.Rmd
rename function

@@
 ##### Prediction function
-predict_tokens <- function(text_original) {
+get_correct_text <- function(text_original) {
   ### Clean text
@@
 
-predict_tokens(text_original)
+get_correct_text(text_original)
 ```

### 2021-01-06 outputs/documents/paper 2.Rmd
Update ngrams paper

--- 2020-08-21 outputs/documents/paper.Rmd
@@
 
+To help others use and build on our work, we construct a Shiny app available at: **(Rohan add link)**.
+That app computes internal and external consistency scores for corpus excerpts.
+We have developed an R Package that allows our approach to be used on larger datasets, which is available at: **(Rohan add link)**.
 The remainder of our paper is structured as follows: **(Rohan add structure)**.
-Additionally, we construct a Shiny app available at: **(Rohan add link)**.
-That app computes internal and external consistency scores for corpus excerpts, and have developed an R Package that allows our approach to be used on larger datasets, which is available at: **(Rohan add link)**.
 
@@
 
-## Language Model
+## Language models
 

### 2021-01-06 outputs/documents/paper_ngrams.Rmd
minor update in ngram paper
Update ngrams paper

@@
     fig_caption: true
+    numbering: true
     latex_engine: pdflatex
@@
 Datasets can have errors, be biased, incomplete, or messy.
-In any case, it is the underlying statistical process of which the dataset is an artifact that is typically of interest.
+In any case, it is the underlying statistical process, of which the dataset is an artifact, that is typically of interest.
 In order to use statistical models to understand that process, we typically need to clean and prepare the dataset in some way.
@@
 Hence if the sentence was actually 'the cat in the bat', then that child would know something was likely wrong.
-The consistency score would likely be lower than if the sentence were 'the cat in the hat'.
+The consistency score would likely be lower than if the sentence were 'the cat in the hat' because we expected 'hat', but actually had 'bat'.
 After we correct this error, the consistency score would likely increase.
 By examining how consistency scores evolve in response to changes made to the text during the data preparation and cleaning stages we can better understand the effect of the changes.
-Including consistency scores and their evolution when text corpuses are shared allows researchers to be more transparent about their corpus.
+Including consistency scores and their evolution when datasets are shared allows researchers to be more transparent about the effect of their decisions in these steps of the analysis workflow.
 And finally, the use of consistency scores allows for an increased level of automation in the cleaning process.
@@
 For instance, if using a sequence of three words, or tri-grams, then we identify the first two words in the sequence and evaluate if the last word in the sequence is 'as expected' by comparing it to the corresponding tri-gram in a comparison dataset.
-The sequences we employed are a mixture of five, four and three consecutive words also known as 5-grams, 4-grams and tri-grams.
+The sequences we employed are a mixture of three, four, and five, consecutive words also known as tri-grams, 4-grams, and 5-grams.
 
@@
 The reason that there is such a variety of statistical language models is that there are various ways of dealing with this fundamental problem.
-N-gram is one of the statistical language models that works by taking the co-occurrence of words in sequence into account.
+N-gram models are statistical language models that work by taking the co-occurrence of words in sequence into account.
 
@@
 The foundation of an n-gram language model is the conditional probability set-up introduced above.
-An n-gram model is a probabilistic language model that predicts the next word in a sequence of words (@bengio2003neural).
+An n-gram model is a probabilistic language model that predicts the next word in a sequence of words [@bengio2003neural].
 The $n$ in n-gram refers to the number of words in that sequence.
@@
 The former is likely to be more frequently encountered in a training corpus.
-Thus, an n-gram would assign a higher probability to the next word being 'wandering' than 'wangling', given the sequence 'We had been'.
+Thus, a 4-gram would assign a higher probability to the next word being 'wandering' than 'wangling', given the sequence 'We had been'.
 
@@
 
-In our application, we use a model that has a mixture of 5-grams, 4-grams and tri-grams to provide word replacements to a text error.
+In our application, we use a model that has a mixture of tri-grams, 4-grams, and 5-grams to provide word replacements to a text error.
 Longer n-grams such as 5-grams capture more context of word sequences and potentially generate more accurate predictions.
-However, there is less probability for a longer sequence of words to fit a specific combination and leads to data sparsity.
+However, there is less chance that a longer sequence of words will fit a specific combination which can lead to sparsity.
 Therefore, shorter n-grams should still be included to preserve generalizability of the predictions.
 
-Consider the common English idiom "elephant in the room".
-If by mistake, the idiom is written as "elephant in the roon" and correction is needed for the last word, we can apply both a tri-gram model or a 4-gram model to predict the candidate word:
+Consider the idiom "elephant in the room".
+If by mistake, this is written as "elephant in the roon" and correction is needed for the last word, we can apply both a tri-gram model or a 4-gram model to predict the candidate word:
 
@@
 
-In the 4-gram model, one extra more word "elephant" is included and thus provides more context for the word "room" to be accurately predicted.
-However, there are times people would simply describe the location of an elephant; for example, "elephant in the pond".
-The combination of this sequence as a 4-gram token might not be in a comparison dataset and no prediction would be available if only the 4-gram models are used.
-Therefore, we have to also rely on the tri-gram models to provide other word candidates including "in the pond".
+In the 4-gram model, one more word, 'elephant', is included, which provides more context and increased chance of 'room' being accurately predicted.
+However, there are times when people may literally be describing the location of an elephant; for instance, 'elephant in the pond'.
+The combination of this sequence as a 4-gram token might not be in a training dataset and no prediction would be available if only 4-gram models are used.
+Hence the use of tri-gram models to provide other word candidates including "in the pond".
 
-# 3. The aRianna R package
+# 3. The `aRianna` R package
 
@@
 
-There are a variety of ways to implement an n-gram model within R @citeR including using R packages such as `Quanteda` [@citequanteda], `tidyText` [@citetidytext] and `tm` [@citetm].
-In our package, we used the `Quanteda` R package because it has a comprehensive set of functions for conducting text analysis.
-We also employed `dplyr` [@2018dplyr] for data frame manipulation and `tibble` [@citeTibble] for transforming data frames to the tibble format:
+There are a variety of ways to implement an n-gram model within R [@citeR] including packages such as `quanteda` [@citequanteda], `tidytext` [@citetidytext] and `tm` [@citetm].
+We use `quanteda` because it has a comprehensive set of functions for conducting text analysis.
+We also use `tidyverse` [@citetidyverse] for data manipulation:
 
@@
 install.packages("quanteda")
-install.packages("dplyr")
-install.packages("tibble")
+install.packages("tidyverse")
+library(quanteda)
+library(tidyverse)
 ```
@@
 
-To obtain the internal consistency score, we will construct an internal consistency dataset from a body of text that is the source of the text being evaluated.
-The function `make_internal_consistency_dataset` is created for this purpose.
-The function turns the body_of_text into an internal consistency dataset which the text being evaluated will be compared with.
-To do so, the function first transforms the body of text into word tokens with the punctuation removed and the letters turned to lowercase:
+To obtain the internal consistency score, we will first construct a dataset from the text that is itself the text being evaluated.
+The function `aRianna::make_internal_consistency_dataset()` is created for this purpose.
+The function turns a `body_of_text` into a dataset to compare the text with.
+To do so, the function first transforms the `body_of_text` into word tokens with the punctuation removed and the letters turned to lowercase:
 
@@
 
-Next, the `make_internal_consistency_dataset` function turns the individual tokens into 5-grams, 4-grams and tri-grams, and keeps only the n-grams that appears in the data for more than once.
-The common n-grams are preserved in a tibble:
+Next, the `aRianna::make_internal_consistency_dataset()` function turns the individual tokens into tri-grams, 4-grams, and 5-grams, keeping only the n-grams that appear in the data for more than once.
+These common n-grams are stored in a tibble:
 
@@
 
-Finally, the `make_internal_consistency_dataset` function splits the n-grams to two parts: the previous $n-1$ words and the last word.
+Finally, the `aRianna::make_internal_consistency_dataset()` function splits the n-grams into two parts: the first $n-1$ words and the last word.
 As a result, a tibble is created with each n-gram containing its original sequence, the first words, and the last word:
@@
 
-After we have the internal consistency dataset, we need a function to compare the text being evaluated and the internal dataset in order to retrieve the internal consistency score — the `generate_internal_consistency_score` function.
-Similar to the `make_internal_consistency_dataset` function, the function first transforms the text being evaluated into 5-grams, 4-grams and tri-grams tokens, then splits the n-grams to their first words and last words:
+After we have the internal consistency dataset, we need a function to compare the text being evaluated and the internal dataset in order to retrieve the internal consistency score.
+We define the `aRianna::generate_internal_consistency_score` function to do this.
+Similar to the `aRianna::make_internal_consistency_dataset` function, this function first transforms the text being evaluated into  tri-grams, 4-grams, and 5-grams tokens, then splits the n-grams to their first words and last words:
 
@@
 
-Next, the function combines the text tokens tibble and the internal consistency tibble by the `left_join()` function of `dplyr`.
-The `left_join()` function returns all the rows from the text tokens tibble and all the columns from both tibbles [@2018dplyr].
+Next, the function uses `dplyr::left_join()` to combine the tibble of text tokens and the internal consistency tibble.
+The `dplyr::left_join()` function returns all the rows from the text tokens tibble and all the columns from both tibbles [@2018dplyr].
 By combining the two tibbles, the function generates two additional columns — the `last_word` column are the last words of all n-grams, and the `last_word_expected` contains the last words that are in the internal consistency dataset.
-By comparing the words in `last_word` and `last_word_expected`, we can identify the "unexpected" words.
+By comparing the words in `last_word` and `last_word_expected`, we can identify the 'unexpected' words.
 If a last word is in `last_word` but not `last_word_expected`, it is regarded as an unexpected word and the corresponding word in `last_word_expected` serves as a replacement candidate:
@@
 
-The consistency score is calculated by the number of words that can be predicted by the model divided by the total number of words in the data.
+The consistency score is calculated by the number of words that have been predicted by the model divided by the total number of words in the data.
 The function identifies the counts of words that are predicted by the model and divided by the total number of words in the internal consistency dataset:
@@
 
-The `generate_internal_consistency_score` function also lists the identified text errors and generates replacement candidates to the text errors:
+The `aRianna::generate_internal_consistency_score` function also lists the identified text errors and generates replacement candidates to the text errors:
 
@@
 
-The `generate_external_consistency_score` function works identically as the `generate_internal_consistency_score` does.
-The only difference lies in that it compares the text being evaluated with an external consistency dataset that is larger and more general.
+The `aRianna::generate_external_consistency_score` function works identically to `aRianna::generate_internal_consistency_score`.
+The only difference is that it compares the text being evaluated with an external consistency dataset that is larger and more general.
 The external consistency dataset we employed in the application is constructed using a subset text data from Open WebText Corpus by @Gokaslan2019OpenWeb.
+In general, any dataset could be used.
 
@@
 
-For demonstration, we use the first paragraph of Jane Eyre as the internal text data.
+For this demonstration, we use the first paragraph of Jane Eyre as the internal text data.
 A sentence within the paragraph is modified with the intention to contain an error and serves as the text to be evaluated:
@@
 The first step is to turn the body of text into the internal consistency dataset.
-The generated internal consistency dataset is a tibble that contains the 5-grams, 4-grams and tri-grams that appear in the internal text data more than once.
-Since only `there_was_no` has more than one occurrence, it is the only reference n-gram in the internal consistency dataset:
+The generated internal consistency dataset is a tibble that contains the tri-grams, 4-grams and 5-grams that appear in the internal text data more than once.
+Since only 'there_was_no' has more than one occurrence, it is the only n-gram in the internal consistency dataset:
 
@@
 The next step is to compare the text to be evaluated with the internal consistency dataset that we created in the previous step.
-The function `generate_internal_consistency_score` takes in two arguments — the text to evaluate and the internal consistency dataset.
-The function identifies the word "na" as an unexpected word, and generates the internal consistency score as zero.
-The function also lists "no" as the replacement of "na":
+The function `aRianna::generate_internal_consistency_score()` takes two arguments — the text to evaluate and the internal consistency dataset.
+The function identifies the word 'na' as an unexpected word, and generates the internal consistency score as zero.
+The function also lists 'no' as the replacement of 'na':
 
@@
 
-To get the external consistency score, we will employ the `generate_external_consistency_score` function.
-The function `generate_internal_consistency_score` takes in only one argument — the text to evaluate, and compares it with the external consistency dataset.
-Here, the text to evaluate is "there was no possibiliti", with the word "possibility" being wrongly spelled.
-The function identifies the word "possibiliti" as an unexpected word, and generates the external consistency score 0.833.
-Because the external dataset is larger, it is capable of providing more replacement candidates for "possibiliti".
-The replacements based on 4-grams are displayed first followed by the replacements based on tri-grams:
+To get the external consistency score, we use the `aRianna::generate_external_consistency_score()` function.
+The function `aRianna::generate_internal_consistency_score()` takes in only one argument — the text to evaluate, and compares it with the external consistency dataset.
+Here, the text to evaluate is 'there was no possibiliti', with the word "possibility" being wrongly spelled.
+The function identifies the word 'possibiliti' as an unexpected word, and generates the external consistency score 0.833.
+Because the external dataset is larger, it is capable of providing more replacement candidates for 'possibiliti'.
+There are no replacements based on 5-grams, so the replacements based on 4-grams are displayed first, followed by the replacements based on tri-grams:
 
 ```{r, eval=FALSE}
+text_to_evaluate <- "there was no possibiliti"
 aRianna::generate_external_consistency_score(text_to_evaluate)
@@
 
+Outside of textbook examples, almost all real-world datasets need to be cleaned and prepared before they can be analyzed.
+While there are many tools to help with reproducibility of analysis, options for ensuring reproducibility in earlier stages of a typicaly data science workflow are more limited.
+Our approach discussed in this paper can help.
+Our notion of consistency can be tracked at many points while the dataset is being cleaned and prepared.
+Stages in which it changes substantially can be investigated.
+While there may be good reasons for such as change, by identifying them it is easier for other researchers to understand the effects of cleaning and preparation on eventual analysis results.
+
+Our approaches are applicable to any dataset, but in this paper we applied them to one based on text.
+The high-dimensionality of text, and the messiness of it when it comes from real-world sources, means that often considerable cleaning and preparation is done to text datasets.
+We demonstrated how sequences of words can be used to forecast the very next word.
+That forecast can be compared with the actual word observed and this difference expressed as a proportion, that we call a 'consistency score'.
+
+The language model that we used is n-grams.
 Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation [@brown1992class].
@@
 Hence next-word prediction using n-gram based language models can be limited.
-Here think of a two-gram involving the word 'good' 'good morning'.
+Here think of a two-gram involving the word 'good', for instance, 'good morning'.
 At scale these can identify missing or unusual words, and work quickly, but they lack nuance.
 For instance, an equally reasonable two-gram involving the word 'good' is 'good work'.
-However, because n-gram models does not take the surrounding text into account, it has not capacity of judging if 'morning' or 'work' is to be suggested as the next word after 'good'.
+However, because n-gram models do not take the surrounding text into account, they do not have the capacity to judge if 'morning' or 'work' should be suggested as the next word after 'good'.
 
-In future works, we intend to investigate in more advanced language models such as word embeddings and Transformers.
-We consider pre-trained word embedding models including Word2Vec [@mikolov2013efficient] [@mikolov2013distributed] and GloVe [@pennington2014glove], which place each word in a multi-dimensional space such that distance between words can illustrate their relationship [@bengio2003neural].
+In future work we intend to investigate in more advanced language models such as word embeddings and transformers.
+We will consider pre-trained word embedding models including Word2Vec [@mikolov2013efficient] [@mikolov2013distributed] and GloVe [@pennington2014glove], which place each word in a multi-dimensional space such that distance between words can illustrate their relationship [@bengio2003neural].
 This approach of representing words in vectors has a long history, but the implementation of @bengio2003neural has become the foundation for much subsequent work.
-Here, terms are encoded into feature vectors that represent different aspects of the term, and so each term is represented by a position in the vector space.
 
-We also consider incorporating generative pre-trained transformer models such as GPT-2 [@radford2019language], GPT-3 [@brown2020language], and BERT [@devlin2018bert].
-GPT-2, GPT-3 and BERT are pre-trained language models that are built based on the Transformer architecture [@vaswani2017attention].
-Proposed by @vaswani2017attention from Google in 2017, the Transformer is a novel network architecture for neural network that outperforms RNN-based and CNN-based models in computational efficiency [@vaswani2017attention].
-Since the emergence of the Transformer model, most of the representative pre-trained models are built based on this architecture [@hanretty2018comparing].
-Therefore, we can say that Transformer marks a new era of language models.
-In our future works, it will be important for us to implement different language models compare their effectiveness in the tasks of generating consistency score and providing suggested text corrections.
+We will also consider incorporating generative pre-trained transformer models such as GPT-2 [@radford2019language], GPT-3 [@brown2020language], and BERT [@devlin2018bert].
+These models are based on the transformer architecture [@vaswani2017attention].
+Proposed by @vaswani2017attention in 2017, the transformer network architecture for neural networks has been found to outperform both  recurrent neural network-based and convolutional neural network-based models in computational efficiency [@vaswani2017attention].
+Since the emergence of the transformer model, most of the representative pre-trained models are built based on this architecture [@hanretty2018comparing].
+
+Although the consistency score itself is not of major importance, compared with its changes, improving the score by using more sophisticated models would be considerable benefits in terms of identifying incorrect words.
+In our future work we are interested in implementing different language models to compare their effectiveness in the task of generating consistency scores and providing suggested text corrections.
 We intend to use our future findings to continuously improve the `aRianna` package.

### 2021-01-06 model starters/BERT.R

@@ -1,0 +1,60 @@
+# Source tutorial: https://blogs.rstudio.com/ai/posts/2019-09-30-bert-r/
+Sys.setenv(TF_KERAS=1) 
+# to see python version
+reticulate::py_config()
+setwd("~/Desktop/repos/consistency_scores_in_text_datasets/model starters")
+
+# 1. Read text from image
+library(tesseract)
+eng <- tesseract("eng")
+text_original <- tesseract::ocr("BERT Python/image1.png", engine = eng)
+cat(text_original)
+
+# 2. Collect incorrect words
+library(textclean)
+text_stripped <- strip(text_original, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
+                       lower.case = FALSE)
+library(hunspell)
+incorrectwords <- hunspell(text_stripped)
+
+# 3. Configure paths to the BERT pre-trained files variables
+### The zip file is available through the tutorial link at the first line
+pretrained_path = 'uncased_L-12_H-768_A-12'
+config_path = file.path(pretrained_path, 'bert_config.json')
+checkpoint_path = file.path(pretrained_path, 'bert_model.ckpt')
+vocab_path = file.path(pretrained_path, 'vocab.txt')
+seq_length = NULL
+
+# 4. Load BERT and tokenize the text
+library(reticulate)
+k_bert = import('keras_bert')
+token_dict = k_bert$load_vocabulary(vocab_path)
+tokenizer = k_bert$Tokenizer(token_dict)
+model = k_bert$load_trained_model_from_checkpoint(
+  config_path,
+  checkpoint_path,
+  seq_len=seq_length,
+  training=T)
+
+# 5. Tokenize the text and get the encoding
+tokens <- tokenizer$tokenize(text_original)
+encodes <- tokenizer$encode(text_original)
+
+# 6. Get encoded inputs for the model: train, segments, targets
+indices = encodes[1]
+segments = encodes[2]
+## Manually flag word 13 and word 33 as masked words to be predicted 
+masks = list(list(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0,
+                  0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
+                  0, 0, 0, 0, 0, 0))
+train = do.call(cbind,indices) %>% t()
[…]

### 2021-01-06 model starters/BERT2.R

@@ -1,0 +1,54 @@
+# Source tutorial: https://blogs.rstudio.com/ai/posts/2019-09-30-bert-r/
+Sys.setenv(TF_KERAS=1) 
+# to see python version
+reticulate::py_config()
+setwd("~/Desktop/repos/consistency_scores_in_text_datasets/model starters")
+
+# Read text from image
+library(tesseract)
+eng <- tesseract("eng")
+text_original <- tesseract::ocr("BERT Python/image1.png", engine = eng)
+cat(text_original)
+
+# Collect incorrect words
+library(textclean)
+text_stripped <- strip(text_original, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
+      lower.case = FALSE)
+library(hunspell)
+incorrectwords <- hunspell(text_stripped)
+
+# Configure paths to the pre-trained files variables
+pretrained_path = 'uncased_L-12_H-768_A-12'
+config_path = file.path(pretrained_path, 'bert_config.json')
+checkpoint_path = file.path(pretrained_path, 'bert_model.ckpt')
+vocab_path = file.path(pretrained_path, 'vocab.txt')
+seq_length = NULL
+
+# Load BERT and tokenize the text
+library(reticulate)
+k_bert = import('keras_bert')
+token_dict = k_bert$load_vocabulary(vocab_path)
+tokenizer = k_bert$Tokenizer(token_dict)
+encodes <- tokenizer$encode(text_original)
+model = k_bert$load_trained_model_from_checkpoint(
+  config_path,
+  checkpoint_path,
+  seq_len=seq_length,
+  training=T)
+
+tokens <- tokenizer$tokenize(text_original)
+length(tokens)
+
+# Get encoded inputs for the model: train, segments, targets
+indices = encodes[1]
+segments = encodes[2]
+masks = list(list(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0,
+                  0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
+                  0, 0, 0, 0, 0, 0))
+train = do.call(cbind,indices) %>% t()
+segments = do.call(cbind,segments) %>% t()
+targets = do.call(cbind,masks) %>% t()
+
+# Get the prediction matrix
+prediction <- model$predict(c(train))
+

### 2021-01-06 model starters/GPT2.R

@@ -1,0 +1,6 @@
+# Source tutorial: https://blogs.rstudio.com/ai/posts/2019-10-23-gpt-2/
+remotes::install_github("r-tensorflow/gpt2")
+gpt2::install_gpt2(envname = "r-gpt2")
+library(gpt2)
+gpt2(prompt = "GPT-3 thinks GPT-2 is", model = "124M", seed = NULL, batch_size = 1, total_tokens = NULL,
+     temperature = 1, top_k = 0, top_p = 1)

### 2021-01-06 model starters/Word2Vec.R

@@ -1,0 +1,68 @@
+# Source tutorial: https://blogs.rstudio.com/ai/posts/2017-12-22-word-embeddings-with-keras/
+
+library(readr)
+library(stringr)
+
+# Using a short text for demonstration
+training_corpus <- "There was no possibility of taking a walk that day. We had been wandering, indeed, in the leafless shrubbery an hour in the morning. 
+But since dinner (Mrs Reed, when there was no company, dined early). The cold winter wind had brought with it clouds so sombre, and a rain so penetrating, 
+that further out-door exercise was now out of the question."
+
+library(keras)
+tokenizer <- text_tokenizer(num_words = 20000)
+tokenizer %>% fit_text_tokenizer(training_corpus)
+
+library(reticulate)
+library(purrr)
+skipgrams_generator <- function(text, tokenizer, window_size, negative_samples) {
+  gen <- texts_to_sequences_generator(tokenizer, sample(text))
+  function() {
+    skip <- generator_next(gen) %>%
+      skipgrams(
+        vocabulary_size = tokenizer$num_words, 
+        window_size = window_size, 
+        negative_samples = 1
+      )
+    x <- transpose(skip$couples) %>% map(. %>% unlist %>% as.matrix(ncol = 1))
+    y <- skip$labels %>% as.matrix(ncol = 1)
+    list(x, y)
+  }
+}
+
+embedding_size <- 128  # Dimension of the embedding vector.
+skip_window <- 5       # How many words to consider left and right.
+num_sampled <- 1       # Number of negative examples to sample for each word.
+
+input_target <- layer_input(shape = 1)
+input_context <- layer_input(shape = 1)
+
+embedding <- layer_embedding(
+  input_dim = tokenizer$num_words + 1, 
+  output_dim = embedding_size, 
+  input_length = 1, 
+  name = "embedding"
+)
+
+target_vector <- input_target %>% 
+  embedding() %>% 
+  layer_flatten()
+
+context_vector <- input_context %>%
+  embedding() %>%
+  layer_flatten()
+
+dot_product <- layer_dot(list(target_vector, context_vector), axes = 1)
+output <- layer_dense(dot_product, units = 1, activation = "sigmoid")
+
+model <- keras_model(list(input_target, input_context), output)
+model %>% compile(loss = "binary_crossentropy", optimizer = "adam")
+
+summary(model)
+
+#model <- keras::train_on_batch(model)
+
+model %>%
+  fit(
[…]

### 2021-01-06 shinyapp/functions.R

@@ -1,0 +1,120 @@
+library(tidyverse)
+library(tau)
+library(tm)
+library(stringr)
+library(stats)
+library(reshape2)
+library(RWeka)
+
+#setwd("~/Desktop/repos/consistency_scores_in_text_datasets/shinyapp")
+path <- getwd()
+demoTraining <- readLines(paste(path,"/samples/demo.txt", sep=""), warn=FALSE)
+
+tokenizer <- function(corpus) {
+  lines <- vector()
+  # Read the corpus line by line
+  for (line in corpus) {
+    # add two start-of-sentence tokens to the beginning of a sentence
+    #line <- paste("<s> <s>", line)
+    lines <-c(lines, line)
+  }
+  lines <- tolower(lines)
+  lines <- gsub("[.!?]$|[.!?] |$", " </s>", lines) # add end-of-sentence token to the end of a sentence
+  tokens <- unlist(strsplit(lines, "[^a-z<>/]"))
+  tokens <- tokens[tokens != ""]
+  return(tokens)
+}
+
+tokens <- tokenizer(demoTraining)
+
+library(data.table)
+unigrams <- textcnt( tokens, n=1, method = "string", split = "[[:space:][:digit:]]+", decreasing = TRUE)
+bigrams  <- textcnt( tokens, n=2, method = "string", split = "[[:space:][:digit:]]+", decreasing = TRUE)
+trigrams <- textcnt( tokens, n=3, method = "string", split = "[[:space:][:digit:]]+", decreasing = TRUE)
+
+# Remove trigram that has "</s>" in it, so the trigrams are constructed within a sentence
+unigrams <- unigrams[!grepl("</s>", names(unigrams))]
+bigrams <-   bigrams[!grepl("</s>", names(bigrams))]
+trigrams <- trigrams[!grepl("</s>", names(trigrams))]
+
+### ngram probability
+getLastWords <- function(string, words) {
+  pattern <- paste("[a-z']+( [a-z']+){", words - 1, "}$", sep="")
+  return(substring(string, str_locate(string, pattern)[,1]))
+}
+
+removeLastWord <- function(string) {
+  sub(" [a-z']+$", "", string)
+}
+
+# Kneser-Ney Smoothing
+kneserNay <- function(ngrams, d) {
+  n <- length(strsplit(names(ngrams[1]), " ")[[1]])
+  # Special case for unigrams
+  if(n==1) {
+    noFirst <- unigrams[getLastWords(names(bigrams), 1)]
[…]

### 2021-01-06 shinyapp/text-input.R

@@ -1,0 +1,1 @@
+demoInput <- "There was no possibility of taking a walk that day. We had been wandering, indeed, in tbe leafless shrubbery an hour in tbe morning; but since dinner (Mrs. Reeal, when tbere was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, tbat further out-door exercise was now out of the question."

### 2021-01-06 model starters/BERT.R
repo clean-up

@@ -1,60 +1,0 @@
-# Source tutorial: https://blogs.rstudio.com/ai/posts/2019-09-30-bert-r/
-Sys.setenv(TF_KERAS=1) 
-# to see python version
-reticulate::py_config()
-setwd("~/Desktop/repos/consistency_scores_in_text_datasets/model starters")
-
-# 1. Read text from image
-library(tesseract)
-eng <- tesseract("eng")
-text_original <- tesseract::ocr("BERT Python/image1.png", engine = eng)
-cat(text_original)
-
-# 2. Collect incorrect words
-library(textclean)
-text_stripped <- strip(text_original, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
-                       lower.case = FALSE)
-library(hunspell)
-incorrectwords <- hunspell(text_stripped)
-
-# 3. Configure paths to the BERT pre-trained files variables
-### The zip file is available through the tutorial link at the first line
-pretrained_path = 'uncased_L-12_H-768_A-12'
-config_path = file.path(pretrained_path, 'bert_config.json')
-checkpoint_path = file.path(pretrained_path, 'bert_model.ckpt')
-vocab_path = file.path(pretrained_path, 'vocab.txt')
-seq_length = NULL
-
-# 4. Load BERT and tokenize the text
-library(reticulate)
-k_bert = import('keras_bert')
-token_dict = k_bert$load_vocabulary(vocab_path)
-tokenizer = k_bert$Tokenizer(token_dict)
-model = k_bert$load_trained_model_from_checkpoint(
-  config_path,
-  checkpoint_path,
-  seq_len=seq_length,
-  training=T)
-
-# 5. Tokenize the text and get the encoding
-tokens <- tokenizer$tokenize(text_original)
-encodes <- tokenizer$encode(text_original)
-
-# 6. Get encoded inputs for the model: train, segments, targets
-indices = encodes[1]
-segments = encodes[2]
-## Manually flag word 13 and word 33 as masked words to be predicted 
-masks = list(list(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0,
-                  0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
-                  0, 0, 0, 0, 0, 0))
-train = do.call(cbind,indices) %>% t()
[…]

### 2021-01-06 model starters/BERT2.R

@@ -1,54 +1,0 @@
-# Source tutorial: https://blogs.rstudio.com/ai/posts/2019-09-30-bert-r/
-Sys.setenv(TF_KERAS=1) 
-# to see python version
-reticulate::py_config()
-setwd("~/Desktop/repos/consistency_scores_in_text_datasets/model starters")
-
-# Read text from image
-library(tesseract)
-eng <- tesseract("eng")
-text_original <- tesseract::ocr("BERT Python/image1.png", engine = eng)
-cat(text_original)
-
-# Collect incorrect words
-library(textclean)
-text_stripped <- strip(text_original, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
-      lower.case = FALSE)
-library(hunspell)
-incorrectwords <- hunspell(text_stripped)
-
-# Configure paths to the pre-trained files variables
-pretrained_path = 'uncased_L-12_H-768_A-12'
-config_path = file.path(pretrained_path, 'bert_config.json')
-checkpoint_path = file.path(pretrained_path, 'bert_model.ckpt')
-vocab_path = file.path(pretrained_path, 'vocab.txt')
-seq_length = NULL
-
-# Load BERT and tokenize the text
-library(reticulate)
-k_bert = import('keras_bert')
-token_dict = k_bert$load_vocabulary(vocab_path)
-tokenizer = k_bert$Tokenizer(token_dict)
-encodes <- tokenizer$encode(text_original)
-model = k_bert$load_trained_model_from_checkpoint(
-  config_path,
-  checkpoint_path,
-  seq_len=seq_length,
-  training=T)
-
-tokens <- tokenizer$tokenize(text_original)
-length(tokens)
-
-# Get encoded inputs for the model: train, segments, targets
-indices = encodes[1]
-segments = encodes[2]
-masks = list(list(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0,
-                  0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
-                  0, 0, 0, 0, 0, 0))
-train = do.call(cbind,indices) %>% t()
-segments = do.call(cbind,segments) %>% t()
-targets = do.call(cbind,masks) %>% t()
-
-# Get the prediction matrix
-prediction <- model$predict(c(train))
-

### 2021-01-06 model starters/GPT2.R

@@ -1,6 +1,0 @@
-# Source tutorial: https://blogs.rstudio.com/ai/posts/2019-10-23-gpt-2/
-remotes::install_github("r-tensorflow/gpt2")
-gpt2::install_gpt2(envname = "r-gpt2")
-library(gpt2)
-gpt2(prompt = "GPT-3 thinks GPT-2 is", model = "124M", seed = NULL, batch_size = 1, total_tokens = NULL,
-     temperature = 1, top_k = 0, top_p = 1)

### 2021-01-06 model starters/GloVe.R

@@ -1,38 +1,0 @@
-# Source tutorial: https://keras.rstudio.com/articles/examples/pretrained_word_embeddings.html
-library(keras)
-
-GLOVE_DIR <- 'glove.6B'
-TEXT_DATA_DIR <- '20_newsgroup'
-MAX_SEQUENCE_LENGTH <- 1000
-MAX_NUM_WORDS <- 20000
-EMBEDDING_DIM <- 100
-VALIDATION_SPLIT <- 0.2
-
-# download data if necessary
-download_data <- function(data_dir, url_path, data_file) {
-  if (!dir.exists(data_dir)) {
-    download.file(paste0(url_path, data_file), data_file, mode = "wb")
-    if (tools::file_ext(data_file) == "zip")
-      unzip(data_file, exdir = tools::file_path_sans_ext(data_file))
-    else
-      untar(data_file)
-    unlink(data_file)
-  }
-}
-download_data(GLOVE_DIR, 'http://nlp.stanford.edu/data/', 'glove.6B.zip')
-
-# first, build index mapping words in the embeddings set
-# to their embedding vector
-
-cat('Indexing word vectors.\n')
-
-embeddings_index <- new.env(parent = emptyenv())
-lines <- readLines(file.path(GLOVE_DIR, 'glove.6B.100d.txt'))
-for (line in lines) {
-  values <- strsplit(line, ' ', fixed = TRUE)[[1]]
-  word <- values[[1]]
-  coefs <- as.numeric(values[-1])
-  embeddings_index[[word]] <- coefs
-}
-
-cat(sprintf('Found %s word vectors.\n', length(embeddings_index)))

### 2021-01-06 model starters/Word2Vec.R

@@ -1,68 +1,0 @@
-# Source tutorial: https://blogs.rstudio.com/ai/posts/2017-12-22-word-embeddings-with-keras/
-
-library(readr)
-library(stringr)
-
-# Using a short text for demonstration
-training_corpus <- "There was no possibility of taking a walk that day. We had been wandering, indeed, in the leafless shrubbery an hour in the morning. 
-But since dinner (Mrs Reed, when there was no company, dined early). The cold winter wind had brought with it clouds so sombre, and a rain so penetrating, 
-that further out-door exercise was now out of the question."
-
-library(keras)
-tokenizer <- text_tokenizer(num_words = 20000)
-tokenizer %>% fit_text_tokenizer(training_corpus)
-
-library(reticulate)
-library(purrr)
-skipgrams_generator <- function(text, tokenizer, window_size, negative_samples) {
-  gen <- texts_to_sequences_generator(tokenizer, sample(text))
-  function() {
-    skip <- generator_next(gen) %>%
-      skipgrams(
-        vocabulary_size = tokenizer$num_words, 
-        window_size = window_size, 
-        negative_samples = 1
-      )
-    x <- transpose(skip$couples) %>% map(. %>% unlist %>% as.matrix(ncol = 1))
-    y <- skip$labels %>% as.matrix(ncol = 1)
-    list(x, y)
-  }
-}
-
-embedding_size <- 128  # Dimension of the embedding vector.
-skip_window <- 5       # How many words to consider left and right.
-num_sampled <- 1       # Number of negative examples to sample for each word.
-
-input_target <- layer_input(shape = 1)
-input_context <- layer_input(shape = 1)
-
-embedding <- layer_embedding(
-  input_dim = tokenizer$num_words + 1, 
-  output_dim = embedding_size, 
-  input_length = 1, 
-  name = "embedding"
-)
-
-target_vector <- input_target %>% 
-  embedding() %>% 
-  layer_flatten()
-
-context_vector <- input_context %>%
-  embedding() %>%
-  layer_flatten()
-
-dot_product <- layer_dot(list(target_vector, context_vector), axes = 1)
-output <- layer_dense(dot_product, units = 1, activation = "sigmoid")
-
-model <- keras_model(list(input_target, input_context), output)
-model %>% compile(loss = "binary_crossentropy", optimizer = "adam")
-
-summary(model)
-
-#model <- keras::train_on_batch(model)
-
-model %>%
-  fit(
[…]

### 2021-01-06 scripts/00-try-BERT-text-prediction.R

@@ -1,83 +1,0 @@
-# Source tutorial: https://blogs.rstudio.com/ai/posts/2019-09-30-bert-r/
-Sys.setenv(TF_KERAS=1)
-# to see python version
-reticulate::py_config()
-setwd("~/Desktop/repos/consistency_scores_in_text_datasets/scripts/model_starters")
-
-#### 1. Read text from image
-library(tesseract)
-eng <- tesseract("eng")
-text_original <- tesseract::ocr("BERT_Python/image1.png", engine = eng)
-cat(text_original)
-
-#### 2. Collect incorrect words
-library(textclean)
-text_stripped <- strip(text_original, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
-                       lower.case = FALSE)
-
-library(hunspell)
-incorrectwords <- hunspell(text_stripped)
-
-tokens <- unlist(strsplit(text_stripped, "[^aA-zZ<>/]"))
-mask_indices <- match(incorrectwords[[1]],tokens)
-for(i in mask_indices){
-  tokens[i] = "MASK"
-}
-str <- toString(tokens)
-str <- strip(str)
-#### 3. Configure paths to the BERT pre-trained files variables
-# The zip file is available through the tutorial link at the first line
-pretrained_path = '../../../uncased_L-12_H-768_A-12'
-config_path = file.path(pretrained_path, 'bert_config.json')
-checkpoint_path = file.path(pretrained_path, 'bert_model.ckpt')
-vocab_path = file.path(pretrained_path, 'vocab.txt')
-seq_length = NULL
-
-#### 4. Load BERT and tokenize the text
-library(reticulate)
-k_bert = import('keras_bert')
-np = import('numpy')
-token_dict = k_bert$load_vocabulary(vocab_path)
-tokenizer = k_bert$Tokenizer(token_dict)
-model = k_bert$load_trained_model_from_checkpoint(
-  config_path,
-  checkpoint_path,
-  seq_len=seq_length,
-  training=T)
-
-library(zeallot)
-#### 5. Tokenize the text and get the encoding
-tokens <- tokenizer$tokenize(str)
-encodes <- tokenizer$encode(str)
-
-for(i in mask_indices){
-  encodes[[1]][i+1] = 103
-}
-
-#### 6. Get encoded inputs for the model: train, segments, targets
-indices <- encodes[1]
-indices <- as.matrix(indices)
[…]

### 2021-01-06 scripts/00-try_external_consistency.R

@@ -1,140 +1,0 @@
-library(tidyverse)
-library(tau)
-library(tm)
-library(stringr)
-library(stats)
-library(reshape2)
-library(RWeka)
-
-#setwd("~/Desktop/repos/consistency_scores_in_text_datasets/scripts/")
-path <- getwd()
-demoTraining <- readLines(paste(path,"/input-sample_en_US.blogs.txt", sep=""), warn=FALSE)
-
-#### OCR Read text from image ####
-# To-do: make an UI for this to upload PDF
-library(tesseract)
-eng <- tesseract("eng")
-input_original <- tesseract::ocr("input-image.png", engine = eng)
-cat(input_original)
-
-#### Collect incorrect words from the OCR ####
-library(textclean)
-input_stripped <- strip(input_original, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
-                       lower.case = FALSE)
-# Tokenizer function
-tokenizer <- function(corpus) {
-  lines <- vector()
-  # Read the corpus line by line
-  for (line in corpus) {
-    lines <-c(lines, line)
-  }
-  lines <- tolower(lines)
-  lines <- gsub("[.!?]$|[.!?] |$", " </s>", lines) # add end-of-sentence token to the end of a sentence
-  tokens <- unlist(strsplit(lines, "[^a-z<>/]"))
-  tokens <- tokens[tokens != ""]
-  return(tokens)
-}
-
-library(hunspell)
-# Use hunspell spell checker to get text errors
-incorrectwords <- hunspell(input_stripped)
-incorrectwords <- lapply(incorrectwords, tolower)
-# Split the input text
-input_tokens <- tokenizer(input_stripped)
-# Get the indices of the incorrect words in the tokens
-incorrectwords_indices <- match(as.vector(incorrectwords[[1]]),input_tokens)
-# Get the previous two words of the text error and put them in the list of bigrams
-bigrams_for_prediction <- vector()
-for(i in incorrectwords_indices) {
-  previous_two_words <- paste(input_tokens[i-2], input_tokens[i-1])
-  bigrams_for_prediction <- append(bigrams_for_prediction, previous_two_words)
-}
-
-#### Training text tokenization ####
-tokens <- tokenizer(demoTraining)
-
-#### Make data table for the ngrams ####
-library(data.table)
[…]

### 2021-01-06 scripts/00-try_ngrams.R

@@ -1,107 +1,0 @@
-#### Preamble ####
-# Contact: Rohan Alexander
-# Email: rohan.alexander@utoronto.ca
-# Date: 19 August 2020
-# Purpose: This script attempts to use n-grams to identify words that may be incorrect and 
-# need to be corrected.
-
-
-#### Workspace setup ####
-library(quanteda)
-library(tidyverse)
-
-
-#### Generate comparison dataset ####
-# First we need a collection of n-grams that we will search for, so we are creating
-# that in this section. We will just look at 3-grams, so once we have this list 
-# we can search for the first two tokens and then compare the third with what we
-# expect.
-example_paragraph <- "I should like to bury something precious in every place 
-where I've been happy and then, when I'm old and ugly and miserable, I could 
-come back and dig it up and remember. I should like to also ensure that my children
-and grandchildren knew where they were. For when I am old and ugly, I should like 
-that they too dig it up and imagine me as I am now - happy and content. I should lIke 
-also to be to be happy and content then, but really, who can promise that?"
-
-# Create tokens
-tokens_from_example <- quanteda::tokens(example_paragraph, remove_punct = TRUE)
-
-# Create ngrams from the tokens
-toks_ngram <- quanteda::tokens_ngrams(tokens_from_example, n = 3)
-head(toks_ngram[[1]])
-
-# Convert to tibble so we can use our familiar verbs
-all_tokens <- tibble(tokens = toks_ngram[[1]])
-
-# We only want the common ones, not every one.
-all_tokens <- 
-  all_tokens %>% 
-  group_by(tokens) %>% 
-  count() %>% 
-  filter(n > 1) %>% 
-  ungroup()
-
-# Create a tibble that has the first two words in one column then the third
-all_tokens <- 
-  all_tokens %>% 
-  mutate(tokens = str_replace_all(tokens, "_", " "),
-         first_words = word(tokens, start = 1, end = 2),
-         last_word = word(tokens, -1),
-         tokens = str_replace_all(tokens, " ", "_"),
-         first_words = str_replace_all(first_words, " ", "_")
-         ) %>% 
-  rename(last_word_expected = last_word) %>% 
-  select(-n, -tokens)
-
-
-#### Apply to the dataset to create a consistency score ####
-# Now that we have our collection of n-grams (this will be internal consistency
[…]

### 2021-01-06 scripts/02-starter-Word2Vec.R

@@ -1,0 +1,68 @@
+# Source tutorial: https://blogs.rstudio.com/ai/posts/2017-12-22-word-embeddings-with-keras/
+
+library(readr)
+library(stringr)
+
+# Using a short text for demonstration
+training_corpus <- "There was no possibility of taking a walk that day. We had been wandering, indeed, in the leafless shrubbery an hour in the morning. 
+But since dinner (Mrs Reed, when there was no company, dined early). The cold winter wind had brought with it clouds so sombre, and a rain so penetrating, 
+that further out-door exercise was now out of the question."
+
+library(keras)
+tokenizer <- text_tokenizer(num_words = 20000)
+tokenizer %>% fit_text_tokenizer(training_corpus)
+
+library(reticulate)
+library(purrr)
+skipgrams_generator <- function(text, tokenizer, window_size, negative_samples) {
+  gen <- texts_to_sequences_generator(tokenizer, sample(text))
+  function() {
+    skip <- generator_next(gen) %>%
+      skipgrams(
+        vocabulary_size = tokenizer$num_words, 
+        window_size = window_size, 
+        negative_samples = 1
+      )
+    x <- transpose(skip$couples) %>% map(. %>% unlist %>% as.matrix(ncol = 1))
+    y <- skip$labels %>% as.matrix(ncol = 1)
+    list(x, y)
+  }
+}
+
+embedding_size <- 128  # Dimension of the embedding vector.
+skip_window <- 5       # How many words to consider left and right.
+num_sampled <- 1       # Number of negative examples to sample for each word.
+
+input_target <- layer_input(shape = 1)
+input_context <- layer_input(shape = 1)
+
+embedding <- layer_embedding(
+  input_dim = tokenizer$num_words + 1, 
+  output_dim = embedding_size, 
+  input_length = 1, 
+  name = "embedding"
+)
+
+target_vector <- input_target %>% 
+  embedding() %>% 
+  layer_flatten()
+
+context_vector <- input_context %>%
+  embedding() %>%
+  layer_flatten()
+
+dot_product <- layer_dot(list(target_vector, context_vector), axes = 1)
+output <- layer_dense(dot_product, units = 1, activation = "sigmoid")
+
+model <- keras_model(list(input_target, input_context), output)
+model %>% compile(loss = "binary_crossentropy", optimizer = "adam")
+
+summary(model)
+
+#model <- keras::train_on_batch(model)
+
+model %>%
[…]

### 2021-01-06 scripts/02-starter_BERT.R

@@ -1,0 +1,86 @@
+# Load all dependencies
+library(reticulate)
+use_condaenv("my-torch", required = TRUE) # set Python env to conda
+library(textclean)
+library(hunspell)
+library(stringr)
+
+library(reticulate)
+install_bert <- function() {
+  reticulate::py_install("pytorch")
+  reticulate::py_install("torchvision")
+  reticulate::py_install("transformers")
+}
+
+install_bert()
+
+library(rTorch)
+transformer <- reticulate::import("transformers")
+
+BertForMaskedLM <- transformer$BertForMaskedLM
+BertTokenizer <- transformer$BertTokenizer
+tokenizer <- BertTokenizer$from_pretrained('bert-base-uncased')
+model <- BertForMaskedLM$from_pretrained('bert-base-uncased')
+
+
+
+text_original <- c("His Majesty Government wish to add that they have no im@@@l@fon of requesting the establishment of military bases in peace time within the area of Palestine now united to the Kis@@®m of Jordan.")
+
+
+##### Prediction function
+predict_tokens <- function(text_original) {
+  ### Clean text
+  text_stripped <- strip(text_original, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
+                         lower.case = FALSE)
+  ### Collect text errors
+  errors <- hunspell(text_stripped)
+
+  ### Replace errors to [MASK]
+  for(e in errors){
+    pats <- paste(e, collapse = '|')
+  }
+  text <- str_replace_all(text_stripped, pats, '[MASK]')
+
+  ### Tokenize the text
+  tokenized_text <- tokenizer$tokenize(text)
+  indexed_tokens <- tokenizer$convert_tokens_to_ids(tokenized_text)
+
+  ### Encode the text and create segments tensors
+  MASKIDS <- which(tokenized_text == "[MASK]")
+  SEGS <- which(tokenized_text == ".")
+
+  segments_ids <- rep(as.integer(0), times = SEGS)
+
+  segments_tensors <- torch$tensor(list(segments_ids))
+  tokens_tensor <- torch$tensor(list(indexed_tokens))
+
+  ### Generate prediction
+  py <- torch$no_grad()
+  with(py, {
+    predictions <- model(tokens_tensor, segments_tensors)
+  })
+
+  predicted_token <- list()
+  for(i in 1:length(MASKIDS)){
[…]

### 2021-01-06 scripts/02-starter_GloVe.R

@@ -1,0 +1,19 @@
+# Source tutorial: https://keras.rstudio.com/articles/examples/pretrained_word_embeddings.html
+# This example shows how one can quickly load glove vectors
+# and train a Keras model in R
+
+library(keras)
+library(dplyr)
+
+# Download Glove vectors if necessary
+if (!file.exists('glove.6B.zip')) {
+  download.file('http://nlp.stanford.edu/data/glove.6B.zip',destfile = 'glove.6B.zip')
+  unzip('glove.6B.zip')
+}
+
+# load glove vectors into R
+vectors = data.table::fread('glove.6B.300d.txt', data.table = F,  encoding = 'UTF-8')
+colnames(vectors) = c('word',paste('dim',1:300,sep = '_'))
+
+# structure of the vectors
+as_tibble(vectors)

### 2021-01-06 scripts/model_starters/R/.Rprofile.R

@@ -1,8 +1,0 @@
-transformers <- import("transformers")
-DistilBertForMaskedLM <- transformers$DistilBertForMaskedLM
-DistilBertModel <- transformers$DistilBertModel
-DistilBertTokenizer <- transformers$DistilBertTokenizer
-tokenizer <- DistilBertTokenizer$from_pretrained('distilbert-base-uncased')
-
-# pip install "transformers==3.1.0"
-model <- DistilBertModel$from_pretrained('distilbert-base-uncased', return_dict=TRUE)

### 2021-01-06 scripts/model_starters/R/BERT.R

@@ -1,25 +1,0 @@
-# Source tutorial: https://blogs.rstudio.com/ai/posts/2019-09-30-bert-r/
-Sys.setenv(TF_KERAS=1) 
-# to see python version
-reticulate::py_config()
-setwd("~/Desktop/repos/consistency_scores_in_text_datasets/model_starters")
-
-
-# 1. Configure paths to the BERT pre-trained files variables
-### The zip file is available through the tutorial link at the first line
-pretrained_path = '../../../uncased_L-12_H-768_A-12'
-config_path = file.path(pretrained_path, 'bert_config.json')
-checkpoint_path = file.path(pretrained_path, 'bert_model.ckpt')
-vocab_path = file.path(pretrained_path, 'vocab.txt')
-seq_length = NULL
-
-# 2. Load BERT and tokenize the text
-library(reticulate)
-k_bert = import('keras_bert')
-token_dict = k_bert$load_vocabulary(vocab_path)
-tokenizer = k_bert$Tokenizer(token_dict)
-model = k_bert$load_trained_model_from_checkpoint(
-  config_path,
-  checkpoint_path,
-  seq_len=seq_length,
-  training=T)

### 2021-01-06 scripts/model_starters/R/BERT_pytorch.R

@@ -1,75 +1,0 @@
-library(reticulate)
-use_condaenv("my-torch", required = TRUE)
-
-pytorch_pretrained_bert <- import("pytorch_pretrained_bert")
-BertModel <- pytorch_pretrained_bert$BertModel
-BertForMaskedLM <- pytorch_pretrained_bert$BertForMaskedLM
-BertTokenizer <- pytorch_pretrained_bert$BertTokenizer
-tokenizer <- BertTokenizer$from_pretrained('bert-base-uncased')
-
-# pip install "transformers==3.1.0"
-model <- BertForMaskedLM$from_pretrained('bert-base-uncased')
-
-#### 1. Get text
-text_original <- c("His Majesty Government wish to add that they have no im@@@l@fon of requesting the establishment of military bases in peace time within the area of Palestine now united to the Kis@@®m of Jordan.")
-
-#### 2. Clean text
-library(textclean)
-text_stripped <- strip(text_original, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
-                      lower.case = FALSE)
-
-#### 3. Collect text errors
-library(hunspell)
-errors <- hunspell(text_stripped)
-
-### 4. Replace errors to [MASK]
-for(e in errors){
-  pats <- paste(e, collapse = '|')
-}
-library(stringr)
-text <- str_replace_all(text_stripped, pats, '[MASK]')
-
-### 5. Tokenize the text
-tokenized_text <- tokenizer$tokenize(text)
-indexed_tokens <- tokenizer$convert_tokens_to_ids(tokenized_text)
-
-### 6. Encode the text and create segments tensors
-MASKIDS <- which(tokenized_text == "[MASK]")
-SEGS <- which(tokenized_text == ".")
-
-segments_ids <- rep(as.integer(0), times = SEGS)
-
-library(rTorch)
-segments_tensors <- torch$tensor(list(segments_ids))
-tokens_tensor <- torch$tensor(list(indexed_tokens))
-
-### 7. Generate prediction
-py <- torch$no_grad()
-with(py, {
-  predictions <- model(tokens_tensor, segments_tensors)
-})
-
-### 8. Prediction function
-predict_tokens <- function(text, predictions, MASKIDS) {
-  predicted_token <- list()
-  for(i in 1:length(MASKIDS)){
[…]

### 2021-01-06 scripts/model_starters/R/GloVe.R

@@ -1,19 +1,0 @@
-# Source tutorial: https://keras.rstudio.com/articles/examples/pretrained_word_embeddings.html
-# This example shows how one can quickly load glove vectors
-# and train a Keras model in R
-
-library(keras)
-library(dplyr)
-
-# Download Glove vectors if necessary
-if (!file.exists('glove.6B.zip')) {
-  download.file('http://nlp.stanford.edu/data/glove.6B.zip',destfile = 'glove.6B.zip')
-  unzip('glove.6B.zip')
-}
-
-# load glove vectors into R
-vectors = data.table::fread('glove.6B.300d.txt', data.table = F,  encoding = 'UTF-8')
-colnames(vectors) = c('word',paste('dim',1:300,sep = '_'))
-
-# structure of the vectors
-as_tibble(vectors)

### 2021-01-06 scripts/model_starters/R/Word2Vec.R

@@ -1,68 +1,0 @@
-# Source tutorial: https://blogs.rstudio.com/ai/posts/2017-12-22-word-embeddings-with-keras/
-
-library(readr)
-library(stringr)
-
-# Using a short text for demonstration
-training_corpus <- "There was no possibility of taking a walk that day. We had been wandering, indeed, in the leafless shrubbery an hour in the morning. 
-But since dinner (Mrs Reed, when there was no company, dined early). The cold winter wind had brought with it clouds so sombre, and a rain so penetrating, 
-that further out-door exercise was now out of the question."
-
-library(keras)
-tokenizer <- text_tokenizer(num_words = 20000)
-tokenizer %>% fit_text_tokenizer(training_corpus)
-
-library(reticulate)
-library(purrr)
-skipgrams_generator <- function(text, tokenizer, window_size, negative_samples) {
-  gen <- texts_to_sequences_generator(tokenizer, sample(text))
-  function() {
-    skip <- generator_next(gen) %>%
-      skipgrams(
-        vocabulary_size = tokenizer$num_words, 
-        window_size = window_size, 
-        negative_samples = 1
-      )
-    x <- transpose(skip$couples) %>% map(. %>% unlist %>% as.matrix(ncol = 1))
-    y <- skip$labels %>% as.matrix(ncol = 1)
-    list(x, y)
-  }
-}
-
-embedding_size <- 128  # Dimension of the embedding vector.
-skip_window <- 5       # How many words to consider left and right.
-num_sampled <- 1       # Number of negative examples to sample for each word.
-
-input_target <- layer_input(shape = 1)
-input_context <- layer_input(shape = 1)
-
-embedding <- layer_embedding(
-  input_dim = tokenizer$num_words + 1, 
-  output_dim = embedding_size, 
-  input_length = 1, 
-  name = "embedding"
-)
-
-target_vector <- input_target %>% 
-  embedding() %>% 
-  layer_flatten()
-
-context_vector <- input_context %>%
-  embedding() %>%
-  layer_flatten()
-
-dot_product <- layer_dot(list(target_vector, context_vector), axes = 1)
-output <- layer_dense(dot_product, units = 1, activation = "sigmoid")
-
-model <- keras_model(list(input_target, input_context), output)
-model %>% compile(loss = "binary_crossentropy", optimizer = "adam")
-
-summary(model)
-
-#model <- keras::train_on_batch(model)
-
-model %>%
[…]

### 2021-01-07 outputs/documents/paper_ngrams.Rmd
Changed transformer to Transformer

@@
 
-In future work we intend to investigate in more advanced language models such as word embeddings and transformers.
+In future work we intend to investigate in more advanced language models such as word embeddings and Transformers.
 We will consider pre-trained word embedding models including Word2Vec [@mikolov2013efficient] [@mikolov2013distributed] and GloVe [@pennington2014glove], which place each word in a multi-dimensional space such that distance between words can illustrate their relationship [@bengio2003neural].
@@
 
-We will also consider incorporating generative pre-trained transformer models such as GPT-2 [@radford2019language], GPT-3 [@brown2020language], and BERT [@devlin2018bert].
-These models are based on the transformer architecture [@vaswani2017attention].
-Proposed by @vaswani2017attention in 2017, the transformer network architecture for neural networks has been found to outperform both  recurrent neural network-based and convolutional neural network-based models in computational efficiency [@vaswani2017attention].
-Since the emergence of the transformer model, most of the representative pre-trained models are built based on this architecture [@hanretty2018comparing].
+We will also consider incorporating generative pre-trained Transformer models such as GPT-2 [@radford2019language], GPT-3 [@brown2020language], and BERT [@devlin2018bert].
+These models are based on the Transformer architecture [@vaswani2017attention].
+Proposed by @vaswani2017attention in 2017, the Transformer network architecture for neural networks has been found to outperform both recurrent neural network-based and convolutional neural network-based models in computational efficiency [@vaswani2017attention].
+Since the emergence of the Transformer model, most of the representative pre-trained models are built based on this architecture [@hanretty2018comparing].
 

### 2021-01-07 outputs/documents/paper_ngrams.Rmd
Deal with Monicas 1 and 2 pionts.

@@
 In order to use statistical models to understand that process, we typically need to clean and prepare the dataset in some way.
-This is especially the case when we work with text data.
-But this cleaning and preparation requires us to make many decisions.
+This is especially the case when we work with text data because they are typically not designed for analysis.
+This cleaning and preparation requires us to make many decisions.
 To what extent should we correct obvious errors?
 What about slightly-less-obvious errors?
+Again, this is especially the case for text datasets because they encode such an enormous amount of information, a feature may only be an error in context, and the high-dimensionality of the data at even its most basic level (26 English characters, compared with 10 numbers).
 Although cleaning and preparation is a necessary step, we may be concerned about the extent to which have we introduced new errors, and the possibility that we have made decisions that have affected, or even driven, our results.
 
-In this paper we introduce the concept of consistency in a text corpus.
-Consistency refers to the proportion of words that are able to be forecast by a statistical model, based on preceding words and surrounding context.
-Further, we define internal consistency as when the model is trained on the corpus itself, and external consistency as when the model is trained on a more general corpus.
-Together, these concepts provide a guide to the cleanliness and reasonableness of a text dataset.
+In this paper we introduce the concept of consistency and implement it for a text corpus.
+Consistency refers to the proportion of data that are able to be forecast by a statistical model, based on preceding and surrounding data.
+Further, we define internal consistency as when the model is trained on the dataset itself, and external consistency as when the model is trained on a more general dataset.
+Together, these concepts provide a guide to the cleanliness and reasonableness of a dataset.
 This can be important when deciding whether a dataset is fit for purpose; when carrying out data cleaning and preparation tasks; and as a comparison between datasets.
+For instance, if the Canadian population is known to be 38 million from a census, but the sum of age-groups is initially 35 million, rising to 38 million only after a inputation error is corrected, then that post-cleaning dataset would have higher consistency than the pre-cleaning dataset.
+When applied to a text dataset, this means that we construct a model that forecasts the next word, given the previous words.
+If the forecast word is always the same as the actual word then the consistency is 1.
+That said, it is not the absolute level of consistency that is our focus, but instead measuring how it changes during the cleaning and preparation process.
 
@@
 Hence the use of tri-gram models to provide other word candidates including "in the pond".
+
+## Text preparation and cleaning
+
+TBD
 

### 2021-01-08 outputs/documents/paper_ngrams.Rmd
Add comments to function code chunks

@@
 ```{r, eval=FALSE}
+make_internal_consistency_dataset <- function(body_of_text) {
   tokens_from_example <- 
-  quanteda::tokens(body_of_text, remove_punct = TRUE, )
+    quanteda::tokens(body_of_text, remove_punct = TRUE)
   tokens_from_example <- 
     quanteda::tokens_tolower(tokens_from_example)
+
+# Continuous code...
+}
 ```
@@
 ```{r, eval=FALSE}
+make_internal_consistency_dataset <- function(body_of_text) {
+# ... Previous code
+
   # Create ngrams from the tokens
@@
     dplyr::ungroup()
+
+# Continuous code...
+}
 ```
@@
 ```{r, eval=FALSE}
+make_internal_consistency_dataset <- function(body_of_text) {
+# ... Previous code
+
   # Create a tibble that has the first two words in one column then the third
@@
     dplyr::select(-n)
+}
 ```
@@
 ```{r, eval=FALSE}
+generate_internal_consistency_score<-function(text_to_check, consistency_dataset){
   # Create tokens with errors
@@
                   first_words = stringr::str_replace_all(first_words, " ", "_"))
+
+# Continuous code...
+}
 ```
@@
 ```{r, eval=FALSE}
+generate_internal_consistency_score<-function(text_to_check, consistency_dataset){
+#... Previous code
+
 all_tokens_with_errors <-
@@
                    by = c("first_words"))
+
+# Continuous code...
+}
 ```
@@
 ```{r, eval=FALSE}
+generate_internal_consistency_score<-function(text_to_check, consistency_dataset){
+#... Previous code
+
   # Calculate the internal consistency score:
@@
     dplyr::mutate(consistency = true_count/sum(n))
+
+# Continuous code...
+}
 ```
@@
 ```{r, eval=FALSE}
+generate_internal_consistency_score<-function(text_to_check, consistency_dataset){
+#... Previous code
+
 # Identify which words were unexpected
@@
   dplyr::select(-tokens)
+}
 ```

### 2021-01-08 outputs/documents/paper_ngrams.Rmd
corrected demo external consistency code chunk output

@@
 Here, the text to evaluate is 'there was no possibiliti', with the word "possibility" being wrongly spelled.
-The function identifies the word 'possibiliti' as an unexpected word, and generates the external consistency score 0.833.
+The function identifies the word 'possibiliti' as an unexpected word, and generates the external consistency score 0.333.
 Because the external dataset is larger, it is capable of providing more replacement candidates for 'possibiliti'.
@@
 ##   <lgl>       <int>       <dbl>
-## 1 FALSE           2       0.833
-## 2 TRUE           10       0.833
+## 1 FALSE           2       0.333
+## 2 TRUE            1       0.333
 

### 2021-01-08 outputs/documents/paper_ngrams.Rmd
Add subsection on maths and subsection on common cleaning

@@
 
-TBD
+In general text requires a great deal of preparation and cleaning before it can be analysed.
+For instance, `compositr` [@citecompositr] is an R package that runs through 13 common steps: remove non-UTF8 characters; insert missing comma spaces; replace misspelled words with most likely equivalent; remove HTML tags; convert case; remove URLs; remove numbers; remove punctuation; replace elongated words; remove extra white space; tokenization; stopword removal; and finally stem/lemmatize.
+
+These are just some of the steps that may be taken.
+And in general this assumes a fairly clean text dataset in the first instance.
+If the dataset had been created from Twitter then hashtags and emoji would need to be considered.
+And if the dataset had been created from OCR then there would likely be systematic errors to be considered.
+One benefit of using packages such as `compositr` is that these decisions become repeatable and so their effect on subsequent analysis can be easily understood through repeated running of the full workflow.
+The notion of consistency provides a measure of this without the need to run the full workflow.
+
+## Consistency
+
+Here we define, mathematically, what we mean by consistency in the case of text.
+Consider some text extract for which we want a consistency score.
+If there are $n$ words, of which $k$ are forecast correctly by a model, then the consistency score is the proportion: k/n.
+We implement this in dataframes using a binary column.
+Each word is a row, and a column tracks the forecast, classifying each word as 1 if it was correctly forecast and 0 otherwise.
+Then the internal consistency estimator is given by Equation (1):
+
+\begin{equation}
+c_i = \frac{\mbox{sum correctly forecast word column}}{\mbox{length of dataset}}.
+\end{equation}
+
+We distinguish between a model trained on the dataset for which consistency is being estimated, which results in an estimate of internal consistency, $c_i$, and a model trained on another dataset, which results in an estimate of external consistency, $c_e$.
+We consider the forecast of each word independently, allowing for correlation in future work.
 

### 2021-01-08 outputs/documents/paper_ngrams.Rmd
Deal with Monicas last point

@@
 
+The choice of what do now depends on a variety of factors including the purpose of the analysis, and the size of the corpus.
+One option is to consider only a subset of potential replacements, or to consider only replacements over a certain probability threshold.
+These are parameters that can be tuned in more sophisticated language models.
+The choice of which word to change 'possibili' to could affect the consistency score if it is not what the model expects.
+Future work, with more sophisticated language models, could allow tuning in this regard.
+
 # 4. Discussion

### 2021-01-08 outputs/documents/paper_ngrams.Rmd
corrected model outputs

@@
   toks_ngram <- 
-    toks_ngram <- quanteda::tokens_ngrams(tokens_from_example, n = 5:3)
+    quanteda::tokens_ngrams(tokens_from_example, n = 5:3)
 
@@
     all_tokens_with_errors %>%
-    dplyr::mutate(tokens = stringr::str_replace_all(tokens, "_", " "),
+    dplyr::mutate(ngram = sapply(strsplit(tokens, "_"), length),
+                  tokens = stringr::str_replace_all(tokens, "_", " "),
                   first_words = stringr::word(tokens, start = 1, end = -2),
@@
   # Calculate the internal consistency score:
-  internal_consistency <-
-    internal_consistency %>%
-    dplyr::ungroup()%>%
-    dplyr::filter(!is.na(as_expected)) %>%
-    dplyr::count(as_expected) %>%
-    dplyr::mutate(consistency = true_count/sum(n))
+  false_count <- length(internal_consistency[which(internal_consistency$as_expected == FALSE & 
+                                                   internal_consistency$ngram == 3)])
+  word_count <- sapply(strsplit(text_to_check, " "), length)
+  true_count <- word_count - false_count
+
+  internal_consistency <- tibble::tibble(
+    "as_expected" = true_count,
+    "unexpected"  = false_count,
+    "consistency" = true_count/word_count
+  )
 
@@
     dplyr::filter(as_expected == FALSE) %>%
-  dplyr::select(-tokens)
+    dplyr::select(-tokens, -ngram, -as_expected)
 }
@@
 ## $`internal consistency`
-## # A tibble: 1 x 3
-##   as_expected     n consistency
-##   <lgl>       <int>       <dbl>
-## 1 FALSE           1           0
+##   as_expected unexpected consistency
+## 1           4          1         0.8
 
@@
 ## 1 there_was   na        no                 
-
 ```
@@
 ## # A tibble: 1 x 3
-##   as_expected     n consistency
-##   <lgl>       <int>       <dbl>
-## 1 FALSE           2       0.333
-## 2 TRUE            1       0.333
+##   as_expected unexpected consistency
+##         <int>      <int>       <dbl>
+## 1           3          1        0.75
 

### 2021-01-08 outputs/documents/paper_ngrams.Rmd
minor update

@@
 The consistency score is calculated by the number of words that have been predicted by the model divided by the total number of words in the data.
-The function identifies the counts of words that are predicted by the model and divided by the total number of words in the internal consistency dataset:
+The function identifies the counts of words that are predicted by the model and divided by the total number of words in the input text:
 
@@
 The function `aRianna::generate_internal_consistency_score()` takes two arguments — the text to evaluate and the internal consistency dataset.
-The function identifies the word 'na' as an unexpected word, and generates the internal consistency score as zero.
+The function identifies the word 'na' as an unexpected word, and generates the internal consistency score as 0.8.
 The function also lists 'no' as the replacement of 'na':
@@
 Here, the text to evaluate is 'there was no possibiliti', with the word "possibility" being wrongly spelled.
-The function identifies the word 'possibiliti' as an unexpected word, and generates the external consistency score 0.333.
+The function identifies the word 'possibiliti' as an unexpected word, and generates the external consistency score 0.75.
 Because the external dataset is larger, it is capable of providing more replacement candidates for 'possibiliti'.

### 2021-01-08 shinyapp/external_consistency_func.R
upate shiny app

@@ -1,61 +1,0 @@
-#### Create bigrams from text errros' previous two words ####
-library(textclean)
-library(hunspell)
-library(tidyverse)
-library(rdrop2)
-trigramsDF <- drop_read_csv("Arianna/trigramsDF.csv")
-#trigramsDF <- read_csv("trigramsDF.csv")
-
-# Tokenizer function
-tokenizer <- function(corpus) {
-  lines <- vector()
-  # Read the corpus line by line
-  for (line in corpus) {
-    lines <-c(lines, line)
-  }
-  lines <- tolower(lines)
-  lines <- gsub("[.!?]$|[.!?] |$", " </s>", lines) # add end-of-sentence token to the end of a sentence
-  tokens <- unlist(strsplit(lines, "[^a-z<>/]"))
-  tokens <- tokens[tokens != ""]
-  return(tokens)
-}
-
-####  Ngram probability ####
-getLastWords <- function(string, words) {
-  pattern <- paste("[a-z']+( [a-z']+){", words - 1, "}$", sep="")
-  return(substring(string, str_locate(string, pattern)[,1]))
-}
-create_bigrams_for_predictions <- function(input_original) {
-  
-  input_stripped <- strip(input_original, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
-                          lower.case = FALSE)
-  # Use hunspell spell checker to get text errors
-  incorrectwords <- hunspell(input_stripped)
-  incorrectwords <- lapply(incorrectwords, tolower)
-  # Split the input text
-  input_tokens <- tokenizer(input_stripped)
-  # Get the indices of the incorrect words in the tokens
-  incorrectwords_indices <- match(as.vector(incorrectwords[[1]]),input_tokens)
-  # Get the previous two words of the text error and put them in the list of bigrams
-  bigrams_for_prediction <- vector()
-  for(i in incorrectwords_indices) {
-    previous_two_words <- paste(input_tokens[i-2], input_tokens[i-1])
-    bigrams_for_prediction <- append(bigrams_for_prediction, previous_two_words)
-  }
-  split_trigrams <- data.frame(bigrams_for_prediction, "incorrectwords" = incorrectwords[[1]])
-  
-  return(split_trigrams)
-}
-
-#### Function to predict words ####
-predict_words <- function(split_trigrams) {
-  expected_words <- c()
-  for (i in 1: nrow(split_trigrams)) {
[…]

### 2021-01-08 shinyapp/functions.R

@@ -1,120 +1,0 @@
-library(tidyverse)
-library(tau)
-library(tm)
-library(stringr)
-library(stats)
-library(reshape2)
-library(RWeka)
-
-#setwd("~/Desktop/repos/consistency_scores_in_text_datasets/shinyapp")
-path <- getwd()
-demoTraining <- readLines(paste(path,"/samples/demo.txt", sep=""), warn=FALSE)
-
-tokenizer <- function(corpus) {
-  lines <- vector()
-  # Read the corpus line by line
-  for (line in corpus) {
-    # add two start-of-sentence tokens to the beginning of a sentence
-    #line <- paste("<s> <s>", line)
-    lines <-c(lines, line)
-  }
-  lines <- tolower(lines)
-  lines <- gsub("[.!?]$|[.!?] |$", " </s>", lines) # add end-of-sentence token to the end of a sentence
-  tokens <- unlist(strsplit(lines, "[^a-z<>/]"))
-  tokens <- tokens[tokens != ""]
-  return(tokens)
-}
-
-tokens <- tokenizer(demoTraining)
-
-library(data.table)
-unigrams <- textcnt( tokens, n=1, method = "string", split = "[[:space:][:digit:]]+", decreasing = TRUE)
-bigrams  <- textcnt( tokens, n=2, method = "string", split = "[[:space:][:digit:]]+", decreasing = TRUE)
-trigrams <- textcnt( tokens, n=3, method = "string", split = "[[:space:][:digit:]]+", decreasing = TRUE)
-
-# Remove trigram that has "</s>" in it, so the trigrams are constructed within a sentence
-unigrams <- unigrams[!grepl("</s>", names(unigrams))]
-bigrams <-   bigrams[!grepl("</s>", names(bigrams))]
-trigrams <- trigrams[!grepl("</s>", names(trigrams))]
-
-### ngram probability
-getLastWords <- function(string, words) {
-  pattern <- paste("[a-z']+( [a-z']+){", words - 1, "}$", sep="")
-  return(substring(string, str_locate(string, pattern)[,1]))
-}
-
-removeLastWord <- function(string) {
-  sub(" [a-z']+$", "", string)
-}
-
-# Kneser-Ney Smoothing
-kneserNay <- function(ngrams, d) {
-  n <- length(strsplit(names(ngrams[1]), " ")[[1]])
-  # Special case for unigrams
-  if(n==1) {
-    noFirst <- unigrams[getLastWords(names(bigrams), 1)]
[…]

### 2021-01-08 shinyapp/generate_trigramsDF_csv.R

@@ -1,123 +1,0 @@
-library(tidyverse)
-library(tau)
-library(tm)
-library(stringr)
-library(stats)
-library(reshape2)
-
-path <- getwd()
-external_training_data <- readLines(paste(path,"/samples/input-sample_en_US.blogs.txt", sep=""), warn=FALSE)
-
-library(textclean)
-library(hunspell)
-
-#### Tokenizer function ####
-tokenizer <- function(corpus) {
-  lines <- vector()
-  # Read the corpus line by line
-  for (line in corpus) {
-    lines <-c(lines, line)
-  }
-  lines <- tolower(lines)
-  lines <- gsub("[.!?]$|[.!?] |$", " </s> ", lines) # add end-of-sentence token to the end of a sentence
-  tokens <- unlist(strsplit(lines, "[^a-z<>/]"))
-  tokens <- tokens[tokens != ""]
-  return(tokens)
-}
-
-#### Create bigrams from text errros' previous two words ####
-create_bigrams_for_predictions <- function(input_original) {
-  
-  input_stripped <- strip(input_original, char.keep = c("?", ".", "’", "~~"), digit.remove = TRUE, apostrophe.remove = FALSE,
-                          lower.case = FALSE)
-  # Use hunspell spell checker to get text errors
-  incorrectwords <- hunspell(input_stripped)
-  incorrectwords <- lapply(incorrectwords, tolower)
-  # Split the input text
-  input_tokens <- tokenizer(input_stripped)
-  # Get the indices of the incorrect words in the tokens
-  incorrectwords_indices <- match(as.vector(incorrectwords[[1]]),input_tokens)
-  # Get the previous two words of the text error and put them in the list of bigrams
-  bigrams_for_prediction <- vector()
-  for(i in incorrectwords_indices) {
-    previous_two_words <- paste(input_tokens[i-2], input_tokens[i-1])
-    bigrams_for_prediction <- append(bigrams_for_prediction, previous_two_words)
-  }
-  split_trigrams <- data.frame(bigrams_for_prediction, "incorrectwords" = incorrectwords[[1]])
-  
-  return(split_trigrams)
-}
-
-#### Training text tokenization ####
-tokens <- tokenizer(external_training_data)
-
-#### Make data table for the ngrams ####
-library(data.table)
[…]

### 2021-01-08 shinyapp/internal_consistency_func.R

@@ -1,87 +1,0 @@
-#### Preamble ####
-# Contact: Rohan Alexander
-# Email: rohan.alexander@utoronto.ca
-# Date: 19 August 2020
-# Purpose: This script attempts to use n-grams to identify words that may be incorrect and 
-# need to be corrected.
-
-#### Workspace setup ####
-library(quanteda)
-library(tidyverse)
-
-#### Generate comparison dataset ####
-# First we need a collection of n-grams that we will search for, so we are creating
-# that in this section. We will just look at 3-grams, so once we have this list 
-# we can search for the first two tokens and then compare the third with what we
-# expect.
-
-create_ngrams_tibble <- function(corpus, is_training) {
-  # Create tokens
-  tokens <- quanteda::tokens(corpus, remove_punct = TRUE)
-  # Create ngrams from the tokens
-  toks_ngram <- quanteda::tokens_ngrams(tokens, n = 3)
-  # Convert to tibble so we can use our familiar verbs
-  all_tokens <- tibble(tokens = toks_ngram[[1]])
-  
-  # If it's a training corpus, we only want the common ones, not every one.
-  if (is_training) {
-    all_tokens <- 
-      all_tokens %>% 
-      group_by(tokens) %>% 
-      count() %>% 
-      filter(n > 1) %>% 
-      ungroup()
-  }
-  
-  # Create a tibble that has the first two words in one column then the third
-  all_tokens <- 
-    all_tokens %>% 
-    mutate(tokens = str_replace_all(tokens, "_", " "),
-           first_words = word(tokens, start = 1, end = 2),
-           last_word = word(tokens, -1),
-           tokens = str_replace_all(tokens, " ", "_"),
-           first_words = str_replace_all(first_words, " ", "_")
-    ) 
-  
-  # If it's a training corpus, drop n column and tokens column
-  if (is_training) {
-    all_tokens <- all_tokens %>% 
-      rename(last_word_expected = last_word) %>% 
-      select(-n, -tokens)
-  }
-  
-  return(all_tokens)
-}
-
-# Get training n-gram tibble
-#training_tokens<- create_ngrams_tibble(example_paragraph, is_training = TRUE)
-# Get example n-gram tibble
-#input_tokens<- create_ngrams_tibble(example_sentence_with_errors, is_training = FALSE)
-
-#### Apply to the dataset to create a consistency score ####
-# Now that we have our collection of n-grams (this will be internal consistency
[…]

### 2021-01-08 shinyapp/server.R

@@ -1,4 +1,4 @@
-library(aRianna)
+# library(aRianna)
 source("training_data.R")
 source("OCR_text_func.R")
 
@@ -12,7 +12,7 @@ server <- function(input, output, session) {
   # Displays input text
   output$unexpected_words <- renderTable({
     unexpected_words <- getInternalResults()[[1]]
-    colnames(unexpected_words) <- c("As Expected", "Count", "Consistency Score")
+    colnames(unexpected_words) <- c("As Expected", "Unexpected", "Consistency Score")
     unexpected_words
   })
 
@@ -69,7 +69,7 @@ server <- function(input, output, session) {
   # Load demo data when the corresponding button is clicked
   observe({
     input$loadDemo
-    updateTextInput(session, "myTextData", value = demoTextData)
+    updateTextInput(session, "myTextData", value = demoInternalTextData)
   })
 
   # Load demo data when the corresponding button is clicked in external consistency panel

### 2021-01-08 shinyapp/training_data.R

@@ -1,2 +1,3 @@
-demoTextData <- "we had no idea that you had no idae"
-demoInternalTrainingData <- "we had no idea that you had no idea that they had no idea"
+demoTextData <- "There was no possibili"
+demoInternalTextData <- "when there was na company"
+demoInternalTrainingData <- "There was no possibility of taking a walk that day. We had been wandering, indeed, in the leafless shrubbery an hour in the morning; but since dinner (Mrs. Reed, when there was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, that further out-door exercise was now out of the question."

### 2021-01-13 outputs/documents/paper.Rmd
paper focusing on ngrams
As sent to arXiv

--- 2020-08-21 outputs/documents/paper.Rmd
@@
     citation_package: natbib
-    keep_tex: false
+    keep_tex: true
     toc: false
     fig_caption: true
+    numbering: true
     latex_engine: pdflatex
@@
   -  \usepackage{hyperref}
+  -  \usepackage{amsmath}
 biblio-style: apalike
-title: "Consistency scores in text data"
-thanks: "Thank you to **X, Y and Z** for helpful comments.
+title: "On consistency scores in text data with an implementation in R"
+thanks: "Thank you to Monica Alexander and Alex Luscombe for helpful comments.
 A Shiny application for interactive and small-scale examples is available: https://kelichiu.shinyapps.io/Arianna/.
-An R package for larger scale applications is available: **HERE**.
+An R package for larger scale applications is available: https://github.com/RohanAlexander/arianna.
 Our code and datasets are available: https://github.com/RohanAlexander/consistency_scores_in_text_datasets.
@@
   affiliation: University of Toronto
-abstract: "In this paper we introduce a process to clean the text extracted from PDFs using various methods from natural language processing.
-Our approach compares originally extracted text with the text generated from, or expected by, these methods using earlier text as stimulus.
-To guide this process, we introduce the notion of a consistency score, which refers to the proportion of text that is unchanged by the model.
-This is used to monitor changes during the cleaning process, and to compare the messiness of different texts.
-The methods that we consider are: n-grams, Word2Vec, GloVe, and the GPT models.
-We illustrate our process on text from the Canadian Hansard and introduce both a Shiny application and an R package to make our process easier for others to adopt."
-keywords: "text-as-data; natural language processing; quantitative analysis."
+abstract: "In this paper, we introduce a reproducible cleaning process for the text extracted from PDFs using n-gram models.
+Our approach compares the originally extracted text with the text generated from, or expected by, these models using earlier text as stimulus.
+To guide this process, we introduce the notion of a consistency score, which refers to the proportion of text that is expected by the model.
+This is used to monitor changes during the cleaning process, and across different corpuses.
+We illustrate our process on text from the book Jane Eyre and introduce both a Shiny application and an R package to make our process easier for others to adopt."
+keywords: "text-as-data; natural language processing; quantitative analysis; optical character recognition"
 date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
 Datasets can have errors, be biased, incomplete, or messy.
-In any case, it is the underlying statistical process of which the dataset is an artifact that is typically of interest.
+In any case, it is the underlying statistical process, of which the dataset is an artifact, that is typically of interest.
 In order to use statistical models to understand that process, we typically need to clean and prepare the dataset in some way.
-This is especially the case when we work with text data.
-But this cleaning and preparation requires us to make many decisions.
+This is especially the case when we work with text data because they are typically not designed for analysis.
+This cleaning and preparation requires us to make many decisions.
 To what extent should we correct obvious errors?
 What about slightly-less-obvious errors?
+Again, this is especially the case for text datasets because: they encode a large amount of information; an aspect may only be an error in context; and the high-dimensionality of the data at even its most basic level (26 English characters, compared with 10 numbers).
 Although cleaning and preparation is a necessary step, we may be concerned about the extent to which have we introduced new errors, and the possibility that we have made decisions that have affected, or even driven, our results.
 
-In this paper we introduce the concept of consistency in a text corpus.
-Consistency refers to the proportion of words that are able to be forecast by a statistical model, based on preceding words and surrounding context.
-Further, we define internal consistency as when the model is trained on the corpus itself, and external consistency as when the model is trained on a more general corpus.
-Together, these concepts provide a guide to the cleanliness and consistency of a text dataset.
+In this paper we introduce the concept of consistency and implement it for a text corpus.
+Consistency refers to the proportion of data that are able to be forecast by a statistical model, based on preceding and surrounding data.
+Further, we define internal consistency as when the model is trained on the dataset itself, and external consistency as when the model is trained on a more general dataset.
+Together, these concepts provide a guide to the cleanliness and reasonableness of a dataset.
 This can be important when deciding whether a dataset is fit for purpose; when carrying out data cleaning and preparation tasks; and as a comparison between datasets.
+For instance, if the Canadian population is known to be 38 million from a census, but the sum of age-groups is initially 35 million, rising to 38 million only after an inputation error is corrected, then that post-cleaning dataset would have higher consistency than the pre-cleaning dataset.
+When applied to a text dataset, this means that we construct a model that forecasts the next word, given the previous words.
+If the forecast word is always the same as the actual word, then the consistency is 1.
+That said, it is not the absolute level of consistency that is our focus, but instead measuring how it changes during the cleaning and preparation process.
 
@@
 Hence if the sentence was actually 'the cat in the bat', then that child would know something was likely wrong.
-The consistency score would likely be lower than if the sentence were 'the cat in the hat'.
+The consistency score would likely be lower than if the sentence were 'the cat in the hat' because we expected 'hat', but actually found 'bat'.
 After we correct this error, the consistency score would likely increase.
-By examining how the consistency scores evolve in response to changes made to the text during the data preparation and cleaning stages we can better understand the effect of the changes.
-Including consistency scores and their evolution when text corpuses are shared allows researchers to be more transparent about their corpus.
+By examining how consistency scores evolve in response to changes made to the text during the data preparation and cleaning stages we can better understand the effect of the changes.
+Including consistency scores and their evolution when datasets are shared allows researchers to be more transparent about the effect of their decisions in these steps of the analysis workflow.
 And finally, the use of consistency scores allows for an increased level of automation in the cleaning process.
 
-We consider various natural language processing methods.
-These include n-grams, Word2Cec, GloVe, GPT models and BERT.
-The n-gram approach involves identifying two, three, or more, words that are commonly found together.
-Here think of a two-gram involving the word 'good' 'good morning'.
-At scale these can identify missing or unusual words, and work quickly, but they lack nuance.
-For instance, an equally reasonable two-gram involving the word 'good' is 'good work'.
-For that reason, we consider pre-trained word embedding models.
-These include Word2Vec and GloVe, which place each word in a multi-dimensional space such that distance between words can illustrate their relationship.
-We also consider pre-trained generative models such as GPT-2 and GPT-3 and BERT.
-GPT-2, GPT-3 and BERT are pre-trained generative unsupervised language models.
-GPT-2 and GPT-3 are the two generations of the same model that differ in the number of parameters.
-GPT-2 has 1.5 billion parameters and GPT-3 has more than 175 billion parameters.
-While GPT-2 is more limited, it can be run on smaller machines, whereas GPT-3 cannot and is accessed via an API.
-BERT has 340 millions parameters which is smaller than both GPT-2 and GPT-3.
-BERT also differs from GPT models because its encoding process is bidirectional instead of unidirectional.
+We employ n-gram models to calculate consistency scores and generate word candidates for text corrections.
+The n-gram approach involves identifying words that are commonly found together.
+In our application, we use these sequences to calculate consistency scores and generate word candidates for text corrections.
+For instance, if we are using a sequence of three words, or tri-grams, then we identify the first two words in the sequence and evaluate if the last word in the sequence is 'as expected' by comparing it to the corresponding tri-gram in a comparison dataset.
+The sequences we employ in this paper are a mixture of three, four, and five, consecutive words also known as tri-grams, 4-grams, and 5-grams.
 
-We apply our approach to the Canadian Hansard.
-The Canadian Hansard was digitised by @beelen2017digitization.
-This was a process by which PDFs were scanned, put through optical character recognition (OCR), and then corrected to a reasonable extent.
-The dataset is extensive, fit-for-purpose, and, appropriately, it has been used considerably, for instance @rheault2020word.
-However, there are many issues with the dataset in terms of reading it.
-We focus on one particular year - **(Rohan pick one)** - and show that our approach can be used to relatively quickly improve the quality of the dataset in a reproducible and consistent manner.
+The comparison dataset can be internal, meaning that the dataset is from the same source as the text being evaluated.
+On the other hand, a comparison dataset can be external, which means the dataset is from a different source, typically a larger and more universal dataset.
+The external dataset we use in the application discussed in this paper is constructed using a subset of text data sourced from Open WebText Corpus [@Gokaslan2019OpenWeb].
+Since the content of Open WebText Corpus is from the internet, the variety of the text data is promising for meeting the generalizability and universality that an external dataset should possess, although it will also have extensive bias.
 
-The remainder of our paper is structured as follows: **(Rohan add structure)**.
-Additionally, we construct a Shiny app available at: **(Rohan add link)**.
-That app computes internal and external consistency scores for corpus excerpts, and have developed an R Package that allows our approach to be used on larger datasets, which is available at: **(Rohan add link)**.
+The resources that we have developed to support this paper include a Shiny app available at: https://kelichiu.shinyapps.io/aRianna/.
+That app computes internal and external consistency scores for corpus excerpts.
+We have also developed an R Package, `aRianna`, that allows our approach to be used on larger datasets, which is available at: https://github.com/RohanAlexander/aRianna.
+
+In this paper, we introduce consistency scores and an R package that implements them.
+In Section \@ref(background), we provide an overview of language models and n-grams in particular.
+In Section \@ref(arianna), we deconstruct the package functions and provide demonstration example of the package usage.
+Finally, in Section \@ref(discussion), we discuss our findings, limitations, and future directions.
 
@@
 This is the underlying principle of all language models.
-Essentially, a language model is a probability distribution over sequences of words.
-The goal of statistical language modeling is to estimate probability distributions over different linguistic units — words, sentences, and even documents [@bengio2003neural].
+Essentially, the goal of statistical language modeling is to estimate probability distributions over different linguistic units — terms, words, sentences, and even documents [@bengio2003neural].
 However, this is difficult as language is categorical.
-If we consider each word in a vocabulary as a category, then the dimensionality of language becomes large [@rosenfeld2000two].
-The reason that there is such a variety of statistical language models is that there are various ways of dealing with this fundamental problem.
+If we consider each word in a vocabulary as a category, then the dimensionality of a language dataset often becomes large quickly [@rosenfeld2000two].
+One reason there is such a variety of statistical language models is that there are various ways of dealing with this fundamental problem.
+N-gram models are statistical language models that work by taking the co-occurrence of words in a sequence into account.
 
-## PDF text extraction and n-grams
-
-The Portable Document Format (PDF) is typically used to render documents in a relatively fixed way.
-PDFs contain a lot of information, but often that information is not able to be analysed directly by statistical language models.
-The information first needs to be extracted from the PDF, often by optical character recognition (OCR), and then cleaned, prepared, and made into some type of dataset.
-One issue in this process is that OCR is not perfect.
-Hence there is a need to correct the output before it is analysed.
-One way to do this is to correct errors one-by-one and independent of context.
-We all know that 'tbe' is wrong and should be 'the'.
-But statistical language models have the advantage of correction that can occur within context (@rosenfeld2000two).
-This allows significant improvements, for instance sometimes 'bat' should be 'hat' and sometimes it really should be 'bat'.
+## N-gram models
 
 The foundation of an n-gram language model is the conditional probability set-up introduced above.
-An n-gram model is a probabilistic language model that predicts the next word in a sequence of words (@bengio2003neural).
-The $n$ in n-gram refers to the number of words in that sequence.
+An n-gram model is a probabilistic language model that predicts the next term in a sequence of terms [@bengio2003neural].
+The $n$ in n-gram refers to the number of terms in that sequence.
 Consider the following excerpt from *Jane Eyre*: 'We had been wandering'.
@@
 The probability of a sequence appearing in a corpus follows the chain rule:
-
-$$\mbox{Pr(We,had,been,wandering)} = \mbox{Pr(We)} \times \mbox{Pr(had|We)} \times \mbox{Pr(been|We,had)} \\  
-\times \mbox{Pr(wandering|We,had,been)}.$$
+\begin{align*} 
+\mbox{Pr(We, had, been, wandering)} = &  \mbox{Pr(We)} \times  \mbox{Pr(had|We)} \times  \mbox{Pr(been|We, had)} \\ 
+ &  \times  \mbox{Pr(wandering|We, had, been)}.
+\end{align*}
 
@@
 This is known as the 'Markov assumption' and it allows us to approximate the probability using only the last $n$ words [@brown1992class]:
-
-$$\mbox{Pr(We,had,been,wandering)} \approx \mbox{Pr(We)} \times \mbox{Pr(had|We)} \times \mbox{Pr(been|had)} \\ 
+$$\mbox{Pr(We, had, been, wandering)} \approx \mbox{Pr(We)} \times \mbox{Pr(had|We)} \times \mbox{Pr(been|had)} \\ 
 \times \mbox{Pr(wandering|been)}.$$
@@
 
-Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation [@brown1992class].
-In our application, we use a tri-gram based model to detect both real-word and non-word errors and correct them with the candidate word that has the highest probability.
+In our application, we use a model that has a mixture of tri-grams, 4-grams, and 5-grams to provide term replacement options for unexpected terms.
+Longer n-grams such as 5-grams capture more context and potentially generate more accurate predictions.
+However, there is less chance that a longer sequence of words will fit a specific combination which can lead to sparsity.
+Therefore, shorter n-grams should still be included to preserve the generalizability of the predictions.
 
-However, n-gram models do not take the linguistic structure of language into account.
-For instance, @rosenfeld2000two [p.
-1] discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'.
-The prediction of the next word is based on only a few preceding words and broader context is not taken into account.
-Hence next-word prediction using n-gram based language models can be limited.
+Consider the idiom 'elephant in the room'.
+If by mistake, this is written as 'elephant in the roon' and correction is needed for the last word, then we can apply both a tri-gram model or a 4-gram model to predict the candidate word:
 
-There are a variety of ways to implement an n-gram model within R @citeR including using R packages such as `Quanteda` [@citequanteda], `tidyText` [@citetidytext] and `tm` [@citetm].
-We used the Text Analysis Utilities, `tau`, R package @citetau because of we can separate text-cleaning and n-gram making in two different steps.
-We want to have control to manually clean the text because we want to add an end-of-sentence token to each sentence int he corpus.
-To illustrate the use of `tau`, we will first install and load the library, and create a small corpus.
+> Using tri-gram model: 'in the `<prediction>`'
+
+> Using 4-gram model: 'elephant in the `<prediction>`'
+
+In the 4-gram model, one more word, 'elephant', is included, which provides more context and increased chance of 'room' being accurately predicted.
+However, there are times when the literal location of an elephant may be being described; for instance, 'elephant in the pond'.
+The combination of this sequence as a 4-gram might not be in a training dataset and no prediction would be available if only 4-gram models are used.
+Hence the use of tri-grams to provide other word candidates including 'in the pond'.
+
+## Text preparation and cleaning
+
+In general, text requires a great deal of preparation and cleaning before it can be analysed.
+For instance, `compositr` [@citecompositr] is an R package that runs through 13 common steps: remove non-UTF8 characters; insert missing comma spaces; replace misspelled words with most likely equivalent; remove HTML tags; convert case; remove URLs; remove numbers; remove punctuation; replace elongated words; remove extra white space; tokenization; stopword removal; and finally stem/lemmatize.
+And these are just some of the steps that may be taken.
+
+Text preparation and cleaning steps and the uncertainty they bring, assume a fairly clean text dataset in the first instance.
+For instance, if a dataset has been created from Twitter then hashtags and emoji would need to be considered.
+And if the dataset had been created from OCR then there would likely be systematic errors to be addressed.
+One benefit of using packages such as `compositr` is that these decisions become repeatable and so their effect on subsequent analysis can be easily understood through repeated running of the full workflow.
+The notion of consistency provides a measure of this without the need to run the full workflow.
+
+## Consistency
+
+Here we define, mathematically, what we mean by consistency in the case of text.
+Consider some text extract for which we want a consistency score.
+If there are $n$ words, of which $k$ are forecast correctly by a model, then the consistency score is the proportion: $k/n$.
+We implement this in a dataframe using a binary column.
+Each word is a row, and a column, 'correctly_forecast', tracks the forecast, classifying each word as 1 if it was correctly forecast, and 0 otherwise.
+Then the internal consistency estimator is given by Equation \@ref(eq:model):
+
+\begin{equation} 
+(\#eq:model)
+c_i = \frac{\mbox{sum 'correctly_forecast' column}}{\mbox{length of dataset}}. 
+\end{equation}
+
+We distinguish between a model trained on the dataset for which consistency is being estimated, which results in an estimate of internal consistency, $c_i$, and a model trained on another dataset, which results in an estimate of external consistency, $c_e$, although the procedure is the same once the forecast has been made.
+We consider the forecast of each word independently, allowing for correlation in future work.
+
+# The `aRianna` R package {#arianna}
+
+## Package dependencies
+
+There are a variety of ways to implement an n-gram model within R [@citeR] including packages such as `quanteda` [@citequanteda], `tidytext` [@citetidytext] and `tm` [@citetm].
+We use `quanteda` because it has a comprehensive set of functions for conducting text analysis.
+We also use `tidyverse` [@citetidyverse] for data manipulation:
 
 ```{r, eval = FALSE}
-install.packages("tau")
+install.packages("quanteda")
+install.packages("tidyverse")
+library(quanteda)
+library(tidyverse)
 ```
 
+## Function to make the internal dataset
+
+To obtain the internal consistency score, we first construct a dataset from the text that is itself the text being evaluated.
+The function `aRianna::make_internal_consistency_dataset()` is created for this purpose.
+The function turns a `body_of_text` into a dataset to compare the text with.
+To do so, the function first transforms the `body_of_text` into word tokens with the punctuation removed and the letters turned to lowercase:
+
 ```{r, eval=FALSE}
-library(tau)
-corpus <- "There was no possibility of taking a walk that day. We had been 
-wandering, indeed, in the leafless shrubbery an hour in the morning; but 
-since dinner (Mrs. Reed, when there was no company, dined early) the cold 
-winter wind had brought with it clouds so sombre, and a rain so penetrating, 
-that further out-door exercise was now out of the question."
+make_internal_consistency_dataset <- 
+  function(body_of_text) {
+  tokens_from_example <- 
+    quanteda::tokens(body_of_text, remove_punct = TRUE)
+  tokens_from_example <- 
+    quanteda::tokens_tolower(tokens_from_example)
+  # code continues...
 ```
 
-Our purpose is to construct tri-grams and so we first split the corpus into sentences by period, exponential mark, or question marks.
-
-First, we remove all punctuation and other non-letter characters, and convert all text to lower case.
+Next, the `aRianna::make_internal_consistency_dataset()` function turns the individual tokens into tri-grams, 4-grams, and 5-grams, keeping only the n-grams that appear in the data for more than once.
+These common n-grams are stored in a tibble:
 
 ```{r, eval=FALSE}
-tokenizer <- function(corpus) {
-  lines <- vector()
-  # Read the corpus line by line
-  for (line in corpus) {
-    # add two start-of-sentence tokens to the beginning of a sentence
-    lines <-c(lines, line)
-  }
-  lines <- tolower(lines)
-  # add end-of-sentence token to the end of a sentence
-  lines <- gsub("[.!?]$|[.!?] |$", " </s>", lines) 
-  tokens <- unlist(strsplit(lines, "[^a-z<>/]"))
-  tokens <- tokens[tokens != ""]
-  return(tokens)
+  # ...continue from previous
+  # Create ngrams from the tokens
+  toks_ngram <- 
+    quanteda::tokens_ngrams(tokens_from_example, n = 5:3)
+
+  # Convert to tibble so we can use our familiar verbs
+  all_tokens <- 
+    tibble::tibble(tokens = toks_ngram[[1]])
+
+  # We only want the common ones, not every one.
+  all_tokens <-
+    all_tokens %>%
+    dplyr::group_by(tokens) %>%
+    dplyr::count() %>%
+    dplyr::filter(n > 1) %>%
+    dplyr::ungroup()
+  # code continues...
+```
+
+Finally, the `aRianna::make_internal_consistency_dataset()` function splits the n-grams into two parts: the first $n-1$ words and the last word.
+As a result, a tibble is created with each n-gram containing its original sequence, the first words, and the last word:
+
+```{r, eval=FALSE}
+  # ...continue from previous
+  # Create a tibble that has the first two words in one column then the third
+  all_tokens <-
+    all_tokens %>%
+    dplyr::mutate(tokens = stringr::str_replace_all(tokens, "_", " "),
+                  first_words = stringr::word(tokens, start = 1, end = -2),
+                  last_word = stringr::word(tokens, -1),
+                  tokens = stringr::str_replace_all(tokens, " ", "_"),
+                  first_words = stringr::str_replace_all(first_words, " ", "_")
+    ) %>%
+    dplyr::rename(last_word_expected = last_word) %>%
+    dplyr::select(-n)
   }
@@
 
-We will use the `textcnt()` function to turn the training corpus into tri-grams.
-The most important argument of this function is _corpus_, as well as _n_ to set the number of words in a sequence, _method_ to determine which type of counts to compute, and _split_ to designate the regular expression pattern for splitting.
-The function outputs a `textcnt` object, which is essentially a table with each tri-gram and their counts [@citetau].
+## Function to generate internal consistency
+
+After we have the internal consistency dataset, we need a function to compare the text being evaluated and the internal dataset in order to retrieve the internal consistency score.
+We define the `aRianna::generate_internal_consistency_score` function to do this.
+Similar to the `aRianna::make_internal_consistency_dataset` function, this function first transforms the text being evaluated into tri-grams, 4-grams, and 5-grams tokens, then splits the n-grams to their first words and last words:
 
 ```{r, eval=FALSE}
-trigrams <- textcnt( tokens, n=3, method = "string", 
-                     split = "[[:space:][:digit:]]+", decreasing = TRUE)
+generate_internal_consistency_score <-
+  function(text_to_check, consistency_dataset){
+  # Create tokens with errors
+  tokens_from_example_with_errors <- 
+    quanteda::tokens(text_to_check, remove_punct = TRUE)
+  tokens_from_example_with_errors <- 
+    quanteda::tokens_tolower(tokens_from_example_with_errors)
+
+  # Create ngrams from the tokens with errors
+  toks_ngram_with_errors <- 
+    quanteda::tokens_ngrams(tokens_from_example_with_errors, n = 5:3)
+
+  all_tokens_with_errors <- 
+    tibble::tibble(tokens = toks_ngram_with_errors[[1]])
+
+  all_tokens_with_errors <-
+    all_tokens_with_errors %>%
+    dplyr::mutate(ngram = sapply(strsplit(tokens, "_"), length),
+                  tokens = stringr::str_replace_all(tokens, "_", " "),
+                  first_words = stringr::word(tokens, start = 1, end = -2),
+                  last_word = stringr::word(tokens, -1),
+                  tokens = stringr::str_replace_all(tokens, " ", "_"),
+                  first_words = stringr::str_replace_all(first_words, " ", "_"))
+  # code continues...
 ```
 
-We then manually added an end of sentence token, `</s>`, to the end of each sentence.
+Next, the function uses `dplyr::left_join()` to combine the tibble of text tokens and the internal consistency tibble.
+The `dplyr::left_join()` function returns all the rows from the text tokens tibble and all the columns from both tibbles [@2018dplyr].
+By combining the two tibbles, the function generates two additional columns — the `last_word` column are the last words of all n-grams, and the `last_word_expected` contains the last words that are in the internal consistency dataset.
+By comparing the words in `last_word` and `last_word_expected`, we can identify the 'unexpected' words.
+If a last word is in `last_word` but not `last_word_expected`, it is regarded as an unexpected word and the corresponding word in `last_word_expected` serves as a replacement candidate:
 
 ```{r, eval=FALSE}
-trigrams <- trigrams[!grepl("</s>", names(trigrams))]
+  # ...continue from previous
+all_tokens_with_errors <-
+    all_tokens_with_errors %>%
+    dplyr::left_join(dplyr::select(consistency_dataset, -tokens), 
+                     by = c("first_words"))
+  # code continues...
 ```
 
-At this point it is possible to take other measures around fine-tuning the tri-grams.
-For instance we may like to remove the tri-grams that contain an end-of-sentence token.
-This would mean that tri-grams that include a word from a new sentence will be removed.
-For example, the tri-gram 'day </s> we' will be removed because it has words from two different sentences.
-Table 1.
-shows four sample tri-grams from the small sample corpus after this process.
+The consistency score is calculated by the number of words that have been predicted by the model divided by the total number of words in the data.
+The function identifies the counts of words that are predicted by the model and divided by the total number of words in the input text:
 
-```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
-tabl <- "
-| there was no  | a rain so | a walk that | an hour in  |
-|---------------|-----------|-------------|-------------|
-| 2             | 1         | 1           | 1           |
+```{r, eval=FALSE}
+  # ...continue from previous
+  # Calculate the internal consistency score:
+  false_count <- 
+    length(internal_consistency[which(internal_consistency$as_expected == FALSE & 
+                                        internal_consistency$ngram == 3)])
+  word_count <- sapply(strsplit(text_to_check, " "), length)
+  true_count <- word_count - false_count
 
-Table 1. Four sample tri-grams generated from the corpus
-"
-cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
+  internal_consistency <- tibble::tibble(
+    "as_expected" = true_count,
+    "unexpected"  = false_count,
+    "consistency" = true_count/word_count
+  )
+  # code continues...
 ```
 
-The tri-grams can be accessed by calling the names of this table object and the counts can be used to calculate probabilities.
-There are a variety of options around constructing these probabilities.
-One is Kneser–Ney smoothing, which is the smoothing technique recommended by @chen1999empirical to improve the accuracy of n-gram models.
-
-Finally, it is helpful to create a data table that contains the first two words and the last of each tri-gram in separate column and the smoothed probability in another column.
-For instance, Table 2.
-provides and example of this set-up.
-
-```{r table2, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
-tabl <- "
-| First Words   | Last Word | Probability |
-|---------------|-----------|-------------|
-| there was     | no        | 0.6278947   |
-
-Table2. The smoothed probability table
-"
-cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
-```
-
-## Word embedding models
-
-Although, from an implementation perspective, there is an inherent numerical representation of language in n-gram models, it is not overly tractable.
-The numeric representation typically relies on each term being represented in a vector, which can quickly result in quite high-dimensional objects.
-In contrast, word embedding models represent terms by vectors of real numbers where each vector is a 'feature', hence the final object has considerably lower dimensionality [@bengio2003neural].
-
-This approach has a long history, but the implementation of @bengio2003neural has become the foundation for much subsequent work.
-Here, terms are encoded into feature vectors that represent different aspects of the term, and so each term is represented by a position in the vector space.
-Terms that are used in similar ways will have similar representations in the vector space.
-This will be due to their co-occurrence in usage.
-For example, 'cat' will be closer to 'dog' than to 'car', because we are more likely to talk about cats and dogs, than cats and cars.
-See how they might be close when considered along a 'pets' dimension, but far when considered along a 'related-to-tigers' dimension.
-
-Terms that are often used in similar contexts will have similar vector representations.
-Hence, word embedding models can identify similar terms, calculate the similarity between a pair of terms, and find the odd term in a group of terms [@mikolov2013efficient].
-There are a variety of ways to achieve this dimensionality reduction, including Word2Vec and GloVe.
-
-### Word2vec
-
-Word2vec is a word embedding model that uses a neural network to learn word associations [@mikolov2013efficient] [@mikolov2013distributed].
-It takes a corpus as an input and outputs a set of feature vectors that allow the terms in that corpus to be represented.
-@mikolov2013efficient proposed two architectures for determining this distributed representation of terms: continuous bag-of-words (CBOW), and continuous skip-gram.
-Both have an input layer, a projection layer, and an output layer.
-But CBOW uses several surrounding words, or the context, to predict a target word.
-On the other hand, continuous skip-gram takes a single word as input to predict its neighbouring words.
-
-To illustrate the difference, consider the quote from Descartes, 'I think therefore I am'.
-The focus word for which we want to learn the embedding is 'therefore', and the context window is set to two.
-This means that we consider the two words immediately preceding and proceeding 'therefore'.
-The CBOW input is 'I, think, I, am', and the output is 'therefore'.
-In contrast, the continuous skip-gram input is 'therefore', while 'I, think, I, am' is the output.
-The input layer encodes terms within a context window to one-hot vectors.
-Using a projection matrix, the projection layer then project the discrete one-hot vectors to lower-dimension continuous vectors.
-Finally, the output layer is a softmax regression classifier that obtains the posterior distribution of words [@rong2014word2vec].
-It is the input and output vectors that differ between the two architectures.
-
-Word2Vec is trained on the Google News dataset [@citegooglenews].
-This contains six billion tokens and the vocabulary size is restricted to the one million most common words [@mikolov2013efficient].
-As this is a large dataset, it provides an accurate statistical pattern of language, as it is used in this dataset, and a comprehensive mapping between terms and vector positions.
-
-To the extent that the relative positions of terms reflects their semantic relationship, Word2Vec will perform well for word analogies.
-However, terms that have context-specific meaning may not perform as well.
-
-There are a few packages for implementing a word2vec model in R, including `rword2vec` [@citerwordvecchaware] and `word2vec` [@citewordvecwijffels].
-Both packages use the same algorithm.
-Another way to implement a word2vec model in R is through Keras which is an API for deep learning written in Python [@chollet2015keras] for which the `keras` package provides an R interface [@citekeras].
-
-Here, we follow @falbel2017word to implement a skip-gram model of Word2Vec through Keras.
-There are three steps in the Keras deep learning framework: 1) prepare the data; 2) define the model; and 3) train and evaluate the model.
-**(Ke-Li: Can we change this to be a different dataset please?)** Following these steps, we first downloaded the Google News dataset [@citegooglenews] and preprocess the text into tokens.
+The `aRianna::generate_internal_consistency_score` function also lists the identified text errors and generates replacement candidates to the text errors:
 
 ```{r, eval=FALSE}
-tokenizer %>% fit_text_tokenizer(training_corpus)
-```
-
-We then define a model that mimics the skip-gram algorithm of word2vec using the function `skipgrams_generator()`.
-
-```{r, eval=FALSE}
-skipgrams_generator <- function(text, tokenizer, window_size, negative_samples) {
-  gen <- texts_to_sequences_generator(tokenizer, sample(text))
-  function() {
-    skip <- generator_next(gen) %>%
-      skipgrams(
-        vocabulary_size = tokenizer$num_words, 
-        window_size = window_size, 
-        negative_samples = 1
-      )
-    x <- transpose(skip$couples) %>% map(. %>% unlist %>% as.matrix(ncol = 1))
-    y <- skip$labels %>% as.matrix(ncol = 1)
-    list(x, y)
-  }
+  # ...continue from previous
+  # Identify which words were unexpected
+  unexpected <-
+    all_tokens_with_errors_only %>%
+    dplyr::mutate(as_expected = last_word == last_word_expected) %>%
+    dplyr::filter(as_expected == FALSE) %>%
+    dplyr::select(-tokens, -ngram, -as_expected)
   }
@@
 
-The embedding dimensions, the size of the context window and the number of samples are configurable within this function.
+## Function to generate external consistency
 
-Finally, we use the `fit_generator()` function to train the model.
+The `aRianna::generate_external_consistency_score` function works identically.
+The only difference is that it compares the text being evaluated with an external consistency dataset that is larger and more general.
+The external consistency dataset we employed in the application is constructed using a subset of text data from Open WebText Corpus [@Gokaslan2019OpenWeb].
+In general, any dataset could be used.
+
+## Demonstration
+
+To install the package and load the library:
 
 ```{r, eval=FALSE}
-model %>%
-  fit_generator(
-    skipgrams_generator(training_corpus, tokenizer, skip_window, negative_samples), 
-    steps_per_epoch = 10, epochs = 1
-  )
+devtools::install_github("RohanAlexander/arianna")
+library(aRianna)
 ```
 
-**What's supposed to happen:** As a result, a embedding matrix is obtained and serves as the look-up table of the word vectors.
-**What really happened:** However, the training process was stuck in the operation of `fit_generator()` function.
-According to Stack Flow this might be an issue of incompatible versions between Keras and Tensorflow.
-More investigation is needed.
+For this demonstration, we use the first paragraph of Jane Eyre as the internal text data.
+A sentence within the paragraph is modified with the intention to contain an error and serves as the text to be evaluated:
 
-### GloVe
+```{r, eval=FALSE}
+body_of_text <- "There was no possibility of taking a walk that day. 
+  We had been wandering, indeed, in the leafless shrubbery an hour in 
+  the morning; but since dinner (Mrs. Reed, when there was no company, 
+  dined early) the cold winter wind had brought with it clouds so sombre, 
+  and a rain so penetrating, that further out-door exercise was now out 
+  of the question."
 
-The Global Vectors (GloVe) model is a pre-trained word embedding model that emphasises the semantic meaning of words [@pennington2014glove].
-In the Word2Vec model, word representations are learnt locally within the context window.
-On the other hand, GloVe considers the properties of the corpus globally.
-As the model considers the context of an entire corpus the co-occurrence probabilities are based on counts for the entire training dataset.
-
-GloVe additionally considers the relative probability of word co-occurrence.
-For instance, consider the two words 'ice' and 'steam'.
-GloVe considers their co-occurrence with a variety of probe words, $k$.
-For instance $k$ could be 'solid', which is a property of ice; 'gas', which is a property of steam; and 'water', which is related to both.The ratio of co-occurrence probability is:
-
-$$\frac{\mbox{Pr}(k|\mbox{ice})}{\mbox{Pr}(k|\mbox{steam})}.$$
-
-The ratio between the co-occurrence probability of the 'ice, solid' pair and the 'steam, solid' pair is high because the co-occurrence of 'ice' and 'solid' is higher than the co-occurrence of 'steam' and 'solid'.
-Conversely, the ratio between the co-occurrence of the 'ice, gas' pair and the 'steam, gas' pair is low.
-On the other hand, the ratio of the co-occurrence probability of the 'ice, water' pair and the 'steam, water' pair us close to one because water is relevant to both 'ice' and 'steam'.
-
-As @hanretty2018comparing point out, although word embedding models can capture semantic meanings of words, "they are context-free and fail to capture higher-level concepts in context, such as polysemous disambiguation, syntactic structures, semantic roles, anaphora".
-Word embedding models do not represent a homonyms that has difference meanings in different vectors depending on the context.
-For example, "squash" can be a type of hard shelled vegetable or can be referring to a racket and ball sport, but there is only one series of vectors to represent the word "squash", instead of dynamically changing the vectors depending on the context.
-Moreover, from our observation, the word embedding models are unable to handle novel words that are not included in the vocabulary of the training corpus; for this reason, we conclude that the generalizability of word embedding models is sub-optimal.
-
-There are a variety of ways to implement GloVe word embedding models in R.
-One is to load GloVe embeddings into a frozen Keras Embedding layer [@kerasglove] using the `Keras` R package [@kerasR2020].
-This means the embedding vectors will not be updated during the training process.
-Another way, is to use the `text2vec` package [@text2vec].
-
-Here we follow **CITE** to implement GloVe using the `keras` R package [@kerasglove].
-The blog post is aimed to create a word embedding model for text classification.
-However, we will only use the guide create a embedding index and skip the steps of text classification.
-Instead of building a model for word embedding from scratch as we did for the skip-gram model of Word2Vec, here we are merely loading the Glove model and its pre-trained word vectors as an index.
-The word embedding index is then saved to an R environment.
-
-**[As before, we need code and examples here]**
-
-## Pre-trained transformer language models
-
-In recent years, the use of pre-trained language models has played a significant role in advancing natural language processing (NLP) tasks such as reading comprehension, text generation, and even creative writing [@brown2020language].
-Pre-trained models remove the need for researchers to prepare training datasets or allot computational resources to the stage of pre-training.
-Because of this, there has been substantial development in NLP research.
-In this section we discuss the GPT models from OpenAI and the BERT model from Google.
-
-### GPT
-
-OpenAI is an AI research company who openly pursues Artificial General Intelligence — to have machines learn and understand any intellectual tasks as human do [@brundage2018malicious].
-OpenAI's Generative Pre-Trained Transformer (GPT) models are pre-trained language models that are built based on the transformer architecture.
-Proposed by @vaswani2017attention from Google in 2017, the Transformer is a novel network architecture for neural network that outperforms RNN-based and CNN-based models in computational efficiency [@vaswani2017attention].
-Since the emergence of the Transformer model, most of the representative pre-trained models are built based on this architecture [@hanretty2018comparing].
-Therefore, we can say that Transformer marks a new era of language models.
-
-As of August 2020, OpenAI has three generations of GPT models.
-GPT was introduced to the public in June, 2018.
-In the year 2019, its successor GPT-2 was released in three stages.
-Not long after, in June 2020, OpenAI announced GPT-3, which is the latest GPT model as of July 2020 [@brown2020language].
-Since the API of GPT-3 at the moment is only available to a few invited users, public applications of GPT-3 are limited.
-On the other hand, its predecessor, GPT-2, has enabled researchers made breakthroughs in the past years.
-@alt2019fine extended GPT-2 to relation extraction, which is a task of identifying the relationship between concepts appeared in text.
-They had shown a milestone in predicting larger range of distinct relation types with higher confidence [@alt2019fine].
-In the field of speech recognition, data scarcity is often an issue due to the difficulty to collect large amount of data from human for learning [@bird2020lstm].
-@bird2020lstm employed GPT-2 as part of the solution to generate augmented data for classification.
-GPT-2 also helped NLP researchers tackling the task of textual paraphrasing.
-@hegde2020unsupervised used GPT-2 to build an unsupervised paraphrasing model that is not restricted by domain shift within an article, which is a problem faced by the usual supervised models of paraphrasing.
-
-Most of prior language models are trained in a single domain by supervised learning to perform a single task.
-OpenAI, on the other hands, train the GPT models to perform many tasks [@radford2019language].
-GPT-2 is pre-trained on WebText, a comprehensive dataset that is scrapped from millions of webpages and contains total of 40 GB of text [@radford2019language].
-WebText contains all kinds of textual data: dialogues, comments, translations, questions and answers...etc.
-Such diversity is critical for GPT-2 to be a generalized model that can perform multiple tasks with any domain boundary.
-However, GPT-2 has the capacity to adapt to a context and generates custom output with the "fine-tuning" technique [@ziegler2019fine].
-While the pre-training stage is unsupervised and mandatory, fine-tuning is an optional training stage where GPT can learn to perform new tasks in a supervised or reinforced approach such as stylistic continuation and summarization [@ziegler2019fine].
-During the pre-training stage, the model is trained to perform text completion task.
-Give it a prompt to start and it will generate texts that continues what has been started.
-In the fine-tuning stage, @radford2019language provided additional training datasets for GPT-2 to generate the texts with specific sentiment, or be descriptive, or even summarize the prompt text.
-
-OpenAI has made GPT-2 open-sourced through Github with Python and Tensorflow [@tensorflow2015-whitepaper].
-There is also a package in R that wraps the Python code into a R package, created by @keydanaluraschi2019gpt2.
-There are three choices of models in terms of parameter sizes — "124M", "355M" and "774M".
-The R package only permits exploratory experiments with the pre-trained tasks; fine-tuning for new tasks is not available thorough R.
-To proceed with the implementation of GPT-2 in R, theoretically, we will fine-tune the model through Python and create a wrapper code for R.
-This could possibility be achieved by employing the `reticulate` R package [@reticulateR] that allows operation of Python code in R.
-The practicality of this approach is to be further investigated.
-
-GPT-3 has over 175 billions of parameters [@brown2020language].
-In the past few years, the growth of parameters of transformer language models expands rapidly.
-The first generation GPT has 110 million parameters.
-Released in 2018, GPT-2 has 1.6 billion parameters, and within two years, GPT-3 has grown the size to 175 billion, which is over 100 times than its predecessor.
-The more parameters there are the better the model can generalize to new tasks; therefore, the increasing amount of parameters accelerates the improvements in text synthesis and downstream NLP tasks [@brown2020language].
-GPT-3 can learn NLP tasks with just a handful of example; OpenAI refers this ability as few-shot learning, or in-context learning [@brown2020language].
-As a result, we only need to provide GPT-3 a few demonstrations for it to learn a new NLP task.
-For example, OpenAI team trained GPT-3 to generate "news articles".
-The prompt input is a plausible first sentence of a news story written by human, and the output is an entire news article generated by the model itself.
-Without the demonstrations, GPT-3 tends to treat the prompts as "tweets" and generate texts that serve as responses or follow-up tweets.
-The team fed a few previous news articles in the model’s context as demonstrations, and the model learned to produce the outputs that adapt to the style and length of news stories [@brown2020language].
-This few-shot learning ability indicates that the model can learn like human do and would democratize the use of language models to broader fields.
-
-### BERT
-
-BERT stands for Bidirectional Encoder Representations from Transformers [@devlin2018bert].
-It was released by researchers at Google AI Language in 2018.
-The key innovation of BERT is its application of bidirectional training of the Transformer model, which results in a deeper sense of language context and flow.
-What is bidirectional training?
-In the context of English language, a sentence is constructed from left to right.
-The context on the left is used to determine the next word on the right.
-In this sense, GPT models are unidirectional since the probability of the next word is dependent only on the past.
-It is also possible to reverse the direction and model the sentence from right to left, and BERT incorporates both directions of left-to-right and right-to-left to model bidirectional contexts.
-
-BERT is pre-trained by using two unsupervised tasks.
-The first task is Masked Language Model and and second task is Next Sentence Prediction.
-The bidirectional approach of BERT is alleviated by the "Masked Language Model" [@devlin2018bert].
-The Masked Language Model masks random words in the input corpus which forces the model to predict the original words by "filling in the blank".
-This training process blends the representation from the left and right context and therefore increases the model's context sensitivity [@nozza2020mask].
-In the Next Sentence Prediction process, the model receives two sentences in sequence as input.
-50% of the sentences are paired in the right sequence and the ones from the other half have sentences that are paired randomly [@devlin2018bert].
-The goal of this process is to have BERT evaluate if the second sentence is the next sentence of the first sentence in the original text and sharpen its capacity to predict the likelihood of the next sentence.
-
-BERT was pre-trained on BooksCorpus dataset with 800 million words and English Wikipedia with 2,500 million words [@devlin2018bert].
-Because BERT employs the Transformer architecture with the self-attention mechanism, fine-tuning is also available with one additional output layer and a task-specific input [@devlin2018bert].
-Google AI researchers have fine-tuned BERT for 11 NLP tasks including question-answering and common-sense inference.
-
-BERT implementation in R is also possible through Keras.
-@kerasbert created a python module called `keras-bert` and the module can be called by R through `reticulate` package [@reticulateR], which enables the operation of Python modules in R.
-To start, we install the keras-bert model in Python on the machine and load BERT model into R through `reticulate`.
-Once `keras-bert` is loaded, we can use it to fine-tune BERT on a task-specific training dataset.
-Fine-tuning is started by tokenizing the training data.
-The output of the tokenization is then fed to BERT as the input.
-The next step is to define the model by configuring its parameters such as batch size, number of epochs and the learning rate.
-Finally, the model is compiled and we can begin the fine-tuning.
-
-For the task of text cleaning, fine-tuning is not required.
-The process of text cleaning involves detecting and correcting mis-spelled words.
-Correcting the incorrect words can be seen as predicting the correct words.
-The Masked Language Model of BERT can be employed to do such prediction.
-In the pre-training process, the Masked Language Model would mask random words to do the prediction.
-However, we can tell the model what we want to "mask" by assigning the words to be encoded as `[MASK]`.
-Consider the erroneous sentence that has ten words:
-
-> "There was no possibility of taking a wlak that day."
-
-Using a spell-checker, "wlak" is identified as an error and is substituted by `[MASK]`:
-
-> "There was no possibility of taking a [MASK] that day."
-
-The implementation of text correction with BERT in R with the `keras-bert` Python module is not yet completed.
-For the purpose of demonstrating how BERT predicts the missing words, we will show the implementation process in Python using the `pytorch_pretrained_bert` library.
-The reason to use `pytorch_pretrained_bert`(**Keli add citation**) instead of `keras-bert`is that `pytorch_pretrained_bert` has relatively comprehensive functions that makes the prediction of missing words more straightforward.
-
-After we retrieved the new text data with the error replaced by `[MASK]`, we use the `tokenizer()` function in `pytorch_pretrained_bert` to tokenize the text:
-
-```{python, eval=FALSE}
-tokenized_text = tokenizer.tokenize("There was no possibility of taking a [MASK] that day.")
-tokenized_text
->>> ['there', 'was', 'no', 'possibility', 'of', 'taking', 'a', '[MASK]', 'that', 'day', '.']
+text_to_evaluate <- "when there was na company"
 ```
 
-We then get the index of `[MASK]` within the sentence and save it to a list. Since we only have one `[MASK]` token, the list has only one element:
+The first step is to turn the body of text into the internal consistency dataset.
+The generated internal consistency dataset is a tibble that contains the tri-grams, 4-grams and 5-grams that appear in the internal text data more than once.
+Since only 'there_was_no' has more than one occurrence, it is the only n-gram in the internal consistency dataset:
 
-```{python, eval=FALSE}
-mask_index = 7
+```{r, eval=FALSE}
+internal_consistency_dataset <- 
+  aRianna::make_internal_consistency_dataset(body_of_text)
+internal_consistency_dataset
+## # A tibble: 1 x 3
+##   tokens       first_words last_word_expected
+##   <chr>        <chr>       <chr>             
+## 1 there_was_no there_was   no                
 ```
 
-`pytorch_pretrained_bert` has a dictionary of tokens generated from the pre-training stage.
-The tokens include the vocabulary, puncuations and special symbols.
-Each token in the dictionary is given an index, represented as `{token: index}`.
-The function `convert_tokens_to_ids()` encodes the tokens we got previously to their positional indices in the tokens dictionary and collects them into a list:
+The next step is to compare the text to be evaluated with the internal consistency dataset that we created in the previous step.
+The function `aRianna::generate_internal_consistency_score()` takes two arguments: the text to evaluate, and the internal consistency dataset.
+The function identifies the word 'na' as an unexpected word and generates the internal consistency score as 0.8.
+The function also lists 'no' as the replacement of 'na':
 
-```{python, eval=FALSE}
-indexed_tokens = tokenizer.convert_tokens_to_ids(tokenized_text)
-indexed_tokens
->>> [2045, 2001, 2053, 6061, 1997, 2635, 1037, 103, 2008, 2154, 1012]
+```{r, eval=FALSE}
+aRianna::generate_internal_consistency_score(
+  text_to_evaluate, internal_consistency_dataset)
+
+## $`internal consistency`
+##   as_expected unexpected consistency
+## 1           4          1         0.8
+
+## $`unexpected words`
+## # A tibble: 1 x 3
+##   first_words last_word last_word_expected 
+##   <chr>       <chr>     <chr>                 
+## 1 there_was   na        no                 
 ```
 
-The next step is to encode the text by its segments.
-The model encodes a sentence of representing each word in the sentence with the positional index of the sentence in the provided corpus.
-For example, the example text has only one sentence "There was no possibility of taking a [MASK] that day." Therefore, the position of the sentence is "0", and the encoding for each word in the sentence, including the period is a series of "0" to the length of the sentence:
+To get the external consistency score, we use the `aRianna::generate_external_consistency_score()` function.
+The function takes only one argument, the text to evaluate, and compares it with the external consistency dataset.
+Here, the text to evaluate is 'there was no possibiliti', with the word "possibility" being wrongly spelled.
+The function identifies the word 'possibiliti' as an unexpected word, and generates the external consistency score 0.75.
+Because the external dataset is larger, it is capable of providing more replacement candidates for 'possibiliti'.
+There are no replacements based on 5-grams, so the replacements based on 4-grams are displayed first, followed by the replacements based on tri-grams:
 
-```{python, eval=FALSE}
-segments_tokens = [0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]
+```{r, eval=FALSE}
+text_to_evaluate <- "there was no possibiliti"
+aRianna::generate_external_consistency_score(text_to_evaluate)
+
+## aRianna::generate_external_consistency_score(text_to_evaluate)
+## $`external consistency`
+## # A tibble: 1 x 3
+##   as_expected unexpected consistency
+##         <int>      <int>       <dbl>
+## 1           3          1        0.75
+
+## $`unexpected words`
+## # A tibble: 11 x 3
+##    first_words  last_word    last_word_expected 
+##    <chr>        <chr>        <chr>                    
+##  1 there_was_no possibiliti  evidence          
+##  2 there_was_no possibiliti  immediate         
+##  3 there_was_no possibiliti  infrastructure    
+##  4 there_was_no possibiliti  one               
+##  5 there_was_no possibiliti  possibility       
+##  6 there_was_no possibiliti  sound             
+##  7 there_was_no possibiliti  way               
+##  8 was_no       possibiliti  good              
+##  9 was_no       possibiliti  longer            
+## 10 was_no       possibiliti  more              
+## 11 was_no       possibiliti  wonder 
 ```
 
-The previous two steps give us two set of encoding — `indexed_tokens` that contains the encoding for each token, and `segments_tokens` that contains the encoding for each sentence.
-These two encoding layers that we feed to BERT Masked Language Model.
-The model takes tensor objects as inputs [@paszke2017automatic], therefore we transformed the two lists into tensor objects before hands.
-These two tensor objects has two dimensions.
-The first dimension is the container, and the second dimension is the length of the token lists.
+The choice of what do now depends on a variety of factors including the purpose of the analysis, and the size of the corpus.
+One option is to consider only a subset of potential replacements, or to consider only replacements over a certain probability threshold.
+These are parameters that can be tuned in more sophisticated language models.
+The choice of which word to change 'possibili' to could affect the consistency score if it is not what the model expects.
+Future work, with more sophisticated language models, could allow tuning in this regard.
 
-```{python, eval=FALSE}
-tokens_tensor = torch.tensor([indexed_tokens])
-tokens_tensor
-
-segments_tensor = torch.tensor([segments_encodes])
-segments_tensor
->>> [[0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]]
-```
-
-The encoding layers can be represented as the following:
-
-```{r table3, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
-tabl <- "
-|        |'there' |'was' |'no' |'possibility'|'of'|'taking'|'a' |'[MASK]'|'that'|'day'| '.'|
-|--------|--------|------|-----|-------------|----|--------|----|--------|------|-----|----|
-|token   |2045    |2001  |2053 |6061         |1997|2635    |1037|103     |2008  |2154 |1012|
-|segment |0       |0     |0    |0            | 0  |0       |0   |0       |0     |0    |0   |
-
-Table2. The smoothed probability table
-"
-cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
-```
-
-We then load the pre-trained BertForMaskedLM and feed the two tensor objects to it.
-The model then returns a tensor object that has three dimensions.
-The `predictions` tensor objects contains the prediction score of all the words in the vocabulary.
-
-```{python, eval=FALSE}
-model = BertForMaskedLM.from_pretrained('bert-base-uncased')
-predictions = model(tokens_tensor, segments_tensor)
-predictions
-```
-
-To get the prediction, we will get the element with the highest prediction score and retrieve its index.
-The retrieved index is then used to find the corresponding token in the tokens dictionary.
-The predicted word is "walk", which matches with the original sentence in **Jane Eyre**.
-
-```{python, eval=FALSE}
-predicted_word_index = torch.argmax(predictions[0, mask_index]).item()
-predicted_word_index
->>> 3328
-predicted_word_token = tokenizer.convert_ids_to_tokens([predicted_index])[0]
-predicted_word_token
->>> walk
-```
-
-# Stylized example
-
-**(TODO: This stylized example need to be changed and updated to actually use GPT-3 at some point.)**
-
-In our application, we deploy OpenAI GPT-3 to generate text and ...
-
-As a stylized example, let's consider the following actual paragraph from Jane Eyre, by Charlotte Bronte:
-
-> There was no possibility of taking a walk that day.
-We had been wandering, indeed, in the leafless shrubbery an hour in the morning; but since dinner (Mrs.
-Reed, when there was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, that further out-door exercise was now out of the question.
-
-Let's pretend that this text had been created from optical character recognition and that it had the following errors: some 'h' were replaced with 'b'; and some 'd' have been replaced with 'al':
-
-> There was no possibility of taking a walk that day.
-We had been wandering, indeed, in tbe leafless shrubbery an hour in tbe morning; but since dinner (Mrs.
-Reeal, when tbere was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, tbat further out-door exercise was now out of the question.
-
-Assume a model that is trained to perfectly forecast the next word in Jane Eyre.
-For this fragment there are 62 words, comprising 5 errors and 57 correct words.
-So the internal consistency score of this fragment would be: 57/62 = 0.919.
-When we recognise and correct the errors, this consistency score would increase to 1.
-
-Similarly, assume a model that is trained on an external data source.
-This means that it will recognise the the cases where some 'h' were replaced with 'b', but not recognise that 'Reed' has become 'Reeal'.
-Hence, the external consistency score would be 58/62 = 0.935.
-
-# Data
-
-Various data can be used....
-
-# Application
+As the dataset increases in size, the number of possible replacements will also likely increase.
+Various options for decreasing the number of possible replacements exist, starting with string distance algorithms and increasing in complexity.
+It is important to recognise that we are focused on the fact that they were not forecast, rather than what they should be replaced with.
+However, improving this aspect helps to make the consistency score more relevant and useful, especially in text settings where the cleaning and preparation aspect is likely to be iterative.
 
@@
 
-Internal validity vs external- one is their own words the other is a general set of words.
+Outside of textbook examples, almost all real-world datasets need to be cleaned and prepared before they can be analyzed.
+While there are many tools to help with reproducibility of analysis, options for ensuring reproducibility in earlier stages of a typical data science workflow are more limited.
+Our approach discussed in this paper can help.
+Our notion of consistency can be tracked at many points while the dataset is being cleaned and prepared.
+Stages in which it changes substantially can be investigated.
+While there may be good reasons for such as change, by identifying them it is easier for other researchers to understand the effects of cleaning and preparation on eventual analysis results.
 
-In the same way that precision and recall provide important measures…
+Our approaches are applicable to any dataset, but in this paper we applied them to one based on text.
+The high-dimensionality of text, and the messiness of it when it comes from real-world sources, means that often considerable cleaning and preparation is done to text datasets.
+We demonstrated how sequences of words can be used to forecast the very next word.
+That forecast can be compared with the actual word observed and this difference expressed as a proportion, that we call a 'consistency score'.
 
-# Appendix
+The language model that we used is n-grams.
+Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation [@brown1992class].
+As demonstrated in our application, we use a n-gram based model to calculate internal and external consistency scores and generate a list of words as text correction candidates.
+However, n-gram models do not take the linguistic structure of language, nor broader context, into account.
+For instance, @rosenfeld2000two [p.
+1] discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'.
+Hence next-word prediction using n-gram-based language models can be limited.
+Here think of a two-gram involving the word 'good' such as 'good morning'.
+At scale, these can identify missing or unusual words, and work quickly, but they lack nuance.
+For instance, an equally reasonable two-gram involving the word 'good' is 'good work'.
+However, because n-gram models do not take the surrounding text into account, they do not have the capacity to judge if 'morning' or 'work' should be suggested as the next word after 'good'.
 
-## Literature summaries
+In future work we intend to investigate more advanced language models such as word embeddings and Transformers.
+We will consider pre-trained word embedding models including Word2Vec [@mikolov2013efficient] [@mikolov2013distributed] and GloVe [@pennington2014glove], which place each word in a multi-dimensional space such that distance between words can illustrate their relationship [@bengio2003neural].
+This approach of representing words in vectors has a long history, but the implementation of @bengio2003neural has become the foundation for much subsequent work.
+For instance, @stoltz2019concept use word embeddings to identify how documents relate to concepts.
 
-(These are just here for now - we'll remove later.)
+We will also consider incorporating generative pre-trained Transformer models such as GPT-2 [@radford2019language], GPT-3 [@brown2020language], and BERT [@devlin2018bert].
+These models are based on the Transformer architecture [@vaswani2017attention].
+Proposed by @vaswani2017attention in 2017, the Transformer network architecture for neural networks has been found to outperform both recurrent neural network-based and convolutional neural network-based models in computational efficiency [@vaswani2017attention].
+Since the emergence of the Transformer model, most of the representative pre-trained models are built on this architecture [@hanretty2018comparing].
 
-### Article: Improving Language Understanding by Generative Pre-Training (2018)
-Introduction of a semi-supervised approach for language understanding machine learning tasks. The approach is the combinaiton of unsupervised pre-training and supervised fine-tuning.
+Although the level of the consistency score itself is not of major importance compared with its changes, improving the score by using more sophisticated models would be beneficial in terms of identifying incorrect words.
+In our future work we are interested in implementing different language models to compare their effectiveness in the task of generating consistency scores and providing suggested text corrections.
+We intend to use our future findings to continuously improve the `aRianna` package.
+We are also interested in expanding the notion of consistency beyond text and applying it to other datasets.
 
-### Article: Language Models are Few-Shot Learners (2020)
-Official paper from OpenAI to introduce OpenAI GPT-3. Emphasizes on the "fewer-shots" and "task-agnostic" aspects of the latest langual model compared to its predecessor GPT-2.
-
-## Discussion of how GPT-3 works
-
-**(Ke-Li - it would be good for us to go through how it actually works in as much detail as we can manage (in our own words).
-This is mostly just to force us to learn how it works and may be taken out of the final version of the paper, but it's still an important exercise.)**
+\newpage
 

### 2021-01-13 outputs/documents/paper.tex
As sent to arXiv

\documentclass[11pt,]{article}
\usepackage[left=1in,top=1in,right=1in,bottom=1in]{geometry}
\newcommand*{\authorfont}{\fontfamily{phv}\selectfont}
\usepackage[]{mathpazo}

  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}

\usepackage{abstract}
\renewcommand{\abstractname}{}    % clear the title
\renewcommand{\absnamepos}{empty} % originally center

\renewenvironment{abstract}
 {{%
    \setlength{\leftmargin}{0mm}
    \setlength{\rightmargin}{\leftmargin}%
  }%
  \relax}
 {\endlist}

\makeatletter
\def\@maketitle{%
  \newpage
%  \null
%  \vskip 2em%
%  \begin{center}%
  \let \footnote \thanks
    {\fontsize{18}{20}\selectfont\raggedright  \setlength{\parindent}{0pt} \@title \par}%
}
%\fi
\makeatother

% \setcounter{secnumdepth}{0}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs}

\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}

\title{On consistency scores in text data with an implementation in R \thanks{Thank you to Monica Alexander and Alex Luscombe for helpful comments. A Shiny application for interactive and small-scale examples is available: \url{https://kelichiu.shinyapps.io/Arianna/}. An R package for larger scale applications is available: \url{https://github.com/RohanAlexander/arianna}. Our code and datasets are available: \url{https://github.com/RohanAlexander/consistency_scores_in_text_datasets}. Comments on the 13 January 2021 version of this paper are welcome at: \href{mailto:rohan.alexander@utoronto.ca}{\nolinkurl{rohan.alexander@utoronto.ca}}.}  }

\author{\Large Ke-Li Chiu\vspace{0.05in} \newline\normalsize\emph{University of Toronto}   \and \Large Rohan Alexander\vspace{0.05in} \newline\normalsize\emph{University of Toronto}  }

\date{}

\usepackage{titlesec}

% \titleformat*{\section}{\normalsize\bfseries}
% \titleformat*{\subsection}{\normalsize\itshape}
% \titleformat*{\subsubsection}{\normalsize\itshape}
% \titleformat*{\paragraph}{\normalsize\itshape}
% \titleformat*{\subparagraph}{\normalsize\itshape}

\usepackage{natbib}
\bibliographystyle{apalike}
\usepackage[strings]{underscore} % protect underscores in most circumstances

\newtheorem{hypothesis}{Hypothesis}
\usepackage{setspace}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{hyperref}
\usepackage{amsmath}

% move the hyperref stuff down here, after header-includes, to allow for - \usepackage{hyperref}

\makeatletter
\@ifpackageloaded{hyperref}{}{%
\ifxetex
  \PassOptionsToPackage{hyphens}{url}\usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \PassOptionsToPackage{hyphens}{url}\usepackage[draft,unicode=true]{hyperref}
\fi
}

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Ke-Li Chiu (University of Toronto) and Rohan Alexander (University of Toronto)},
             pdfkeywords = {text-as-data; natural language processing; quantitative analysis; optical character recognition},  
            pdftitle={On consistency scores in text data with an implementation in R},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

% Add an option for endnotes. -----

% add tightlist ----------
\providecommand{\tightlist}{%
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% add some other packages ----------

% \usepackage{multicol}
% This should regulate where figures float
% See: https://tex.stackexchange.com/questions/2275/keeping-tables-figures-close-to-where-they-are-mentioned
\usepackage[section]{placeins}

\begin{document}

% \pagenumbering{arabic}% resets `page` counter to 1 
%
% \maketitle

{% \usefont{T1}{pnc}{m}{n}
\setlength{\parindent}{0pt}
\thispagestyle{plain}
{\fontsize{18}{20}\selectfont\raggedright 
\maketitle  % title \par

}

{
   \vskip 13.5pt\relax \normalsize\fontsize{11}{12} 
\textbf{\authorfont Ke-Li Chiu} \hskip 15pt \emph{\small University of Toronto}   \par \textbf{\authorfont Rohan Alexander} \hskip 15pt \emph{\small University of Toronto}

}

}

\begin{abstract}

    \hbox{\vrule height .2pt width 39.14pc}

    \vskip 8.5pt % \small

\noindent In this paper, we introduce a reproducible cleaning process for the text extracted from PDFs using n-gram models. Our approach compares the originally extracted text with the text generated from, or expected by, these models using earlier text as stimulus. To guide this process, we introduce the notion of a consistency score, which refers to the proportion of text that is expected by the model. This is used to monitor changes during the cleaning process, and across different corpuses. We illustrate our process on text from the book Jane Eyre and introduce both a Shiny application and an R package to make our process easier for others to adopt.

\vskip 8.5pt \noindent \emph{Keywords}: text-as-data; natural language processing; quantitative analysis; optical character recognition \par

    \hbox{\vrule height .2pt width 39.14pc}

\end{abstract}

\vskip -8.5pt

 % removetitleabstract

\noindent

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

When we think of quantitative analysis, we may like to think that our job is to `let the data speak'. But this is rarely the case in practise. Datasets can have errors, be biased, incomplete, or messy. In any case, it is the underlying statistical process, of which the dataset is an artifact, that is typically of interest. In order to use statistical models to understand that process, we typically need to clean and prepare the dataset in some way. This is especially the case when we work with text data because they are typically not designed for analysis. This cleaning and preparation requires us to make many decisions. To what extent should we correct obvious errors? What about slightly-less-obvious errors? Again, this is especially the case for text datasets because: they encode a large amount of information; an aspect may only be an error in context; and the high-dimensionality of the data at even its most basic level (26 English characters, compared with 10 numbers). Although cleaning and preparation is a necessary step, we may be concerned about the extent to which have we introduced new errors, and the possibility that we have made decisions that have affected, or even driven, our results.

In this paper we introduce the concept of consistency and implement it for a text corpus. Consistency refers to the proportion of data that are able to be forecast by a statistical model, based on preceding and surrounding data. Further, we define internal consistency as when the model is trained on the dataset itself, and external consistency as when the model is trained on a more general dataset. Together, these concepts provide a guide to the cleanliness and reasonableness of a dataset. This can be important when deciding whether a dataset is fit for purpose; when carrying out data cleaning and preparation tasks; and as a comparison between datasets. For instance, if the Canadian population is known to be 38 million from a census, but the sum of age-groups is initially 35 million, rising to 38 million only after an inputation error is corrected, then that post-cleaning dataset would have higher consistency than the pre-cleaning dataset. When applied to a text dataset, this means that we construct a model that forecasts the next word, given the previous words. If the forecast word is always the same as the actual word, then the consistency is 1. That said, it is not the absolute level of consistency that is our focus, but instead measuring how it changes during the cleaning and preparation process.

To provide an example, consider the sentence, `the cat in the\dots'. A child who has read this book could tell you that the next word should be `hat'. Hence if the sentence was actually `the cat in the bat', then that child would know something was likely wrong. The consistency score would likely be lower than if the sentence were `the cat in the hat' because we expected `hat', but actually found `bat'. After we correct this error, the consistency score would likely increase. By examining how consistency scores evolve in response to changes made to the text during the data preparation and cleaning stages we can better understand the effect of the changes. Including consistency scores and their evolution when datasets are shared allows researchers to be more transparent about the effect of their decisions in these steps of the analysis workflow. And finally, the use of consistency scores allows for an increased level of automation in the cleaning process.

We employ n-gram models to calculate consistency scores and generate word candidates for text corrections. The n-gram approach involves identifying words that are commonly found together. In our application, we use these sequences to calculate consistency scores and generate word candidates for text corrections. For instance, if we are using a sequence of three words, or tri-grams, then we identify the first two words in the sequence and evaluate if the last word in the sequence is `as expected' by comparing it to the corresponding tri-gram in a comparison dataset. The sequences we employ in this paper are a mixture of three, four, and five, consecutive words also known as tri-grams, 4-grams, and 5-grams.

The comparison dataset can be internal, meaning that the dataset is from the same source as the text being evaluated. On the other hand, a comparison dataset can be external, which means the dataset is from a different source, typically a larger and more universal dataset. The external dataset we use in the application discussed in this paper is constructed using a subset of text data sourced from Open WebText Corpus \citep{Gokaslan2019OpenWeb}. Since the content of Open WebText Corpus is from the internet, the variety of the text data is promising for meeting the generalizability and universality that an external dataset should possess, although it will also have extensive bias.

The resources that we have developed to support this paper include a Shiny app available at: \url{https://kelichiu.shinyapps.io/aRianna/}. That app computes internal and external consistency scores for corpus excerpts. We have also developed an R Package, \texttt{aRianna}, that allows our approach to be used on larger datasets, which is available at: \url{https://github.com/RohanAlexander/aRianna}.

In this paper, we introduce consistency scores and an R package that implements them. In Section \ref{background}, we provide an overview of language models and n-grams in particular. In Section \ref{arianna}, we deconstruct the package functions and provide demonstration example of the package usage. Finally, in Section \ref{discussion}, we discuss our findings, limitations, and future directions.

\hypertarget{background}{%
\section{Background}\label{background}}

\hypertarget{language-model}{%
\subsection{Language model}\label{language-model}}

Given the nature of human languages, some combinations of words tend to occur more frequently than others. Think of `good', which is more often followed by `morning', than `duck'. As such, we could consider English text production as a conditional probability, \(\mbox{Pr}(w_{k} | w^{k-1}_{1})\), where \(k\) is the number of words in a sentence, \(w_{k}\) is the predicted word, and \(w^{k-1}_{1}\) is the history of the word occurring in a sequence \citep{brown1992class}. In this way, the generation of some prediction, \(w_{k}\), is based on the history, \(w^{k-1}_{1}\). This is the underlying principle of all language models. Essentially, the goal of statistical language modeling is to estimate probability distributions over different linguistic units --- terms, words, sentences, and even documents \citep{bengio2003neural}. However, this is difficult as language is categorical. If we consider each word in a vocabulary as a category, then the dimensionality of a language dataset often becomes large quickly \citep{rosenfeld2000two}. One reason there is such a variety of statistical language models is that there are various ways of dealing with this fundamental problem. N-gram models are statistical language models that work by taking the co-occurrence of words in a sequence into account.

\hypertarget{n-gram-models}{%
\subsection{N-gram models}\label{n-gram-models}}

The foundation of an n-gram language model is the conditional probability set-up introduced above. An n-gram model is a probabilistic language model that predicts the next term in a sequence of terms \citep{bengio2003neural}. The \(n\) in n-gram refers to the number of terms in that sequence. Consider the following excerpt from \emph{Jane Eyre}: `We had been wandering'. `We' is a uni-gram, `We had' is a bi-gram, `We had been' is a tri-gram, and `We had been wandering' is a 4-gram. Notice that the two tri-grams in this excerpt, `We had been', and `had been wandering', overlap. The use of n-gram models enables us to assign probabilities to both the next sequence of words and just the next word. For instance, consider the two sentences: `We had been wandering', and `We had been wangling'. The former is likely to be more frequently encountered in a training corpus. Thus, a 4-gram would assign a higher probability to the next word being `wandering' than `wangling', given the sequence `We had been'.

To predict the next word, we have to take the sequence of preceding words into account, which requires knowing the probability of the sequence of words. The probability of a sequence appearing in a corpus follows the chain rule:
\begin{align*} 
\mbox{Pr(We, had, been, wandering)} = &  \mbox{Pr(We)} \times  \mbox{Pr(had|We)} \times  \mbox{Pr(been|We, had)} \\ 
 &  \times  \mbox{Pr(wandering|We, had, been)}.
\end{align*}

However, the likelihood that more and more words will occur next to each other in an identical sequence becomes smaller and smaller, making prediction difficult. Alternatively, we can approximate the probability of a word depending on only the previous word. This is known as the `Markov assumption' and it allows us to approximate the probability using only the last \(n\) words \citep{brown1992class}:
\[\mbox{Pr(We, had, been, wandering)} \approx \mbox{Pr(We)} \times \mbox{Pr(had|We)} \times \mbox{Pr(been|had)} \\ 
\times \mbox{Pr(wandering|been)}.\]

As a bi-gram model only considers the immediately preceding word, under the Markov assumption, an \(n\)-gram model can be reduced to a bi-gram model with \(n\) being any number:

\[\mbox{Pr}(w_{n} | w^{n-1}_{1}) \approx \mbox{Pr}(w_{n} | w_{n -1})\]

In our application, we use a model that has a mixture of tri-grams, 4-grams, and 5-grams to provide term replacement options for unexpected terms. Longer n-grams such as 5-grams capture more context and potentially generate more accurate predictions. However, there is less chance that a longer sequence of words will fit a specific combination which can lead to sparsity. Therefore, shorter n-grams should still be included to preserve the generalizability of the predictions.

Consider the idiom `elephant in the room'. If by mistake, this is written as `elephant in the roon' and correction is needed for the last word, then we can apply both a tri-gram model or a 4-gram model to predict the candidate word:

\begin{quote}
Using tri-gram model: `in the \texttt{\textless{}prediction\textgreater{}}'
\end{quote}

\begin{quote}
Using 4-gram model: `elephant in the \texttt{\textless{}prediction\textgreater{}}'
\end{quote}

In the 4-gram model, one more word, `elephant', is included, which provides more context and increased chance of `room' being accurately predicted. However, there are times when the literal location of an elephant may be being described; for instance, `elephant in the pond'. The combination of this sequence as a 4-gram might not be in a training dataset and no prediction would be available if only 4-gram models are used. Hence the use of tri-grams to provide other word candidates including `in the pond'.

\hypertarget{text-preparation-and-cleaning}{%
\subsection{Text preparation and cleaning}\label{text-preparation-and-cleaning}}

In general, text requires a great deal of preparation and cleaning before it can be analysed. For instance, \texttt{compositr} \citep{citecompositr} is an R package that runs through 13 common steps: remove non-UTF8 characters; insert missing comma spaces; replace misspelled words with most likely equivalent; remove HTML tags; convert case; remove URLs; remove numbers; remove punctuation; replace elongated words; remove extra white space; tokenization; stopword removal; and finally stem/lemmatize. And these are just some of the steps that may be taken.

Text preparation and cleaning steps and the uncertainty they bring, assume a fairly clean text dataset in the first instance. For instance, if a dataset has been created from Twitter then hashtags and emoji would need to be considered. And if the dataset had been created from OCR then there would likely be systematic errors to be addressed. One benefit of using packages such as \texttt{compositr} is that these decisions become repeatable and so their effect on subsequent analysis can be easily understood through repeated running of the full workflow. The notion of consistency provides a measure of this without the need to run the full workflow.

\hypertarget{consistency}{%
\subsection{Consistency}\label{consistency}}

Here we define, mathematically, what we mean by consistency in the case of text. Consider some text extract for which we want a consistency score. If there are \(n\) words, of which \(k\) are forecast correctly by a model, then the consistency score is the proportion: \(k/n\). We implement this in a dataframe using a binary column. Each word is a row, and a column, `correctly\_forecast', tracks the forecast, classifying each word as 1 if it was correctly forecast, and 0 otherwise. Then the internal consistency estimator is given by Equation \eqref{eq:model}:

\begin{equation} 
\label{eq:model}
c_i = \frac{\mbox{sum 'correctly_forecast' column}}{\mbox{length of dataset}}. 
\end{equation}

We distinguish between a model trained on the dataset for which consistency is being estimated, which results in an estimate of internal consistency, \(c_i\), and a model trained on another dataset, which results in an estimate of external consistency, \(c_e\), although the procedure is the same once the forecast has been made. We consider the forecast of each word independently, allowing for correlation in future work.

\hypertarget{arianna}{%
\section{\texorpdfstring{The \texttt{aRianna} R package}{The aRianna R package}}\label{arianna}}

\hypertarget{package-dependencies}{%
\subsection{Package dependencies}\label{package-dependencies}}

There are a variety of ways to implement an n-gram model within R \citep{citeR} including packages such as \texttt{quanteda} \citep{citequanteda}, \texttt{tidytext} \citep{citetidytext} and \texttt{tm} \citep{citetm}. We use \texttt{quanteda} because it has a comprehensive set of functions for conducting text analysis. We also use \texttt{tidyverse} \citep{citetidyverse} for data manipulation:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{install.packages}\NormalTok{(}\StringTok{"quanteda"}\NormalTok{)}
\KeywordTok{install.packages}\NormalTok{(}\StringTok{"tidyverse"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(quanteda)}
\KeywordTok{library}\NormalTok{(tidyverse)}
\end{Highlighting}
\end{Shaded}

\hypertarget{function-to-make-the-internal-dataset}{%
\subsection{Function to make the internal dataset}\label{function-to-make-the-internal-dataset}}

To obtain the internal consistency score, we first construct a dataset from the text that is itself the text being evaluated. The function \texttt{aRianna::make\_internal\_consistency\_dataset()} is created for this purpose. The function turns a \texttt{body\_of\_text} into a dataset to compare the text with. To do so, the function first transforms the \texttt{body\_of\_text} into word tokens with the punctuation removed and the letters turned to lowercase:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{make_internal_consistency_dataset <-}\StringTok{ }
\StringTok{  }\ControlFlowTok{function}\NormalTok{(body_of_text) \{}
\NormalTok{  tokens_from_example <-}\StringTok{ }
\StringTok{    }\NormalTok{quanteda}\OperatorTok{::}\KeywordTok{tokens}\NormalTok{(body_of_text, }\DataTypeTok{remove_punct =} \OtherTok{TRUE}\NormalTok{)}
\NormalTok{  tokens_from_example <-}\StringTok{ }
\StringTok{    }\NormalTok{quanteda}\OperatorTok{::}\KeywordTok{tokens_tolower}\NormalTok{(tokens_from_example)}
  \CommentTok{# code continues...}
\end{Highlighting}
\end{Shaded}

Next, the \texttt{aRianna::make\_internal\_consistency\_dataset()} function turns the individual tokens into tri-grams, 4-grams, and 5-grams, keeping only the n-grams that appear in the data for more than once. These common n-grams are stored in a tibble:

\begin{Shaded}
\begin{Highlighting}[]
  \CommentTok{# ...continue from previous}
  \CommentTok{# Create ngrams from the tokens}
\NormalTok{  toks_ngram <-}\StringTok{ }
\StringTok{    }\NormalTok{quanteda}\OperatorTok{::}\KeywordTok{tokens_ngrams}\NormalTok{(tokens_from_example, }\DataTypeTok{n =} \DecValTok{5}\OperatorTok{:}\DecValTok{3}\NormalTok{)}

  \CommentTok{# Convert to tibble so we can use our familiar verbs}
\NormalTok{  all_tokens <-}\StringTok{ }
\StringTok{    }\NormalTok{tibble}\OperatorTok{::}\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{tokens =}\NormalTok{ toks_ngram[[}\DecValTok{1}\NormalTok{]])}

  \CommentTok{# We only want the common ones, not every one.}
\NormalTok{  all_tokens <-}
\StringTok{    }\NormalTok{all_tokens }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{group_by}\NormalTok{(tokens) }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{count}\NormalTok{() }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{filter}\NormalTok{(n }\OperatorTok{>}\StringTok{ }\DecValTok{1}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{ungroup}\NormalTok{()}
  \CommentTok{# code continues...}
\end{Highlighting}
\end{Shaded}

Finally, the \texttt{aRianna::make\_internal\_consistency\_dataset()} function splits the n-grams into two parts: the first \(n-1\) words and the last word. As a result, a tibble is created with each n-gram containing its original sequence, the first words, and the last word:

\begin{Shaded}
\begin{Highlighting}[]
  \CommentTok{# ...continue from previous}
  \CommentTok{# Create a tibble that has the first two words in one column then the third}
\NormalTok{  all_tokens <-}
\StringTok{    }\NormalTok{all_tokens }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{tokens =}\NormalTok{ stringr}\OperatorTok{::}\KeywordTok{str_replace_all}\NormalTok{(tokens, }\StringTok{"_"}\NormalTok{, }\StringTok{" "}\NormalTok{),}
                  \DataTypeTok{first_words =}\NormalTok{ stringr}\OperatorTok{::}\KeywordTok{word}\NormalTok{(tokens, }\DataTypeTok{start =} \DecValTok{1}\NormalTok{, }\DataTypeTok{end =} \DecValTok{-2}\NormalTok{),}
                  \DataTypeTok{last_word =}\NormalTok{ stringr}\OperatorTok{::}\KeywordTok{word}\NormalTok{(tokens, }\DecValTok{-1}\NormalTok{),}
                  \DataTypeTok{tokens =}\NormalTok{ stringr}\OperatorTok{::}\KeywordTok{str_replace_all}\NormalTok{(tokens, }\StringTok{" "}\NormalTok{, }\StringTok{"_"}\NormalTok{),}
                  \DataTypeTok{first_words =}\NormalTok{ stringr}\OperatorTok{::}\KeywordTok{str_replace_all}\NormalTok{(first_words, }\StringTok{" "}\NormalTok{, }\StringTok{"_"}\NormalTok{)}
\NormalTok{    ) }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{rename}\NormalTok{(}\DataTypeTok{last_word_expected =}\NormalTok{ last_word) }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{n)}
  \ErrorTok{\}}
\end{Highlighting}
\end{Shaded}

\hypertarget{function-to-generate-internal-consistency}{%
\subsection{Function to generate internal consistency}\label{function-to-generate-internal-consistency}}

After we have the internal consistency dataset, we need a function to compare the text being evaluated and the internal dataset in order to retrieve the internal consistency score. We define the \texttt{aRianna::generate\_internal\_consistency\_score} function to do this. Similar to the \texttt{aRianna::make\_internal\_consistency\_dataset} function, this function first transforms the text being evaluated into tri-grams, 4-grams, and 5-grams tokens, then splits the n-grams to their first words and last words:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{generate_internal_consistency_score <-}
\StringTok{  }\ControlFlowTok{function}\NormalTok{(text_to_check, consistency_dataset)\{}
  \CommentTok{# Create tokens with errors}
\NormalTok{  tokens_from_example_with_errors <-}\StringTok{ }
\StringTok{    }\NormalTok{quanteda}\OperatorTok{::}\KeywordTok{tokens}\NormalTok{(text_to_check, }\DataTypeTok{remove_punct =} \OtherTok{TRUE}\NormalTok{)}
\NormalTok{  tokens_from_example_with_errors <-}\StringTok{ }
\StringTok{    }\NormalTok{quanteda}\OperatorTok{::}\KeywordTok{tokens_tolower}\NormalTok{(tokens_from_example_with_errors)}

  \CommentTok{# Create ngrams from the tokens with errors}
\NormalTok{  toks_ngram_with_errors <-}\StringTok{ }
\StringTok{    }\NormalTok{quanteda}\OperatorTok{::}\KeywordTok{tokens_ngrams}\NormalTok{(tokens_from_example_with_errors, }\DataTypeTok{n =} \DecValTok{5}\OperatorTok{:}\DecValTok{3}\NormalTok{)}

\NormalTok{  all_tokens_with_errors <-}\StringTok{ }
\StringTok{    }\NormalTok{tibble}\OperatorTok{::}\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{tokens =}\NormalTok{ toks_ngram_with_errors[[}\DecValTok{1}\NormalTok{]])}

\NormalTok{  all_tokens_with_errors <-}
\StringTok{    }\NormalTok{all_tokens_with_errors }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{ngram =} \KeywordTok{sapply}\NormalTok{(}\KeywordTok{strsplit}\NormalTok{(tokens, }\StringTok{"_"}\NormalTok{), length),}
                  \DataTypeTok{tokens =}\NormalTok{ stringr}\OperatorTok{::}\KeywordTok{str_replace_all}\NormalTok{(tokens, }\StringTok{"_"}\NormalTok{, }\StringTok{" "}\NormalTok{),}
                  \DataTypeTok{first_words =}\NormalTok{ stringr}\OperatorTok{::}\KeywordTok{word}\NormalTok{(tokens, }\DataTypeTok{start =} \DecValTok{1}\NormalTok{, }\DataTypeTok{end =} \DecValTok{-2}\NormalTok{),}
                  \DataTypeTok{last_word =}\NormalTok{ stringr}\OperatorTok{::}\KeywordTok{word}\NormalTok{(tokens, }\DecValTok{-1}\NormalTok{),}
                  \DataTypeTok{tokens =}\NormalTok{ stringr}\OperatorTok{::}\KeywordTok{str_replace_all}\NormalTok{(tokens, }\StringTok{" "}\NormalTok{, }\StringTok{"_"}\NormalTok{),}
                  \DataTypeTok{first_words =}\NormalTok{ stringr}\OperatorTok{::}\KeywordTok{str_replace_all}\NormalTok{(first_words, }\StringTok{" "}\NormalTok{, }\StringTok{"_"}\NormalTok{))}
  \CommentTok{# code continues...}
\end{Highlighting}
\end{Shaded}

Next, the function uses \texttt{dplyr::left\_join()} to combine the tibble of text tokens and the internal consistency tibble. The \texttt{dplyr::left\_join()} function returns all the rows from the text tokens tibble and all the columns from both tibbles \citep{2018dplyr}. By combining the two tibbles, the function generates two additional columns --- the \texttt{last\_word} column are the last words of all n-grams, and the \texttt{last\_word\_expected} contains the last words that are in the internal consistency dataset. By comparing the words in \texttt{last\_word} and \texttt{last\_word\_expected}, we can identify the `unexpected' words. If a last word is in \texttt{last\_word} but not \texttt{last\_word\_expected}, it is regarded as an unexpected word and the corresponding word in \texttt{last\_word\_expected} serves as a replacement candidate:

\begin{Shaded}
\begin{Highlighting}[]
  \CommentTok{# ...continue from previous}
\NormalTok{all_tokens_with_errors <-}
\StringTok{    }\NormalTok{all_tokens_with_errors }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{left_join}\NormalTok{(dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(consistency_dataset, }\OperatorTok{-}\NormalTok{tokens), }
                     \DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"first_words"}\NormalTok{))}
  \CommentTok{# code continues...}
\end{Highlighting}
\end{Shaded}

The consistency score is calculated by the number of words that have been predicted by the model divided by the total number of words in the data. The function identifies the counts of words that are predicted by the model and divided by the total number of words in the input text:

\begin{Shaded}
\begin{Highlighting}[]
  \CommentTok{# ...continue from previous}
  \CommentTok{# Calculate the internal consistency score:}
\NormalTok{  false_count <-}\StringTok{ }
\StringTok{    }\KeywordTok{length}\NormalTok{(internal_consistency[}\KeywordTok{which}\NormalTok{(internal_consistency}\OperatorTok{$}\NormalTok{as_expected }\OperatorTok{==}\StringTok{ }\OtherTok{FALSE} \OperatorTok{&}\StringTok{ }
\StringTok{                                        }\NormalTok{internal_consistency}\OperatorTok{$}\NormalTok{ngram }\OperatorTok{==}\StringTok{ }\DecValTok{3}\NormalTok{)])}
\NormalTok{  word_count <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(}\KeywordTok{strsplit}\NormalTok{(text_to_check, }\StringTok{" "}\NormalTok{), length)}
\NormalTok{  true_count <-}\StringTok{ }\NormalTok{word_count }\OperatorTok{-}\StringTok{ }\NormalTok{false_count}

\NormalTok{  internal_consistency <-}\StringTok{ }\NormalTok{tibble}\OperatorTok{::}\KeywordTok{tibble}\NormalTok{(}
    \StringTok{"as_expected"}\NormalTok{ =}\StringTok{ }\NormalTok{true_count,}
    \StringTok{"unexpected"}\NormalTok{  =}\StringTok{ }\NormalTok{false_count,}
    \StringTok{"consistency"}\NormalTok{ =}\StringTok{ }\NormalTok{true_count}\OperatorTok{/}\NormalTok{word_count}
\NormalTok{  )}
  \CommentTok{# code continues...}
\end{Highlighting}
\end{Shaded}

The \texttt{aRianna::generate\_internal\_consistency\_score} function also lists the identified text errors and generates replacement candidates to the text errors:

\begin{Shaded}
\begin{Highlighting}[]
  \CommentTok{# ...continue from previous}
  \CommentTok{# Identify which words were unexpected}
\NormalTok{  unexpected <-}
\StringTok{    }\NormalTok{all_tokens_with_errors_only }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{as_expected =}\NormalTok{ last_word }\OperatorTok{==}\StringTok{ }\NormalTok{last_word_expected) }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{filter}\NormalTok{(as_expected }\OperatorTok{==}\StringTok{ }\OtherTok{FALSE}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{tokens, }\OperatorTok{-}\NormalTok{ngram, }\OperatorTok{-}\NormalTok{as_expected)}
  \ErrorTok{\}}
\end{Highlighting}
\end{Shaded}

\hypertarget{function-to-generate-external-consistency}{%
\subsection{Function to generate external consistency}\label{function-to-generate-external-consistency}}

The \texttt{aRianna::generate\_external\_consistency\_score} function works identically. The only difference is that it compares the text being evaluated with an external consistency dataset that is larger and more general. The external consistency dataset we employed in the application is constructed using a subset of text data from Open WebText Corpus \citep{Gokaslan2019OpenWeb}. In general, any dataset could be used.

\hypertarget{demonstration}{%
\subsection{Demonstration}\label{demonstration}}

To install the package and load the library:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{devtools}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"RohanAlexander/arianna"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(aRianna)}
\end{Highlighting}
\end{Shaded}

For this demonstration, we use the first paragraph of Jane Eyre as the internal text data. A sentence within the paragraph is modified with the intention to contain an error and serves as the text to be evaluated:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{body_of_text <-}\StringTok{ "There was no possibility of taking a walk that day. }
\StringTok{  We had been wandering, indeed, in the leafless shrubbery an hour in }
\StringTok{  the morning; but since dinner (Mrs. Reed, when there was no company, }
\StringTok{  dined early) the cold winter wind had brought with it clouds so sombre, }
\StringTok{  and a rain so penetrating, that further out-door exercise was now out }
\StringTok{  of the question."}

\NormalTok{text_to_evaluate <-}\StringTok{ "when there was na company"}
\end{Highlighting}
\end{Shaded}

The first step is to turn the body of text into the internal consistency dataset. The generated internal consistency dataset is a tibble that contains the tri-grams, 4-grams and 5-grams that appear in the internal text data more than once. Since only `there\_was\_no' has more than one occurrence, it is the only n-gram in the internal consistency dataset:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{internal_consistency_dataset <-}\StringTok{ }
\StringTok{  }\NormalTok{aRianna}\OperatorTok{::}\KeywordTok{make_internal_consistency_dataset}\NormalTok{(body_of_text)}
\NormalTok{internal_consistency_dataset}
\CommentTok{## # A tibble: 1 x 3}
\CommentTok{##   tokens       first_words last_word_expected}
\CommentTok{##   <chr>        <chr>       <chr>             }
\CommentTok{## 1 there_was_no there_was   no                }
\end{Highlighting}
\end{Shaded}

The next step is to compare the text to be evaluated with the internal consistency dataset that we created in the previous step. The function \texttt{aRianna::generate\_internal\_consistency\_score()} takes two arguments: the text to evaluate, and the internal consistency dataset. The function identifies the word `na' as an unexpected word and generates the internal consistency score as 0.8. The function also lists `no' as the replacement of `na':

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{aRianna}\OperatorTok{::}\KeywordTok{generate_internal_consistency_score}\NormalTok{(}
\NormalTok{  text_to_evaluate, internal_consistency_dataset)}

\CommentTok{## $`internal consistency`}
\CommentTok{##   as_expected unexpected consistency}
\CommentTok{## 1           4          1         0.8}

\CommentTok{## $`unexpected words`}
\CommentTok{## # A tibble: 1 x 3}
\CommentTok{##   first_words last_word last_word_expected }
\CommentTok{##   <chr>       <chr>     <chr>                 }
\CommentTok{## 1 there_was   na        no                 }
\end{Highlighting}
\end{Shaded}

To get the external consistency score, we use the \texttt{aRianna::generate\_external\_consistency\_score()} function. The function takes only one argument, the text to evaluate, and compares it with the external consistency dataset. Here, the text to evaluate is `there was no possibiliti', with the word ``possibility'' being wrongly spelled. The function identifies the word `possibiliti' as an unexpected word, and generates the external consistency score 0.75. Because the external dataset is larger, it is capable of providing more replacement candidates for `possibiliti'. There are no replacements based on 5-grams, so the replacements based on 4-grams are displayed first, followed by the replacements based on tri-grams:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{text_to_evaluate <-}\StringTok{ "there was no possibiliti"}
\NormalTok{aRianna}\OperatorTok{::}\KeywordTok{generate_external_consistency_score}\NormalTok{(text_to_evaluate)}

\CommentTok{## aRianna::generate_external_consistency_score(text_to_evaluate)}
\CommentTok{## $`external consistency`}
\CommentTok{## # A tibble: 1 x 3}
\CommentTok{##   as_expected unexpected consistency}
\CommentTok{##         <int>      <int>       <dbl>}
\CommentTok{## 1           3          1        0.75}

\CommentTok{## $`unexpected words`}
\CommentTok{## # A tibble: 11 x 3}
\CommentTok{##    first_words  last_word    last_word_expected }
\CommentTok{##    <chr>        <chr>        <chr>                    }
\CommentTok{##  1 there_was_no possibiliti  evidence          }
\CommentTok{##  2 there_was_no possibiliti  immediate         }
\CommentTok{##  3 there_was_no possibiliti  infrastructure    }
\CommentTok{##  4 there_was_no possibiliti  one               }
\CommentTok{##  5 there_was_no possibiliti  possibility       }
\CommentTok{##  6 there_was_no possibiliti  sound             }
\CommentTok{##  7 there_was_no possibiliti  way               }
\CommentTok{##  8 was_no       possibiliti  good              }
\CommentTok{##  9 was_no       possibiliti  longer            }
\CommentTok{## 10 was_no       possibiliti  more              }
\CommentTok{## 11 was_no       possibiliti  wonder }
\end{Highlighting}
\end{Shaded}

The choice of what do now depends on a variety of factors including the purpose of the analysis, and the size of the corpus. One option is to consider only a subset of potential replacements, or to consider only replacements over a certain probability threshold. These are parameters that can be tuned in more sophisticated language models. The choice of which word to change `possibili' to could affect the consistency score if it is not what the model expects. Future work, with more sophisticated language models, could allow tuning in this regard.

As the dataset increases in size, the number of possible replacements will also likely increase. Various options for decreasing the number of possible replacements exist, starting with string distance algorithms and increasing in complexity. It is important to recognise that we are focused on the fact that they were not forecast, rather than what they should be replaced with. However, improving this aspect helps to make the consistency score more relevant and useful, especially in text settings where the cleaning and preparation aspect is likely to be iterative.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Outside of textbook examples, almost all real-world datasets need to be cleaned and prepared before they can be analyzed. While there are many tools to help with reproducibility of analysis, options for ensuring reproducibility in earlier stages of a typical data science workflow are more limited. Our approach discussed in this paper can help. Our notion of consistency can be tracked at many points while the dataset is being cleaned and prepared. Stages in which it changes substantially can be investigated. While there may be good reasons for such as change, by identifying them it is easier for other researchers to understand the effects of cleaning and preparation on eventual analysis results.

Our approaches are applicable to any dataset, but in this paper we applied them to one based on text. The high-dimensionality of text, and the messiness of it when it comes from real-world sources, means that often considerable cleaning and preparation is done to text datasets. We demonstrated how sequences of words can be used to forecast the very next word. That forecast can be compared with the actual word observed and this difference expressed as a proportion, that we call a `consistency score'.

The language model that we used is n-grams. Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation \citep{brown1992class}. As demonstrated in our application, we use a n-gram based model to calculate internal and external consistency scores and generate a list of words as text correction candidates. However, n-gram models do not take the linguistic structure of language, nor broader context, into account. For instance, \citet[p.~1]{rosenfeld2000two} discusses language in this context, saying that `\ldots it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'. Hence next-word prediction using n-gram-based language models can be limited. Here think of a two-gram involving the word `good' such as `good morning'. At scale, these can identify missing or unusual words, and work quickly, but they lack nuance. For instance, an equally reasonable two-gram involving the word `good' is `good work'. However, because n-gram models do not take the surrounding text into account, they do not have the capacity to judge if `morning' or `work' should be suggested as the next word after `good'.

In future work we intend to investigate more advanced language models such as word embeddings and Transformers. We will consider pre-trained word embedding models including Word2Vec \citep{mikolov2013efficient} \citep{mikolov2013distributed} and GloVe \citep{pennington2014glove}, which place each word in a multi-dimensional space such that distance between words can illustrate their relationship \citep{bengio2003neural}. This approach of representing words in vectors has a long history, but the implementation of \citet{bengio2003neural} has become the foundation for much subsequent work. For instance, \citet{stoltz2019concept} use word embeddings to identify how documents relate to concepts.

We will also consider incorporating generative pre-trained Transformer models such as GPT-2 \citep{radford2019language}, GPT-3 \citep{brown2020language}, and BERT \citep{devlin2018bert}. These models are based on the Transformer architecture \citep{vaswani2017attention}. Proposed by \citet{vaswani2017attention} in 2017, the Transformer network architecture for neural networks has been found to outperform both recurrent neural network-based and convolutional neural network-based models in computational efficiency \citep{vaswani2017attention}. Since the emergence of the Transformer model, most of the representative pre-trained models are built on this architecture \citep{hanretty2018comparing}.

Although the level of the consistency score itself is not of major importance compared with its changes, improving the score by using more sophisticated models would be beneficial in terms of identifying incorrect words. In our future work we are interested in implementing different language models to compare their effectiveness in the task of generating consistency scores and providing suggested text corrections. We intend to use our future findings to continuously improve the \texttt{aRianna} package. We are also interested in expanding the notion of consistency beyond text and applying it to other datasets.

\newpage

\newpage
\singlespacing 
\renewcommand\refname{References}
\bibliography{references.bib}

\end{document}
